### 2015-10-24 docs/main.tex
first commit

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[11pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
%\usepackage{natbib}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xspace}
% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

\usepackage{sectsty}
\subsectionfont{\normalsize}
\subsubsectionfont{\mdseries\itshape}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}
\usepackage{caption}
\usepackage{subcaption}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{/Users/stephens/Dropbox/Documents/stylefiles/plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **
\usepackage{enumerate}
\usepackage{multirow} 
\usepackage{url}
\usepackage{xr} %for cross-referencing
%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW
\newtheorem{algorithm}{Algorithm}
\newtheorem{proposition}{Proposition}
\newtheorem{restateproposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{result}{Result}
\newtheorem{note}{Note}
\newtheorem{definition}{Definition}

\def\lfdr{\textit{lfdr}}
\def\lfsr{\textit{lfsr}}
\def\x{\mbf{x}}
\def\y{\mbf{y}}
\def\e{\mbf{e}}
\def\g{\mbf{g}}
\def\p{p}
\def\Yb{\hat{Y}_1^{\text{BAYES}}}
\def\Xperp{X_{1 \perp 0}}
\def\Pperp{P_{1 \perp 0}}
\def\Pperpb{\Pperp^{\text{B}}}
\def \hatd { \widehat{ d } }
\def \dhatin {\hatd^{\mbox{\tiny IN}}}
\def \dhatsw {\hatd^{\mbox{\tiny SW}}}
\def \mvn{\text{MVN}}
\def\mn{\text{MN}}
\def \iw{\text{W}^{-1}}
\def \fp {fastPHASE\ }
\def\bhat{\hat{\beta}}
\def\shat{\hat{s}}
\def\var{\rm Var}
\def\mvr{\text{BMVR}}
\def\bfuni{\text{BF}_\text{uni}}
\def\one{\mbf{1}}
\def\zero{\mbf{0}}
\def\hall{H_{\text{all}}}
\def\pall{\p_{\text{all}}}
\def\bfall{\text{BF}_\text{all}}
\def\bfav{\text{BF}_\text{av}}
\def\gall{\gamma_\text{all}}
\def\limbfall{\bfall^\rightarrow}
\def\limbfav{\bfav^\rightarrow}
\def\limbfgamma{\BF_\gamma^\rightarrow}
\def\sxx{S_{xx}}
\def\syy{S_{yy}}
\def\vyy{V_{yy}}
\def\vyx{V_{yx}}
\def\vxx{{V_{xx}}}
\def\rss{\text{\rm RSS}}
\def\dss{\text{\rm $\Delta$SS}}
\def\U{U}
\def\D{D}
\def\I{I}
\def\luu{L_{UU}}
\def\indep{\perp}
\def\qvalue{{\tt qvalue}\xspace}
\def\locfdr{{\tt locfdr}\xspace}
\def\mixfdr{{\tt mixfdr}\xspace}
\def\ashr{{\tt ashr}\xspace}

\newcommand{\mbf}[1]{\mbox{\boldmath$#1$}} % true mathbf works in math
                                % mode too
\newcommand\height{{\it height}}
\newcommand\weight{{\it weight}}
\def\BF{\text{BF}}

%% END MACROS SECTION
%\externaldocument{plos_one_rev1_suppinfo}

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{False Discovery Rates: A New Deal}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matthew Stephens$^{1*}$, 
\\
\bf{1} Department of Statistics and Department of Human Genetics, University of Chicago, Chicago, IL, USA
\\
$\ast$ E-mail: Corresponding mstephens@uchicago.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

We introduce a novel Empirical Bayes approach for performing large-scale hypothesis testing, including  
estimating False Discovery Rates (FDRs), and estimating effect sizes. Compared with
existing commonly-used approaches to FDR analysis, the method has two key differences. First, it
assumes that the distribution of the actual (unobserved) effects being tested is unimodal, with a mode at 0.
This ``unimodal assumption" (UA), which is common in other related contexts, but very different from assumptions usually made
in FDR analyses, yields more accurate estimates of FDR than existing methods, provided the UA holds.
Second, the method takes as it input two numbers for each test (an effect size estimate, and corresponding standard error),
rather than the one number usually used ($p$ value, or $z$ score). Provided these two numbers are available,
using them allows our method to better account for
variations in measurement precision across tests. It also facilitates the estimation of actual 
effect sizes, and our approach provides ``shrunken" interval estimates (credible regions) for each effect in addition to measures of ``significance".
To provide a bridge between interval estimates and signicance measures we introduce the term ``local false sign rate"
to refer to the probability of getting the sign of an effect wrong, and argue that it is a superior
measure of significance than the local FDR because it is both more generally applicable, and
can be more robustly estimated. Our methods are implemented in an R package {\tt ashr}
available from \url{http://github.com/stephens999/ashr}.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
%\section*{Author Summary}

\section*{Introduction}

Since its introduction in 1995 by Benjamini and Hochberg 
\cite{benjamini1995controlling}, the ``False Discovery Rate" (FDR) has quickly established itself
as a key concept in modern statistics, and the primary tool by which most practitioners handle large-scale multiple testing
in which the goal is to identify the non-zero ``effects" among a large number of imprecisely-measured effects.

Here we consider an Empirical Bayes (EB) approach to FDR. This idea is, of course, far from new:
indeed, the notion that EB approaches could be helpful in handling multiple comparisons predates introduction of
the FDR  (e.g.~\cite{greenland1991empirical}). More recently, EB approaches to the FDR
have been extensively studied by several authors, 
especially B.~Efron and co-authors 
\cite{efron2001empirical, efron2002empirical,efron2003robbins,efron2008microarrays,efron2010large}; 
see also \cite{kendziorski2003parametric,newton2004detecting, 
datta2005empirical,muralidharan2010empirical} for example.
So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
existing widely-used FDR pipelines) and can substantially affect
inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
 and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.

In addition to these two new ideas, we highlight a third idea that is old, but which remains under-used in practice:
the idea that it may be preferable to focus on estimation rather than on testing.
In principle, Bayesian approaches can naturally unify testing and estimation into a single framework -- testing is
simply estimation with some positive prior probability that the effect is exactly zero.
However, despite ongoing interest in this area from both frequentist \cite{benjamini2005false} and Bayesian \cite{zhao2012empirical,gelman2012we} 
perspectives, in practice large-scale studies that assess many effects almost invariably focus on testing significance and
controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the $\lfsr$, particularly its increased robustness to modelling assumptions.

%terminology
%``signed discovery" for a discovery (an effect declared to be non-zero) for which the sign of the effect (positive or negative) is declared,
%and correspondingly define the ``false signed discovery rate", or simply ``false sign rate" (FSR), in a natural way.

%The methods we provide here can address both problems, and we provide results that highlight important benefits of focussing on estimation, and particularly on assessing the correctness of the {\it sign} of the estimated effects, rather than on testing for non-zero effects.

Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
(The names comes from the fact that EB methods in general, and -- due to the UA -- our EB method in particular, is a form of shrinkage estimation,
and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation. 
This will be discussed further in Xing and Stephens (in preparation).)

%Measurement precision will often vary among genes --  for example, log-fold change estimate for highly expressed genes may have greater precision than those for lower-expressed genes -- and one goal  of our methods is to allow for thiis variation.

%Let $\mu^0_j$ and $\mu^1_j$ denote the mean expression 
%of gene $j$ ($j=1,\dots,J$) in the two conditions, and define $\beta_j:= \mu^0_j - \mu^1_j$ to be the difference. Typically expression
%measurements are made on only a small number of
%samples in each condition - sometimes as few as one
%sample in each condition. Thus the error in estimates
%of $\mu^0_j$ and $\mu^1_j$ is appreciable, and the error
%in estimates of $\beta_j$ still greater.
%

%Many tools have been developed for estimating or controlling the FDR and related quantities \cite{}, and some are widely used.
%Here we present new tools that bring two important ideas to bear on the problem.
%First, these tools are designed to take account of variations in the {\it measurement precision} of different effects, reducing the influence of low-precision measurements.
%For example, in the example of detecting differentially expressed genes, measurements of log fold change for high-expressed genes 
%may have better precision than those for low-expressed genes; by incorporating this information our methods can avoid the tendancy for 
%imprecise measurements at low-expressed genes to dilute the signal at higher-expressed genes.
%Second, our tools explicitly assume that the non-zero effects have a unimodal distribution centered at zero. This assumption 
%is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
%demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
%a substantial increase in the number of discoveries at a given FDR).

%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
%As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero. Such problems are usually tackled by computing a $p$ value for each $j$, often by applying a $t$ test to $\bhat_j/s_j$,
%and then applying a generic procedure, such as that of \cite{benjamini1995controlling} or \cite{storey2003statistical}, designed to control or
%estimate the false discovery rate (FDR) or the positive FDR \cite{storey.03}. In essence we aim to provide analagous
%generic methods that work directly with two numbers for each 
%measurement $(\bhat_j,s_j$), rather than a single number (e.g.~ the $p$ value, or $t$ statistic). Working with these two numbers has two important benefits: first, it permits estimation and not only testing; second, the 
%uncertainty in each measurement $\bhat_j$ can be more fully accounted for, reducing the impact of ``high-noise" measurements (large $s_j$) that can reduce the effectiveness of a standard FDR analysis.

\def\df{df}
\def\FDR{\text{FDR}}
\def\fdr{\text{\lfdr}}
\def\FSR{\text{FSR}}
\def\fsr{\text{lfsr}}

 \section*{Methods}

 \subsection*{Model Outine}

% Here we consider one setting in which FDR-based methods are widely used:
%an experimenter has measured a large numbers of ``effects", each with an associated
%measurement precision (standard error), 
%and the goal is to identify which of the effects we can declare to be non-zero, with some prescribed level of confidence.
%Although the usual focus of FDR-based methods is on distinguishing zero and non-zero effects,
%here we also focus on providing estimates -- both point estimates and interval estimates -- for these effects.
%This setting arises commonly in many areas of statistics, and particularly in genomics applications.
%For example, in genomics
%a common goal is to identify ``differentially-expressed genes" -- that is, genes
%whose mean expression (activity) level differs in two conditions. Here the ``effect" of interest is the difference in (log) 
%mean expression level between the two conditions, often referred to as the ``log-fold change". 
Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.

Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
more general problem of estimating, and assessing uncertainty in, $\beta_j$.

Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
which by Bayes theorem can be written as 
\begin{equation}
p(\beta | \bhat, \shat) \propto p(\beta | \shat) p(\bhat | \beta, \shat).
\end{equation}

For $p(\beta | \shat)$ we assume that the $\beta_j$ are independent from a unimodal distribution $g$.
This unimodal assumption (UA) is a key assumption that distinguishes our approach from previous EB approaches to FDR analysis.
A simple way to implement the UA is to assume that $g$ is
a mixture of a point mass at 0 and a mixture of {\it zero-mean} normal distributions:
\begin{align} \label{eqn:beta}
p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\   
g(\cdot; \pi) & = \pi_0 \delta_0(\cdot) + \sum_{k=1}^K \pi_k N(\cdot; 0, \sigma_k^2), \label{eqn:mixnorm}
\end{align}
where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
although of course in practice we use a finite grid -- see Implementation Details.)

For the likelihood $p(\bhat | \beta, \shat)$ we assume  
\begin{equation} \label{eqn:normlik}
p(\bhat | \beta, \shat) = \prod_j N(\bhat_j; \beta_j, \shat_j^2).
\end{equation}
Here, in additional to some conditional independence assumptions, we are 
effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.

This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
However, this simple model also has several additional assumptions that can be relaxed. Specifically,
\begin{enumerate}
\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
\item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation. We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
\end{enumerate}
These embellishments are detailed in Detailed Methods.
Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.

%We assume that the observations $\bhat$ are conditionally independent given $\beta,\shat$, and that the distribution of
%$\bhat_j$ depends only on the corresponding $\beta_j,\shat_j$, so
%\begin{equation}
%p(\bhat | \beta, \shat) = \prod_j p(\bhat_j | \beta_j, \shat_j).
%\end{equation}
%Initially we assume that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation,
%\begin{equation} \label{eqn:normlik}
%p(\bhat_j | \beta_j, \shat_j) =  N(\bhat_j; \beta_j, \shat^2_j),
%\end{equation}
%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and variance $\sigma^2$.
%
%We assume that the (unobserved) effects $\beta_j$ come from some distribution $g$, which is to be estimated, and which we write 
%as a mixture of null and alternative components:
%\begin{equation} \label{eqn:betaj0}
%g(\cdot) = \pi_0 \delta_0(\cdot) + (1-\pi_0) g_1(\cdot)
%\end{equation}
%where $\pi_0$ denotes the proportion of effects that are null, $\delta_0(\cdot)$ denotes a point mass at 0, and $g_1(\cdot)$ denotes the distribution of the non-zero $\beta_j$.
%We also assume that (given $g$) the $\beta_j$ are independent, and independent of $\shat$ (a point we return to later), so
%\begin{equation} \label{eqn:beta-iid}
%p(\beta | \shat) = \prod_j g(\beta_j).
%\end{equation}
%
%Our key further assumption, which distinguishes our approach from previous EB approaches to FDR analysis, is that $g_1$ is unimodal,
%with a mode at 0. One simple way to implement this ``Unimodal Assumption" (UA)  is
%to use a mixture of zero-mean normal distributions:
%\begin{equation} \label{eqn:mixnorm}
%g_1(\cdot) = \sum_{k=1}^K w_k N(\cdot; 0, \sigma_k^2)
%\end{equation}
%where $N(\cdot; \mu,\sigma^2)$ denotes the density of a normal distribution with mean $\mu$ and standard deviation $\sigma$.
%Here the $w_k$ are non-negative mixture weights that sum to $1$ and are to be estimated, and the component standard deviations $\sigma_1,\dots,\sigma_K$ 
%represent a large and dense grid of fixed positive numbers spanning a range from very small to very big (so $K$ is fixed and large).
% The symmetry of the normal distribution implies that (\ref{eqn:mixnorm}) is not only unimodal, but also symmetric about zero. We introduce
% more flexible alternatives based on using mixtures of uniform distributions, and further details (such as choice of gridpoints $\{\sigma_k:k=1,\dots,K\}$) below. 
% 
% Substituting (\ref{eqn:mixnorm}) for $g_1$ in (\ref{eqn:betaj0}), and defining $\pi_k := (1-\pi_0)w_k$ for $k=1,\dots,K$ (and also $\sigma_k:=0$, and $N(\cdot;0,0):=\delta_0(\cdot)$)
% yields a notationally-convenient form for $g$, with hyperparameters $\pi:=(\pi_0,\pi_1,\dots,\pi_K)$ whose elements are non-negative and sum to 1:
% \begin{equation} \label{eqn:g}
% g(\cdot; \pi) = \sum_{k=0}^K \pi_k N(\cdot; 0, \sigma_k^2).
% \end{equation}
% 
% Putting all this together, and making the dependence on hyperparameters $\pi$ explicit, we have:
% \begin{align} \label{eqn:betaj}
% p(\beta | \shat, \pi) & = \prod_j g(\beta_j; \pi), \\  \label{eqn:normlik2}
% p(\bhat | \beta, \shat, \pi) &= \prod_j N(\bhat_j; \beta_j, \shat_j^2), 
% \end{align}
% with $g$ given by the mixture of normals $(\ref{eqn:g})$.

\subsection*{Fitting the model}

In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
\begin{equation} \label{eqn:marginallik}
p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
\end{equation}
where we define $\sigma_0:=0$.

The usual EB approach to fitting this model would involve two simple steps:
\begin{enumerate}
\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
\item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
\end{enumerate}
Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)

Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.

\subsection*{The local False Discovery Rate and local False Sign Rate}

As noted above, the posterior distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ have a simple analytic form.
In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
In other words it is the posterior probability that $\beta_j$ is actually zero:
\begin{equation} \label{eqn:fdr}
\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
\end{equation}

The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$. Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
\begin{quote}
...the more meaningful question: ``is the evidence strong enough to support a belief that the observed difference has the correct sign?"
\end{quote}
Along the same lines, Gelman and co-authors \cite{gelman2000type, gelman2012we} suggest  
focussing on ``type S errors", meaning errors in sign, rather than the more traditional type I errors.

Motivated by these suggestions, we define the ``local False Sign Rate" for effect $j$, $\lfsr_j$, to be the probability that we
would make an error in the sign of effect $j$ if we were forced to declare it either positive or negative. Specifically,
\begin{equation} \label{eqn:lfsr}
\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
\end{equation}
To illustrate, suppose that 
\begin{gather*}
p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
\end{gather*} 
Then 
from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.

As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
(this follows from the definition because both the events $\beta_j \geq 0$
and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.

From these ``local" measures of significance, we can also compute average error rates over subsets of observations
$\Gamma \subset \{1,\dots,J\}$. For example,   
\begin{equation} \label{eqn:FDRhat}
\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
\end{equation}
estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
And 
\begin{equation} \label{eqn:qvalue}
q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
\end{equation}
provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.

%Because our methods involve explicitly modelling effects $\beta_j$, rather than 
%$p$ values or $z$ scores, they can provide estimates of the local false sign rates.

%Indeed, even in settings where some
%$\beta_j$ are exactly zero, separately identifying which are positive and which negative may be useful. For example, when identifying differentially expressed genes,
%analysts will often separate the genes that are ``upregulated" from those that ``downregulated" in a condition.

\subsection*{Related work}

\subsubsection*{Previous approaches focussed on FDR}

Among previous methods that explicitly consider the FDR and related quantities,
our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
 (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
  %Here we highlight the key differences between our approach and these three existing approaches.

There are two key differences between our approach and all of these three existing methods. First, 
whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
how it can improve performance by taking better account of variation in measurement precision among observations.
%(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
% Or more generally Section xx illustrates the potential practical effects

%allows for different modelling assumptions to be compared through likelihood comparisons.
Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}). Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
from the UA as they can be.

Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
decreasingly plausible, and so the distribution of the effects will be unimodal about 0. To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero). It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.

%Tukey "peri-null"

%Interestingly, although the unimodal assumption seems natural, much previous related work on estimating False Discovery Rates  
%has assumed, either explicitly or implicitly, that the distribution of effects is multimodal - see Figure \ref{fig:ZA}.

In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
The UA assumption on regression coefficients is directly analagous to our UA here, and so its
widespread use in the regression context supports its use here.

Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).

Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
and statistically stable. We illustrate these features in Results.

\subsubsection*{Other work}

There is also a very considerable literature that does not directly focus on the FDR problem, but which involves similar ideas and methods. 
 Among these, a paper about deconvolution \cite{cordy1997deconvolution} is most similar, methodologically, to our work here:
indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
Also, they provide no software implementation.  More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic". Some pointers to recent papers in which other relevant
citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}. Much of the literature focusses on the homoskedastic case (i.e. $\shat_j$ all equal) whereas we allow for heteroskedasticity. And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential. Several recent papers consider more flexible non-parametric assumptions on 
$g$ than the UA assumption we make here. In particular, \cite{jiang2009general,koenker2014convex} consider the unconstrained non-parametric maximum likelihood
estimate (NPMLE) for $g$. These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive. However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
Indeed, one way of thinking about the UA is as a way to regularize $g$.

%The conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$ 
%encapsulate uncertainty in $\beta_j$, combining information across {\it all} observations 
%$j=1,\dots,J$. Specifically, the combining of the information occurs through estimation of
%$\pi$, which involves all of the data. Importantly, when estimating $\pi$ our approach accounts for
%the precision of each measurement through the likelihood (\ref{eqn:normlik}). For example, 
%observations with a large $\shat_j$ have an essentially flat likelihood, and so have little effect on estimation of $\pi$, and
%corresondingly little effect on the estimated FDR of other observations. 
%We highlight this because it contrasts with approaches that model the $p$ values or $z$ scores directly: observations with large $\shat_j$ will tend to produce 
%$p$ values with approximately $U[0,1]$ distribution, and $z$ scores with an approximately $N(0,1)$ distribution, under both the null and the alternative hypothesis (assuming
%that the effect $\beta_j$ is small compared with $\shat_j$). Such noisy observations therefore do affect inference; specifically they add to the
%estimated weight of the null component $\pi_0$, and inflate estimates of the FDR. 
%Because of this our approach can take better account of the informativeness of each measurement, as illustrated in Results below.

\def\dt{\text{dt}}

%
%assume that the likelihood $L(\beta) := p(D | \beta)$ can be approximated by 
%$\hat{L}(\beta) := \prod_j \hat{L}_j(\beta_j)$ where
% $\hat{L}_j(\beta_j) := p(\bhat_j | \beta_j, \shat_j)$ is given by 
%  \begin{equation} \label{eqn:lik}
% \hat{L}_j(\beta_j) \propto \dt_\nu(\bhat_j; \beta_j, \shat_j)
% \end{equation}
% where $\dt_\nu(\cdot; \mu,\sigma)$ denotes the density of the generalized $t$ distribution on $\nu$ degrees of freedom with mean $\mu$ and scale parameter $\sigma$ (i.e. the density of $\mu+\sigma T_\nu$). In the special case $\nu \rightarrow \infty$ this corresponds, of course, to a normal approximation to the likelihood:
%  \begin{equation} \label{eqn:normlik}
% \hat{L}_j(\beta_j) \propto N(\bhat_j; \beta_j, \shat_j).
% \end{equation}
%which, in light of (\ref{eqn:betahat}), is equivalent to
 % \begin{equation} \label{eqn:lik2}
% L(\beta_j) \propto p(\bhat_j | \beta_j, \shat_j).
 %\end{equation}

%The $t$ likelihood (\ref{eqn:lik}) can derived from an assumption of conditional independence of the data on each effect given the true effects, plus two additional assumptions: a) $\bhat_j, \shat_j$ contain most of the information about $\beta_j$ in the data $D_j$ (i.e. they are ``approximately sufficient" for $D_j$); and b) $\shat_j$ alone contains little information about $\beta_j$. The former (a) suggests the approximation $\hat{L}_j(\beta_j) \propto p(\bhat_j, \shat_j | \beta) = p(\bhat | \beta, \shat_j) p(\shat_j | \beta)$, while the latter (b) suggests assuming that $p(\shat_j|\beta_j)$ does not depend on $\beta_j$. Putting these two together yields $\hat{L}_j(\beta_j) \propto p(\bhat_j | \beta, \shat_j)$. If $\bhat_j, \shat_j$ are not sufficient for $D_j$ then there is some loss of efficiency in making this approximation, but all Bayesian inferences remain valid provided they are interpreted as conditional on $\bhat_j,\shat_j$, rather than conditional on $D_j$. See \cite{johnson:2008} and \cite{wakefield:2009} for more general discussion and examples of Bayesian inference based on summary statistics.

%This likelihood approximation can be motivated by assuming that i) $\bhat$
% $(\bhat_j -\beta_j)/\shat_j | \shat_j$ has a $T_\nu$ distribution, and then writing
% $p(D_j | \beta_j) = p(\bhat_j, \shat_j | \beta) = $ as sufficient for $\beta$. 
%

%Alternatively (\ref{eqn:lik}) could be considered definitional for $\bhat_j,\shat_j$: that is, choose $\bhat_j,\shat_j,\nu$ such that (\ref{eqn:lik}) holds.

%While has parallels with approaches like \cite{efron2008microarrays, sun2007oracle}
 %that model the $Z$ scores $Z_j = \bhat_j/\shat_j$ as normally distributed under the null, 
%there is an important difference: here,

%The choice of normal likelihood seems natural, and indeed it can be motivated in multiple ways. For example, we can write $p(\bhat_j, \shat_j | \beta_j) = p(\bhat_j | \shat_j, \beta_j)p(\shat_j | \beta_j)$. Now, if we are willing to assume that $\shat_j$ alone contains no information about $\beta_j$, or equivalently that $p(\shat_j | \beta_j)$ does not depend on $\beta_j$, then $L(\beta_j) \propto p(\bhat_j | \shat_j, \beta_j)$, and if $\bhat_j | \shat_j, \beta_j \sim N(\beta_j, \shat_j^2)$, as is often asymptotically the case, then we obtain the likelihood (**) above.

\section*{Results}

We compare results of \ashr with existing FDR-based methods implemented
in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).  In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
to \locfdr using {\tt nulltype=0} and to \mixfdr using {\tt theonull=TRUE}. Otherwise all packages were used with default options.

\subsection*{Effects of the Unimodal Assumption}

Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
standard error, $s_j=1$ and all methods are provided that information. That is,
$\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$. In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.

To briefly summarize the results in this section:
\begin{enumerate}
\item The UA can produce very different inferences compared with the ZA made by existing methods.
\item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
\end{enumerate}

\subsubsection*{The UA and ZA can produce different inferences}

\begin{figure}
\center \includegraphics[height=6in]{../figures/decomp_ZA.pdf}
\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan). In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition. (In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
\end{figure}

To illustrate the different inferences from the UA and ZA we show results for a single
dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
 under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)

Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA. Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further. This would increase the null component (dark blue) at the expense of
the alternative component (light blue). Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0. The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears. In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0. See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
demonstration.

 \subsubsection*{The UA can produce conservative estimates of $\pi_0$}

The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
This is desirable, provided that these estimates remain conservative: that is,
provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
for $\pi_0$. The alternative distributions are shown in Figure \ref{fig:altdens}, with details in Table \ref{table:scenarios}.
They range from a ``spiky" distribution -- where many non-zero $\beta$ are
too close to  zero to be reliably detected, making reliable estimation of $\pi_0$ essentially impossible -- to a much
flatter distribution, which is a normal distribution with large variance (``big-normal") -- where most non-zero $\beta$ are easily detected
making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.

% How do these differences influence estimates of the FDR? To address this, note that  -- in this simple case where the 
% $\shat_j$ are all identical --  the primary driver of differences in FDR estimates among methods is {\it differences in their estimates
% of $\pi_0$}. To see why, observe that, given an estimate $\hat\pi_0$ for $\pi_0$, equation (\ref{eqn:FDRest}) provides a 
% strikingly natural non-parametric estimate for the FDR (provided the denominator is not very small). Thus, 
% although the methods considered here differ from one another in many ways, having estimated $\pi_0$
% they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
%

%Thus, differences in FDR estimates among methods are driven primarily by differences in their estimates of $\pi_0$,
% and so to assess differences in FDR estimates among methods we focus on assessing differences in estimates of $\pi_0$.
%Figure \ref{fig:ZA} suggests that, compared with the ZA, the UA will tend to reduce the estimate of $\pi_0$. 
%At the same time, provided the UA holds, we hope still to provide ``conservative" estimates of $\pi_0$ (that is,
%overestimates of $\pi_0$, which will yield overestimates of the FDR).

\begin{figure} 
\begin{center}
\begin{subfigure}{\textwidth}
	\includegraphics[height=1.8in]{../figures/scenarios_density.pdf} 
	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[height=1.8in]{../figures/pi0_est.pdf} 
\caption{Comparison of true and estimated values of $\pi_0$. When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate. When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[height=1.8in]{../figures/lfdr_est.png} 
\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n). Black line is $y=x$ and red line is $y=2x$. Estimates of $\lfdr$ are conservative when UA holds, due to conservative estimates of $\pi_0$.} \label{fig:lfdr}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[height=1.8in]{../figures/lfsr_est.png} 
\caption{As in c), but for $\lfsr$ instead of $\lfdr$. Estimates of $\lfsr$ are consistently less conservative than $\lfdr$ when UA holds, and also less anti-conservative in bimodal scenario.}  \label{fig:lfsr}
\end{subfigure}
\end{center}
\caption{Results of simulation studies (constant precision $s_j=1$).}
\end{figure}

For each simulation scenario we simulated 100 independent data sets, each with $J=1000$ observations. For each data set we simulated
data as follows:
\begin{enumerate}
\item Simulate $\pi_0  \sim U[0,1]$.
\item For $j=1,\dots,J$, simulate $\beta_j \sim \pi_0 \delta_0 + (1-\pi_0) g_1(\cdot)$.
\item For $j=1,\dots,J$, simulate $\bhat_j | \beta_j \sim N(\beta_j,1)$.
\end{enumerate}
%Thus these simulations assume the same
%precision $1$ for each measurement, and for all methods this precision is assumed known (i.e. $\shat_j=1$).

Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
are used. (In addition, the effects on $\lfsr$ estimates turn out to be relatively modest; see below).

%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
%
%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
%and so all effects are non-zero and no nulls hold. This scenario is rather different
%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
%``mostly non-null"
%scenario, and we would agree! On the other hand, the distribution of $p$ values
%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
%Thus it seems desirable to have methods that perform well in both cases, and this is
%an important feature of the methods we introduce here.
%
%TODO: put FDRs on graph? Simulate with 50\% null?

\subsubsection*{The $\lfsr$ is more robust than \lfdr}

The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
Nonetheless, in many scenarios \ashr continues
to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
This is because these scenarios include 
an appreciable fraction of ``small non-null effects" that are essentially indistinguishable from 0, making accurate
estimation of $\pi_0$ impossible. Put another way, and as is well known, $\pi_0$ is not identifiable:
 the data can effectively provide an upper bound on plausible values of $\pi_0$,
but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
be substantially larger than the true value.

Since FDR-related quantities depend quite sensitively on $\pi_0$, the consequence of this 
overestimation of $\pi_0$ is corresponding overestimation of FDR (and $\lfdr$, and $q$ values).
To illustrate, Figure \ref{fig:lfdr} compares the estimated $\lfdr$ from ash.n with the true value (computed using
Bayes rule from the true $g_1$ and $\pi_0$). As predicted, $\lfdr$ is overestimated, especially in scenarios which involve many
non-zero effects that are very near 0 (e.g.~the spiky scenario with $\pi_0$ small) where
$\pi_0$ can be grossly overestimated.  
(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)

The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
 become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
 This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.

Note that, compared with previous debates regarding testing, 
this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
In previous debates authors have argued against testing whether an effect is 0
because it is {\it implausible that effects are exactly 0}. Here we add that {\it even if one believes
that some effects may be exactly zero}, it is still better to focus on the sign, because generally {\it the data are more informative about that question}
and so inferences are more robust to, say, the inevitable mis-estimation of $\pi_0$.
To provide some intuition, consider an observation with a $z$ score of 0. The $\lfdr$ of this observation can range from 0 (if $\pi_0=0$)
to 1 (if $\pi_0=1$). But, assuming a symmetric $g$, the $\lfsr>0.5$ whatever the value of $\pi_0$, because the observation $z=0$ says
nothing about the sign of the effect.
Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.

Given that we argue for using $\lfsr$ rather than $\lfdr$, one might ask whether we even need a point mass on zero in our analysis.
Indeed, one advantage of the $\lfsr$ is that it makes sense even if no effect is exactly zero. And, 
if we are prepared to assume that no effects are exactly zero, then removing the point mass 
yields smaller and more accurate estimates of $\lfsr$ when that assumption is true (Figure \ref{fig:lfsr-nn}). 
However, there is ``no free lunch":  if in fact some effects are exactly zero
then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.

\subsubsection*{Numerical Stability}

The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
even for this largest difference results from the default run (e.g. the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood. See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.

\subsubsection*{The UA helps provide reliable estimates of $g$}

An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
can estimate the {\it size} of each effect $\beta_j$.
Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
threshold $T$.
%Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
%this posterior reflects information from all the observations and not only from the observation $\bhat_j$. 
%In our case, because we assume $g$ to be unimodal about zero, this prior always ``shrinks" the posterior distributions
%towards zero, with the strength of the shrinkage being estimated from the data.
Further, because the posterior distribution is, by definition,
conditional on the observed data, interval estimates based on posterior distributions are also valid Bayesian inferences for any subset of the effects that have
been selected based on the observed data. This kind of ``post-selection" validity is much harder to achieve in the frequentist paradigm.
In particular the posterior distribution solves the (Bayesian analogue of the) ``False Coverage Rate" problem posed by
\cite{benjamini2005false} which \cite{efron2008microarrays} summarizes as follows: ``having applied FDR methods to select a set of nonnull cases,
how can confidence intervals be assigned to the true
effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
%but also notes the difficulties of implementing a stable general algorithm.

The ability of the EB approach to provide valid ``post-selection" interval estimates is extremely attractive in principle.
But its usefulness in practice
depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
which are notoriously difficult in general. Indeed, Efron emphasises
the difficulties of implementing a stable general algorithm, noting in his rejoinder
``the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying
to be overly nonparametric ... but it is hard to imagine a
generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
Our key point here is that the UA greatly simplifies the deconvolution problem.
While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
that the UA can handle many cases of practical interest.

To illustrate this, Figure \ref{fig:egcdf} compares the
estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
The greater reliability of estimates afforded
by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.

Interestingly, even in the ``bimodal" scenario $\ashr$ 
is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
clearly visible in the observed $\bhat$.

We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).

\begin{figure}[h!] 
\begin{subfigure}{\textwidth}
	\includegraphics[height=2in]{../figures/egcdf.pdf}
	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[height=2in]{../figures/mean_cdf.pdf}
\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[height=2in]{../figures/mean_cdf_nopen.pdf}
\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated. Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
\end{subfigure}
\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
\end{figure}

%making the UA provides an attractive and simple way to achieve estimating $g$  large
%subset of cases of interest (those where the UA can reasonably be assumed to hold).

%In practice, the utility of the posterior distributions noted above depends primarily on the accuracy
%of the estimate of the distribution $g(\cdot)$.

 %quote from efron rejoinder , p46:
%MorrisÕ Section 3 is especially pertinent. His formula
%for p(?|z) is related to my discussion of the
%BenjaminiÐYekutieli False Coverage Rate method in
%Section 7, particularly (7.2)Ð(7.4). Originally I had
%hoped to develop an empirical Bayes method for estimating
%such models, but the effort foundered on practical
%difficulties involving the perils of deconvolution.
%Section 6 on the Òone-group modelÓ is the ugly duckling
%of the current paper, but it bears on some important
%points raised in the discussion. Figure 7 concerns
%a fuzzy version of simultaneous hypothesis testing,
%where, as in MorrisÕ hospital example, the usual singlepoint
%null hypotheses seem unequal to the task. The
%development from (6.6) onwards, particularly (6.12),
%bears on the possibility of nonnormal null distributions,
%and is about as far as I can go in answering Professors
%Rice and SpiegelhalterÕs penultimate question.
%With g(?) a normal distribution, model (6.1) returns
%us to the realm of Stein estimation, the scene of my
%happy collaborations with Carl Morris. I continue to
%be surprised at how much harder bumpy, nonnormal
%models like (7.1) are to deal with. JamesÐStein estimation
%works fine with, say, N = 10 component problems,
%while the RobbinsÕ form of empirical Bayes appropriate
%to (7.1) seems to require hundreds or thousands.
%The information calculations in Efron (2008)
%reinforce this gloomy assessment. Maybe I am trying
%to be overly nonparametric in constructing the empirical
%Bayes Fdr estimates, but it is hard to imagine a
%generally satisfactory parametric formulation for (6.1).
%Or perhaps it is just that hypothesis testing is more demanding
%than estimation.
%My Note: (6.1) is the general model (mu \sim g()).
%7.1 is the case where the effects are a mixture of zero and N(-3,1).

\subsubsection*{Calibration of posterior intervals}

To quantify the effects of errors in estimates of $g$ we examine the calibration of the resulting posterior distributions (averaged over 100 simulations in each Scenario). 
Specifically we examine the empirical coverage of nominal lower 95\% credible bounds for a) all observations; b) significant negative discoveries; c) significant positive discoveries.  We examine only lower bounds because the results for upper bounds follow by 
symmetry (except for the one asymmetric scenario). We separately examine positive and negative discoveries because the lower bound plays a different role
in each case: for negative discoveries the lower bound is typically large and negative and limits how big (in absolute value) 
the effect could be; for positive discoveries the lower bound 
is positive, and limits how small (in absolute value) the effect could be. Intuitively, the lower bound for negative discoveries depends on the accuracy of $g$ in its tail,
whereas for positive discoveries it is more dependent on the accuracy of $g$ in the center.

The results are shown in Table \ref{tab:coverage}.  Most of the empirical coverage rates are in the range 0.92-0.96 for nominal coverage of 0.95, which we view
as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.

% As noted above, the penalty term (\ref{eqn:penalty}) causes systematic over-estimation of the mass of $g$ near 0,  which one would expect to overshrink interval estimates towards zero. 
%The effect of overshrinking on the left tail area is different for ``negative discoveries" and ``positive discoveries": for negative discoveries the lower credible bound is negative and overshrinking (moving it too far towards zero) makes the left tail too big; conversely, for positive discoveries, the lower confidence bound is positive, and  
%overshrinking (moving it too far towards zero) makes the tail too small. In the context of reliably identifying ``large" effects we might worry more about under-shrinkage, 
%and be prepared to put up with some overshrinkage to avoid it.

%Coverage rates are given in:
%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}

%Note - this could probably be solved by making the small components symetric - the idea is that the asymetry only comes in with the tail.

\begin{table}[!ht]
\begin{subtable}{\textwidth}
	\centering\input{tables/scoverage_all.tex}
	\caption{All observations. Coverage rates are generally satisfactory, except for the extreme ``spiky" scenario. This is due to the penalty term (\ref{eqn:penalty}) which tends to cause over-shrinking towards zero. Removing this penalty term produces coverage rates closer to the nominal levels for uniform and normal methods (Table \ref{tab:nopen}). Removing the penalty in the half-uniform case is not recommended (see Appendix).}
\end{subtable}

\begin{subtable}{\textwidth}
\centering\input{tables/scoverage_neg.tex}
\caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
and the normal-based method in the flat-top scenario. These results likely reflect inaccurate estimates of the tails of $g$ due to a disconnect between the tail of $g$ and the component distributions in these cases. For example, the uniform methods sometimes substantially underestimate the length of the tail of $g$ in these long-tailed scenarios, 
causing over-shrinkage of the tail toward 0.}
\end{subtable}

\begin{subtable}{\textwidth}
\centering\input{tables/scoverage_pos.tex}
\caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
\end{subtable}

\caption{Table of empirical coverage for nominal 95\% lower credible bounds} \label{tab:coverage}
\end{table}

%
%\subsubsection*{Reverse asymptotics}
%
%Note that if observation precisions are all similar and most observations are null then all the methods likely to give similar results?
%
%Note: we'd like to make sure it behaves ok in the case where most observations are null.
%Perhaps simulate with 999 null and 1 alternative?
%Also maybe look at performance for small sample sizes - e.g. n=1?
%
%Note that, even though the exact value of the point mass $\pi_0$ cannot, in general, be reliably estimated, the actual underlying distribution $g$ 
%can be quite accurately estimated from the data, provided we assess accuracy by a metric that is not sensitive to $\pi_0$: for example
%by measuring the difference between the true and estimated cumulative distribution function (cdf). See supplementary Figure \ref{} for examples.
%

\subsection*{Differing measurement precision across units}

 We turn now to the second important component of our work: allowing for varying
 measurement precision across units. The key to this is the use of a likelihood,
 (\ref{eqn:normlik}) or (\ref{eqn:tlik}), that explicitly incorporates the measurement precision (standard error) of each $\bhat_j$.

  To illustrate, we conduct a simulation where half the measurements are quite precise (standard error $s_j = 1$), and the other half are very poor
 ($s_j=10$).  In both cases,
we assume that half the effects are null
and the other half are normally distributed with standard deviation 1:
\begin{equation}
p(\beta) = 0.5 \delta_0(\beta) + 0.5 N(\beta; 0,1).
\end{equation}
In this setting, the poor-precision measurements $(s_j=10)$ tell us very little, and any sane analysis should effectively ignore them. 
 However, this is not the case in standard FDR-type analyses (Figure \ref{fig:goodpoor}). This is because the poor measurements 
 produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
 which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
 This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).

\begin{figure}
\begin{subfigure}{\textwidth}
\centering\includegraphics[width=4in]{../figures/GOODPOOReg_hist.pdf} 
    \label{fig:goodpoor-hist}
    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
 \end{subfigure}
\begin{subfigure}{\textwidth}
\centering\includegraphics[width=4in]{../figures/GOODPOOReg_scatter.pdf}
    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis). Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
\end{subfigure}
\caption{Simulation illustrating how, for existing FDR methods, 
poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only. In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
\end{figure}

 \subsubsection*{Reordering of significance, and the ``$p$ value prior"}

Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
 with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
 We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
 The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
 and so the $\lfsr$ for these poor-precision measurements is always high.

The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
was previously discussed in \cite{guan.stephens.08}. However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores). He named this prior the ``$p$-value prior" because it can be thought of as the 
implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)

This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))]. Setting $\alpha=0$ implies that $\beta_j$ is
independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
effects (in absolute value). This latter assumption may often be qualitatively plausible: 
for example, in gene expression studies the standard error for gene $j$ depends partly on
the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
which will be dataset-specific. 
Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
which has been implemented in our software by C. Dai.

%\begin{result}
%If the prior distribution on $\beta/s | s$ is a mixture of zero-centered normal distributions, (\ref{eqn:mixnorm}), and the likelihood for $\beta$ is 
%given by $\bhat | s, \beta \sim N(0,s^2)$ then $p(\beta>0 | \bhat, s)$ is a monotonic increasing function of the $z$ score, $\bhat/s$. 
%\end{result}

%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
%Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
%In contrast, existing methods effectively assume $\alpha=1$, which implies that effects tend to scale proportional to their standard error.
%Whether it is better to set $\alpha=0$ or $\alpha=1$ in practice will depend on actual relationship between $\beta_j$ and $s_j$,
%which will be dataset-specific. Indeed, it seems quite likely that, in general, the optimal $\alpha$ may be some other (non-integer) value.
%(Alternatively, since fractional values of $\alpha$ may be tricky to interpret, users
%may prefer to simply choose whichever of $\alpha=0,1$ yields the highest likelihood.)

%\begin{result}
%Treat the standard errors $s_j$ as fixed and given. Under the normal likelihood (\ref{eqn:normlik}), and using a mixture of normals for the prior distribution 
%(\ref{eqn:beta-alpha}) with $\alpha=1$, 
%the (Bayesian) $\lfsr_j$ and $\lfdr_j$ are all monotonic with $|t_j|$ where $t_j = \bhat_j/s_j$, and with the corresponding $p$ value $(p_j)$. Also, the shrinkage factor ($\text{Posterior Mean($\beta_j$)}/\bhat_j$) 
%is montonic with $|t_j|$. 
%\end{result}

%It is straightforward to prove, by performing Bayesian analysis for the standardized b$-J$
%Actually the key requirement is that the prior be such that the $t$ statistics $beta_j/s_j | s_j$ are identically distributed under the alternative.

%Most previous authors of EB approaches to false discovery rates \cite{efron2008microarrays,muralidharan2010empirical}
%have worked with models for $z$ scores $\beta_j/s_j$, rather than models for the effects $\beta_j$.
%In doing this they assume (implicitly or explicitly) that the $z$ scores are identically distributed under the alternative model.
%This is analogous
%to assuming $\alpha=1$ in (ref{eqn:beta-alpha}). And, given the result above, 
%standard frequentist FDR methods, like \qvalue and Benjamini--Hochberg, 
%that rank tests by their $p$ value, could also be viewed as making the same implicit assumption.

\begin{figure}
\centering\includegraphics[width=4in]{../figures/lfsr_vs_pval_GOODPOOR.pdf}
\caption{Figure illustrating affects of prior assumptions on re-ordering of significance. Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$. Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
\end{figure}

%\subsubsection*{Adaptive Shrinkage}
%
%Johnstone and Silverman \cite{johnstone2004needles} (see also \cite{raykar2011empirical}) note the ability of EB methods to adapt to overall signal strength
%when attempting to identify and estimate strong signals. The idea is that, if most $\beta_j$ are truly at or near zero then (given enough data) 
%the estimated prior distribution $\hat{g}$ will reflect this by being concentrated near 0, and the posterior distributions
% $p(\beta_j | \bhat_j, \shat_j,\hat{g})$ will consequently also tend to concentrate around zero (strong shrinkage). 
% In contrast, if most $\beta_j$ are large then $\hat{g}$ will be flatter, resulting in less shrinkage.
% Indeed, in the limit as $\hat{g}$ becomes flatter and flatter the posterior distribution on $\beta_j$ becomes $N(\bhat_j,\shat_j)$,
% and the Bayesian credible intervals match standard confidence intervals, which might be considered ``no shrinkage".
%
%Our methods we build on this idea in two ways. First,
%by using flexible semi-parametric approaches to estimate $g$ we aim to maximise the potential for this adaptive behaviour.
%Second, by incorporating the precision of each measurement into the likelihood, (\ref{eqn:normlik}) or (\ref{eqn:tlik}), we ensure that shrinkage
%adapts to measurement precision: more precise measurements undergo
% less shrinkage than less precise measurements.  This is because observations $\bhat_j$ with larger standard errors have flatter likelihoods 
%than observations with
%small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
%Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
%
%\begin{figure}[!ht]
%\begin{center}
%\includegraphics[width=400px]{../examples/ash.pdf} 
%\end{center}
%\caption{Figure illustrating adaptive shrinkage. Results are shown for two different scenarios: (left) ``big-normal" where the effects have a wide distribution; (right) ``spiky", where the effects are more concentrated near 0. The standard error of each observation, $s_j$, was simulated from Inverse Gamma(5,5), which resulted in $s_j$ varying from 0.34 to 6.7.  The shrunken estimates ($y$ axis) are plotted against the observed estimates ($x$ axis), with color indicating the (square root of the) precision of each measurement (blue=lower-precision; red=higher-precision). The two key features are i) shrinkage is adaptive to signal in data, so shrinkage is stronger for the spiky scenario; ii) shrinkage is adaptive to the precision of each measurement, so less precise observations (blue) are shrunken more strongly.} \label{fig:ash}
%\end{figure}

%DISCUSSION ON ADAPTIVE SHRINKAGE; BEST TO PUT IN RESULTS?
%To understand how this approach leads to ``adaptive shrinkage", think of (\ref{eqn:betaj}) as defining the ``prior distribution" for $\beta_j$ (determined by hyper-parameters $\pi$),  and (\ref{eqn:bhat}) as defining the likelihood for $\beta_j$. Because the prior distribution is unimodal about zero, 
%the posterior distribution of $\beta$ will be ``shrunk" towards zero (compared with the likelihood), with the
%amount of shrinkage being determined by a) how concentrated the prior is near zero (determined by $\hat{\pi}$) and b) how concentrated the likelihood is around $\bhat_j$.
%For example, if most $\beta_j$ are truly at or near zero then $\hat\pi$ will concentrate on zero and very small values of $\sigma_k$, and this will produce stronger shrinkage
%than if $\beta_j$ are mostly large. And observations $\bhat_j$ with large standard error will have flatter likelihoods than observations with
%small standard error, and so their posteriors will be more affected by the prior, and undergo more shrinkage.

%previous work based on normal means:
%http://link.springer.com/article/10.1007/BF02562623

%\subsection{Null behaviour}

%Results of simulations under the null.

%\subsection{Comparison of different underlying component distributions}
%
%Because the mixture of zero-anchored uniforms is more flexible than the mixture of zero-centered uniforms, which is more flexible
%than the mixture of zero-centered normals, we expect that the likelihoods will satisfy $L_{hu} > L_u > L_n$. However, for Scenario A we expect the differences in
%likelihood to be small, because the true model here is a mixture of zero-centered normals, so any gain of $L_{hu}$ and $L_u$ over $L_n$ is necessarily due to ``overfitting".
%For Scenario B, we expect that $L_{hu}$ and $L_{u}$ to be significantly larger than $L_n$, but any 
%increase of $L_{hu}$ over $L_u$ must be due to overfitting (because the truth is symmetric). Finally, for Scenario C we
%expect $L_{hu}$ to be significantly larger than the others.
%
%Figure \ref{fig:rmse_loglik_boxplot} illustrates this.
%
%\begin{figure}[!ht] \label{fig:rmse_loglik_boxplot}
%\begin{center}
%\includegraphics[width=6.5in]{Rcode/figures/simABC_egdens.pdf}
%%\includegraphics[width=6.5in]{Rcode/figures/rmse_loglik_boxplot.pdf}
%\includegraphics[width=6.5in]{Rcode/figures/rmse_boxplot.pdf}
%\includegraphics[width=6.5in]{Rcode/figures/loglik_boxplot.pdf}
%\end{center}
%\caption{}
%\end{figure}

%Notes from Adrian Raftery - 
%Newcombe (1886)
%Dean and Raftery
%Uniform is super-efficient

\section*{Discussion}

We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
While requiring two numbers is slightly more onorous than requiring one, in many settings these
numbers are easily available and if so we argue it makes sense to use them.
Second, we note the potential benefits -- both statistical and computational -- of assuming that the effects come from a unimodal distribution, and provide
flexible implementations for performing inference under this assumption. We also introduce the ``false sign rate" as
an alternative measure of error to the FDR, and illustrate its improved robustness to errors in model fit, particularly mis-estimation
of the proportion of null tests, $\pi_0$.

Multiple testing is often referred to as a ``problem" or a ``burden". In our opinion, EB approaches turn this idea on its head,
treating multiple testing as an {\it opportunity}: specifically, an opportunity to learn about the prior distributions, and other modelling assumptions,
to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
and the FDR remains the same. (If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example. 
The EB approach captures this dependence in an intuitive way:
if there are lots of strong signals then we infer $\pi_0$ to be small, and the estimated FDR (or $\lfdr$, or $\lfsr$) at a given threshold may be low, even if a large number of tests were performed; and conversely
if there are no strong signals then we infer $\pi_0$ to be large and the FDR at the same threshold may be high, even if relatively few tests were performed. 
More generally, overall signal strength is reflected in the estimated $g$, which in turn influences the estimated FDR.

Two important practical issues that we have not addressed here are correlations among tests,
and the potential for deviations from the theoretical null distributions of test statistics. These two
issues are connected: specifically, unmeasured confounding factors can cause both correlations among tests
and deviations from the theoretical null \cite{efron2007correlation,leek:2007}. And although there are certainly other
factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
because they can induce strong correlations among large numbers of tests and profoundly impact results.
Approaches to deal with unmeasured confounders can be largely divided into two types: those
that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
that attempt to infer counfounders using clustering, principal components analysis, or factor models 
\cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:normlik}) using
$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
this immediately raises issues of identifiability and we do not pursue the idea further here.

Another important practical issue is the challenge of small sample sizes. For example, in genomics applications researchers sometimes
attempt to identify differences between two conditions based on only a handful of samples in each. In such settings the
normal likelihood approximation (\ref{eqn:normlik}) will be inadequate. And, although the $t$ likelihood (\ref{eqn:tlik}) partially
addresses this issue, it is also, it turns out, not entirely satisfactory.  The root of the problem is that, with small sample sizes, raw estimated standard
errors $\shat_j$ can be horribly variable. In genomics it is routine to address this issue by applying EB methods \cite{smyth:2004} to ``moderate" (i.e.~shrink) variance estimates,
before computing $p$ values from ``moderated" test statistics. We are currently investigating how
our methods should incorporate such ``moderated" variance estimates to make it applicable to small sample settings.

Our approach involves compromises between flexibility, generality, and simplicity on the one hand, and statistical efficiency and principle on the other.
For example, in using an EB approach that uses a point estimate for $g$, rather than a fully Bayes approach that accounts for uncertainty in $g$,
we have opted for simplicity over statistical principle. And in summarizing every test by
two numbers and making a normal or $t$ approximation to the likelihood, we have aimed to produce generic methods that 
 can be applied whenever such summary data are available -- just as \qvalue can be applied to any set of $p$ values for example -- although
 possibly at the expense of statistical efficiency compared with developing multiple tailored approaches based on context-specific likelihoods. 
 Any attempt to produce generic methods will involve compromise between generality and efficiency.
 In genomics, many analyses -- not only FDR-based analyses -- involve first computing a series of $p$ values before
 subjecting them to some further downstream analysis. An important message here is 
 that working with two numbers ($\bhat_j,\shat_j$), rather than one ($p_j$ or $z_j$), 
 can yield substantial gains in functionality (e.g.~ estimating effect sizes, as well as testing; accounting for variations in
measurement precision across units) while losing only a little in generality.  We hope that our work will
encourate development of methods that exploit this idea in other contexts.

%This is connected with work by \cite{wakefield:2009} (see also \cite{johnson:2008}) who performs Bayesian inference
%from summary statistics, computing $p(\beta_j | \bhat_j, \shat_j)$ for a single $j$,
%with a specific fixed prior on $\beta_j$. Essentially we extend this to multiple $j$, using a hierarchical model to connect the observations.

%
%
%Both \locfdr and \mixfdr allow for deviations from
%the theoretical null distribution by allowing that, under the null, the $z$ scores are $N(0,\sigma^2)$ with $\sigma$ estimated from the data rather
%than fixed to 1. Within our approach this idea could be incorporated by changing the likelihood - for example, to assume $\bhat_j \sim N(0,s_ 
%
%These are tricky issues that, as far as we are aware, can affect
%all approaches to FDR analysis. An important focus of Efron's work, also implemented in \locfdr, 
% is to relax the assumption that the null $z$ scores are $N(0,1)$, to allow for empirical deviations from this
%theoretical null distribution. 
%
%For example, we and others have worked on Bayesian approaches to genetic association testing, using subjective prior specification of
%effect sizes based on external data (e.g.~\cite{stephens:2009}). However, the large numbers of tests brings the opportunity to learn about
%priors, and here we aim to exploit this

%Choice of mixture components. Use of likelihood to compare different mixture components?

%  
%Directly modeling the p values, or z scores, say via non-parametric methods, 
%can lead to unrealistic distributions being fitted. Put another way, because z scores are the result of adding noise to some
%distribution, the range of distributions they can take is limited Using entirely non-parametric methods loses this information.
%The solution is to model beta hat as a convolution of some distribution g and an error component.

%Before proceeding, we feel obliged to correct the common misconception
% that the use of FDR methods is to ``correct" for the number of tests performed. In fact the FDR does
% not depend on the number of tests. 
%  A better way to think of it is that the FDR accounts for the {\it amount of signal} in the tests that were performed: 
% 
% In other words, FDR corrects for the {\it results} of all tests performed, not the {\it number} of all tests performed.

\section*{Detailed Methods} \label{sec:implement}

\subsection*{Embellishments}

\subsubsection*{More flexible unimodal distributions}

Using a mixture of zero-centered normal distributions for $g$ in (\ref{eqn:mixnorm}) implies that $g$ is not only unimodal, but also symmetric.
Furthermore, even some symmetric unimodal distributions, such as those with a flat top, cannot be well approximated by a mixture of zero-centered normals.
 Therefore, we have implemented a more general approach based on 
\begin{equation} \label{eqn:g-general}
g(\cdot; \pi) = \sum_{k=0}^K \pi_k f_k(\cdot) ,
\end{equation}
where $f_0$ is a point mass on 0, and $f_k$ ($k=1,\dots,K$) are pre-specified component distributions with one of the following forms: 
\begin{enumerate}[(i)]
\item $f_k(\cdot) = N(\cdot; 0, \sigma^2_k)$, \qquad (``ash.n")
\item $f_k(\cdot) = U[\cdot; -a_k,a_k]$,  \qquad (``ash.u")
\item $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$,  \qquad (``ash.hu")
\end{enumerate}
where $U[\cdot; a,b]$ denotes the density of a uniform distribution on $[a,b]$.
(In (iii) we include both components in the mixture (\ref{eqn:g-general}), so a grid of values $a_1,\dots,a_K$ defines $2K+1$ mixture component densities, 
and $\pi$ is a $2K+1$ vector that sums to 1.)
The simplest version (\ref{eqn:mixnorm}) corresponds to (i).
Replacing these with uniform components (ii)-(iii) only slightly complicates calculations
under the normal likelihood (\ref{eqn:normlik}), and greatly simplifies the calculations under the $t$ likelihood
(\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)

Moving from (i) to (iii) the representation (\ref{eqn:g-general}) becomes increasingly flexible. Indeed,
using a large dense grid of $\sigma^2_k$ or $a_k$, (i)-(iii) can respectively approximate,
with arbitrary accuracy,
\begin{enumerate}[(i)]
\item any scale mixture of normals, which includes as special cases
the double exponential (Laplace) distribution, any $t$ distribution, and a very large number of other distributions used in  high-dimensional regression settings.
\item any symmetric unimodal distribution about 0.
\item any unimodal distribution about 0.
\end{enumerate}
The latter two claims are related to characterizations of unimodal distributions due to \cite{khintchine1938unimodal} and  \cite{shepp1962symmetric}; see \cite{feller1971introduction}, p158. 
In other words, (ii) and (iii) provide fully non-parametric estimation for $g$ under the constraints that it is (ii) both unimodal and symmetric, or (iii) unimodal only.

Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
See Implementation for details of our software defaults.

%Although it is natural to worry about ``overfitting" if $K$ is large,  in fact the unimodal constraint is sufficient to prevent serious problems with overfitting. For example, note that the unimodal constraint is sufficiently strong that there will exist a non-parametric maximum likelihood estimate (NPMLE) for $g$ under this constraint. (Indeed, if the $\beta_j$ are observed without error, corresponding to the standard errors $\shat_j \rightarrow 0$, this NPMLE has an explicit and elegant form; \cite{grenander1956theory}.)
%Thus allowing $K$ to be large in ii) or iii), and maximizing the likelihood with respect to $\pi$, will approximate the NPMLE for $g$, under the constraints of symmetry and unimodality (ii), or unimodality alone (iii).

\subsubsection*{Dependence of effects on standard errors}

Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
 \begin{equation} \label{eqn:beta-alpha}
 \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
 \end{equation}
for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.

The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
 standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
 Applying the algorithm for $\alpha=0$ to effect estimates $\hat{b}_1,\dots,\hat{b}_J$ with standard errors $\shat'_1,\dots,\shat'_J$
yields a posterior distribution $p(b_j | \shat_j, \hat{b}_j, \hat{\pi}, \alpha)$, which induces a posterior distribution on $\beta_j = b_j \shat_j^\alpha$.
%In the special case $\alpha=1$, $\hat{b}_j$ is the $t$ score $t_j$ and $\shat'{b}=1$, so this involves applying our algorithm to the $t$ scores
%treating them all as having the same standard error. In this case the local fdr from our method becomes monotonic in each direction as
%the $t$ score increases away from zero, just as with \qvalue, and (usually) \mixfdr and \locfdr.

%Different values of $\alpha$ in (\ref{eqn:beta-alpha}) essentially correspond to different models for
%how effect sizes $\beta_j$ scale with the standard error $\shat_j$. Since this scaling is unknown it may be desirable to estimate
%it from the data, and this can be done here by jointly maximizing the likelihood $p(\hat{b} | \shat,\pi,\alpha)$ over $\pi$ and $\alpha$
%rather than just $\pi$.  Although values of $\alpha$ other than 0 and 1 may not be easy interpret,
%and the actual relationship between $\beta_j$ and $\shat_j$ is unlikely to exactly follow (\ref{eqn:beta-alpha}) for any $\alpha$,
%this approach seems preferable in practice than simply assuming $\alpha=0$ or $\alpha=1$. We
%do not investigate this in detail here, but it has been implemented in our software by C.Dai, using a simple grid search over $\alpha$.

\subsubsection*{Replace normal likelihood with $t$ likelihood}

We generalize the normal likelihood (\ref{eqn:normlik}) by replacing it with a $t$ likelihood:
  \begin{equation} \label{eqn:tlik}
 \bhat_j \, | \, \beta_j, \shat_j \sim T_\nu(\beta_j, \shat_j)
 \end{equation}
where $T_\nu(\beta_j,\shat_j)$ denotes the distribution of $\beta_j+\shat_j T_\nu$ where $T_\nu$ has a standard $t$ distribution on $\nu$ degrees of freedom,
and  $\nu$ denotes the degrees of freedom used to estimate $\shat_j$ (assumed known, and for simplicity assumed to be the same for each $j$).
The normal approximation (\ref{eqn:normlik}) corresponds to the limit $\nu \rightarrow \infty$.
This generalization does not complicate inference when the mixture components $f_k$ in (\ref{eqn:g-general}) are uniforms; see Implementation below.
When the $f_k$ are normal the computations with a $t$ likelihood are considerably more difficult and we have not implemented this combination.

Equation (\ref{eqn:tlik}) is, of course, motivated by the standard asymptotic result
\begin{equation} \label{eqn:tdist}
(\bhat_j-\beta_j)/\shat_j \sim T_\nu.
\end{equation}
However (\ref{eqn:tdist}) does not imply (\ref{eqn:tlik}), because in (\ref{eqn:tdist}) $\shat_j$ is random whereas in (\ref{eqn:tlik}) it is conditioned on.
In principle it would be preferable, for a number of reasons, to model the randomness in $\shat_j$; we are
currently pursuing this improved approach (joint work with M.Lu) and results will be published elsewhere. 
%We note here that
%modelling the randomness in $\shat_j$ is particularly helpful when the number of observations used to estimate $\shat_j$ is very small (e.g. $<10$),
%where our simpler approach can produce unsatisfactory results.

\subsubsection*{Non-zero mode}

An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.

\subsection*{Implementation Details}

\subsubsection*{Likelihood for $\pi$}

We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$. [One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]

Using the prior $\beta_j \sim \sum_{k=0}^K \pi_k f_k(\beta_j)$ given by (\ref{eqn:g-general}), and the normal likelihood (\ref{eqn:normlik}), integrating over $\beta_j$ yields
\begin{equation}
p(\bhat_j | \shat,\pi)  = \sum_{k=0}^K \pi_k \tilde{f}_k(\bhat_j)
\end{equation}
where 
\begin{equation}
\tilde{f}_k(\bhat_j) := \int f_k(\beta_j) N(\bhat_j; \beta_j,\shat^2_j) \, d\beta_j
\end{equation}
denotes the convolution of $f_k$ with a normal density.
These convolutions are straightforward to evaluate whether $f_k$ is a normal or uniform density.
Specifically, 
\begin{equation} \label{eqn:uconv}
\tilde{f}_k(\bhat_j)  = 
\begin{cases}
N(\bhat_j;0, \shat_j^2 +  \sigma_k^2) & \text{if $f_k(\cdot) = N(\cdot; 0, \sigma_k^2)$}, \\
\frac{\Psi((\bhat_j-a_k)/\shat_j) - \Psi((\bhat_j-b_k)/\shat_j)}{b_k-a_k} & \text{if $f_k(\cdot) = U(\cdot; a_k,b_k)$},
\end{cases}
\end{equation}
where $\Psi$ denotes the cumulative distribution function (c.d.f.) of the standard normal distribution.
If we replace the normal likelihood with the $t_\nu$ likelihood (\ref{eqn:tlik}) then the convolution 
for $f_k$ uniform the convolution is still given by (\ref{eqn:uconv}) but with $\Psi$ the c.d.f. 
of the $t_\nu$ distribution function. (The convolution for $f_k$ normal is tricky and we have not implemented it.)

\subsubsection*{Penalty term on $\pi$}

To make $\lfdr$ and $\lfsr$ estimates from our method ``conservative" we add a penalty 
term $log(h(\pi;\lambda))$ to the log-likelihood $\log L(\pi)$ to encourage over-estimation of $\pi_0$:
\begin{equation} \label{eqn:penalty}
h(\pi;\lambda) = \prod_{k=0}^K \pi_k^{\lambda_k-1}
\end{equation}
where $\lambda_k \geq 1 \, \forall k$. The default is $\lambda_0=10$ and $\lambda_k=1$, which yielded
consistently conservative estimation of $\pi_0$ in our simulations (Figure \ref{fig:pi0sims}).

Although this penalty is based on a Dirichlet density, we do not interpret this as a ``prior distribution" for $\pi$:
we chose it to provide conservative estimates of $\pi_0$ rather than to represent prior belief.

\subsubsection*{Problems with removing the penalty term in the half-uniform case}

It is straightforward to remove the penalty term by setting $\lambda_k=1$ in (\ref{eqn:penalty}).
We note here an unanticipated problem we came across when using no penalty term in the half-uniform case 
(i.e.~ $f_k(\cdot) = U[\cdot; -a_k,0] \text{ and/or } U[\cdot; 0,a_k]$ in (\ref{eqn:g-general})): when the data
are nearly null, the estimated $g$ converges, as expected and desired, to a distribution where almost all the mass is near 0, but
sometimes all this mass is concentrated almost entirely just to one side (left or right) or 0. This can have a very profound effect on the local false sign rate:
for example, if all the mass is just to the right of 0 then all observations will be assigned a very high probability of being positive (but very small),
and a (misleading) low local false sign rate. For this reason we do not recommend use of the half-uniform with no penalty.

\subsubsection*{EM algorithm}

With this in place, the penalized log-likelihood for $\pi$ is given by:
\begin{equation}
\log L(\pi) + \log h(\pi) = \sum_{j=1}^n \log(\sum_{k=0}^K \pi_k l_{kj}) + \sum_{k=0}^K (\lambda_k-1) \log \pi_k
\end{equation}
where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
one-step updates are:
\begin{align} \label{eqn:w}
w_{kj} & = \pi_k l_{kj} / \sum_{k'} {\pi_{k'} l_{k'j}} \\
n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
\end{align}
Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.

%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
\subsubsection*{Initialization}

By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
but weak signal in the data cannot quickly draw the algorithm towards the null.

In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
data are consistent with the global null.

\subsubsection*{Conditional distributions}

Given $\hat\pi$, we compute the conditional distributions 
\begin{equation}
p(\beta_j | \hat\pi, \bhat, s) \propto g(\beta_j; \pi) L(\beta_j; \bhat_j, \shat_j).
\end{equation} 
Each posterior is a mixture on $K+1$ components:
\begin{equation}
p(\beta_j | \hat\pi, \bhat, s) = \sum_{k=0}^K w_{kj} p_k(\beta_j | \bhat_j, \shat_j)
\end{equation}
where the posterior weights $w_{kj}$ are computed as in (\ref{eqn:w}) with $\pi=\hat\pi$,
and the posterior mixture component $p_k$ is the posterior on $\beta_j$ that would be obtained using prior 
$f_k(\beta_j)$ and likelihood $L(\beta_j; \bhat_j,\shat_j)$.
All these posterior distributions are easily available.
For example, if $f_k$ is uniform and $L$ is $t_\nu$ then this is a truncated $t$ distribution.
If $f_k$ is normal and $L$ is normal, then this is a normal distribution.

\subsubsection*{Choice of grid for $\sigma_k, a_k$} \label{sec:grid}

\def\sigmamax{\sigma_{\text{max}}}
\def\sigmamin{\sigma_{\text{min}}}
When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=\min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$
based on the idea that $\sigmamax$ should be big enough so that  $\sigmamax^2 + \shat_j^2$ should exceed $\bhat_j^2$.   (In rare cases where $\max(\bhat_j^2-\shat_j^2)$
is negative we set $\sigmamax = 8\sigmamin$.)

When the mixture components $f_k$ are uniform, we use the same grid for the parameters $a_k$ as for $\sigma_k$ described above.

Our goal in specifying a grid was to make the limits sufficiently large and small, and the grid sufficiently dense, that results would not change appreciably with 
a larger or denser grid. For a specific data set one can of course check this by experimenting with the grid, but these defaults usually work well in our experience.

\appendix

\begin{figure}
\begin{center}
\begin{subfigure}{\textwidth}
\includegraphics[height=1.8in]{../figures/lfsr_est_s_nn.png} 
\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$). Black line is $y=x$ and red line is $y=2x$. The results illustrate how estimates of $\lfsr$ can be more accurate in this case. That is, assuming there is no point mass can be beneficial if that is indeed true.}  \label{fig:lfsr-nn}
\end{subfigure}
\begin{subfigure}{\textwidth}
\includegraphics[height=1.8in]{../figures/lfsr_est_s.png} 
\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$). Black line is $y=x$ and red line is $y=2x$. The results illustrate how estimates of $\lfsr$ can be anti-conservative if we assume there is no point mass when the truth is that there is a point mass.} \label{fig:lfsr-s}
\end{subfigure}
\end{center}
\caption{Illustration of effects of excluding a point mass from the analysis.} \label{fig:lfsr-nopointmass}
\end{figure}

\begin{table}[!ht]
\centering\begin{tabular}{c c } \toprule
Scenario & Alternative distribution, $g_1$  \\ \midrule
spiky & $0.4 N(0,0.25^2) + 0.2 N(0,0.5^2) + 0.2 N(0,1^2), 0.2 N(0,2^2) $\\
near normal & $2/3 N(0,1^2) + 1/3 N(0,2^2)$ \\
flattop& $(1/7) [N(-1.5,.5^2) + N(-1,.5^2) + N(-.5,.5^2) +$ \\
 &  $N(0,.5^2) +N(0.5,.5^2)  +N(1.0,.5^2) + N(1.5,.5^2)]$  \\
skew & $(1/4) N(-2,2^2) + (1/4) N(-1,1.5^2) +  (1/3) N(0,1^2) + (1/6) N(1,1^2) $\\
big-normal & $N(0,4^2)$ \\ 
bimodal & $0.5 N(-2,1^2) + 0.5 N(2,1^2)$ \\ \bottomrule
\end{tabular}
\caption{Summary of simulation scenarios considered} \label{table:scenarios}
\end{table}

\begin{table}[!ht]
\begin{subtable}{\textwidth}
	\centering\input{tables/scoverage_all_nopen.tex}
	\caption{All observations}
\end{subtable}

\begin{subtable}{\textwidth}
\centering\input{tables/scoverage_neg_nopen.tex}
\caption{``Significant" negative discoveries.}
\end{subtable}

\begin{subtable}{\textwidth}
\centering\input{tables/scoverage_pos_nopen.tex}
\caption{``Significant" positive discoveries.}
\end{subtable}

\caption{Table of empirical coverage for nominal 95\% lower credible bounds for methods {\it without} the penalty term). } \label{tab:nopen}
\end{table}
%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
%\begin{equation}
%q_j = \widehat\FDR(\{k:p_k \leq p_j\})
%\end{equation}
%which is close to the definition in \cite{storey.03}.

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgements}

Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text
prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.

This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.

%\section*{References}
% The bibtex filename
\bibliography{/Users/stephens/Dropbox/Documents/mainbib}

\section*{Figure Legends}

\clearpage

\section*{Tables}

\section*{Supporting Information Legends}

Supplementary material can be found in {\bf Supplementary Information S1.}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}

### 2015-10-26 analysis/template.Rmd
add template.Rmd and chunk options

---
title: "Untitled"
author: "First Last"
date: YYYY-MM-DD
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

This is a template for writing reports with R Markdown, copied from @jdblischak.

## Section title

You can create a new file from the template using the command `cp`.

```bash
cp analysis/template.Rmd analysis/newfile.Rmd
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/summarize_dsc_robust.Rmd
add summarize_dsc_robust.Rmd

---
title: "Summarize results on robustness of EM algorithm"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

Here we summarize the results of the simulation study to look at convergence of ash.
```{r}
  library(dscr)
  library(ashr)
  library(ggplot2)
  library(reshape2)
  library("dplyr")
  load("../output/dsc-robust-files/dsc_robust.RData") 
  res = dsc_robust$res
```

```{r}
  quantile((res %>% filter(method=="ash.n"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
  quantile((res %>% filter(method=="ash.hu"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
  quantile((res %>% filter(method=="ash.u"))$score.diff1,c(0.8,0.9,0.95,0.99,1))

  plot(ecdf((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1),ylim=c(0.8,1))
  plot(ecdf((res %>% filter(method %in% c("ash.n")))$score.diff1),ylim=c(0.8,1))
```

```{r}
  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1<0.02)
  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1>1)
  mean((res %>% filter(method %in% c("ash.n")))$score.diff1<0.02)  
```

Now check out the example that showed the biggest convergence problem
```{r}
#check out which one showed biggest convergence problem
prob = res[which.max(res$score.diff1),]
setwd("../output")
prob_eg = loadExample(dsc_robust,seed=prob$seed,scenario=as.character(prob$scenario),method=as.character(prob$method))
setwd("../analysis")
```

Run ash on the example to check we can recapitulate the convergence problem.
```{r}
set.seed(prob$seed+1) #dscr sets the seed for the method to 1+seed 
ash.res=list()
ash.res[[1]] = ash(prob_eg$input$betahat,prob_eg$input$sebetahat,mixcompdist="uniform",cxx=TRUE)
for(i in 2:11){ash.res[[i]] = ash(prob_eg$input$betahat,prob_eg$input$sebetahat,mixcompdist="uniform",randomstart=TRUE,cxx=TRUE)}
```

```{r}
loglik = mapply(get_loglik,ash.res)
loglik
```

```{r}
best=which.max(loglik)
plot(ash.res[[best]]$lfsr,ash.res[[1]]$lfsr,xlab="lfsr (best run)",ylab="lfsr (default run)",main="Comparison of lfsr (best vs default run)")
abline(a=0,b=1,col=2)
plot(ash.res[[best]]$PosteriorMean,ash.res[[1]]$PosteriorMean,xlab="Posterior Mean (best run)",ylab="Posterior Mean (default run)",main="Comparison of Posterior Means (best vs default run)")
abline(a=0,b=1,col=2)

```

Compare Cdfs:
```{r}
  x = seq(-5,5,length=100)
  cdf_res = data.frame(x=x,
                     best=as.numeric(cdf.ash(ash.res[[best]],x)$y),
                     default=as.numeric(cdf.ash(ash.res[[1]],x)$y))
  cdf_res=melt(cdf_res,id.vars="x",value.name="y",variable.name="run")
  qplot(x=x,y=y,data=cdf_res,geom="line",colour=run,main="comparison of estimated cdf (best run and default run)")
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/summarize_dsc_robust.Rmd
add index

@@
 prob = res[which.max(res$score.diff1),]
-setwd("../output")
-prob_eg = loadExample(dsc_robust,seed=prob$seed,scenario=as.character(prob$scenario),method=as.character(prob$method))
-setwd("../analysis")
+prob_eg = load_example(dsc_robust,seed=prob$seed,scenario=as.character(prob$scenario),method=as.character(prob$method),homedir="../output")
 ```

### 2015-10-26 analysis/index.Rmd
add index

---
title: "Home"
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

## DSC Robust (Simulation study to assess EM convergence)

*  [Check robustness simulation study](summarize_dsc_robust.html)

## DSC Shrink (the main simulation study in the paper)

*  [Make coverage tables](summarize_coverage.html)

### 2015-10-26 analysis/summarize_coverage.Rmd
add index

---
title: "Summarize coverage"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

---
title: "Summarize Coverage"
output: html_document
---

```{r}
load("../output/dsc-shrink-files/res.RData")
coverthresh = 0.05 # threshold at which we look at coverage
findthresh=0.95 #threshold at we define a discovery significant
library("dscr")
library("ashr")
library(reshape2)
library(ggplot2)
library(dplyr)

neglong = 
  res$negprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="negprob") %>%
    filter(negprob > findthresh)

poslong = 
  res$posprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="posprob") %>%
    filter(posprob > findthresh)

reslong = 
  res$cdf_score %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%    
    melt(id.vars=c("method","scenario","seed",".id"))

reslong.pos = inner_join(reslong,poslong)
reslong.neg = inner_join(reslong,neglong)
dim(reslong.pos)
dim(reslong.neg)

```

Overall proportion of negative findings is `r nrow(reslong.neg)/nrow(reslong)`.

Overall proportion of positive findings is `r nrow(reslong.pos)/nrow(reslong)`.

Table of lower tail for all observations
```{r}
xtabs(lt~method+scenario,reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
```

Table of lower tail of positive findings. Because of the unimodal assumption
and the favoritism toward the null this should assess problems
with``over shrinkage" toward 0.
```{r}
xtabs(lt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
```

Table of lower tail of negative findings. This should indicate problems with tail behaviour of $g$. The uniform methods tend to over-shrink.
```{r}
xtabs(lt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
```

```{r}
save_latex_coverage_table=function(df,methodnames,filename,SCENARIONAMES=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),switch=FALSE){
  library(xtable)
  df$method=factor(df$method,levels=methodnames)
  df$scenario=factor(df$scenario,levels=SCENARIONAMES)
  mat <- as.matrix(xtabs(lt~method+scenario,df))
  if(switch){
    mat=1-mat
  }
  mat <- xtable(mat,digits=rep(2,ncol(mat)+1))
  write(print(mat, 
        sanitize.text.function = function(x){x},
        floating=FALSE, 
        hline.after=NULL, 
        add.to.row=list(pos=list(-1,0, nrow(mat)), 
                        command=c('\\toprule ',
                                  '\\midrule ',
                                  '\\bottomrule '))
      ),file=filename)

}

save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_neg.tex")
save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_pos.tex")
save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_all.tex")

save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_neg_nopen.tex")
save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_pos_nopen.tex")
save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_all_nopen.tex")
```

These tables show 1-the left tail instead of the left tail.
```{r}
save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_neg.tex",switch=TRUE)
save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_pos.tex",switch=TRUE)
save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_all.tex",switch=TRUE)

save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_neg_nopen.tex",switch=TRUE)
save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_pos_nopen.tex",switch=TRUE)
save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_all_nopen.tex",switch=TRUE)
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/index.Rmd
add more analyses

@@
 
+*  [Investigate coverage of hu method without penalty](investigate_hu_badcoverage.html)
+*  [Run mixfdr on low noise bimodal situation](check_mixfdr_lownoise.html)
 *  [Make coverage tables](summarize_coverage.html)
+
+## DSC Znull (assess behaviour of ash LR under null)
+
+* [Summarize results of null simulations](summarize_dsc_znull.html)

### 2015-10-26 analysis/summarize_dsc_znull.Rmd
add more analyses

---
title: "Summarize distribution of LR statistic under null"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r}
library(ggplot2)
library(dplyr)
load("../output/dsc-znull-files/res.znull.RData")
temp=ungroup(res %>% group_by(scenario,method) %>% summarise(gt0=mean(logLR>0)))
xtabs(gt0~method+scenario,temp)

```

Note that the ash.n.s basically conforms exactly to the null expectation
of 0.5 chi2_0 +0.5 chi2_1 from Stram and Lee, Biometrics

conjecture: the uniform has the same asymptotic behaviour
and the half uniform is like the sum of two of these?

From this

a 95% procedure would need to check for logLR > `r qchisq(0.9,df=1)/2`.

```{r}
ggplot(res %>% filter(logLR!=0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
ggplot(res %>% filter(logLR!=0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
ggplot(res %>% filter(logLR!=0) %>% filter(grepl("ash.n",method)), aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")

```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/plot_cdf_eg.Rmd
added plotting Rmd files

---
title: "Various examples of cdf plots"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r}
library(plyr)
library(dplyr)
library(dscr)
library(magrittr)
library(ashr)
library(ggplot2)
source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
source("../R/set_plot_colors.R")
load("../output/dsc-shrink-files/res.RData")

#df is a list with components method,seed, scenario
#cdf is evaluated at x
#returns list of x,y,pi0, where cdf values are in y
get_cdf = function(df,dsc=dsc_shrink,x=seq(-6,6,length=100),homedir="../output"){
  m=df$method
  if(m=="truth"){m="ash.n"}
  temp=load_example(dsc,df$seed,df$scenario,m,homedir)
  pi0=temp$meta$pi0

  if(df$method=="truth"){
    s=dsc$scenarios[[df$scenario]]
    galt=s$args$g
    g=normalmix(c(pi0,(1-pi0)*mixprop(galt)),
                c(0,comp_mean(galt)),
                c(0,comp_sd(galt)))
  } else {
    if(grepl("mixfdr",df$method)){
      temp$output= mixfdr2fitted.g(temp$output)
    }
    g=temp$output$fitted.g
  }
  data.frame(x=x,y=as.numeric(mixcdf(g,x)),pi0=pi0)
}

plot_mean_cdf = function(SEEDS,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
      PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),pi0filter=FALSE,...){
  PLOTNAMES=PLOTSCENARIOS
  #set up dataframe with cdf for all methods and all datasets
  df= expand.grid(seed=SEEDS,scenario=PLOTSCENARIOS, method=PLOTMETHODS,stringsAsFactors = FALSE)
  df.cdf=ddply(df,.(seed,scenario,method), get_cdf)

  if(pi0filter==TRUE){
    df.cdf %<>% filter(pi0<0.55 & pi0>0.45)
  }
  if(length(SEEDS)>1){
    df.cdf %<>% group_by(x,method,scenario) %>% dplyr::summarise(y=mean(y))
  }

  df.cdf$scenario=factor(df.cdf$scenario,levels=PLOTSCENARIOS)
  levels(df.cdf$scenario)=PLOTNAMES

  ggplot(df.cdf,aes(x=x,y=y,color=method),...) + colScale + geom_line() + facet_grid(.~scenario) + theme(legend.position = "bottom")

}
```

```{r egcdf}
#pdf("../paper/figures/egcdf.pdf",height=3,width=9)
plot_mean_cdf(1)
```

```{r mean_cdf}
#pdf("../paper/figures/mean_cdf.pdf",height=3,width=9)
plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth"),pi0filter=TRUE)
```

```{r mean_cdf_nopen}
#pdf("../paper/figures/mean_cdf_nopen.pdf",height=3,width=9)
names(myColors) <- c("mixfdr.tnull","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr","truth")
colScale <- scale_colour_manual(name = "method",values = myColors)
plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n.s","ash.u.s","ash.hu.s","truth"),pi0filter=TRUE)
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/plot_egdens.Rmd
added plotting Rmd files

---
title: "Plot densities for simulation studies"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r}
library("ashr")
library("ggplot2")
library("dplyr")
library("dscr")
load("../output/dsc-shrink-files/res.RData")
PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
PLOTNAMES = PLOTSCENARIOS
#PLOTFILE = "../paper/figures/scenarios_density.pdf"

df=data.frame()

for(i in PLOTSCENARIOS){
  s=dsc_shrink$scenarios[[i]]
  g=s$args$g
  x = seq(-6,6,length=100)
  y = as.numeric(dens(g,x))
  df = rbind(df,data.frame(x=x,y=y,scenario=i))
}

df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
levels(df$scenario)= PLOTNAMES
```

```{r scenario_density}
#pdf(PLOTFILE,height=3,width=9)
ggplot(df, aes(x=x,y=y)) + geom_line(size=1.2,linetype=1) + facet_grid(.~scenario) + ylab("density")
#dev.off()
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/plot_lfsr.Rmd
add more figure plotting rmd files; tidy up size and cropping of pdf

---
title: "Plot local false sign rate results"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r}
load("../output/dsc-shrink-files/res.RData")
source("../R/set_plot_colors.R")
library(dplyr)

#' @param df dataframe of scores for many methods/scenrios etc
#' @return tall dataframe with columns of scores for each method and the "goldmethod" against which plot is to be made
process_score_for_plotting_against_gold=function(df,PLOTMETHODS=c("ash.n","ash.u","ash.hu"),
                                                 GOLDMETHOD="bayes",PLOTSEEDS=1:100,
                                                 PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),
                                                 PLOTNAMES=PLOTSCENARIOS){
  df %<>% filter(seed %in% PLOTSEEDS) %>% filter(scenario %in% PLOTSCENARIOS) %>% filter(method %in% c(PLOTMETHODS,GOLDMETHOD))
  df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
  levels(df$scenario)= PLOTNAMES
  #create tall version of dataframe
  df %<>% dplyr::select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    reshape2::melt(id.vars=c("method","scenario","seed",".id"),value.name="val")
  #separate bayes and remainder
  df.bayes = df %>% filter(method==GOLDMETHOD)
  df.rest = df %>% filter(method!=GOLDMETHOD)

#join bayes with others, so each line has both the bayes and the non-bayes version
  return(inner_join(df.bayes, df.rest, by=c("scenario","seed","variable")))
}

plot_lfsr=function(lfsr,xlab="True lfsr",ylab="Estimated lfsr",xlim=c(0,0.2),ylim=c(0,0.2),legend.position="bottom"){
  p=ggplot(lfsr,
         aes(val.x,val.y,colour=method.y)) +
    facet_grid(. ~ scenario) + 
    guides(alpha=FALSE) +
    geom_abline(colour = "black") +
    geom_abline(colour= "red", slope=2) +
    xlab(xlab) +
    ylab(ylab) +
    geom_point(shape=1,size=0.1,alpha=0.2)

  p +scale_y_continuous(limits=ylim) +
        scale_x_continuous(limits=xlim)

}

lfsr = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n")
lfdr = process_score_for_plotting_against_gold(res$lfdr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n")

p1=plot_lfsr(lfsr,ylim=c(0,1),xlim=c(0,0.2))
p2=plot_lfsr(lfdr,ylim=c(0,1),xlim=c(0,0.2),xlab="True lfdr",ylab="Estimated lfdr")

# cowplot command: plot_grid(p2+theme_gray()+theme(legend.position="none"),p1+theme_gray()+theme(legend.position="none"),nrow=2,align="v")

```

```{r plot_lfsr, dev='pdf', fig.width=9, fig.height=3, crop=TRUE}
print(p1+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
      +coord_equal(ratio=1/5) + colScale)
#ggsave("../paper/figures/lfsr_est.pdf",height=3,width=9)
```

```{r plot_lfdr, dev='pdf', fig.width=9, fig.height=3, crop=TRUE}
print(p2+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
      +coord_equal(ratio=1/5) + colScale)
#ggsave("../paper/figures/lfdr_est.png",height=3,width=9)
#ggsave("../paper/figures/lfdr_est.pdf",height=3,width=9)
```

```{r plot_lfsr_s, dev='pdf', fig.width=9, fig.height=3, crop=TRUE}
lfsr.s = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n.s")
p1.s=plot_lfsr(lfsr.s,ylim=c(0,1),xlim=c(0,0.2))

print(p1.s+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
      +coord_equal(ratio=1/5))
#ggsave("../paper/figures/lfsr_est_s.png",height=3,width=9)
#ggsave("../paper/figures/lfsr_est_s.pdf",height=3,width=9)
```

```{r plot_lfsr_s_nn, dev='pdf', fig.width=9, fig.height=3, crop=TRUE}
lfsr.s.nn = process_score_for_plotting_against_gold(
    res$lfsr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n.s",
    PLOTSCENARIOS=paste0(c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),"-nn"))
p1.s.nn=plot_lfsr(lfsr.s.nn,ylim=c(0,1),xlim=c(0,0.2))

print(p1.s.nn+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
      +coord_equal(ratio=1/5))
#ggsave("../paper/figures/lfsr_est_s_nn.png",height=3,width=9)

```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 paper/main.tex
add more figure plotting rmd files; tidy up size and cropping of pdf

--- 2015-10-24 docs/main.tex
@@
 \begin{figure}
-\center \includegraphics[height=6in]{../figures/decomp_ZA.pdf}
+\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
 \caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
@@
 \begin{subfigure}{\textwidth}
-	\includegraphics[height=1.8in]{../figures/scenarios_density.pdf} 
+	\includegraphics[width=\textwidth]{../analysis/figure/plot_egdens.Rmd/scenario_density-1.pdf} 
 	\caption{Densities of non-zero effects, $g_1$, used in simulations.} \label{fig:altdens}
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{../figures/pi0_est.pdf} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_pi0est.Rmd/plot_pi0est-1.pdf} 
 \caption{Comparison of true and estimated values of $\pi_0$.
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{../figures/lfdr_est.png} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfdr-1.pdf} 
 \caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{../figures/lfsr_est.png} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr-1.pdf} 
 \caption{As in c), but for $\lfsr$ instead of $\lfdr$.
@@
 \begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{../figures/egcdf.pdf}
+	\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/egcdf-1.png}
 	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{../figures/mean_cdf.pdf}
+\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf-1.png}
 \caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{../figures/mean_cdf_nopen.pdf}
+\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf_nopen-1.png}
 \caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
@@
 \begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all.tex}
+	\centering\input{../analysis/table/scoverage_all.tex}
 	\caption{All observations.
@@
 \begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg.tex}
+\centering\input{../analysis/table/scoverage_neg.tex}
 \caption{``Significant" negative discoveries. Coverage rates are generally satisfactory, except for the uniform-based methods in the spiky and near-normal scenarios,
@@
 \begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos.tex}
+\centering\input{../analysis/table/scoverage_pos.tex}
 \caption{``Significant" positive discoveries. Coverage rates are generally satisfactory, except for the symmetric methods under the asymmetric (``skew") scenario. }
@@
 \begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{../figures/GOODPOOReg_hist.pdf} 
+\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
     \label{fig:goodpoor-hist}
@@
 \begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{../figures/GOODPOOReg_scatter.pdf}
+\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
     \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
@@
 \begin{figure}
-\centering\includegraphics[width=4in]{../figures/lfsr_vs_pval_GOODPOOR.pdf}
+\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
 \caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{../figures/lfsr_est_s_nn.png} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s_nn-1.pdf} 
 \caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=1.8in]{../figures/lfsr_est_s.png} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s-1.pdf} 
 \caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
@@
 \begin{subtable}{\textwidth}
-	\centering\input{tables/scoverage_all_nopen.tex}
+	\centering\input{../analysis/table/scoverage_all_nopen.tex}
 	\caption{All observations}
@@
 \begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_neg_nopen.tex}
+\centering\input{../analysis/table/scoverage_neg_nopen.tex}
 \caption{``Significant" negative discoveries.}
@@
 \begin{subtable}{\textwidth}
-\centering\input{tables/scoverage_pos_nopen.tex}
+\centering\input{../analysis/table/scoverage_pos_nopen.tex}
 \caption{``Significant" positive discoveries.}

### 2015-10-26 analysis/chunk-options.R

@@ -6,3 +6,5 @@ opts_chunk$set(tidy = FALSE,
                comment = NA,
                fig.align = "center",
                fig.path = paste0("figure/", current_input(), "/"))
+
+knit_hooks$set(crop=hook_pdfcrop)

### 2015-10-26 analysis/index.Rmd
added plotting Rmd files
add more figure plotting rmd files; tidy up size and cropping of pdf
add more figures

@@
 
+## Simple illustrations
+
+*  [Illustrate Zero Assumption](makefig_FDReg.html)
+*  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
+
 ## DSC Robust (Simulation study to assess EM convergence)
@@
 ## DSC Shrink (the main simulation study in the paper)
+
+*  [Plot Densities of simulations](plot_egdens.html)
+*  [Plot estimates of pi0](plot_pi0est.html)
+*  [Plot local false sign rates and lfdr vs truth](plot_lfsr.html)
+*  [Plot CDFs](plot_cdf_eg.html)
 

### 2015-10-26 analysis/make_GOODPOOR_figs.Rmd
add more figures

---
title: "Make figures for simple simulation with good and poor precision observations"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

First perform a simple simulated example to illustrate high and low signal.
True values of beta simulated from $(0.5 N(0,1) + 0.5 delta_0)$.
```{r}
ntest = 10000
set.seed(112)
require(ashr)
require(qvalue)
require(locfdr)
require(mixfdr)
source("../R/plot_FDReg_hist.R")

null_alt = rbinom(ntest,1,0.5)
beta = rnorm(ntest)
beta = ifelse(null_alt==1,beta,0)
GOOD= 1:(ntest/2)
sebetahat = rep(1,ntest)
sebetahat[-GOOD] = 10

betahat = beta + rnorm(ntest, 0, sebetahat)
zscore = betahat/sebetahat
pval = pchisq(zscore^2,df=1,lower.tail=F)
```

Show how poor precision observations dilute good ones:
```{r, GOODPOOReg_hist, dev='pdf',fig.width=6.5, fig.height=3}
par(mai=c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
layout(matrix(1:3,ncol=3,byrow=TRUE))
plot_FDReg_hist(pval[GOOD],1,type=1,title="Good-precision observations",ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
plot_FDReg_hist(pval[-GOOD],1,type=1,title="Poor-precision observations",ylab="",nc=20,yaxt='n',cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
axis(side=2, labels=FALSE,tck=-0.01)
plot_FDReg_hist(pval,1,type=1,title="Combined",yaxt='n',ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
axis(side=2, labels=FALSE,tck=-0.01)
```

Apply different methods
```{r}
res.qvalue = qvalue(pval)
res.locfdr = locfdr(zscore,nulltype=0,plot=0)
res.ash = ash(betahat,sebetahat,method="fdr")

res.qvalue.good = qvalue(pval[GOOD])
res.locfdr.good = locfdr(zscore[GOOD],nulltype=0,plot=0)
res.ash.good = ash(betahat[GOOD],sebetahat[GOOD],method="fdr")
```

```{r}
par(mai=c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
layout(matrix(1:3,ncol=3,byrow=TRUE))
plot(res.qvalue.good$q,res.qvalue$q[GOOD],main="qvalue",xlim=c(0,1),ylim=c(0,1),axes=F)
axis(side=2)
axis(side=1)
abline(a=0,b=1,col=2)
plot(res.locfdr.good$fdr,res.locfdr$fdr[GOOD],main="locfdr",xlim=c(0,1),ylim=c(0,1),axes=F)
axis(side=1)
abline(a=0,b=1,col=2)
plot(res.ash.good$lfsr,res.ash$lfsr[GOOD],main="ashr",xlim=c(0,1),ylim=c(0,1),axes=F)
axis(side=1)
abline(a=0,b=1,col=2)
```

```{r GOODPOOReg_scatter, dev='pdf', fig.width=6.5, fig.height=3}
res = rbind(data.frame(x= res.qvalue.good$qvalues, y = res.qvalue$qvalues[GOOD],type="qvalue"),
            data.frame(x=res.locfdr.good$fdr,y=res.locfdr$fdr[GOOD],type='locfdr'), 
            data.frame(x=res.ash.good$lfsr,y=res.ash$lfsr[GOOD],type="ashr"))

library("ggplot2")
pp= ggplot(data=res,aes(x,y)) +geom_point(shape=1) +
  facet_grid(. ~ type) +
  geom_abline(colour = "red") +
  xlab("Analysing good-precision data only") +
  ylab("Analysing combined data")

print(pp +scale_y_continuous(limits=c(0,1)) +
        scale_x_continuous(limits=c(0,1))  +
        coord_equal(ratio=1))

```

```{r lfsr_vs_pval_GOODPOOR, dev='pdf', fig.width=6.5, fig.height=3}
make_df_for_pval=function(ash,method="default"){
  df = data.frame(p=pnorm(-abs(ash$data$betahat/ash$data$sebetahat)),lfsr=ash$lfsr,s=ash$data$sebetahat,method=method)
}

plot_pval_vs_lfsr=function(df,plot.it=TRUE){
  if(length(unique(df$s))>2){df$s = log(df$s)} else{ df$s=as.factor(df$s)}

  p=ggplot(df,aes(x=p,y=lfsr,color=s)) + geom_point() +facet_grid(. ~ method) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr")

  if(length(unique(df$s))>2){
    p = p+scale_colour_gradient2(midpoint = 1, low="blue", mid="white", high="red", space="Lab")
  } 
  if(plot.it){print(p)}
  return(p)
}

res.ash.ET = ash(betahat,sebetahat,method="fdr",model="ET",mixcompdist="normal")
p=plot_pval_vs_lfsr(rbind(
  make_df_for_pval(res.ash,method="Default prior (alpha=0)"),
  make_df_for_pval(res.ash.ET,method="p-value prior (alpha=1)")))
print(p+ theme(axis.text.x = element_text(size = 8,angle=45)) + 
        scale_colour_manual(values=c("blue","red")))

res.ash.ET$logLR
res.ash$logLR

```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 paper/main.tex
add more figures

@@
 \begin{figure}
-\center \includegraphics[height=6in]{figures/decomp_ZA.pdf}
+\center \includegraphics[height=6in]{../analysis/figure/makefig_FDReg.Rmd/decomp_ZA-1.pdf}
 \caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
@@
 \begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_hist.pdf} 
+\centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/GOODPOOReg_hist-1.pdf} 
     \label{fig:goodpoor-hist}
@@
 \begin{subfigure}{\textwidth}
-\centering\includegraphics[width=4in]{figures/GOODPOOReg_scatter.pdf}
+\centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/GOODPOOReg_scatter-1.pdf}
     \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
@@
 \begin{figure}
-\centering\includegraphics[width=4in]{figures/lfsr_vs_pval_GOODPOOR.pdf}
+\centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/lfsr_vs_pval_GOODPOOR-1.pdf}
 \caption{Figure illustrating affects of prior assumptions on re-ordering of significance.

### 2015-10-26 analysis/adaptive_shrinkage.Rmd
add adaptive shrinkage example

---
title: "Illustration of Adaptive Shrinkage"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

The goal here is to illustrate the "adaptive" nature of the adaptive shrinkage. The shrinkage is adaptive in two senses. First,
the amount of shrinkage depends on the distribution $g$ of the true effects, which is learned from the data: when $g$ is very peaked about zero
then ash learns this and deduces that signals should be more strongly shrunk towards zero than when g is less peaked about zero.
Second, the amount of shrinkage of each observation depends on its standard error: the smaller the standard error, the more informative
the data, and so the less shrinkage that occurs. From an Empirical Bayesian perspective both of these 
points are entirely natural: the posterior depends on both the prior
and the likelihood; the prior, $g$, is learned from the data, and the likelihood incorporates the standard error of each observation.

We simulate from two scenarios, the first with effects more peaked about zero, the second with effects less peaked.
```{r initialize}
set.seed(100)
library(ashr)
library(ggplot2)
library(scales)
#load the simulation function, (rnormmix_datamaker)
source("../code/dsc-shrink/datamakers/datamaker.R")

NSAMP=1000
s = 1/rgamma(NSAMP,5,5)
sim.bignormal = rnormmix_datamaker(args=list(g=normalmix(c(1),c(0),c(4)),
                        min_pi0=0,
                        max_pi0=0,
                        nsamp=NSAMP,
                        betahatsd=s
                      ))

sim.spiky = rnormmix_datamaker(args=list(g=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
                        min_pi0=0,
                        max_pi0=0,
                        nsamp=NSAMP,
                        betahatsd=s
                      ))

```

Now run ash on both:
```{r run-ash}
beta.bignormal.ash = ash(sim.bignormal$input$betahat, s)
beta.spiky.ash = ash(sim.spiky$input$betahat, s)
```

Here we plot the shrunken estimates against the observed values, colored
according to the (square root of) precision: precise esitmates being colored red, and less precise estimates being blue. Two key features of the plots
illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large standard error (blue) are shrunk more than estimates with small standard error (red) illustrating that shrinkage adapts to measurement precision.

```{r plotting}
make_df_for_ashplot = function(sim1,sim2,ash1,ash2,name1="spiky",name2="big-normal"){
  df = data.frame(betahat = c(sim1$input$betahat,sim2$input$betahat),
                beta_est = c(ash1$PosteriorMean, ash2$PosteriorMean),
                lfsr = c(ash1$lfsr, ash2$lfsr),
                s=c(sim1$input$sebetahat,sim2$input$sebetahat),
                scenario=c(rep(name1,NSAMP),rep(name2,NSAMP)))
  df$signif = df$lfsr<0.05
  return(df)
}

ashplot=function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate"){
  ggplot(df,aes(x=betahat,y=beta_est,color=1/s)) + 
    xlab(xlab) +
    ylab(ylab) + geom_point() +  facet_grid(.~scenario) +   geom_abline(intercept=0,slope=1) +
  scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab")  +coord_fixed(ratio=1)
}
df=make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash,beta.bignormal.ash)
ashplot(df)
```

Now plot lfsr against z scores, colored
according to the (square root of) precision

```{r plotting lfsr}

z_lfsr_plot=function(df,ylab="Observed Z score",xlab="lfsr"){
  ggplot(df,aes(x=lfsr,y=betahat/s,color=1/s)) + 
    xlab(xlab) +
    ylab(ylab) + geom_point() +  facet_grid(.~scenario) + scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab") 
}

z_lfsr_plot(df)
```

A related consequence is that significance of each observation is no longer monotonic with $p$ value.

```{r}
pval_plot=function(df){
  ggplot(df,aes(x=pnorm(-abs(betahat/s)),y=lfsr,color=log(s))) + geom_point() +  facet_grid(.~scenario) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr") +
  scale_colour_gradient2(midpoint = 0,low="blue", mid="white",high="red")  
}
pval_plot(df)
```

Let's see how these are affected by changing the modelling assumptions so that
the standardized beta are exchangeable (rather than the beta being exchangeable).
```{r run-ash-ET}
beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,model="ET",mixcompdist="normal")
beta.spiky.ash.ET = ash(sim.spiky$input$betahat, s,model="ET",mixcompdist="normal")
df.ET = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ET,beta.bignormal.ash.ET)
ashplot(df.ET,ylab="Shrunken beta estimate (ET model)")
pval_plot(df.ET)
```

This is a "volcano plot" showing effect size against p value. The blue points are "significant" in that they have lfsr<0.05. In this case the significance by lfsr is not quite the same as cutting off at a given $p$ value (you can see that the decision boundary is not quite the same as drawing a horizontal line), but also not that different, presumably because the standard errors, although varying across observations, do not vary greatly.
```{r volcano}
  ggplot(df,aes(x=betahat,y=-log10(2*pnorm(-abs(betahat/s))),col=signif)) + geom_point(alpha=1,size=1.75) +  facet_grid(.~scenario) + theme(legend.position = "none") +
  xlim(c(-10, 10)) + ylim(c(0, 15)) +
  xlab("Effect (beta)") + ylab("-log10 p-value")
```

```{r sehist}
hist(s,main="histogram of standard errors")
summary(s)
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 code/dsc-shrink/addScenarios.R
added dsc-robust and dsc-shrink code

@@ -1,0 +1,114 @@
+source_dir("datamakers")
+
+
+addNamedScenarios = function(dsc,names,nsamp=1000){
+
+#the -nn suffix indicates no nulls (min_pi0 = max_pi0 = 0)
+
+if("near-normal-nn" %in% names){
+	add_scenario(dsc,name="near-normal-nn",
+                    fn=rnormmix_datamaker,
+                    args=list(
+                      g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
+                      min_pi0=0,
+                      max_pi0=0,
+                      nsamp=nsamp,
+                      betahatsd=1
+                      ),
+                    seed=1:100)
+}
+
+if("flat-top-nn" %in% names){
+	add_scenario(dsc,name="flat-top-nn",
+                    fn=rnormmix_datamaker,
+                    args=list(
+                      g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
+                      min_pi0=0,
+                      max_pi0=0,
+                      nsamp=nsamp,
+                      betahatsd=1
+                    ),
+                    seed=1:100)
+}
+
+add_scenario(dsc,name="skew-nn",
+                      fn=rnormmix_datamaker,
+                      args=list(
+                        g=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
+                        min_pi0=0,
+                        max_pi0=0,
+                        nsamp=nsamp,
+                        betahatsd=1
+                      ),
+                      seed=1:100)
+
+add_scenario(dsc,name="near-normal",
+                    fn=rnormmix_datamaker,
+                    args=list(
+                      g=normalmix(c(2/3,1/3),c(0,0),c(1,2)),
+                      min_pi0=0,
+                      max_pi0=1,
+                      nsamp=nsamp,
+                      betahatsd=1
+                    ),
+                    seed=1:100)
+
+add_scenario(dsc,name="flat-top",
+                    fn=rnormmix_datamaker,
+                    args=list(
+                      g=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
+                      min_pi0=0,
+                      max_pi0=1,
+                      nsamp=nsamp,
+                      betahatsd=1
+                    ),
+                    seed=1:100)
+
+
+add_scenario(dsc,name="skew",
+                      fn=rnormmix_datamaker,
[…]

### 2015-10-26 code/dsc-shrink/addScenarios.null.R

@@ -1,0 +1,17 @@
+source_dir("datamakers")
+
+
+#Now, for each scenario create an element of scenarios of the following form
+add_scenario(dsc_znull,name="znull.1000",
+            fn=null_z_datamaker,
+            args=list(
+              nsamp=1000
+            ),
+            seed=1:500)
+
+add_scenario(dsc_znull,name="znull.100",
+            fn=null_z_datamaker,
+            args=list(
+              nsamp=100
+            ),
+            seed=1:500)

### 2015-10-26 code/dsc-shrink/add_named_scenarios.R

@@ -1,0 +1,91 @@
+source_dir("datamakers")
+
+gdef = list(spiky=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
+            skew=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),
+            bignormal=normalmix(c(1),c(0),c(4)),
+            bimodal=normalmix(c(0.5,0.5),c(-2,2),c(1,1)),
+            flat_top=normalmix(rep(1/7,7),c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7)),
+            near_normal=normalmix(c(2/3,1/3),c(0,0),c(1,2))
+            )
+
+add_named_scenarios = function(dsc,names,nsamp=1000,min_pi0=0,max_pi0=1,suffix=""){
+
+if("near-normal" %in% names){
+  add_scenario(dsc,name=paste0("near-normal",suffix),
+                    fn=rnormmix_datamaker,
+                    args=list(
+                      g=gdef$near_normal,
+                      min_pi0=min_pi0,
+                      max_pi0=max_pi0,
+                      nsamp=nsamp,
+                      betahatsd=1
+                    ),
+                    seed=1:100)
+}
+
+if("flat-top" %in% names){
+  add_scenario(dsc,name=paste0("flat-top",suffix),
+                    fn=rnormmix_datamaker,
+                    args=list(
+                      g=gdef$flat_top,
+                      min_pi0=min_pi0,
+                      max_pi0=max_pi0,
+                      nsamp=nsamp,
+                      betahatsd=1
+                    ),
+                    seed=1:100)
+}
+
+if("skew" %in% names){
+  add_scenario(dsc,name=paste0("skew",suffix),
+                      fn=rnormmix_datamaker,
+                      args=list(
+                        g=gdef$skew,
+                        min_pi0=min_pi0,
+                        max_pi0=max_pi0,
+                        nsamp=nsamp,
+                        betahatsd=1
+                      ),
+                      seed=1:100)
+}
+
+if("spiky" %in% names){
+  add_scenario(dsc,name=paste0("spiky",suffix),
+                      fn=rnormmix_datamaker,
+                      args=list(
+                        g=gdef$spiky,
+                        min_pi0=min_pi0,
+                        max_pi0=max_pi0,
+                        nsamp=nsamp,
+                        betahatsd=1
+                      ),
+                      seed=1:100)
+}
[…]

### 2015-10-26 code/dsc-shrink/datamakers/datamaker.R

@@ -1,0 +1,41 @@
+library(ashr)
+
+#' @title datamaker for FDR DSC
+#'
+#' @description Simulates data for a DSC for methods to compute FDR/q values  
+#' @details None
+#' 
+#' @param args A list of the remaining arguments, which in this case is
+#' \item{nsamp}{The number of samples to create}
+#' \item{g}{An object of class normalmix specifying the mixture distribution from which non-null beta values 
+#' are to be simulated}
+#' \item{min_pi0}{The minimum value of pi0, the proportion of true nulls}
+#' \item{max_pi0}{The maximum value of pi0, the proportion of true null}
+#' \item{betahatsd}{The standard deviation of betahat to use}
+#'
+#' @return a list with the following elements
+#' \item{meta}{A list containing the meta data. In this case beta}
+#' \item{input}{A list containing the input data; in this case the set of betahat values and their standard errors}
+#' 
+rnormmix_datamaker = function(args){
+  #here is the meat of the function that needs to be defined for each dsc to be done
+  pi0 = runif(1,args$min_pi0,args$max_pi0) #generate the proportion of true nulls randomly
+  
+  k = ncomp(args$g)
+  comp = sample(1:k,args$nsamp,mixprop(args$g),replace=TRUE) #randomly draw a component
+  isnull = (runif(args$nsamp,0,1) < pi0)
+  beta = ifelse(isnull, 0,rnorm(args$nsamp,comp_mean(args$g)[comp],comp_sd(args$g)[comp]))
+  sebetahat = args$betahatsd
+  betahat = beta + rnorm(args$nsamp,0,sebetahat)
+  meta=list(g1=args$g,beta=beta,pi0=pi0)
+  input=list(betahat=betahat,sebetahat=sebetahat,df=NULL)
+  
+  #end of meat of function
+  
+  data = list(meta=meta,input=input)
+  
+  return(data)
+  
+}
+
+

### 2015-10-26 code/dsc-shrink/methods/mixfdr.wrapper.R

@@ -1,0 +1,34 @@
+#' @title wrapper for mixfdr for shrinkage DSC
+#'
+#' @description Runs mixfdr to compute effect estimates from z scores (computed from betahat and sebetahat)  
+#' @details None
+#' 
+#' @param input a list with elements betahat and sebetahat
+#' @param args a list containing other additional arguments to ash
+#' 
+#' @return output a list with the following elements
+#' \item{beta_est}{vector containing point estimates for beta}
+#' \item{pi0_est}{scalar estimate of pi0}
+#'
+library(mixfdr)
+
+mixfdr.wrapper=function(input,args=NULL){
+  res = try(do.call(mixFdr, args= c(list(x=input$betahat/input$sebetahat, noiseSD=1, plots=FALSE),args)))
+  if(inherits(res,"try-error")){res=list(effectSize=rep(NA,length(input$betahat)),pi0_est=NA,pi=NA,mu=NA,sigma=NA,noiseSD=NA)}
+  return(list(res=res,input=input))
+}
+
+mixfdr2beta_est = function(output){
+  return (list(beta_est=output$res$effectSize*output$input$sebetahat))
+}
+
+mixfdr2pi0_est = function(output){
+  return (list(pi0_est=output$res$pi[1]))
+}
+
+mixfdr2fitted.g = function(output){
+  return (list(fitted.g=
+    normalmix(output$res$pi,output$res$mu,
+              sqrt(output$res$sigma^2-output$res$noiseSD^2))))
+}
+

### 2015-10-26 analysis/summarize_coverage.Rmd
add .Rprofile

@@
 
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_neg.tex")
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_pos.tex")
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/coverage_all.tex")
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/coverage_neg.tex")
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/coverage_pos.tex")
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/coverage_all.tex")
 
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_neg_nopen.tex")
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_pos_nopen.tex")
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/coverage_all_nopen.tex")
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/coverage_neg_nopen.tex")
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/coverage_pos_nopen.tex")
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/coverage_all_nopen.tex")
 ```
@@
 ```{r}
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_neg.tex",switch=TRUE)
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_pos.tex",switch=TRUE)
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"../paper/tables/scoverage_all.tex",switch=TRUE)
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/scoverage_neg.tex",switch=TRUE)
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/scoverage_pos.tex",switch=TRUE)
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/scoverage_all.tex",switch=TRUE)
 
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_neg_nopen.tex",switch=TRUE)
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_pos_nopen.tex",switch=TRUE)
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"../paper/tables/scoverage_all_nopen.tex",switch=TRUE)
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/scoverage_neg_nopen.tex",switch=TRUE)
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/scoverage_pos_nopen.tex",switch=TRUE)
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/scoverage_all_nopen.tex",switch=TRUE)
 ```

### 2015-10-26 analysis/check_mixfdr_lownoise.Rmd

---
title: "Check mixfdr in low noise simulation"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

Here we check that mixfdr actually does OK at fitting the bimodal distribution when 
noise is small compared with bimodality.

First simulate bimodal data with well-separated peaks. The histogram
shows the multi-modal nature of the underlying g is clear.
```{r}
  source("../code/dsc-shrink/datamakers/datamaker.R")
  sim.bimodal=rnormmix_datamaker(
              args=list(g=normalmix(c(0.5,0.5),c(-5,5),c(1,1)),
              min_pi0=0.4,
              max_pi0=0.4,
              nsamp=1000,
              betahatsd=1)
            )
  hist(sim.bimodal$input$betahat,nclass=20,xlim=c(-10,10),prob=TRUE,main="histogram of simulated betahat values")
```

Now run mixfdr on it. Note that the fit captures the multi-modal nature of g, 
although mixfdr continues to overestimate pi0.
```{r}
  source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
  sim.bimodal.mixfdroutput = mixfdr.wrapper(sim.bimodal$input,args=list(theonull=TRUE))
  g = mixfdr2fitted.g(sim.bimodal.mixfdroutput)$fitted.g
  x=seq(-10,10,length=100)
  plot(ecdf(sim.bimodal$meta$beta),main="Cdf of true beta: empirical (black) and fitted (red)")
  lines(x,mixcdf(g,x),col=2)
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/index.Rmd

@@
 *  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
+*  [Adpative Shrinkage](adaptive_shrinkage.html)
 

### 2015-10-26 analysis/investigate_hu_badcoverage.Rmd

---
title: "Investigate poor coverage for ash.hu"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r}
load("../output/dsc-shrink-files/res.RData")

library("dscr")
library("ashr")
library(reshape2)
library(ggplot2)
library(dplyr)

neglong = 
  res$negprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="negprob") %>%
    filter(negprob > 0.95)

poslong = 
  res$posprob %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="posprob") %>%
    filter(posprob > 0.95)

reslong = 
  res$cdf_score %>% 
    select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%    
    melt(id.vars=c("method","scenario","seed",".id"))

reslong.pos = inner_join(reslong,poslong)
reslong.neg = inner_join(reslong,neglong)

```

```{r}
ash.hu.ft = reslong.pos %>% dplyr::filter(method=="ash.hu.s" & scenario=="flat-top")
head(ash.hu.ft)
```

```{r}
  eg=load_example(dsc_shrink,42,"flat-top",method="ash.hu.s","../output")
  eg$output$fitted.g
```

Notice how almost all the inferred weight is on a small positive component. As a result false sign rate will be small, and there will be a strong tendency
to overestimate 0 effects. This leads to coverage problems observed.

Now let's look at an example with u and spiky which seems to be somewhat badly calibrated for the negative discoveries. 
```{r}
ash.u.s.spiky = reslong.neg %>% dplyr::filter(method=="ash.u.s" & scenario=="spiky")
ash.n.s.spiky = reslong.neg %>% dplyr::filter(method=="ash.n.s" & scenario=="spiky")
hist(ash.u.s.spiky$value,nclass=100,main="histogram of quantile where observation falls in its CI; significant negative discoveries")
hist(ash.n.s.spiky$value,nclass=100,main="histogram of quantile where observation falls in its CI; significant negative discoveries")
```

So what seems to be happening here is that the uniform tail is too short; when observation falls outside of this tail it gets a zero quantile of posterior interval.

Can we find an example?
```{r}
  table((ash.u.s.spiky %>% filter(value==0))$seed)
  eg =load_example(dsc_shrink,seed = 39,method="ash.u.s",scenario="spiky","../output")
```

For comparison, here are the positive discoveries; here there is not too bad.
```{r}
ash.u.s.spiky = reslong.pos %>% dplyr::filter(method=="ash.u.s" & scenario=="spiky")
ash.n.s.spiky = reslong.pos %>% dplyr::filter(method=="ash.n.s" & scenario=="spiky")
hist(ash.n.s.spiky$value)
hist(ash.u.s.spiky$value)
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/makefig_FDReg.Rmd

---
title: "Some figures to illustrate basic ideas in FDR"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

This file makes two figures:

1. illustrates basic idea of way q value does FDR analysis in p value space
2. compares methods in the way they decompose p values and z scores into two groups

```{r}
require(ashr)
require(qvalue)
require(locfdr)
require(mixfdr)
source("../R/plot_FDReg_hist.R") #some plotting functions
source("../R/nullalthist.R")

############################ FDReg_hist.pdf ##################################################################

#simple simulated example
ncz = 100 # number of bins in z score histograms
ntest = 10000
set.seed(111)

#simulate (with all tests alternative, true effects $\beta \sim N(0,1)$)
beta = rnorm(ntest)
sebetahat = 1
betahat = beta + rnorm(ntest, 0, sebetahat)
zscore = betahat/sebetahat
pval = pchisq(zscore^2,df=1,lower.tail=F)

#apply the different methods
res.qvalue = qvalue(pval)
res.locfdr = locfdr(zscore,nulltype=0,plot=0)
res.mixfdr = mixFdr(zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
res.ash = ash(betahat,1,method="fdr")

#roughly compute a local fdr for qvalue (to aid the plotting of the decomposition in zscore space)
temp=hist(pval,nclass=50)
bin_fdr = res.qvalue$pi0/temp$density #compute local fdr in each bin of histogram
qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
qval_fdr = pmin(1,qval_fdr) #threshold at 1, so fdr never bigger than 1
qval_fdr = qvalue::lfdr(pval)
```

Here's an example of how qvalue decomposes $p$ values into null and alternative components
```{r}
#pdf("../figures/FDReg_hist.pdf",width=6,height=3)
par(mai=c(0.5,0.5,0.2,0.2),mgp = c(3, 0.5, 0))
layout(matrix(1:2,ncol=2,byrow=TRUE))
plot_FDReg_hist(pval,res.qvalue$pi0,type=1,title="p values",cex.axis=0.8,cex.main=0.8)
plot_FDReg_hist(pval,res.qvalue$pi0,type=4,cex.axis=0.8,yaxt='n',textsize=0.9,cex.main=0.8,title="Decomposition into null/alternative")
axis(side=2, labels=FALSE,tck=-0.01)
```

############################## decomp_ZA.pdf #############################################
```{r decomp_ZA, dev='pdf', fig.width=6, fig.height=6, crop=TRUE}
#pdf("../figures/decomp_ZA.pdf",width=6,height=6)
layout(matrix(1:12,ncol=3,byrow=FALSE))
plotlabel=function(label,cex=1.5){
  plot(0,0,type="n",axes=FALSE,xlab="",ylab="")
  text(0,0,label,cex=cex)
}
plotlabel("qvalue")
plotlabel("locfdr")
plotlabel("mixfdr")
plotlabel("ashr")

par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))

#p value histograms
altnullhist(pval,qval_fdr,main="p values",ncz=50,xaxt='n',cex.axis=0.8)
#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"p values",line=1)
altnullhist(pval,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.ash$lfdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"p values",line=2)

#z score histograms
nullalthist(zscore,qval_fdr,main="z scores",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"z scores",line=1)
nullalthist(zscore,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.ash$lfdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"z scores",line=2)

#dev.off()

```

This one is a different layout (4 columns, 2 rows) for my Tukey poster

```{r decomp_ZA_poster, dev='pdf', fig.width=6, fig.height=2, crop=TRUE}
#pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
layout(matrix(1:8,ncol=4,byrow=TRUE))
par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))
ncz=25

#p value histograms
altnullhist(pval,qval_fdr,main="p values: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8)
#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"p values",line=1)
altnullhist(pval,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.ash$lfdr,main="ashr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"p values",line=2)

#z score histograms
nullalthist(zscore,qval_fdr,main="z scores: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"z scores",line=1)
nullalthist(zscore,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.ash$lfdr,main="ashr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"z scores",line=2)

dev.off()
```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 analysis/plot_cdf_eg.Rmd

@@
 
-```{r egcdf}
+```{r egcdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
 #pdf("../paper/figures/egcdf.pdf",height=3,width=9)
@@
 
-```{r mean_cdf}
+```{r mean_cdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
 #pdf("../paper/figures/mean_cdf.pdf",height=3,width=9)
@@
 
-```{r mean_cdf_nopen}
+```{r mean_cdf_nopen, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
 #pdf("../paper/figures/mean_cdf_nopen.pdf",height=3,width=9)

### 2015-10-26 analysis/plot_egdens.Rmd

@@
 
-```{r scenario_density}
+```{r scenario_density, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
 #pdf(PLOTFILE,height=3,width=9)

### 2015-10-26 analysis/plot_pi0est.Rmd

---
title: "Plot estimates of pi0"
author: "Matthew Stephens"
date: 2015-10-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r plot_pi0est, dev='pdf', crop=TRUE}
load("../output/dsc-shrink-files/res.RData")
source("../R/set_plot_colors.R")
library(dplyr)

PLOTMETHODS= c("mixfdr.tnull","ash.n","ash.u","qvalue","locfdr")
PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
PLOTNAMES = PLOTSCENARIOS
ALPHALEVEL = 0.8 # controls transparency

df = res$pi0_score %>% filter(scenario %in% PLOTSCENARIOS) %>% filter(method %in% PLOTMETHODS)
df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
levels(df$scenario)= PLOTNAMES

pi0_plot=ggplot(df,
         aes(pi0,pi0_est,colour=method,alpha=ALPHALEVEL)) +geom_point(shape=1) +
  facet_grid(. ~ scenario) + 
  guides(alpha=FALSE) +
  geom_abline(colour = "black") +
  xlab("True pi0") +
  ylab("Estimated pi0") 
print(pi0_plot +scale_y_continuous(limits=c(0,1)) +
        scale_x_continuous(limits=c(0,1)) +
        #scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
        coord_equal(ratio=1) + colScale + theme(legend.position = "top",axis.text.x = element_text(size = 8,angle=45)))

```

## Session information

```{r info}
sessionInfo()
```

### 2015-10-26 paper/main.tex

@@
 \begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/egcdf-1.png}
+	\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/egcdf-1.pdf}
 	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf-1.png}
+\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf-1.pdf}
 \caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf_nopen-1.png}
+\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf_nopen-1.pdf}
 \caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.

### 2015-10-26 R/packages/mixfdr/R/dens.mixture.R
add mixfdr package code

@@ -1,0 +1,9 @@
+`dens.mixture` <-
+function(x,m){
+	mu = m$mu
+	pi = c(m$pi)
+	sig = m$sig
+	phiMat = matDens(x,mu,sig)
+	return((phiMat %*% pi)[,1])
+}
+

### 2015-10-26 R/packages/mixfdr/R/effectSize.R

@@ -1,0 +1,24 @@
+`effectSize` <-
+function(z, m, noiseSD = NA){
+	e = m
+	if(!is.null(m$noiseSD) && is.na(noiseSD)){
+		noiseSD = m$noiseSD
+	}else{
+		if(is.na(noiseSD)) noiseSD = min(m$sig)	
+	}
+	sig = noiseSD
+	if(any(e$sig<sig)) warning(paste("Marginal variance of a group is less than noise variance, this is impossible under the model."))
+	gpProb = groupProbs(z,e)
+	priorVar = pmax(e$sig^2 - sig^2,0)
+	postVars = 1/((1/priorVar) + (1/sig^2)) * (priorVar != 0)
+	postMeans =  matrix(z, length(z),length(e$mu),byrow=FALSE)/sig^2
+	postMeans = postMeans + matrix((e$mu)/(priorVar),length(z),length(e$mu),byrow=TRUE)
+	postMeans = postMeans * matrix(postVars, length(z), length(e$mu), byrow= TRUE)
+	postMeans[is.nan(postMeans)] = 0
+	postMeans = postMeans + matrix((priorVar ==0)*e$mu, length(z), length(e$mu), byrow = TRUE)
+	postMeansSq = postMeans^2 + matrix(postVars, length(z), length(e$mu), byrow = TRUE)
+	EdeltZ = rowSums(postMeans * gpProb)
+	VdeltZ = rowSums(postMeansSq * gpProb) - EdeltZ^2
+	return(cbind(Edz = EdeltZ, Vdz = VdeltZ))
+}
+

### 2015-10-26 R/packages/mixfdr/R/mixfdr-internal.R

@@ -1,0 +1,132 @@
+`.Random.seed` <-
+c(403L, 248L, 1943584748L, -496199485L, -1662213855L, -1827092946L, 
+503020944L, 2119163611L, -435693445L, -79399090L, 272546151L, 
+892623469L, -548975477L, -310206652L, -1057076345L, 365351663L, 
+-1841157002L, 1614520170L, -865830356L, -251601403L, -1126626475L, 
+-1648630717L, -1065959009L, 2062140742L, 1756551190L, 673932921L, 
+246000368L, 2054884061L, -1027835535L, 914892765L, 1443580177L, 
+926191030L, -1450832567L, 109956025L, -319842146L, -318253603L, 
+-1052503858L, -137922542L, -1069537469L, 1631277097L, -280187193L, 
+645345071L, 1472421183L, 399084212L, 440392314L, -1319400811L, 
+-403958896L, -438684076L, 1076551163L, 1471432323L, -1049899524L, 
+826242072L, 2113813555L, 2055453402L, 1840843051L, -852522404L, 
+1503467528L, -451879156L, -1150923382L, -1605148649L, -1068998860L, 
+1276685149L, -248384192L, 1397814493L, 1295850984L, -465239085L, 
+-1968950399L, 902887758L, -2038690635L, -784073061L, -1321936265L, 
+1492105735L, 2032639324L, 1094450198L, -1973440144L, 1797700546L, 
+-2053042223L, -1624730408L, -1831821204L, 1251297935L, -1644939652L, 
+-1481336848L, 1595801398L, 1353398153L, -1290681255L, 1823937610L, 
+-1172797874L, 690003959L, 1981221359L, 1438408164L, -688767515L, 
+1032973279L, -1832275611L, 324056891L, -265082085L, 1493949096L, 
+-2127522055L, -771079036L, 558462968L, -258961249L, -543972256L, 
+-548427812L, 184016945L, 490059559L, -448510139L, 770836711L, 
+-985705915L, 464959828L, 711899353L, 1933082509L, 1540187078L, 
+1696785471L, 444032146L, -82446942L, 20634834L, 1482283067L, 
+-1023037433L, -350898326L, -1043228303L, -1058079456L, 1775229802L, 
+-906141168L, -306096035L, -2043860621L, -1787347563L, -579438570L, 
+-311934740L, 1380177713L, 2094646236L, 1982887694L, 1217439581L, 
[…]

### 2015-10-26 R/packages/mixfdr/R/plot.mix.dens.R

@@ -1,0 +1,16 @@
+`plot.mix.dens` <-
+function(x, plot.subdens = TRUE, ...){
+	z = x$data
+	mi = min(z)
+	ma = max(z)
+	s = seq(mi - 1, ma + 1, by = 0.001)
+	hist(z, pr = TRUE, ...)
+	lines(s, dens.mixture(s,x), lwd = 2)
+	J = length(x$mu)
+	if(plot.subdens == TRUE){
+	for(j in 1:J){
+		lines(s, x$pi[j] * dnorm(s, x$mu[j],x$sigma[j]),col=min(3, j+1), lwd = 2)
+	}	
+	}
+}
+

### 2015-10-26 R/packages/mixfdr/R/tailFDRMixModel.R

@@ -1,0 +1,29 @@
+`tailFDRMixModel` <-
+function(z, m, nullGroups = NA){
+	e = m
+	if(any(is.na(nullGroups))){
+		 nullGroups = rep(FALSE, length(e$mu))
+		 nullGroups[1] = TRUE
+	}
+	# for each group, compute P(|z|>z)
+	# P(Z > z) + P(Z < -z)
+	# P(Z > (z - mu)/sigma) + P(Z < (-z - mu)/sigma)
+	
+	A = matrix(z, length(z), length(e$mu))
+	ri = A - matrix(e$mu, length(z), length(e$mu), byrow = TRUE)
+	ri = ri/matrix(e$sig, length(z), length(e$sig), byrow = TRUE)
+	le = -A - matrix(e$mu, length(z), length(e$mu), byrow = TRUE)
+	le = le/matrix(e$sig, length(z), length(e$sig), byrow = TRUE)
+	pRi = 1 - pnorm(ri)
+	pLe = pnorm(le)
+	pMix2sd = (pRi + pLe)*matrix(e$pi, length(z), length(e$pi), byrow = TRUE)
+	nullProb = rowSums(matrix(pMix2sd[,nullGroups],ncol = sum(nullGroups)))
+	overallProb = rowSums(pMix2sd)
+	twoSided = ((nullProb/overallProb)*(z>=0) + rep(sum(e$pi[nullGroups]), length(z)) *(z<0))
+	ppRi = pRi*matrix(e$pi, length(z), length(e$pi), byrow = TRUE)
+	right = rowSums(matrix(ppRi[,nullGroups],ncol = sum(nullGroups)))/rowSums(ppRi)
+	ppRi = pnorm(ri)*matrix(e$pi, length(z), length(e$pi), byrow = TRUE)
+	left = rowSums(matrix(ppRi[,nullGroups],ncol = sum(nullGroups)))/rowSums(ppRi)
+	return(list(FDRTwoSided = twoSided, FDRleft = left, FDRright = right))
+}
+

### 2015-10-27 R/packages/mixfdr/R/fdrMixModel.R
add mixfdr

@@ -1,0 +1,11 @@
+`fdrMixModel` <-
+function(z,m, nullGroups = NA){
+	e = m
+	if(any(is.na(nullGroups))){
+		 nullGroups = rep(FALSE, length(e$mu))
+		 nullGroups[1] = TRUE
+	}
+	G = groupProbs(z,e)
+	rowSums(matrix(G[,nullGroups],ncol = sum(nullGroups)))
+}
+

### 2015-10-27 R/packages/mixfdr/R/getStartsNew.R

@@ -1,0 +1,23 @@
+`getStartsNew` <-
+function(x,J, noiseSD = 1){
+	# going to have equally
+	# one big group in the middle
+	pi = rep(1/J,J)
+	mid = floor(median(1:J))
+	pi[mid] = 0.75
+	if(J!=1) pi[-mid] = 0.25/(J-1)
+	pi = pi/sum(pi)
+	q = quantile(x, c(0,cumsum(pi)))	
+	mu = sig = 1:J
+	for(i in 1:J){
+		dat = x[x<=q[i+1] & x>=q[i]]
+		mu[i] = mean(dat)
+		sig[i] = noiseSD
+	}
+	starts = cbind(pi,mu,sig)
+	tmp = starts[mid,]
+	starts[mid,]= starts[1,]
+	starts[1,] = tmp
+	return(starts)
+}
+

### 2015-10-27 R/packages/mixfdr/R/matDens.R

@@ -1,0 +1,13 @@
+`matDens` <-
+function(x, mu, sig){
+	n = length(x)
+	J = length(mu)
+	res = matrix(x, n, J) - matrix(mu, n, J, byrow = TRUE)
+	ss = matrix(1/sig,n,J,byrow=TRUE)
+	res = res * ss # divide each column j by sigma[j]
+	res = dnorm(0) * exp(-0.5 * res^2) #stupid pi renaming nonsense
+	res = res * ss # divide again
+	return(res)
+#	return(matDensC(x,mu,sig))
+}
+

### 2015-10-27 R/packages/mixfdr/R/mixModelFitter.R

@@ -1,0 +1,61 @@
+`mixModelFitter` <-
+function(x, J, N, starts, maxiter, tol, p, muInt, sigInt){
+
+				
+		EMstep = function(x, pi, mu, sig, N, p, muInt, sigInt){
+				n = length(x)
+				J = length(mu)
+				
+				# E-step	
+				phiMat = matDens(x,mu,sig)
+				denominators = (phiMat %*% pi)[,1]
+				G = phiMat * ((1/denominators) %o% pi)
+				tG = t(G)
+				
+				# M-step
+				numeratorsMu = (tG %*% x)[,1]
+				zplusj = rowSums(tG)
+				newMu = (numeratorsMu/zplusj)
+				newSig = tG %*% (x*x) / zplusj - newMu^2
+				
+				newMu = pmax(pmin(muInt[,2],newMu),muInt[,1])
+				newSig = pmax(pmin(sigInt[,2], sqrt(newSig)), sigInt[,1])
+				newPi = (N*p - 1 + zplusj)/(sum(zplusj) + N - J)
+				if(N<0) newPi = zplusj/sum(zplusj)
+				newPi = (newPi+0.001)
+				newPi = pmax(0,pmin(newPi,1))
+				newPi = newPi/sum(newPi)
+				if(any(is.na(newPi), is.na(newMu), is.na(newSig))){
+					print(pi)
+					print(mu)
+					print(sig)	
+				}
+				return(list(pi = newPi, mu = newMu, sigma = newSig)) 
+			}
+		
+	# starting points	
+		pi = starts[,1]
+		mu = starts[,2]
+		sigma = starts[,3]
+	# that's it, do the loop	
+		iter = 0
+		converged = FALSE
+		while(!converged && iter<maxiter){
+			res = EMstep(x, pi, mu, sigma, N, p, muInt, sigInt)
+			distance = sum((pi-res$pi)^2)+sum((mu-res$mu)^2) + sum((sigma-res$sigma)^2)
+			if(is.na(distance)){
+				print(starts)
+				print(muInt)
+				print(sigInt)	
+			}
+			if(distance<tol) converged = TRUE
+			iter = iter + 1			
+			pi = res$pi
+			mu = res$mu
+			sigma = res$sigma
+		}
+#		e = list(pi = pi, mu = mu, sigma = sigma, converged = converged, nIter = iter, data = x)
[…]

### 2015-10-27 R/packages/mixfdr/R/dens.mixture.R
update README

@@ -1,9 +1,0 @@
-`dens.mixture` <-
-function(x,m){
-	mu = m$mu
-	pi = c(m$pi)
-	sig = m$sig
-	phiMat = matDens(x,mu,sig)
-	return((phiMat %*% pi)[,1])
-}
-

### 2015-10-27 R/packages/mixfdr/R/fdrMixModel.R

@@ -1,11 +1,0 @@
-`fdrMixModel` <-
-function(z,m, nullGroups = NA){
-	e = m
-	if(any(is.na(nullGroups))){
-		 nullGroups = rep(FALSE, length(e$mu))
-		 nullGroups[1] = TRUE
-	}
-	G = groupProbs(z,e)
-	rowSums(matrix(G[,nullGroups],ncol = sum(nullGroups)))
-}
-

### 2015-10-27 R/packages/mixfdr/R/getStartsNew.R

@@ -1,23 +1,0 @@
-`getStartsNew` <-
-function(x,J, noiseSD = 1){
-	# going to have equally
-	# one big group in the middle
-	pi = rep(1/J,J)
-	mid = floor(median(1:J))
-	pi[mid] = 0.75
-	if(J!=1) pi[-mid] = 0.25/(J-1)
-	pi = pi/sum(pi)
-	q = quantile(x, c(0,cumsum(pi)))	
-	mu = sig = 1:J
-	for(i in 1:J){
-		dat = x[x<=q[i+1] & x>=q[i]]
-		mu[i] = mean(dat)
-		sig[i] = noiseSD
-	}
-	starts = cbind(pi,mu,sig)
-	tmp = starts[mid,]
-	starts[mid,]= starts[1,]
-	starts[1,] = tmp
-	return(starts)
-}
-

### 2015-10-27 R/packages/mixfdr/R/matDens.R

@@ -1,13 +1,0 @@
-`matDens` <-
-function(x, mu, sig){
-	n = length(x)
-	J = length(mu)
-	res = matrix(x, n, J) - matrix(mu, n, J, byrow = TRUE)
-	ss = matrix(1/sig,n,J,byrow=TRUE)
-	res = res * ss # divide each column j by sigma[j]
-	res = dnorm(0) * exp(-0.5 * res^2) #stupid pi renaming nonsense
-	res = res * ss # divide again
-	return(res)
-#	return(matDensC(x,mu,sig))
-}
-

### 2015-10-27 R/packages/mixfdr/R/mixFdr.R

@@ -1,160 +1,0 @@
-`mixFdr` <-
-function(x, J = 3, P = NA, noiseSD = NA, theonull = FALSE, calibrate = FALSE, plots= TRUE, nearlyNull = 0.2, starts = NA, p = NA, maxIter = 1000, tol = 0.001, nocheck = FALSE){
-# try lower sigma threshhold	
-# warn if
-	# uncalibrated
-	# empirical null is substantially different
-	# pi0 is small
- 
-# return
-	# model parameters
-	# effect size estimates
-	# fdr estimates
-	# plot
-	
-# calibration code
-		
-		muInt = sigInt = NA
-		sigma = noiseSD
-		n = length(x)	
-		if(any(is.na(p))) p = c(1, rep(0,J-1))	
-		if(any(is.na(muInt))){
-			muInt = matrix(0,J,2)
-			muInt[,1] = -Inf
-			muInt[,2] = Inf		
-		}  	
-		if(any(is.na(sigInt)) | any(is.null(dim(sigInt)))){
-			left = sigma
-			if(is.na(sigma)) left = 0.1
-			sigInt = matrix(0,J,2)
-			sigInt[,1] = left
-			sigInt[,2] = Inf
-		}
-		if(theonull){
-			muInt[1,1] = muInt[1,2] = 0
-			sigInt[1,1] = sigInt[1,2] = 1	
-		}
-		if(is.na(P)){
-			P = length(x)/5
-			if(!calibrate && !theonull){
-				warning("Using uncalibrated default penalization P,  which can give misleading results for empirical nulls. Consider rerunning with calibrate = TRUE and using the resulting penalization","\n")
-				}
-		}
-		if(!is.na(sigma)){
-			warning(paste("Assuming known noise noiseSD = ",sigma,". If needed rerun with noiseSD = NA to fit noiseSD."))
-			
-		if(!theonull){
-			warning("Note that using known noiseSD constrains the null to have sd at least noiseSD. If underdispersion is suspected, rerun with noiseSD = NA.")
-			}
-		}
-		
-		cat("Fitting preliminary model", "\n")
-		m =  mixModelManyStarts(x, J, P, starts, maxIter, tol, p, muInt, sigInt, theonull)
-		m$noiseSD = min(median(abs(x-m$mu[1]))/0.6745, m$sig)
-		if(calibrate == TRUE){
[…]

### 2015-10-27 R/packages/mixfdr/R/mixfdr-internal.R

@@ -1,132 +1,0 @@
-`.Random.seed` <-
-c(403L, 248L, 1943584748L, -496199485L, -1662213855L, -1827092946L, 
-503020944L, 2119163611L, -435693445L, -79399090L, 272546151L, 
-892623469L, -548975477L, -310206652L, -1057076345L, 365351663L, 
--1841157002L, 1614520170L, -865830356L, -251601403L, -1126626475L, 
--1648630717L, -1065959009L, 2062140742L, 1756551190L, 673932921L, 
-246000368L, 2054884061L, -1027835535L, 914892765L, 1443580177L, 
-926191030L, -1450832567L, 109956025L, -319842146L, -318253603L, 
--1052503858L, -137922542L, -1069537469L, 1631277097L, -280187193L, 
-645345071L, 1472421183L, 399084212L, 440392314L, -1319400811L, 
--403958896L, -438684076L, 1076551163L, 1471432323L, -1049899524L, 
-826242072L, 2113813555L, 2055453402L, 1840843051L, -852522404L, 
-1503467528L, -451879156L, -1150923382L, -1605148649L, -1068998860L, 
-1276685149L, -248384192L, 1397814493L, 1295850984L, -465239085L, 
--1968950399L, 902887758L, -2038690635L, -784073061L, -1321936265L, 
-1492105735L, 2032639324L, 1094450198L, -1973440144L, 1797700546L, 
--2053042223L, -1624730408L, -1831821204L, 1251297935L, -1644939652L, 
--1481336848L, 1595801398L, 1353398153L, -1290681255L, 1823937610L, 
--1172797874L, 690003959L, 1981221359L, 1438408164L, -688767515L, 
-1032973279L, -1832275611L, 324056891L, -265082085L, 1493949096L, 
--2127522055L, -771079036L, 558462968L, -258961249L, -543972256L, 
--548427812L, 184016945L, 490059559L, -448510139L, 770836711L, 
--985705915L, 464959828L, 711899353L, 1933082509L, 1540187078L, 
-1696785471L, 444032146L, -82446942L, 20634834L, 1482283067L, 
--1023037433L, -350898326L, -1043228303L, -1058079456L, 1775229802L, 
--906141168L, -306096035L, -2043860621L, -1787347563L, -579438570L, 
--311934740L, 1380177713L, 2094646236L, 1982887694L, 1217439581L, 
--1899829717L, 1057235554L, 1694052062L, -472436044L, 1548185765L, 
[…]

### 2015-10-27 R/packages/mixfdr/R/plot.mix.dens.R

@@ -1,16 +1,0 @@
-`plot.mix.dens` <-
-function(x, plot.subdens = TRUE, ...){
-	z = x$data
-	mi = min(z)
-	ma = max(z)
-	s = seq(mi - 1, ma + 1, by = 0.001)
-	hist(z, pr = TRUE, ...)
-	lines(s, dens.mixture(s,x), lwd = 2)
-	J = length(x$mu)
-	if(plot.subdens == TRUE){
-	for(j in 1:J){
-		lines(s, x$pi[j] * dnorm(s, x$mu[j],x$sigma[j]),col=min(3, j+1), lwd = 2)
-	}	
-	}
-}
-

### 2015-10-27 R/packages/mixfdr/R/tailFDRMixModel.R

@@ -1,29 +1,0 @@
-`tailFDRMixModel` <-
-function(z, m, nullGroups = NA){
-	e = m
-	if(any(is.na(nullGroups))){
-		 nullGroups = rep(FALSE, length(e$mu))
-		 nullGroups[1] = TRUE
-	}
-	# for each group, compute P(|z|>z)
-	# P(Z > z) + P(Z < -z)
-	# P(Z > (z - mu)/sigma) + P(Z < (-z - mu)/sigma)
-	
-	A = matrix(z, length(z), length(e$mu))
-	ri = A - matrix(e$mu, length(z), length(e$mu), byrow = TRUE)
-	ri = ri/matrix(e$sig, length(z), length(e$sig), byrow = TRUE)
-	le = -A - matrix(e$mu, length(z), length(e$mu), byrow = TRUE)
-	le = le/matrix(e$sig, length(z), length(e$sig), byrow = TRUE)
-	pRi = 1 - pnorm(ri)
-	pLe = pnorm(le)
-	pMix2sd = (pRi + pLe)*matrix(e$pi, length(z), length(e$pi), byrow = TRUE)
-	nullProb = rowSums(matrix(pMix2sd[,nullGroups],ncol = sum(nullGroups)))
-	overallProb = rowSums(pMix2sd)
-	twoSided = ((nullProb/overallProb)*(z>=0) + rep(sum(e$pi[nullGroups]), length(z)) *(z<0))
-	ppRi = pRi*matrix(e$pi, length(z), length(e$pi), byrow = TRUE)
-	right = rowSums(matrix(ppRi[,nullGroups],ncol = sum(nullGroups)))/rowSums(ppRi)
-	ppRi = pnorm(ri)*matrix(e$pi, length(z), length(e$pi), byrow = TRUE)
-	left = rowSums(matrix(ppRi[,nullGroups],ncol = sum(nullGroups)))/rowSums(ppRi)
-	return(list(FDRTwoSided = twoSided, FDRleft = left, FDRright = right))
-}
-

### 2015-10-27 R/packages/mixfdr/R/calibrateP.R
add mixfdr package

@@ -1,0 +1,73 @@
+`calibrateP` <-
+function(m, B = 25){
+	getSig0 = function(x, N, J, p, muInt, sigInt){
+		starts = array(getStartsNew(x,J), dim = c(2, J, 3));
+		starts[1,,] = getStartsNew(x,J)		
+		starts[2,,] = cbind(pi = c(0.99, rep(0.01,J-1)/(J-1)), mu = c(0, sample(x,J-1)), sig = rep(sigInt[1,1],J))
+		m = mixModelManyStarts(x, J = J, N = N, p = p, muInt = muInt, sigInt = sigInt, starts = starts)
+		return(m$sig[1])	
+	}
+	
+	rMixModel = function(n,m,pi0=NA, sig0 = NA){
+	J = length(m$pi)	
+	pi = m$pi
+	if(!is.na(pi0)){
+		pi[1] = pi0
+		pi[-1] = pi[-1]*(1-pi0)/sum(pi[-1])	
+	}
+	mu = m$mu
+	sig = m$sig
+	if(!is.na(sig0)){
+		 mu = mu*sqrt(sig0/sig[1])
+		 sig[1] = sig0
+	}	 
+	gp = sample(J, n, rep = TRUE, prob = pi)
+	z = rnorm(n)
+	data = z * sig[gp] + mu[gp]
+	return(data)
+}
+
+
+	N = length(m$data)	
+	noiseSD = m$noiseSD
+	sig0 = m$sig[1]
+	sig0s = pmax(noiseSD,sig0*c(0.8,1,1.3,1.5))
+	sig0s = unique(sig0s)
+
+	Ps = seq(100, N/2, length = 20)
+	
+	data = array(0, dim = c(length(Ps), length(sig0s), B, N))
+	cat("Calibration: Generating random data...")
+	for(i in 1:length(sig0s)){
+		a = rMixModel(N*B, m, sig0 = sig0s[i])	
+		for(j in 1:length(Ps))
+		data[j,i,,] = a
+	}
+				J = length(m$sig)
+				p = c(1, rep(0,J-1))	
+			
+				muInt = matrix(0,J,2)
+				muInt[,1] = -Inf
+				muInt[,2] = Inf		
+			  	
+			
+				sigInt = matrix(0,J,2)
+				sigInt[,1] = noiseSD
+				sigInt[,2] = Inf
+
+	
+	cat("done\n")
[…]

### 2015-10-27 R/packages/mixfdr/R/groupProbs.R

@@ -1,0 +1,11 @@
+`groupProbs` <-
+function(z,m){
+	e = m
+	mu = e$mu
+	pi = c(e$pi)
+	sig = e$sig
+	phiMat = matDens(z,mu,sig)
+	dens = (phiMat %*% pi)[,1]
+	gpProbs = phiMat / matrix(dens, nrow(phiMat), ncol(phiMat), byrow = FALSE) * matrix(pi, nrow(phiMat), ncol(phiMat), byrow =TRUE)
+	}
+

### 2015-10-27 R/packages/mixfdr/R/mixFdr.R

@@ -1,0 +1,160 @@
+`mixFdr` <-
+function(x, J = 3, P = NA, noiseSD = NA, theonull = FALSE, calibrate = FALSE, plots= TRUE, nearlyNull = 0.2, starts = NA, p = NA, maxIter = 1000, tol = 0.001, nocheck = FALSE){
+# try lower sigma threshhold	
+# warn if
+	# uncalibrated
+	# empirical null is substantially different
+	# pi0 is small
+ 
+# return
+	# model parameters
+	# effect size estimates
+	# fdr estimates
+	# plot
+	
+# calibration code
+		
+		muInt = sigInt = NA
+		sigma = noiseSD
+		n = length(x)	
+		if(any(is.na(p))) p = c(1, rep(0,J-1))	
+		if(any(is.na(muInt))){
+			muInt = matrix(0,J,2)
+			muInt[,1] = -Inf
+			muInt[,2] = Inf		
+		}  	
+		if(any(is.na(sigInt)) | any(is.null(dim(sigInt)))){
+			left = sigma
+			if(is.na(sigma)) left = 0.1
+			sigInt = matrix(0,J,2)
+			sigInt[,1] = left
+			sigInt[,2] = Inf
+		}
+		if(theonull){
+			muInt[1,1] = muInt[1,2] = 0
+			sigInt[1,1] = sigInt[1,2] = 1	
+		}
+		if(is.na(P)){
+			P = length(x)/5
+			if(!calibrate && !theonull){
+				warning("Using uncalibrated default penalization P,  which can give misleading results for empirical nulls. Consider rerunning with calibrate = TRUE and using the resulting penalization","\n")
+				}
+		}
+		if(!is.na(sigma)){
+			warning(paste("Assuming known noise noiseSD = ",sigma,". If needed rerun with noiseSD = NA to fit noiseSD."))
+			
+		if(!theonull){
+			warning("Note that using known noiseSD constrains the null to have sd at least noiseSD. If underdispersion is suspected, rerun with noiseSD = NA.")
+			}
+		}
+		
+		cat("Fitting preliminary model", "\n")
+		m =  mixModelManyStarts(x, J, P, starts, maxIter, tol, p, muInt, sigInt, theonull)
+		m$noiseSD = min(median(abs(x-m$mu[1]))/0.6745, m$sig)
+		if(calibrate == TRUE){
[…]

### 2015-10-27 R/packages/mixfdr/R/mixModelManyStarts.R

@@ -1,0 +1,47 @@
+`mixModelManyStarts` <-
+function(x, J, N, starts = NA, maxiter = 1000, tol = 0.001, p = NA, muInt = NA, sigInt = 1, theonull = FALSE){
+	
+# try different starting points
+	# getStartsNew
+	# all null
+	# random
+	# choose highest penalized likelihood
+
+	if(any(is.na(starts))){
+	starts = array(0, dim=c(5, J, 3))
+	sigMAD = median(abs(x))
+	sigST = sigInt[1,1]
+	if(sigInt[1,1]/sigMAD < 0.25) sigST = sigMAD/2
+	starts[1,,] = getStartsNew(x,J, sigST)
+	starts[2,,] = cbind(pi = c(0.99, rep(0.01,J-1)/(J-1)), mu = c(0, sample(x,J-1)), sig = rep(sigST,J))
+	starts[3,,] = cbind(pi = c(0.99, rep(0.01,J-1)/(J-1)), mu = c(0, sample(x,J-1)), sig = rep(sigST,J))
+	for(i in 4:5){
+		starts[i,,1] = rep(1,J)/J
+		starts[i,,2] = sample(x, J)
+		starts[i,,3] = rep(sigST,J)	
+	}
+	}
+	res = array(0, dim = c(dim(starts)[1], J, 3))
+#	print(starts)
+	for(i in 1:dim(starts)[1]){
+		m = mixModelFitter(x, J, N, starts[i,,], maxiter, tol, p, muInt, sigInt)
+		res[i,,1] = m$pi
+		res[i,,2] = m$mu
+		res[i,,3] = m$sig
+	}
+	
+	penLogLik = function(params, x, N, p){
+	pi = params[,1]
+	mu = params[,2]
+	sig = params[,3]
+	ll = sum(log(dens.mixture(x, list(pi = pi, mu = mu, sig = sig))))
+	alpha = N*p
+	pen = sum((alpha-1)*log(pi))
+	return(ll + pen)
+	}
+	
+	plls = apply(res, 1, penLogLik, x, N, p)
+	bestInd = which.max(plls)
+	return(list(pi = res[bestInd,,1], mu = res[bestInd,,2], sigma = res[bestInd,,3], data = x))
+}
+

### 2015-10-30 code/dsc-shrink/run_dsc_znull.R
fix name of files for dsc_znull

@@ -3,7 +3,7 @@
 
 library(dscr)
 
-dsc_znull=new.dsc("znull","../../output/dsc-znull-files")
+dsc_znull=new_dsc("znull","../../output/dsc-znull-files")
 source("add_scenarios.null.R")
 source("add_methods.null.R")
 

### 2015-11-05 analysis/investigate_hu_badcoverage.Rmd
fixed loading of example

@@
 ```{r}
-  eg=load_example(dsc_shrink,42,"flat-top",method="ash.hu.s","../output")
+  eg=load_example(dsc_shrink,42,"flat-top",method="ash.hu.s","../output/dsc-shrink-files/")
   eg$output$fitted.g
@@
   table((ash.u.s.spiky %>% filter(value==0))$seed)
-  eg =load_example(dsc_shrink,seed = 39,method="ash.u.s",scenario="spiky","../output")
+  eg =load_example(dsc_shrink,seed = 39,method="ash.u.s",scenario="spiky","../output/dsc-shrink-files")
 ```

### 2015-11-05 analysis/plot_cdf_eg.Rmd
correct file directory

@@
 #returns list of x,y,pi0, where cdf values are in y
-get_cdf = function(df,dsc=dsc_shrink,x=seq(-6,6,length=100),homedir="../output"){
+get_cdf = function(df,dsc=dsc_shrink,x=seq(-6,6,length=100),homedir="../output/dsc-shrink-files"){
   m=df$method

### 2015-11-05 analysis/investigate_hu_badcoverage.Rmd
small update

--- 2015-10-26 analysis/investigate_hu_badcoverage.Rmd
@@
 ```{r}
-  eg=load_example(dsc_shrink,42,"flat-top",method="ash.hu.s","../output")
+  eg=load_example(dsc_shrink,42,"flat-top",method="ash.hu.s","../code/dsc-shrink")
   eg$output$fitted.g
@@
   table((ash.u.s.spiky %>% filter(value==0))$seed)
-  eg =load_example(dsc_shrink,seed = 39,method="ash.u.s",scenario="spiky","../output")
+  eg =load_example(dsc_shrink,seed = 39,method="ash.u.s",scenario="spiky","../code/dsc-shrink")
 ```

### 2015-11-05 analysis/plot_cdf_eg.Rmd
small update

--- 2015-10-26 analysis/plot_cdf_eg.Rmd
@@
 #returns list of x,y,pi0, where cdf values are in y
-get_cdf = function(df,dsc=dsc_shrink,x=seq(-6,6,length=100),homedir="../output"){
+get_cdf = function(df,dsc=dsc_shrink,x=seq(-6,6,length=100),homedir="../code/dsc-shrink"){
   m=df$method

### 2015-11-05 analysis/summarize_dsc_robust.Rmd
small update

@@
 ```{r}
-  quantile((res %>% filter(method=="ash.n"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
-  quantile((res %>% filter(method=="ash.hu"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
-  quantile((res %>% filter(method=="ash.u"))$score.diff1,c(0.8,0.9,0.95,0.99,1))
+  quantile((res %>% filter(method=="ash.n"))$diff1,c(0.8,0.9,0.95,0.99,1))
+  quantile((res %>% filter(method=="ash.hu"))$diff1,c(0.8,0.9,0.95,0.99,1))
+  quantile((res %>% filter(method=="ash.u"))$diff1,c(0.8,0.9,0.95,0.99,1))
 
-  plot(ecdf((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1),ylim=c(0.8,1))
-  plot(ecdf((res %>% filter(method %in% c("ash.n")))$score.diff1),ylim=c(0.8,1))
+  plot(ecdf((res %>% filter(method %in% c("ash.u","ash.hu")))$diff1),ylim=c(0.8,1))
+  plot(ecdf((res %>% filter(method %in% c("ash.n")))$diff1),ylim=c(0.8,1))
 ```
@@
 ```{r}
-  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1<0.02)
-  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$score.diff1>1)
-  mean((res %>% filter(method %in% c("ash.n")))$score.diff1<0.02)  
+  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$diff1<0.02)
+  mean((res %>% filter(method %in% c("ash.u","ash.hu")))$diff1>1)
+  mean((res %>% filter(method %in% c("ash.n")))$diff1<0.02)  
 ```
@@
 #check out which one showed biggest convergence problem
-prob = res[which.max(res$score.diff1),]
-prob_eg = load_example(dsc_robust,seed=prob$seed,scenario=as.character(prob$scenario),method=as.character(prob$method),homedir="../output")
+prob = res[which.max(res$diff1),]
+prob_eg = load_example(dsc_robust,seed=prob$seed,scenario=as.character(prob$scenario),method=as.character(prob$method),homedir="../code/dsc-robust")
 ```

### 2015-11-17 code/dsc-shrink/run_dsc_znull.R
removed dsc-znull for paper branch

@@ -1,16 +1,0 @@
-#the aim of this dsc is to assess the distribution of the logLR 
-#value for ash under the null that all effects are 0
-
-library(dscr)
-
-dsc_znull=new_dsc("znull","../../output/dsc-znull-files")
-source("add_scenarios.null.R")
-source("add_methods.null.R")
-
-source("score.R")
-add_score(dsc_znull,score_logLR,"logLR","ash_output")
-
-
-res=run_dsc(dsc_znull)
-save(res,dsc_znull,file="../../output/dsc-znull-files/res.znull.RData")
-

### 2015-11-18 R/packages/mixfdr/R/calibrateP.R
install Rmosek, REBayes; correct dscr::source_dir

@@ -1,73 +1,0 @@
-`calibrateP` <-
-function(m, B = 25){
-	getSig0 = function(x, N, J, p, muInt, sigInt){
-		starts = array(getStartsNew(x,J), dim = c(2, J, 3));
-		starts[1,,] = getStartsNew(x,J)		
-		starts[2,,] = cbind(pi = c(0.99, rep(0.01,J-1)/(J-1)), mu = c(0, sample(x,J-1)), sig = rep(sigInt[1,1],J))
-		m = mixModelManyStarts(x, J = J, N = N, p = p, muInt = muInt, sigInt = sigInt, starts = starts)
-		return(m$sig[1])	
-	}
-	
-	rMixModel = function(n,m,pi0=NA, sig0 = NA){
-	J = length(m$pi)	
-	pi = m$pi
-	if(!is.na(pi0)){
-		pi[1] = pi0
-		pi[-1] = pi[-1]*(1-pi0)/sum(pi[-1])	
-	}
-	mu = m$mu
-	sig = m$sig
-	if(!is.na(sig0)){
-		 mu = mu*sqrt(sig0/sig[1])
-		 sig[1] = sig0
-	}	 
-	gp = sample(J, n, rep = TRUE, prob = pi)
-	z = rnorm(n)
-	data = z * sig[gp] + mu[gp]
-	return(data)
-}
-
-
-	N = length(m$data)	
-	noiseSD = m$noiseSD
-	sig0 = m$sig[1]
-	sig0s = pmax(noiseSD,sig0*c(0.8,1,1.3,1.5))
-	sig0s = unique(sig0s)
-
-	Ps = seq(100, N/2, length = 20)
-	
-	data = array(0, dim = c(length(Ps), length(sig0s), B, N))
-	cat("Calibration: Generating random data...")
-	for(i in 1:length(sig0s)){
-		a = rMixModel(N*B, m, sig0 = sig0s[i])	
-		for(j in 1:length(Ps))
-		data[j,i,,] = a
-	}
-				J = length(m$sig)
-				p = c(1, rep(0,J-1))	
-			
-				muInt = matrix(0,J,2)
-				muInt[,1] = -Inf
-				muInt[,2] = Inf		
-			  	
-			
-				sigInt = matrix(0,J,2)
-				sigInt[,1] = noiseSD
-				sigInt[,2] = Inf
-
-	
-	cat("done\n")
[…]

### 2015-11-18 R/packages/mixfdr/R/effectSize.R

@@ -1,24 +1,0 @@
-`effectSize` <-
-function(z, m, noiseSD = NA){
-	e = m
-	if(!is.null(m$noiseSD) && is.na(noiseSD)){
-		noiseSD = m$noiseSD
-	}else{
-		if(is.na(noiseSD)) noiseSD = min(m$sig)	
-	}
-	sig = noiseSD
-	if(any(e$sig<sig)) warning(paste("Marginal variance of a group is less than noise variance, this is impossible under the model."))
-	gpProb = groupProbs(z,e)
-	priorVar = pmax(e$sig^2 - sig^2,0)
-	postVars = 1/((1/priorVar) + (1/sig^2)) * (priorVar != 0)
-	postMeans =  matrix(z, length(z),length(e$mu),byrow=FALSE)/sig^2
-	postMeans = postMeans + matrix((e$mu)/(priorVar),length(z),length(e$mu),byrow=TRUE)
-	postMeans = postMeans * matrix(postVars, length(z), length(e$mu), byrow= TRUE)
-	postMeans[is.nan(postMeans)] = 0
-	postMeans = postMeans + matrix((priorVar ==0)*e$mu, length(z), length(e$mu), byrow = TRUE)
-	postMeansSq = postMeans^2 + matrix(postVars, length(z), length(e$mu), byrow = TRUE)
-	EdeltZ = rowSums(postMeans * gpProb)
-	VdeltZ = rowSums(postMeansSq * gpProb) - EdeltZ^2
-	return(cbind(Edz = EdeltZ, Vdz = VdeltZ))
-}
-

### 2015-11-18 R/packages/mixfdr/R/groupProbs.R

@@ -1,11 +1,0 @@
-`groupProbs` <-
-function(z,m){
-	e = m
-	mu = e$mu
-	pi = c(e$pi)
-	sig = e$sig
-	phiMat = matDens(z,mu,sig)
-	dens = (phiMat %*% pi)[,1]
-	gpProbs = phiMat / matrix(dens, nrow(phiMat), ncol(phiMat), byrow = FALSE) * matrix(pi, nrow(phiMat), ncol(phiMat), byrow =TRUE)
-	}
-

### 2015-11-18 R/packages/mixfdr/R/mixModelFitter.R

@@ -1,61 +1,0 @@
-`mixModelFitter` <-
-function(x, J, N, starts, maxiter, tol, p, muInt, sigInt){
-
-				
-		EMstep = function(x, pi, mu, sig, N, p, muInt, sigInt){
-				n = length(x)
-				J = length(mu)
-				
-				# E-step	
-				phiMat = matDens(x,mu,sig)
-				denominators = (phiMat %*% pi)[,1]
-				G = phiMat * ((1/denominators) %o% pi)
-				tG = t(G)
-				
-				# M-step
-				numeratorsMu = (tG %*% x)[,1]
-				zplusj = rowSums(tG)
-				newMu = (numeratorsMu/zplusj)
-				newSig = tG %*% (x*x) / zplusj - newMu^2
-				
-				newMu = pmax(pmin(muInt[,2],newMu),muInt[,1])
-				newSig = pmax(pmin(sigInt[,2], sqrt(newSig)), sigInt[,1])
-				newPi = (N*p - 1 + zplusj)/(sum(zplusj) + N - J)
-				if(N<0) newPi = zplusj/sum(zplusj)
-				newPi = (newPi+0.001)
-				newPi = pmax(0,pmin(newPi,1))
-				newPi = newPi/sum(newPi)
-				if(any(is.na(newPi), is.na(newMu), is.na(newSig))){
-					print(pi)
-					print(mu)
-					print(sig)	
-				}
-				return(list(pi = newPi, mu = newMu, sigma = newSig)) 
-			}
-		
-	# starting points	
-		pi = starts[,1]
-		mu = starts[,2]
-		sigma = starts[,3]
-	# that's it, do the loop	
-		iter = 0
-		converged = FALSE
-		while(!converged && iter<maxiter){
-			res = EMstep(x, pi, mu, sigma, N, p, muInt, sigInt)
-			distance = sum((pi-res$pi)^2)+sum((mu-res$mu)^2) + sum((sigma-res$sigma)^2)
-			if(is.na(distance)){
-				print(starts)
-				print(muInt)
-				print(sigInt)	
-			}
-			if(distance<tol) converged = TRUE
-			iter = iter + 1			
-			pi = res$pi
-			mu = res$mu
-			sigma = res$sigma
-		}
-#		e = list(pi = pi, mu = mu, sigma = sigma, converged = converged, nIter = iter, data = x)
[…]

### 2015-11-18 R/packages/mixfdr/R/mixModelManyStarts.R

@@ -1,47 +1,0 @@
-`mixModelManyStarts` <-
-function(x, J, N, starts = NA, maxiter = 1000, tol = 0.001, p = NA, muInt = NA, sigInt = 1, theonull = FALSE){
-	
-# try different starting points
-	# getStartsNew
-	# all null
-	# random
-	# choose highest penalized likelihood
-
-	if(any(is.na(starts))){
-	starts = array(0, dim=c(5, J, 3))
-	sigMAD = median(abs(x))
-	sigST = sigInt[1,1]
-	if(sigInt[1,1]/sigMAD < 0.25) sigST = sigMAD/2
-	starts[1,,] = getStartsNew(x,J, sigST)
-	starts[2,,] = cbind(pi = c(0.99, rep(0.01,J-1)/(J-1)), mu = c(0, sample(x,J-1)), sig = rep(sigST,J))
-	starts[3,,] = cbind(pi = c(0.99, rep(0.01,J-1)/(J-1)), mu = c(0, sample(x,J-1)), sig = rep(sigST,J))
-	for(i in 4:5){
-		starts[i,,1] = rep(1,J)/J
-		starts[i,,2] = sample(x, J)
-		starts[i,,3] = rep(sigST,J)	
-	}
-	}
-	res = array(0, dim = c(dim(starts)[1], J, 3))
-#	print(starts)
-	for(i in 1:dim(starts)[1]){
-		m = mixModelFitter(x, J, N, starts[i,,], maxiter, tol, p, muInt, sigInt)
-		res[i,,1] = m$pi
-		res[i,,2] = m$mu
-		res[i,,3] = m$sig
-	}
-	
-	penLogLik = function(params, x, N, p){
-	pi = params[,1]
-	mu = params[,2]
-	sig = params[,3]
-	ll = sum(log(dens.mixture(x, list(pi = pi, mu = mu, sig = sig))))
-	alpha = N*p
-	pen = sum((alpha-1)*log(pi))
-	return(ll + pen)
-	}
-	
-	plls = apply(res, 1, penLogLik, x, N, p)
-	bestInd = which.max(plls)
-	return(list(pi = res[bestInd,,1], mu = res[bestInd,,2], sigma = res[bestInd,,3], data = x))
-}
-

### 2015-11-18 code/dsc-robust/run_dsc_robust.R

@@ -1,21 +1,21 @@
 library(dscr)
-source_dir("methods")
+dscr::source_dir("methods")
 
 ###### Initialize #######
 
 dsc_robust=new_dsc("robust","../../output/dsc-robust-files")
 
 ###### Add Scenarios #####
-source_dir("../dsc-shrink/datamakers")
+dscr::source_dir("../dsc-shrink/datamakers")
 source("../dsc-shrink/add_named_scenarios.R")
 add_named_scenarios(dsc_robust,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"))
 
 
 ###### Add Methods #####
 
-add_method(dsc_robust,"ash.u",ash.wrapper,args=list(mixcompdist="uniform",cxx=TRUE))
-add_method(dsc_robust,"ash.hu",ash.wrapper,args=list(mixcompdist="halfuniform",cxx=TRUE))
-add_method(dsc_robust,"ash.n",ash.wrapper,args=list(mixcompdist="normal",cxx=TRUE))
+add_method(dsc_robust,"ash.u",ash.wrapper,args=list(mixcompdist="uniform"))
+add_method(dsc_robust,"ash.hu",ash.wrapper,args=list(mixcompdist="halfuniform"))
+add_method(dsc_robust,"ash.n",ash.wrapper,args=list(mixcompdist="normal"))
 
 
 

### 2015-11-18 code/dsc-shrink/add_methods.R

@@ -1,28 +1,22 @@
-source_dir("methods")
+dscr::source_dir("methods")
 
-add_method(dsc_shrink,name="ash.hu.nocxx",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=FALSE),outputtype = "ash_output")
-add_method(dsc_shrink,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=TRUE),outputtype = "ash_output")
-add_method(dsc_shrink,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif",cxx=TRUE),outputtype = "ash_output")
-add_method(dsc_shrink,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal",cxx=TRUE),outputtype = "ash_output")
+#add_method(dsc_shrink,name="ash.hu.nocxx",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=FALSE),outputtype = "ash_output")
+add_method(dsc_shrink,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
+add_method(dsc_shrink,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")
+add_method(dsc_shrink,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal"),outputtype = "ash_output")
 
-#add "truth" method
-#for(i in c("spiky")){ #},"near-normal","flat-top","skew","big-normal","bimodal")){
-#  s=dsc_shrink$scenarios[[i]]
-#  g=s$args$g
-#  add_method(dsc_shrink,name=paste0("truth.",i),fn =ash.wrapper,args=list(g=g,cxx=TRUE,control=list(maxiter=0)),outputtype = "ash_output")
-#}
 
 add_method(dsc_shrink,name="bayes",fn=bayes.wrapper,gold_flag=TRUE,outputtype = "ash_output")
 
 # try smaller null weight
-add_method(dsc_shrink,name="ash.hu.nw2",fn =ash.wrapper,args=list(mixcompdist="halfunif",nullweight=2,cxx=TRUE),outputtype = "ash_output")
-add_method(dsc_shrink,name="ash.u.nw2",fn =ash.wrapper,args=list(mixcompdist="unif",nullweight=2,cxx=TRUE),outputtype = "ash_output")
[…]

### 2015-11-18 code/dsc-shrink/add_named_scenarios.R

@@ -1,4 +1,4 @@
-source_dir("datamakers")
+dscr::source_dir("datamakers")
 
 gdef = list(spiky=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
             skew=normalmix(c(1/4,1/4,1/3,1/6),c(-2,-1,0,1),c(2,1.5,1,1)),

### 2015-11-18 code/dsc-shrink/add_methods.R
put nocxx method back

@@ -1,6 +1,6 @@
 dscr::source_dir("methods")
 
-#add_method(dsc_shrink,name="ash.hu.nocxx",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=FALSE),outputtype = "ash_output")
+add_method(dsc_shrink,name="ash.hu.nocxx",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=FALSE),outputtype = "ash_output")
 add_method(dsc_shrink,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
 add_method(dsc_shrink,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")
 add_method(dsc_shrink,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal"),outputtype = "ash_output")

### 2015-11-18 code/dsc-shrink/add_methods.null.R
fix dscr::source_dir

@@ -1,4 +1,4 @@
-source_dir("methods")
+dscr::source_dir("methods")
 
 add_method(dsc_znull,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
 add_method(dsc_znull,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")

### 2015-11-18 code/dsc-shrink/add_scenarios.null.R

@@ -1,4 +1,4 @@
-source_dir("datamakers")
+dscr::source_dir("datamakers")
 
 
 #Now, for each scenario create an element of scenarios of the following form

### 2015-11-20 code/dsc-opt/methods/ash.multiopt.wrapper.R
change dsc-robust to dsc-opt

@@ -1,0 +1,37 @@
+#' @title wrapper for ash for shrinkage DSC
+#'
+#' @description Runs ash to compute betahat values  
+#' @details None
+#' 
+#' @param input a list with elements betahat and sebetahat
+#' @param args a list containing other additional arguments to ash
+#' 
+#' @return output a list containing a vector of loglikelihoods from multiple runs of ash
+#'
+library(ashr)
+
+ash.multiopt.wrapper=function(input,args=NULL){
+  if(is.null(args)){  #set shrink so that the likelihood is unpenalized
+    args=list(mixcompdist="halfuniform",method="shrink")
+  }
+  loglik=rep(0,2)
+  
+  #first run is with interior point
+  res=do.call(ash, args= c(list(betahat=input$betahat,sebetahat=input$sebetahat,optmethod="mixIP"),args))
+  loglik[1]= get_loglik(res)
+  
+  #now with EM
+  res=do.call(ash, args= c(list(betahat=input$betahat,sebetahat=input$sebetahat,optmethod="cxxMixSquarem"),args))
+  loglik[2]= get_loglik(res)
+  
+  
+  #first run is with non-random start (ie default start)
+  #res=do.call(ash, args= c(list(betahat=input$betahat,sebetahat=input$sebetahat,randomstart=FALSE),args))
+  #loglik[1]= get_loglik(res)
+    
+  #for(i in 2:11){
+  #  res = do.call(ash, args= c(list(betahat=input$betahat,sebetahat=input$sebetahat,randomstart=TRUE),args))
+  #  loglik[i]= get_loglik(res)
+  #}
+  return(list(loglik=loglik))
+}

### 2015-11-20 code/dsc-opt/run_dsc_opt.R

@@ -1,0 +1,43 @@
+library(dscr)
+dscr::source_dir("methods")
+
+###### Initialize #######
+
+dsc_opt=new_dsc("opt","../../output/dsc-opt-files")
+
+###### Add Scenarios #####
+dscr::source_dir("../dsc-shrink/datamakers")
+source("../dsc-shrink/add_named_scenarios.R")
+add_named_scenarios(dsc_opt,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"))
+
+
+###### Add Methods #####
+
+add_method(dsc_opt,"ash.u",ash.multiopt.wrapper,args=list(mixcompdist="uniform",method="shrink"))
+add_method(dsc_opt,"ash.hu",ash.multiopt.wrapper,args=list(mixcompdist="halfuniform",method="shrink"))
+add_method(dsc_opt,"ash.n",ash.multiopt.wrapper,args=list(mixcompdist="normal",method="shrink"))
+
+
+
+####### Define Score and Add it #######
+
+score = function(data, output){
+  x=output$loglik
+  #names(x)=paste0('C',1:length(x))
+  #class(x)<-'data.frame'
+  #return(list(diff1 = max(x)-x[1],diff2=(max(x)-min(x))))
+  return(list(diff1 = x[1]-x[2],diff2=0))
+}
+
+add_score(dsc_opt,score)
+
+######## Run the DSC #################
+
+res_opt=run_dsc(dsc_opt)
+save(dsc_opt,file="../../output/dsc-opt-files/dsc_opt.RData")
+
+
+
+
+
+

### 2015-11-20 code/dsc-robust/run_dsc_robust.R

@@ -1,42 +1,0 @@
-library(dscr)
-dscr::source_dir("methods")
-
-###### Initialize #######
-
-dsc_robust=new_dsc("robust","../../output/dsc-robust-files")
-
-###### Add Scenarios #####
-dscr::source_dir("../dsc-shrink/datamakers")
-source("../dsc-shrink/add_named_scenarios.R")
-add_named_scenarios(dsc_robust,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"))
-
-
-###### Add Methods #####
-
-add_method(dsc_robust,"ash.u",ash.wrapper,args=list(mixcompdist="uniform"))
-add_method(dsc_robust,"ash.hu",ash.wrapper,args=list(mixcompdist="halfuniform"))
-add_method(dsc_robust,"ash.n",ash.wrapper,args=list(mixcompdist="normal"))
-
-
-
-####### Define Score and Add it #######
-
-score = function(data, output){
-  x=output$loglik
-  #names(x)=paste0('C',1:length(x))
-  #class(x)<-'data.frame'
-  return(list(diff1 = max(x)-x[1],diff2=(max(x)-min(x))))
-}
-
-add_score(dsc_robust,score)
-
-######## Run the DSC #################
-
-res_robust=run_dsc(dsc_robust)
-save(dsc_robust,file="../../output/dsc-robust-files/dsc_robust.RData")
-
-
-
-
-
-

### 2015-11-22 code/dsc-opt/run_dsc_opt.R
simplify score

@@ -23,10 +23,7 @@ add_method(dsc_opt,"ash.n",ash.multiopt.wrapper,args=list(mixcompdist="normal",m
 
 score = function(data, output){
   x=output$loglik
-  #names(x)=paste0('C',1:length(x))
-  #class(x)<-'data.frame'
-  #return(list(diff1 = max(x)-x[1],diff2=(max(x)-min(x))))
-  return(list(diff1 = x[1]-x[2],diff2=0))
+  return(list(diff1 = x[1]-x[2])) 
 }
 
 add_score(dsc_opt,score)

### 2015-11-22 analysis/checkIP.Rmd
add checkIP analysis

---
title: "Check interior point method"
author: "Matthew Stephens"
date: 2015-11-18
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

We want to compare the likelihoods achieved by the newly implemented interior point method with the EM algorithm.

## Log-likelihood comparison

The 'dsc-opt' simply runs both methods (IP and EM) on the same data sets, 
and compares the difference in the log-likelihoods. We can see that
in >90% of cases the difference is negligible. Then there is a tail of
cases where the IP loglik is higher - presumably where the EM doesn't really converge so well.

```{r}
  load("../output/dsc-opt-files/dsc_opt.RData")
  plot(ecdf(dsc_opt$res$diff1),main="ECDF of log-likelihood difference (IP-EM)")
```

## lfsr comparison

However, what we really care about is inference quantities, such as the lfsr, rather than the log-likelihood. Here we compare the lfsr for the simulations where
we ran both the IP method and EM method (nocxx) in dsc-shrink.

```{r}
  load("../output/dsc-shrink-files/res.RData")
  library("ggplot2")
  library("dplyr")
  library("reshape2")
  # select out the lfsr estimates for each method

  df1 = res$lfsr %>% dplyr::filter(method %in% c("ash.hu.nocxx")) %>% select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="lfsr.nocxx")
  df2 = res$lfsr %>% dplyr::filter(method %in% c("ash.hu")) %>% select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
    melt(id.vars=c("method","scenario","seed",".id"),value.name="lfsr.IP")

  df = inner_join(df1,df2,by=c("scenario","seed","variable"))
  ggplot(df,aes(lfsr.nocxx,lfsr.IP)) +geom_point(shape=1) +
  facet_wrap(~ scenario, nrow=2) + 
  geom_abline(colour = "black") +
  xlab("EM algorithm (no cxx)") +
  ylab("IP method")
```

## Session information

```{r info}
sessionInfo()
```

### 2015-12-14 paper/main.tex
update paper to detail convex optimization

@@
 in FDR analyses, yields more accurate estimates of FDR than existing methods, provided the UA holds.
-Second, the method takes as it input two numbers for each test (an effect size estimate, and corresponding standard error),
-rather than the one number usually used ($p$ value, or $z$ score). Provided these two numbers are available,
-using them allows our method to better account for
-variations in measurement precision across tests. It also facilitates the estimation of actual 
-effect sizes, and our approach provides ``shrunken" interval estimates (credible regions) for each effect in addition to measures of ``significance".
-To provide a bridge between interval estimates and signicance measures we introduce the term ``local false sign rate"
+It also facilitates efficient and robust computation because estimating the unimodal distribution involves solving a simple convex optimization problem.
+Second, the method takes as its input two numbers for each test (an effect size estimate, and corresponding standard error),
+rather than the one number usually used ($p$ value, or $z$ score). When available, using two numbers instead of one allows 
+for better accounting of variations in measurement precision across tests. It also facilitates the estimation of actual 
+effect sizes, and our approach provides interval estimates (credible regions) for each effect in addition to measures of significance.
+To provide a bridge between interval estimates and significance measures we introduce the term ``local false sign rate"
 to refer to the probability of getting the sign of an effect wrong, and argue that it is a superior
@@
 in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
-the $\lfsr$ and $\lfdr$ can be quite different, and emphasise benefits of the $\lfsr$, particularly its increased robustness to modelling assumptions.
+the $\lfsr$ and $\lfdr$ can be quite different, and emphasize benefits of the $\lfsr$, particularly its increased robustness to modeling assumptions.
 
@@
 Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(The names comes from the fact that EB methods in general, and -- due to the UA -- our EB method in particular, is a form of shrinkage estimation,
+(The names comes from the fact that EB methods in general -- and, due to the UA, our EB method in particular -- perform shrinkage estimation,
 and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation. 
@@
 %mean expression level between the two conditions, often referred to as the ``log-fold change". 
-Here we describe the simplest version of our method, and briefly discuss embellishments we have also implemented.
+Here we describe the simplest version of the method, and briefly discuss embellishments we have also implemented.
 
@@
 where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyperparameters, which are non-negative and sum to one, and are to be estimated,
+Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyper-parameters, which are non-negative and sum to one, and are to be estimated,
 while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
 represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourate the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
+(We encourage the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
 although of course in practice we use a finite grid -- see Implementation Details.)
@@
 
-This simple model features both the key ideas we want to emphasise in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
+This simple model features both the key ideas we want to emphasize in this paper: the UA is encapsulated in (\ref{eqn:mixnorm}) while
 the different measurement precision of different observations is encapsulated in the likelihood (\ref{eqn:normlik}) -- specifically, observations
@@
 \begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) also implies that $g$ is symmetric about 0; more flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see (\ref{eqn:g-general})].
+\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
+More flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see equation (\ref{eqn:g-general})]; indeed this 
+can allow $g$ to approximate {\it any} unimodal distribution. 
 \item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
@@
 \begin{enumerate}
-\item Estimate the hyperparameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
+\item Estimate the hyper-parameters $\pi$ by maximizing the likelihood $L(\pi)$, given by (\ref{eqn:marginallik}), yielding $\hat{\pi} := \arg \max L(\pi)$.
 \item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
@@
 \end{enumerate}
-Both steps 1 and 2 are very straightforward: $\hat{\pi}$ can be obtained using a simple EM algorithm \cite{dempster77}, and the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of the EM algorithm in step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
+Both steps 1 and 2 are very straightforward.
+Step 1 is a convex optimization problem \cite{}, and can be solved very quickly using interior point methods \cite{boyd2004convex,koenker2013convex}.
+And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
+(The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
 which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
 
-Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximise a penalized
+Here we slightly modify this usual procedure: instead of obtaining $\hat\pi$ by maximizing the likelihood, we maximize a penalized
 likelihood [see (\ref{eqn:penalty})], where the penalty encourages $\hat\pi_0$ to be as big as possible whilst remaining consistent with the observed data. 
-We introduce this penalty because in FDR appllications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
+We introduce this penalty because in FDR applications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
 
@@
 small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
-Of coure, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
+Of course, being confident in the sign of an effect logically implies that we are confident it is non-zero, and
 this is reflected in the fact that $\lfsr_j \geq \lfdr_j$ 
@@
 In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modelling assumptions than $\lfdr$.
+$\lfsr$ can be substantially more robust to modeling assumptions than $\lfdr$.
 
@@
 To paraphrase Tukey, ``All we know about the world teaches us that large effects are rare, whereas small effects abound."
-We emphasise that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
+We emphasize that the UA relates to the distribution of {\it all} effects, and not only the {\it detectable} effects (i.e.~those that are significantly different from zero).
 It is very likely that the distribution of {\it detectable} non-zero effects will be multimodal, with one mode for detectable positive effects and another for detectable negative effects, and the UA does not contradict this.
@@
 
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficents - 
+In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficients -- 
 common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
@@
 Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
-and statistically stable. We illustrate these features in Results.
+and statistically stable (see Results).
 
@@
 However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it seems possible it will not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
+may perform well for point estimation, it may not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
 Indeed, one way of thinking about the UA is as a way to regularize $g$.
@@
 provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
-The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behaviour. To check its effectiveness
+The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behavior. To check its effectiveness
 we performed simulations under various alternative scenarios (i.e. various distributions for the non-zero effects, which we denote $g_1$), and values
@@
  the data can effectively provide an upper bound on plausible values of $\pi_0$,
-but not a lower bound (because the data cannot rule out that everything is non-null, but with miniscule effects).
-To obtain conservative behaviour we must estimate $\pi_0$ by this upper bound, which can
+but not a lower bound (because the data cannot rule out that everything is non-null, but with minuscule effects).
+To obtain conservative behavior we must estimate $\pi_0$ by this upper bound, which can
 be substantially larger than the true value.
@@
 be conservative, overestimating the truth, the overestimation is substantially less pronounced than for the $\lfdr$, especially for
-the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behaviour is less pronounced in $\lfsr$ than $\lfdr$.
+the ``spiky" scenario. Further, in the bi-modal scenario, the anti-conservative behavior is less pronounced in $\lfsr$ than $\lfdr$.
 
@@
 We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
-
-\subsubsection*{Numerical Stability}
-
-The EM algorithm, which we use here to fit our model, is notorious for convergence to local optima. 
-However, in this case, over hundreds of applications of the procedure, we observed no obvious serious problems 
-caused by such behaviour. To quantify this, we ran \ashr 10 times on each of the 600 simulated datasets above using a random initialization for $\pi$,
-in addition running it using our default initialization procedure (see Implementation details). We then compared the largest
-log-likelihood achieved across all 11 runs with the log-likelihood achieved by the default run.
-When using a mixture of normals (ash.n) the results were extremely stable: 96\% showed a negligible log-likelihood difference ($< 0.02$), 
-and the largest difference was 0.8. When using mixtures of uniforms (ash.u,ash.hu) results were slightly less stable: 89\% showed a negligible log-likelihood difference ($<0.02$),
-and 6\% of runs showed an appreciable log-likelihood difference ($>1$), with the largest difference being 5.0. However, perhaps suprisingly, 
-even for this largest difference results from the default run (e.g.
-the $\lfsr$ values, and the posterior means) were in other ways virtually indistinguishable from the results from the run with the highest log-likelihood.
-See \url{http://github.com/stephens999/ash/blob/master/dsc-robust/summarize_dsc_robust.rmd} for further details.
 
@@
 depends on reliably estimating the distribution $g$. Estimating $g$ is a ``deconvolution problem",
-which are notoriously difficult in general. Indeed, Efron emphasises
+which are notoriously difficult in general. Indeed, Efron emphasizes
 the difficulties of implementing a stable general algorithm, noting in his rejoinder
@@
 segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behaviour, effectively requiring $g$ to be somewhat smooth.
+The UA prevents this kind of ``irregular" behavior, effectively requiring $g$ to be somewhat smooth.
 While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
@@
 a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behaviour directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
+The bottom panel (b) illustrates this behavior directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
 In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
@@
  with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measuments that have the same $p$ value.
+ We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measurements that have the same $p$ value.
  The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
@@
 
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to downweight imprecise observations,
+The potential for Bayesian analyses to re-order the significance of observations, and specifically to down-weight imprecise observations,
 was previously discussed in \cite{guan.stephens.08}.
@@
 %small standard error, and so their posteriors will be more affected by the prior, which, being unimodal at 0, tends to shrink estimates towards 0. 
-%To emphasise these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
+%To emphasize these two key features we refer to our method as ``{\bf a}daptive {\bf sh}rinkage".
 %Figure \ref{fig:ash} illustrates the idea by contrasting results for two simulation scenarios with different signal strengths, where observations vary in their measurment precision.
@@
 
-We have presented an Empirical Bayes approach to large-scale multiple testing that emphasises two ideas.
-First, we emphasise the potential benefits of using two numbers ($\bhat$, and its standard error)
+We have presented an Empirical Bayes approach to large-scale multiple testing that emphasizes two ideas.
+First, we emphasize the potential benefits of using two numbers ($\bhat$, and its standard error)
 rather than just one number (a $p$ value or $z$ score)  to summarize the information on each test.
@@
 to improve inference and make informed decisions about significance (see also \cite{greenland1991empirical}). This view
-also emphasises that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
+also emphasizes that, what matters in multiple testing settings is {\it not} the number of tests, but the {\it results} of the tests.
 Indeed, the FDR at a given fixed threshold does not depend on the number of tests: as the number of tests increases, both the true positives and false positives increase linearly,
 and the FDR remains the same.
-(If this intuitve argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
+(If this intuitive argument does not convince, see \cite{storey.03}, and note that the FDR at a given $p$ value threshold does not depend on the number of tests $m$.) Conversely, the FDR {\it does} depend on the overall distribution of effects, and particularly on $\pi_0$ for example.
 The EB approach captures this dependence in an intuitive way:
@@
 that simply attempt to correct for the resulting inflation of test statistics \cite{devlin1999genomic,efron2004large}, and those
-that attempt to infer counfounders using clustering, principal components analysis, or factor models 
+that attempt to infer confounders using clustering, principal components analysis, or factor models 
 \cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
 and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
-they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alterntively, our methods
+they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alternatively, our methods
 could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:normlik}) using
@@
 measurement precision across units) while losing only a little in generality.  We hope that our work will
-encourate development of methods that exploit this idea in other contexts.
+encourage development of methods that exploit this idea in other contexts.
 
@@
 (\ref{eqn:tlik}) introduced below. The use of uniform components here closely mirrors \cite{cordy1997deconvolution}.
-(In fact our implementation can handle {\it any} prespecified uniform or normal distributions for $f_k$ provided they are all from the same family;
+(In fact our implementation can handle {\it any} pre-specified uniform or normal distributions for $f_k$ provided they are all from the same family;
 however, we restrict our attention here to (i)-(iii) which imply a unimodal $g$.)
@@
 
-Although our discussion above emphasises the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
+Although our discussion above emphasizes the use of large $K$, in practice modest values of $K$ can provide reasonable performance. 
 The key point is that the value of $K$ is not critical provided it is sufficiently large, and the grid of $\sigma_k$ or $a_k$ values suitably chosen.
@@
 We define the likelihood for $\pi$ to be the probability of the observed data $\bhat$ conditional on $\shat$: $L(\pi) := p(\bhat | \shat,\pi),$ which
-by our conditional indendence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
+by our conditional independence assumptions is equal to the product $\prod_j p(\bhat_j | \shat, \pi)$.
 [One might prefer to define the likelihood as $p(\bhat, \shat | \pi) = p(\bhat | \shat, \pi) p(\shat | \pi)$, in which case our definition comes down to assuming that the term $p(\shat | \pi)$ does not depend on $\pi$.]
@@
 
-\subsubsection*{EM algorithm}
+\subsubsection*{Optimization}
 
@@
 \end{equation}
-where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This can be maximized using an EM algorithm \cite{dempster77}, whose
-one-step updates are:
+where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This is a convex optimization problem, which
+can be solved very quickly and reliably using interior point methods.
+We used the {\tt KWdual} function from the {\tt R} package {\tt REBayes} \cite{REBayes}, which uses {\tt Rmosek} \cite{Rmosek}.
+We also found a simple EM algorithm \cite{dempster77}, accelerated using the elegant {\tt R} package {\tt SQUAREM} \cite{varadhan2008simple},
+to provide adequate performance. We initialized the EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$,
+and the one-step updates are:
 \begin{align} \label{eqn:w}
@@
 n_k & = \sum_j w_{kj} + \lambda_k - 1 \quad \text{[E Step]} \\
-\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}
+\pi_k &= n_k/\sum_{k'} n_{k'} \quad \text{[M step]}.
 \end{align}
-Note that $\pi_k$ can be interpreted as the prior probability that $\beta_j$ arose from component $k$,
-and $l_{kj}$ is the likelihood for $\beta_j$ given that it arose from component $k$, so $w_{kj}$ is the posterior 
-probability that $\beta_j$ arose from component $k$, given $\bhat, \shat, \pi$. (The $w_{kj}$ are sometimes referred to as the ``responsibilities".)
-Thus $n_k$ is the expected number of $\beta_j$ that arose from component $k$, plus pseudo-counts $\lambda_k-1$ from the penalty term.
-We used the elegant R package {\tt SQUAREM} \cite{varadhan2008simple} to accelerate convergence of this EM algorithm.
-
-%[Note: an alternative Variational Bayes EM approach to estimating $\pi$ has been implemented in our R package by Mengyin Lu.]
-\subsubsection*{Initialization}
-
-By default we initialize our EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$. (In all our simulations here $K<<n$ so this initializes with most mass on $\pi_0$.)  
-Our rationale for initializing ``near the null" like this is that we expect that strong signal in the data can quickly draw the EM algorithm away from the null (in a single iteration),
-but weak signal in the data cannot quickly draw the algorithm towards the null.
-
-In addition, once the EM algorithm has converged, we check that the (penalized) log-likelihood attained is higher than that achieved by the global null solution
-$\pi_0=1; \pi_k=0 (k>0)$. If not then we replace the EM solution with the global null solution. This aims to guard against errors due to convergence to a local optimum when the
-data are consistent with the global null.
+The only benefit to the EM algorithm is fewer software dependencies.
 
@@
 When $f_k$ is $N(0,\sigma_k)$ we specify our grid by specifying: i) a maximum and minimum value $(\sigmamin,\sigmamax)$; ii) a multiplicative factor $m$ to be used
-in going from one gridpoint to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
+in going from one grid-point to the other, so that $\sigma_{k} = m \sigma_{k-1}$. The multiplicative factor affects the density of the grid; we used $m=\sqrt{2}$
 as a default. We chose $\sigmamin$ to be small compared with the measurement precision ($\sigmamin=\min(\shat_j)/10$) and $\sigmamax= 2\sqrt{\max(\bhat_j^2-\shat_j^2)}$

### 2015-12-14 paper/main.tex
update versions

@@
 We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.0 from Bioconductor), \locfdr (v1.1-7 from \url{https://cran.r-project.org/src/contrib/Archive/locfdr/}), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
+in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).  
 In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this

### 2015-12-15 paper/main.tex
minor update to wording

@@
 
-We introduce a novel Empirical Bayes approach for performing large-scale hypothesis testing, including  
+We introduce a novel Empirical Bayes approach for large-scale hypothesis testing, including  
 estimating False Discovery Rates (FDRs), and estimating effect sizes. Compared with
-existing commonly-used approaches to FDR analysis, the method has two key differences. First, it
+existing approaches to FDR analysis, the method has two key differences. First, it
 assumes that the distribution of the actual (unobserved) effects being tested is unimodal, with a mode at 0.
 This ``unimodal assumption" (UA), which is common in other related contexts, but very different from assumptions usually made
-in FDR analyses, yields more accurate estimates of FDR than existing methods, provided the UA holds.
-It also facilitates efficient and robust computation because estimating the unimodal distribution involves solving a simple convex optimization problem.
+in FDR analyses, yields more accurate inferences than existing methods provided that the UA holds.
+The UA also facilitates efficient and robust computation because estimating the unimodal distribution involves solving a simple convex optimization problem.
 Second, the method takes as its input two numbers for each test (an effect size estimate, and corresponding standard error),
-rather than the one number usually used ($p$ value, or $z$ score). When available, using two numbers instead of one allows 
-for better accounting of variations in measurement precision across tests. It also facilitates the estimation of actual 
+rather than the one number usually used ($p$ value, or $z$ score). When available, using two numbers instead of one 
+helps account for variation in measurement precision across tests. It also facilitates the estimation of actual 
 effect sizes, and our approach provides interval estimates (credible regions) for each effect in addition to measures of significance.
@@
 and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation. 
-This will be discussed further in Xing and Stephens (in preparation).)
+This will be discussed further in Xing and Stephens (in preparation).) Code and instructions for reproducing analyses and figures in this paper are at
+\url{https://github.com/stephenslab/ash}.
 
@@
 Step 1 is a convex optimization problem \cite{}, and can be solved very quickly using interior point methods \cite{boyd2004convex,koenker2013convex}.
+Our {\tt R} package typically takes a few seconds for $p=100,000$.
 And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
@@
 to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://github.com/stephens999/ash/blob/master/dsc-shrink/check_mixfdr_lownoise.rmd}). Of course, we would not advocate the UA in settings where multi-modality is 
+(\url{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}). Of course, we would not advocate the UA in settings where multi-modality is 
 clearly visible in the observed $\bhat$.
@@
 factors that could cause dependence among tests, unmeasured confounders are perhaps the most worrisome in practice
-because they can induce strong correlations among large numbers of tests and profoundly impact results.
+because they can induce strong correlations among large numbers of tests and profoundly impact results, ultimately resulting
+in too many hypotheses being rejected and a failure to control FDR.
+We are acutely aware that, because our method is less conservative than existing methods, it may unwittingly
+exacerbate these issues if they are not adequately dealt with.
 Approaches to deal with unmeasured confounders can be largely divided into two types: those
@@
 \cite{pritchard.stephens.rosenberg.donnelly.00,price:2006,leek:2007,gagnon2012using},
-and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable,
+and then correct for them in computation of the test statistics (in our case, $\bhat,\shat$). When these latter approaches are viable
 they provide perhaps the most satisfactory solution, and are certainly a good fit for our framework. Alternatively, our methods
-could also be modified to allow for test statistic inflation, perhaps by incorporating the inflation into the likelihood (\ref{eqn:normlik}) using
-$\bhat_j | \shat_j \sim N(0,\lambda_1 \shat_j + \lambda_2)$, where $\lambda_1,\lambda_2$ are to be estimated. However,
-this immediately raises issues of identifiability and we do not pursue the idea further here.
+could be modified to allow for test statistic inflation, an idea that may be worth pursuing in future work.
 
@@
 An addition to our software implementation, due to C.Dai, allows the mode to be estimated from the data by maximum likelihood, rather than fixed to 0.
+This involves a simple grid search.
 

### 2015-12-16 paper/main.tex
remove pdf files

@@
 
+\usepackage{endfloat}
+
 \usepackage{sectsty}
@@
 % Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
+\usepackage{setspace} 
+\doublespacing
 
@@
 Both steps 1 and 2 are very straightforward.
-Step 1 is a convex optimization problem \cite{}, and can be solved very quickly using interior point methods \cite{boyd2004convex,koenker2013convex}.
+Step 1 is a convex optimization problem, and can be solved very quickly using interior point methods \cite{boyd2004convex,koenker2013convex}.
 Our {\tt R} package typically takes a few seconds for $p=100,000$.
@@
 
+I thank J.~Lafferty for pointing out the convexity of the likelihood function, which lead to our improved implementation with interior point methods. 
 Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text

### 2015-12-16 code/dsc-shrink/methods/ash.wrapper.R

@@ -29,7 +29,7 @@ bayes.wrapper = function(input, meta, args=NULL){
   g$mean = c(0,g1$mean)
   g$sd = c(0,g1$sd)
   #do computations for bayes rule by running ash with g, with 0 iterations
-  res = do.call(ash, args= list(betahat=input$betahat,sebetahat=input$sebetahat,df=input$df,g=g,control=list(maxiter=0)))
+  res = do.call(ash, args= list(betahat=input$betahat,sebetahat=input$sebetahat,df=input$df,g=g,fixg=TRUE))
   return(res)
 }
 

### 2016-01-20 paper/main.tex
update paper

@@
 assumes that the distribution of the actual (unobserved) effects being tested is unimodal, with a mode at 0.
-This ``unimodal assumption" (UA), which is common in other related contexts, but very different from assumptions usually made
-in FDR analyses, yields more accurate inferences than existing methods provided that the UA holds.
+This ``unimodal assumption" (UA), although natural in many contexts, is very different from assumptions usually made
+in FDR analyses, and yields more accurate inferences than existing methods provided that it holds.
 The UA also facilitates efficient and robust computation because estimating the unimodal distribution involves solving a simple convex optimization problem.
@@
 existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}. This yields a very simple, fast, and stable computer 
-implementation, as well as improving inference of FDR when the unimodal assumption is correct.  
+inference. The first idea is to {\it assume that the distribution of effects is unimodal}.  We provide a very simple, fast, and stable computer 
+implementation for perfoming EB inference under this assumption, and illustrate how it can improve inference of FDR when the unimodal assumption is correct.  
 The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
@@
 
+Although we have focussed here on applications to FDR, the idea of performing Empirical Bayes inference by flexibly esimating an underlying unimodal prior distribution
+could be useful more generally - for example in shrinkage estimation contexts such as wavelet denoising \cite{donoho:1995}. 
+Importantly, our work demonstrates that EB inference under a general unimodal assumption is, if anything, computationally simpler  
+than commonly-used more restrictive assumptions -- such as a spike and slab or Laplace prior distribution \cite{johnstone2004needles} --  as well as being more flexible.
 Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
-(The names comes from the fact that EB methods in general -- and, due to the UA, our EB method in particular -- perform shrinkage estimation,
-and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation. 
-This will be discussed further in Xing and Stephens (in preparation).) Code and instructions for reproducing analyses and figures in this paper are at
+(The name comes from the fact that our EB approach using a unimodal prior naturally results in shrinkage estimation,
+and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation.) 
+Code and instructions for reproducing analyses and figures in this paper are at
 \url{https://github.com/stephenslab/ash}.
@@
 we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}.
+for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}; see also \cite{efron1993bayes}.
 Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 

### 2016-01-26 analysis/index.Rmd
update README
change dsc-robust to dsc-opt
update all html for latest ash

--- 2015-10-26 analysis/index.Rmd
@@
 *  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
-
-## DSC Robust (Simulation study to assess EM convergence)
-
-*  [Check robustness simulation study](summarize_dsc_robust.html)
+*  [Adaptive Shrinkage](adaptive_shrinkage.html)
+*  [Comparing models: EE vs ET models](CompareEEvsETmodel.html)
 
@@
 
+## Optimization method: EM vs Interior Point
+*  [Compare loglik achieved by EM vs Interior Point](checkIP.html)
+
 ## DSC Znull (assess behaviour of ash LR under null)

### 2016-01-26 analysis/make_GOODPOOR_figs.Rmd
update all html for latest ash

@@
 First perform a simple simulated example to illustrate high and low signal.
-True values of beta simulated from $(0.5 N(0,1) + 0.5 delta_0)$.
+True values of beta simulated from $(0.5 N(0,1) + 0.5 \delta_0)$.
 ```{r}
@@
 res.locfdr = locfdr(zscore,nulltype=0,plot=0)
-res.ash = ash(betahat,sebetahat,method="fdr")
+res.ash = ash(betahat,sebetahat,method="fdr",outputlevel=4)
 
@@
 
-res.ash.ET = ash(betahat,sebetahat,method="fdr",model="ET",mixcompdist="normal")
+res.ash.ET = ash(betahat,sebetahat,method="fdr",model="ET",mixcompdist="normal",outputlevel=4)
 p=plot_pval_vs_lfsr(rbind(

### 2016-01-26 code/dsc-shrink/run_dsc.R

@@ -1,4 +1,6 @@
 library(dscr)
+library(ashr)
+sessionInfo()
 
 dsc_shrink=new_dsc("shrink","../../output/dsc-shrink-files")
 source("add_named_scenarios.R")

### 2016-01-26 code/dsc-shrink/score.R

@@ -51,5 +51,5 @@ score_betahat = function(data, output){
 }
 
 score_logLR = function(data,output){
-  return(c(logLR=output$logLR))
+  return(c(logLR=ashr::calc_logLR(output,data$input$betahat,data$input$sebetahat,NULL)))
 }

### 2016-01-27 analysis/CompareEEvsETmodel.Rmd
add Rmd

---
title: "Comparing Exchangeable Effects (EE) and Exchangeable Standardized Effects (ET) models"
output: html_document
---

The ashr package implements a shrinkage-based Empirical Bayes
method for estimating the values of effects $\beta_j$ based
on estimates of the effects ($\hat\beta_j$) and their standard errors ($s_j$).

A key modelling assumption is that the effects $\beta_j$ are identically distributed
from a unimodal distribution $g$. 
In particular, ashr assumes that $\beta_j$ is independent of the standard error $s_j$.
That is,
$$\beta_j | s_j \sim g(\cdot)     \quad (*).$$

Here we consider an alternative modelling assumption:
$$\beta_j/s_j | s_j \sim g(\cdot)   \quad (**).$$
Under this alternative assumption
the *t statistics* $\betahat_j/s_j$ are identically distributed from a unimodal
distribution. Under this assumption the expected size of the *unstandardized* effect $\beta_j$ depends on the standard error $s_j$: the larger $s_j$ is, the larger (in absolute value) $\beta_j$ is expected to be.

By analogy with Wen and Stephens (AoAS), we refer to the model (*) as the EE model ("exchangeable effects"), and the model (**)
as the ET model ("exchangeable T statistics model")

What might motivate the ET model? 
We can provide two distinct motivations. First, suppose for concreteness that the effects
$\beta_j$ reflect differences in gene expression between two conditions. One factor that affects
$s_j$ is the variance of gene $j$ within each condition. One could imagine that, perhaps, genes with a larger variance within each condition are less tightly regulated, and therefore more likely to show a large difference between conditions (i.e. large $\beta_j$) than genes with a small variance. This provides
a biological motivation for the possibility that larger $s_j$ might 
correlate with larger $\beta_j$ (although of course not for the exact functional form above).
A second motivation is more statistical: it turns out that this assumption is in some sense
the implicit assumption made by existing methods to fdr analysis based on p values. 
More specifically, under this alternative assumption, when 
attempting to identify ``significant" effects, the empirical Bayes approach will rank the genes 
in the same way as the usual $p$ values computed from $\hat\beta_j/s_j$.

It is straightforward to use the ashr package to perform analysis under the ET model:
simply specify model="ET". (Internally, this causes ash to 
replaces betahat with the standardized betahat, $\hat\beta_j/s_j$, and the standard errors
for these standarized betahat values with 1.) It is also straightforward to compare
the two competing modelling assumptions (EE vs ET) by computing the log likelihood ratio, $log[p_{EE}(\hat\beta | s, \hat g_{EE})/p_{ET}(\hat\beta | s, \hat g_{ET})]$.

We now illustrate by a simulated example. We assume that the standard errors come
from a gamma distribution, and then generate the effects $\beta_j$ under the ET model so that genes with bigger standard errors tend to have bigger effects.
```{r}
  set.seed(1234)
  nsamp=1000
  betahat.se = rgamma(nsamp,1,1)
  beta = betahat.se * rnorm(nsamp) #simulate effects under the alternative assumption 
  betahat = rnorm(nsamp,beta,betahat.se)
  zscore = betahat/betahat.se
  pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
  plot(betahat, -log(pval))
```

Here is the EE analysis.
```{r}
  library(ashr)
  ashEE.res = ash(betahat, betahat.se,method="fdr",model="EE")
```

And here is how we can perform the ET analysis,
by replacing betahat with the standardized betahat, and replacing the standard error with 1:
```{r}
  ashET.res = ash(betahat, betahat.se,method="fdr", model="ET")
```
Note that the posterior distributions computed by ash will now be for the standardized
effects, and not for the effects $\beta_j$. However, the local false sign rate (lfsr) and
related quantities are unaffected by this because the standardized effect is positive if and only if the effect is positive.

Now we compare the EE vs ET models:
```{r}
  ashEE.res$loglik-ashET.res$loglik
```

Then the log likelihood ratio is loglikEE-loglikET=`r ashEE.res$loglik-ashET.res$loglik`. This highly negative
loglikelihood indicate that the data strongly favor 
the ET model, which is expected because the data were generated under this model.
(One might be tempted to ask whether the log likelihood ratio is ``significant". We don't 
know how to address this question, but suggest in practice it doesn't matter: if the loglikelihood ratio is positive then use EE, if it is negative then use ET.)

The above illustrates these ideas on simulations from the ET model. For comparison,
we now provide results for simulations under the EE model.
```{r}
  set.seed(1234)
  nsamp=1000
  betahat.se = rgamma(nsamp,1,1)
  beta = rnorm(nsamp) #simulate effects under the original assumption 
  betahat = rnorm(nsamp,beta,betahat.se)
  zscore = betahat/betahat.se
  pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
  ashEE.res = ash(betahat, betahat.se,method="fdr",model="EE")
  ashET.res = ash(betahat, betahat.se, method="fdr",model="ET")
  ashEE.res$loglik-ashET.res$loglik
```
So here the log likelihood ratio is positive, indicating that the EE model is preferred
(which is expected since the data were generated under that model).

### 2016-01-27 analysis/IPvsEM.Rmd
add Rmd

---
title: "A brief comparison of speed of Interior Point and EM algorithms"
author: "Matthew Stephens"
date: 2016-01-26
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

# Set up simulation

The following simulation assumes $\beta \sim N(0,betasd^2)$ and with $\hat{\beta}$ having standard deviation 1.
```{r}
timed_sims = function(ash.args,nsim=20,nmin = 100,nmax = 1000,betasd=1){
  n=10^seq(log10(nmin),log10(nmax),length=nsim)
  elapsed.time= rep(0,nsim)
  for(i in 1:nsim){
    set.seed(i)
    betahat = rnorm(n[i],sd=sqrt(1+betasd^2))
    elapsed.time[i] = system.time(do.call(ashr::ash,args = modifyList(ash.args,list(betahat = betahat, sebetahat=1))))[3]
  }
  return(data.frame(elapsed.time=elapsed.time,seed=1:nsim,n = n))
}
```

Now run a simulation study, for $n$ (number of tests; $p$ in paper) in range 10 to 100,000.
Note that the warnings are being generated by the EM algorithm in big problems due to lack of convergence.

```{r}
  df=data.frame()
  for(method in c("fdr","shrink")){
    for(optmethod in c("mixIP","cxxMixSquarem")){
      df = rbind(df, data.frame(method=method, optmethod=optmethod,  timed_sims(list(method=method,optmethod=optmethod),nsim=50,nmin=10,nmax=100000)))
    }
  }

```

# Results

Now plot the time as a function of $n$
```{r}
  ggplot2::qplot(x=n,y=elapsed.time,data=df,col=optmethod,facets=.~method,ylab="elapsed time (s)")
```

And a zoom-in on "small" problems with $n<5000$.
```{r}
  ggplot2::qplot(x=n,y=elapsed.time,data=dplyr::filter(df,n<5000),col=optmethod,facets=.~method,ylab="elapsed time (s)")
```

# Summary

The IP method clearly scales to large problem much better than EM. It is faster, and also more reliable (sometimes reaches higher log-likelihood than EM, never smaller); see
[checkIP.html](checkIP.html).

However, for small problems (n<5000) the EM is adequate for many practical purposes, solving within a few seconds. This is particularly true for the penalty term (method=fdr), which
 helps the EM converge, presumably because it is helping identifiability, removing the large flat parts of the likelihood objective function.

Indeed, for small problems with the penalty term (n<2000) EM with squarem acceleration is a little faster in these comparisons than IP.

## Session information

```{r info}
sessionInfo()
```

### 2016-01-27 analysis/summarize_dsc_znull.Rmd
update dsc_znull

@@
 
-Note that the ash.n.s basically conforms exactly to the null expectation
+Note that the ash.n.s conforms closely to the null expectation
 of 0.5 chi2_0 +0.5 chi2_1 from Stram and Lee, Biometrics
@@
 
-From this
-
-a 95% procedure would need to check for logLR > `r qchisq(0.9,df=1)/2`.
+From this a 95% procedure would need to check for logLR > `r qchisq(0.9,df=1)/2`.
 
 ```{r}
-ggplot(res %>% filter(logLR!=0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
+ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
   stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
-ggplot(res %>% filter(logLR!=0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
+ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
   stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
-ggplot(res %>% filter(logLR!=0) %>% filter(grepl("ash.n",method)), aes(sample = logLR)) + facet_grid(scenario~method) +
+ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.n",method)), aes(sample = logLR)) + facet_grid(scenario~method) +
   stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")

### 2016-01-27 paper/main.tex
add hyperrefs

@@
 
-\usepackage{endfloat}
+%\usepackage{endfloat}
+\usepackage{hyperref}
 
@@
 can be more robustly estimated. Our methods are implemented in an R package {\tt ashr}
-available from \url{http://github.com/stephens999/ashr}.
+available from \href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr}.
 
@@
 than commonly-used more restrictive assumptions -- such as a spike and slab or Laplace prior distribution \cite{johnstone2004needles} --  as well as being more flexible.
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \url{http://github.com/stephens999/ashr}.
+Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr}.
 (The name comes from the fact that our EB approach using a unimodal prior naturally results in shrinkage estimation,
@@
 Code and instructions for reproducing analyses and figures in this paper are at
-\url{https://github.com/stephenslab/ash}.
+\href{https://github.com/stephenslab/ash}{https://github.com/stephenslab/ash}.
 
@@
 Both steps 1 and 2 are very straightforward.
-Step 1 is a convex optimization problem, and can be solved very quickly using interior point methods \cite{boyd2004convex,koenker2013convex}.
-Our {\tt R} package typically takes a few seconds for $p=100,000$.
+Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}.
+(Alternatively a simple EM algorithm can be used, and we found this to perform adequately for moderate-sized problems, say $J<2000$).
 And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
@@
 We introduce this penalty because in FDR applications it is considered desirable to avoid underestimating $\pi_0$ so as to avoid underestimating the FDR.
+
+Our {\tt R} package implementation typically takes about $~20$ seconds on a modern laptop for $p=100,000$, and scales linearly with $J$.
 
@@
 We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).  
+in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \href{https://cran.r-project.org/src/contrib/Archive/mixfdr/}{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).  
 In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
@@
 In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
-See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
+See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
 demonstration.
@@
 % they should, if sensible, produce estimates of FDR that agree closely with (\ref{eqn:FDRest}). And, indeed, limited
-% empirical checks suggest that this is indeed the case (\url{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
+% empirical checks suggest that this is indeed the case (\href{https://www.github.com/stephens999/ash/examples/qvalue.nonp.rmd}).
 %
@@
 to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}). Of course, we would not advocate the UA in settings where multi-modality is 
+(\href{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}{http://stephenslab.github.io/ash/analysis/check\_mixfdr\_lownoise.html}). 
+Of course, we would not advocate the UA in settings where multi-modality is 
 clearly visible in the observed $\bhat$.
@@
 %Coverage rates are given in:
-%\url{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
+%\href{file:///Users/stephens/Dropbox/Documents/git/ash/dsc-shrink/investigate_cdf.html}
 
@@
 where the $l_{kj}:= \tilde{f}_k(\bhat_j)$ are known. This is a convex optimization problem, which
-can be solved very quickly and reliably using interior point methods.
+can be solved very quickly and reliably using interior point (IP) methods.
 We used the {\tt KWdual} function from the {\tt R} package {\tt REBayes} \cite{REBayes}, which uses {\tt Rmosek} \cite{Rmosek}.
 We also found a simple EM algorithm \cite{dempster77}, accelerated using the elegant {\tt R} package {\tt SQUAREM} \cite{varadhan2008simple},
-to provide adequate performance. We initialized the EM algorithm with $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$,
+to provide adequate performance. In our EM implementation we initialized $\pi_k=1/n$ for $k=1,\dots,K$, with $\pi_0=1-\pi_1-\dots-\pi_K$,
 and the one-step updates are:
@@
 \end{align}
-The only benefit to the EM algorithm is fewer software dependencies.
+One benefit to the EM algorithm is fewer software dependencies.
+Both EM and IP methods are implemented in the {\tt ashr} package; results shown here are from the IP method, but graphs from EM are essentially the same.
+See \href{http://stephenslab.github.io/ash/analysis/checkIP.html}{http://stephenslab.github.io/ash/analysis/checkIP.html}  and \href{http://stephenslab.github.io/ash/analysis/IPvsEM.html}{http://stephenslab.github.io/ash/analysis/checkIP.html}  for comparisons.
 
@@
 prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
+git and github. The {\tt ashr} R package is available from \href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr} and includes contributions 
 from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.

### 2016-01-27 paper/main.tex
add conversion of png files to make files more manageable size

@@
 
-Our {\tt R} package implementation typically takes about $~20$ seconds on a modern laptop for $p=100,000$, and scales linearly with $J$.
+Our {\tt R} package implementation typically takes about $20$ seconds on a modern laptop for $p=100,000$, and scales linearly with $J$.
 
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfdr-1.pdf} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfdr-1.png} 
 \caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr-1.pdf} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr-1.png} 
 \caption{As in c), but for $\lfsr$ instead of $\lfdr$.
@@
 nothing about the sign of the effect.
-Thus, are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
+Thus, there are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
 whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s_nn-1.pdf} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s_nn-1.png} 
 \caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
@@
 \begin{subfigure}{\textwidth}
-\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s-1.pdf} 
+\includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s-1.png} 
 \caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).

### 2016-01-29 paper/main.tex
update paper

@@
 Both EM and IP methods are implemented in the {\tt ashr} package; results shown here are from the IP method, but graphs from EM are essentially the same.
-See \href{http://stephenslab.github.io/ash/analysis/checkIP.html}{http://stephenslab.github.io/ash/analysis/checkIP.html}  and \href{http://stephenslab.github.io/ash/analysis/IPvsEM.html}{http://stephenslab.github.io/ash/analysis/checkIP.html}  for comparisons.
+See \href{http://stephenslab.github.io/ash/analysis/checkIP.html}{http://stephenslab.github.io/ash/analysis/checkIP.html}  and \href{http://stephenslab.github.io/ash/analysis/IPvsEM.html}{http://stephenslab.github.io/ash/analysis/IPvsEM.html}  for comparisons.
 

### 2016-01-30 paper/main.tex
Fix minor typos in paper/main.tex

--- 2016-01-27 paper/main.tex
@@
 \end{equation}
-Here, in additional to some conditional independence assumptions, we are 
+Here, in addition to some conditional independence assumptions, we are 
 effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
@@
 we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation \ref{eqn:normlik}; see also \cite{efron1993bayes}.
+for individual null hypotheses using these two numbers, using the normal approximation (\ref{eqn:normlik}); see also \cite{efron1993bayes}.
 Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
@@
 
-The model \ref{eqn:beta-alpha} for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
+The model (\ref{eqn:beta-alpha}) for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
 To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with

### 2016-02-02 paper/main.tex
fix minor typos from David Poznick

@@
 controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
+``local false sign rate" (\lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
 in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
@@
 \begin{equation} \label{eqn:marginallik}
-p(\bhat | \shat, \pi) =   \prod_j  [\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2)],
+p(\bhat | \shat, \pi) =   \prod_j  \left[\sum_{k=0}^K \pi_k N(\bhat_j; 0, \sigma_k^2 + \shat_j^2) \right],
 \end{equation}
@@
 \item Compute quantities of interest from the conditional distributions $p(\beta_j | \bhat, \shat, \hat{\pi})$. For example, the evidence
-against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | | \bhat, \shat,\hat{\pi})$.
+against the null hypothesis $\beta_j=0$ can be summarized by $p(\beta_j \neq 0 | \bhat, \shat,\hat{\pi})$.
 \end{enumerate}
@@
 whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
-the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey  (\cite{tukey1991philosophy}) argued that
+the null hypothesis $H_j: \beta_j=0$ is often implausible. For example, Tukey \cite{tukey1991philosophy} argued that
 ``All we know about the world teaches us that the effects of $A$ and $B$ are always different -- in some decimal place -- for any $A$ and $B$.
-Thus asking `Are the effects different?' is foolish." Instead, Tukey suggested (\cite{tukey1962future}, p32,) that one should address
+Thus asking `Are the effects different?' is foolish." Instead, Tukey (\cite{tukey1962future}, p32) suggested  that one should address
 \begin{quote}
@@
 \begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
+\lfsr_j := \min[ p(\beta_j \geq 0| \hat\pi, \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
 \end{equation}
@@
 \caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
-The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
+The figure shows, for a single simulated dataset, the ways different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
 In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
@@
 This would increase the null component (dark blue) at the expense of
-the alternative component (light blue).
-Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the light-blue histogram at 0.
+the alternative component (cyan).
+Because the null component is $N(0,1)$, and so is biggest at 0, this would eventually create a ``dip" in the cyan histogram at 0.
 The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
+In contrast the ZA pushes the dark blue component until the cyan component {\it disappears} at 0.
 See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
@@
 
-The key point we want to make here is estimation of $\pi_0$, and the accompanying identifiability issues,
+The key point we want to make here is that estimation of $\pi_0$, and the accompanying identifiability issues,
  become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
@@
 effect size for each selected case?". \cite{efron2008microarrays} notes the potential for EB approaches to tackle this problem,
-and \cite{zhao2012empirical} consider in detail the case where the non-null effects are normally distributed.
+and \cite{zhao2012empirical} considers in detail the case where the non-null effects are normally distributed.
 %but also notes the difficulties of implementing a stable general algorithm.

### 2016-05-03 analysis/index.Rmd
update index
add simulation from referee report

@@
 *  [Comparing models: EE vs ET models](CompareEEvsETmodel.html)
+*  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)
 

### 2016-05-03 paper/main.tex
add simulation from referee report

--- 2016-01-30 paper/main.tex
@@
 The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
-to summarize each measurement. This idea allows variations in measurement precision to be better accounted for,
- and avoids a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
+to summarize each measurement. Here we use this idea to allow variation in measurement precision to be better accounted for,
+ avoiding a problem with standard pipelines that poor-precision measurements can inflate estimated FDR.
+(\cite{ploner2006multidimensional} also suggest using more than one number in FDR analyses, taking a
+rather different approach to the one we use here.)
 
@@
 controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" (lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
+``local false sign rate" ($\lfsr$), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
 in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
@@
 
-Although we have focussed here on applications to FDR, the idea of performing Empirical Bayes inference by flexibly esimating an underlying unimodal prior distribution
+Although we have focussed here on applications to FDR, the idea of performing EB inference by flexibly esimating an underlying unimodal prior distribution
 could be useful more generally - for example in shrinkage estimation contexts such as wavelet denoising \cite{donoho:1995}. 
@@
 than commonly-used more restrictive assumptions -- such as a spike and slab or Laplace prior distribution \cite{johnstone2004needles} --  as well as being more flexible.
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at \href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr}.
+Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at 
+\href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr}.
 (The name comes from the fact that our EB approach using a unimodal prior naturally results in shrinkage estimation,
@@
 
-Suppose that we are interested in the values of $J$ ``effects" $\beta=(\beta_1,\dots,\beta_J)$. For example, in a typical genomics application
-that aims to identify differentially expressed genes,  $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
+Let $\beta=(\beta_1,\dots,\beta_J)$ denote $J$ ``effects" of interest. For example, in a genomics application 
+$\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
 In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, in testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
+that is, testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
 more general problem of estimating, and assessing uncertainty in, $\beta_j$.
@@
 \end{align}
-where $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
+where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
 Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyper-parameters, which are non-negative and sum to one, and are to be estimated,
@@
 represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
-(We encourage the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
-although of course in practice we use a finite grid -- see Implementation Details.)
-
-For the likelihood $p(\bhat | \beta, \shat)$ we assume  
+We encourage the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
+although of course in practice we use a finite grid -- see Implementation Details.
+
+For the likelihood $p(\bhat | \beta, \shat)$ we assume a normal approximation:
 \begin{equation} \label{eqn:normlik}
@@
 \end{equation}
-Here, in addition to some conditional independence assumptions, we are 
-effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
+%Here, in addition to some conditional independence assumptions, we are 
+%effectively  assuming that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
 
@@
 with larger standard error will have a flatter likelihood, and therefore have less impact on inference.
-However, this simple model also has several additional assumptions that can be relaxed. Specifically,
+However, the model also has several additional assumptions that can be relaxed. Specifically,
 \begin{enumerate}
 \item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
-More flexibility can be obtained by replacing the mixture of normals with mixtures of uniforms [see equation (\ref{eqn:g-general})]; indeed this 
+More flexibility can be obtained using mixtures of uniforms [see equation (\ref{eqn:g-general})]; indeed this 
 can allow $g$ to approximate {\it any} unimodal distribution. 
 \item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
-We can relax this to allow for a relationship between these quantities [see (\ref{eqn:beta-alpha})].
-\item The likelihood (\ref{eqn:normlik}) effectively assumes that the number of observations used to compute $\bhat_j,\shat_j$ are sufficiently large to justify a normal approximation.
-We can generalize this likelihood using a $t$ likelihood [see (\ref{eqn:tlik})].
+This can be relaxed [see (\ref{eqn:beta-alpha})].
+\item The normal likelihood (\ref{eqn:normlik}) can be generalized to a $t$ likelihood [see (\ref{eqn:tlik})].
 \end{enumerate}
 These embellishments are detailed in Detailed Methods.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions encapsulated in our model
-(which are also made by most existing EB approaches to this problem). Correlations among tests certainly arise in practice, either due to genuine correlations
-in the system of study, or due to unmeasured confounders, and their potential to impact results of an FDR analysis is important to consider
-whatever analysis methods are used: see \cite{efron2007correlation,leek:2007} for relevant discussion.
+Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions
+(which are also made by most existing approaches). Correlations among tests certainly arise in practice, either due to genuine correlations 
+or due to unmeasured confounders, and their potential impact on estimated FDRs is important to consider
+whatever analysis methods are used \cite{efron2007correlation,leek:2007}.
 
@@
 
-In words, the model above assumes that the effects $\beta_j$ are independent and identically distributed from a mixture of zero-centered normal distributions, and 
-each observation $\bhat_j$ is a noisy measurement of $\beta_j$ with standard error $\shat_j$.
-Together, these assumptions imply that the observations $\bhat_j$ are also independent observations, each from a mixture of normal distributions:
+Together, (\ref{eqn:beta})-(\ref{eqn:normlik}) imply that $\bhat_j$ are independent observations from a mixture of normal distributions:
 \begin{equation} \label{eqn:marginallik}
@@
 Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}.
-(Alternatively a simple EM algorithm can be used, and we found this to perform adequately for moderate-sized problems, say $J<2000$).
+(Alternatively a simple EM algorithm can also work well, particularly for modest $J$, say $J<2000$).
 And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
@@
 
-Our {\tt R} package implementation typically takes about $~20$ seconds on a modern laptop for $p=100,000$, and scales linearly with $J$.
+Our {\tt R} package implementation typically takes $~20s$ on a modern laptop for $J=100,000$, and scales linearly with $J$.
 
@@
 
-The $\lfdr$, like most other measures of significance (e.g. $p$ values and $q$ values), is rooted in the hypothesis testing paradigm which focuses on 
+The $\lfdr$, like most other measures of significance, is rooted in the hypothesis testing paradigm which focuses on 
 whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
@@
 \begin{equation} \label{eqn:lfsr}
-\lfsr_j := \min[ p(\beta_j \geq 0| \bhat, s), p(\beta_j \leq 0| \hat\pi, \bhat, s) ].
+\lfsr_j := \min[ \Pr(\beta_j \geq 0| \hat\pi, \bhat, s), \Pr(\beta_j \leq 0| \hat\pi, \bhat, s) ].
 \end{equation}
@@
 \begin{gather*}
-p(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
-p(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
-p(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
+\Pr(\beta_j < 0| \bhat, s, \hat\pi)=0.95, \\
+\Pr(\beta_j =0| \bhat, s, \hat\pi) = 0.03, \\
+\Pr(\beta_j >0| \bhat, s, \hat\pi) = 0.02.
 \end{gather*} 
@@
 from (\ref{eqn:lfsr}) $\lfsr_j=\min(0.05,0.98)=0.05$ (and, from (\ref{eqn:fdr}), $\lfdr_j=0.03$). This $\lfsr$ corresponds to the fact that, given these results, 
-our best guess for the sign of $\beta_j$ is that it is negative, and the probability that this guess is wrong would be $0.05$.
-
-As our notation suggests, $\lfsr_j$ is intended to be compared and contrasted with $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
+we would guess that $\beta_j$ is negative, with probability $0.05$ of being wrong.
+
+As our notation suggests, $\lfsr_j$ is analogous to $\lfdr_j$: whereas small values of $\lfdr_j$ indicate that we can be {\it confident that $\beta_j$ is non-zero}, 
 small values of $\lfsr_j$ indicate that we can be {\it confident in the sign of $\beta_j$}. 
@@
 In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ can be substantially more robust to modeling assumptions than $\lfdr$.
+$\lfsr$ is more robust to modeling assumptions.
 
@@
 $\Gamma \subset \{1,\dots,J\}$. For example, 
-\begin{equation} \label{eqn:FDRhat}
-\widehat\FDR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \fdr_j. 
-\end{equation}
-estimates the FDR we would obtain if we were to declare all tests in $\Gamma$ significant.
-And 
+\begin{equation} \label{eqn:FSRhat}
+\widehat\FSR(\Gamma):= (1/|\Gamma|) \sum_{j \in \Gamma} \lfsr_j 
+\end{equation}
+is an estimate of the total proportion of errors made if we were to estimate the sign of all effects in $\Gamma$
+(an error measure analogous to the usual (tail) FDR).
+And, we can define the ``$s$-value" 
 \begin{equation} \label{eqn:qvalue}
-q_j := \widehat \FDR(\{k: \fdr_k \leq \fdr_j\}) 
-\end{equation}
-provides a measure of significance analogous to Storey's $q$ value \cite{storey.03}.
+s_j := \widehat \FSR(\{k: \lfsr_k \leq \lfsr_j\}) 
+\end{equation}
+as a measure of significance analogous to Storey's $q$-value \cite{storey.03}.
+(Replacing $\lfsr_j$ with $\lfdr_j$ in these definitions gives estimates for the usual $\FDR(\Gamma)$, and a $q$ value for each observation respectively.)
 
@@
 we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses using these two numbers, using the normal approximation (\ref{eqn:normlik}); see also \cite{efron1993bayes}.
-Using two numbers instead of one clearly has the potential to be more informative, and indeed, results later (Figure \ref{fig:goodpoor}) illustrate 
-how it can improve performance by taking better account of variation in measurement precision among observations.
+for individual null hypotheses in a similar way; see also \cite{efron1993bayes}.
+Using two numbers instead of one clearly has the potential to be more informative, and this is illustrated in Results later (Figure \ref{fig:goodpoor}).
+
 %(And in the special case where $\shat_j$ are equal for all $j$, our approach essentially comes down to the same thing as these existing methods.)
@@
 %allows for different modelling assumptions to be compared through likelihood comparisons.
-Second, our unimodal assumption (UA) that the effects are unimodal about zero is quite different from assumptions made by $\qvalue, \locfdr$ or $\mixfdr$.
-Indeed, $\locfdr$ assumes that all $z$ scores near 0 are null (Efron calls this the Zero Assumption; ZA), which implies that 
-under the alternative hypothesis the distribution of $z$ scores has {\it no mass at 0}; this contrasts strikingly with the UA, which implies that this distribution has its peak at 0! 
-Similarly, $\qvalue$ assumes that all $p$ values near 1 are null,
-which is the same as the ZA because $p$ values near 1 correspond to $z$ scores near 0. And although \mixfdr does not formally make the ZA,
-we have found that in practice, with default settings, the results often approximately satisfy the ZA (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
-Thus, not only do these existing methods not make the UA, they actually make assumptions that are, in some sense, as different
-from the UA as they can be.
-
-Given that the UA and ZA are so different, it seems worth discussing why we generally favor the UA. 
-Although the UA will not apply to all situations, we believe that it will often be reasonable, especially in FDR-related contexts
-that have traditionally focussed on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis,  
-it seems reasonable to expect that ``$\beta_j$ very near 0" is also plausible. Further, it seems reasonable to expect that larger effects become
+%\def\sZA{ZA}
+Second, our unimodal assumption (UA) that the effects are unimodal about zero, is an assumption 
+not made by existing methods, and one that we argue to be both plausible and beneficial in many contexts.
+Although the UA will not hold in all settings, it will often be reasonable, especially in FDR-related contexts
+that focus on rejecting the null hypotheses $\beta_j=0$. This is because if ``$\beta_j=0$" is a plausible null hypothesis
+then ``$\beta_j$ very near 0" should also be plausible. Further, it seems reasonable to expect that larger effects become
 decreasingly plausible, and so the distribution of the effects will be unimodal about 0. 
@@
 
+% Indeed, neither $\locfdr$ nor $\qvalue$ explicitly model the effect distribution at all:  they effectively avoid this by simply subtracting a null distribution 
+%(of $z$ scores or $p$ values) from the observed empirical distribution, with the remains necessarily 
+%representing the alternative distribution. When doing this they assume that all the $z$ scores near zero (or, equivalently,
+%all $p$ values near 1) are null, which results in a ``hole" in the alternative distribution
+
+%Indeed, $\locfdr$ \cite{efron2008microarrays} assumes that all $z$ scores near 0 are null 
+%(which we call the Zero Assumption, \sZA), which implies that 
+%under the alternative hypothesis the distribution of $z$ scores has {\it no support at 0}.  
+%(Although \cite{efron2008microarrays} motivates and defines the \sZA as ``most $z$ scores near 0 are null",
+%results from $\locfdr$ actually obey the stronger version of the \sZA  that we use here: {\it all} $z$ scores near 0 are null.)
+%Similarly, $\qvalue$ assumes -- when estimating $\pi_0$ -- that all $p$ values near 1 are null.
+%which, if the $p$ values were derived from $z$ scores, is the ZA because $p$ values near 1 correspond to $z$ scores near 0.
+%And although \mixfdr does not formally make the \sZA,
+%we have found that in practice, with default settings, the results often approximately satisfy this assumption (due, we believe, to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}).
 %Tukey "peri-null"
@@
 
-In further support of the UA for FDR applications, we note that almost all analogous work in sparse regression models make the UA for the regression coefficients -- 
-common choices of uni-modal distribution being the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than the approach we take here, which provides for general uni-modal distributions, and 
+In further support of the UA, note that large-scale regression methods almost always make the UA for the regression coefficients,
+which are analogous to the ``effects" we estimate here. 
+Common choices of uni-modal distribution for regression coefficients 
+include the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
+These are all less flexible than our approach, which provides for general uni-modal distributions, and 
 it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-The UA assumption on regression coefficients is directly analagous to our UA here, and so its
-widespread use in the regression context supports its use here.
-
-Alternatively, we could motivate the UA by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
+Additionally, the UA can be motivated by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
 such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
@@
 
-Finally, the UA also has a considerable practical benefit: it yields simple algorithms that are both computationally
+Besides its plausibility, the UA has two important practical benefits: it facilitates more accurate estimates
+of FDR-related quantities, and it yields simple algorithms that are both computationally
 and statistically stable (see Results).
@@
 Also, they provide no software implementation.
+
 More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
@@
 
-Here we consider the effects of making the UA. To isolate these effects we consider the simplest case, where every observation has the same
+Here we consider the effects of making the UA. 
+To isolate these effects we consider the simplest case, where every observation has the same
 standard error, $s_j=1$ and all methods are provided that information. That is,
 $\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the only difference between our method and methods like $\locfdr$ and $\mixfdr$ are in how they estimate $g$.
+In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the primary difference among methods lies in how they estimate $g$.
 
@@
 \begin{enumerate}
-\item The UA can produce very different inferences compared with the ZA made by existing methods.
+\item The UA can produce quite different results from existing methods.
 \item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
@@
 
-\subsubsection*{The UA and ZA can produce different inferences}
+\subsubsection*{The UA can produce quite different results from existing methods.}
 
@@
 In the $z$ score space the alternative distribution is placed on the bottom to highlight the differences in its shape among methods.
-The three existing methods (\qvalue, \locfdr, \mixfdr) all effectively make the Zero Assumption, which results in a ``hole" in the alternative $z$ score distribution around 0.
+The three existing methods (\qvalue, \locfdr, \mixfdr) all produce a ``hole" in the alternative $z$ score distribution around 0.
 In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
@@
 
-To illustrate the different inferences from the UA and ZA we show results for a single
-dataset simulated with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
+We illustrate the effects of the UA, and the difference with existing methods, using a simple simulated dataset
+ with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
 but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
 \qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-The results (Figure \ref{fig:ZA}) illustrate the clear difference between the existing methods and our method. The effects of the ZA made by \qvalue and \locfdr
-are visually clear, producing a ``hole"  in the alternative $z$ score distribution around 0. Although \mixfdr does not formally make the ZA, its decomposition
-exhibits a similar hole. In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
-effectively ``filling in" the hole.  (Of course the null distribution also has a peak at 0, and the local fdr 
+(Here we are using the fact that these methods all provide an estimate of the \lfdr{} for each observation, which
+implies such a decomposition;  specifically the average \lfdr{} within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component.) The results (Figure \ref{fig:ZA}) illustrate a clear difference between the existing methods and our method:
+all the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ashr, due to the UA, has a mode near 0.
+%Although \mixfdr does not formally make the \sZA, its decomposition
+%exhibits a similar hole. 
+%In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
+%effectively ``filling in" the hole.  
+(Of course the null distribution also has a peak at 0, and the local fdr 
  under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
+
+This qualitative difference among methods in the inferred alternative distribution of the $z$ scores is quite general, and 
+not limited to this simple simulation. To understand why the alternative distribution of $z$ scores from $\locfdr$ and $\qvalue$ has a hole at zero, note that neither
+of these methods explicitly models the alternative distribution: instead they simply subtract a null distribution (of $z$ scores or $p$ values) from the observed empirical distribution, letting the alternative
+distribution be defined implicitly, by what remains.
+ In deciding how much null distribution to subtract -- that is, in estimating the null proportion, $\pi_0$ -- 
+ both methods assume that all $z$ scores near zero (or, equivalently,
+all $p$ values near 1) are null. The consequence of this is that their (implicitly-defined) distribution for the alternative $z$ scores has 
+a ``hole" at 0 -- quite different from our assumption of a mode at zero. (Why \mixfdr exhibits similar behaviour is less clear, since it does explicitly
+model the alternative distribution; however we believe it may be due to the default choice of penalty term $\beta$ described in \cite{muralidharan2010empirical}.)
 
@@
 The role of the penalty term is to push the dark blue component as far as possible, right up to (or, to be conservative, just past) the point where this dip appears.
-In contrast the ZA pushes the dark blue component until the light-blue component {\it disappears} at 0.
+In contrast the
+existing methods effectively push the dark blue component until the light-blue component {\it disappears} at 0.
 See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
@@
 
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than the ZA.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs than existing methods that make the ZA. 
+The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than existing methods.
+Consequently \ashr will estimate smaller $\lfdr$s and FDRs. 
 This is desirable, provided that these estimates remain conservative: that is,

### 2016-05-04 paper/main.tex
update

@@
 as a measure of significance analogous to Storey's $q$-value \cite{storey.03}.
-(Replacing $\lfsr_j$ with $\lfdr_j$ in these definitions gives estimates for the usual $\FDR(\Gamma)$, and a $q$ value for each observation respectively.)
+(Replacing $\lfsr_j$ with $\lfdr_j$ in these definitions gives an estimate for the usual $\FDR(\Gamma)$, and an estimated $q$ value for each observation respectively.)
 
@@
 \qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-(Here we are using the fact that these methods all provide an estimate of the \lfdr{} for each observation, which
-implies such a decomposition;  specifically the average \lfdr{} within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component.) The results (Figure \ref{fig:ZA}) illustrate a clear difference between the existing methods and our method:
+Here we are using the fact that these methods all provide an estimate of the \lfdr{} for each observation, which
+implies such a decomposition;  specifically the average \lfdr{} within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component.
+The results (Figure \ref{fig:ZA}) illustrate a clear difference between the existing methods and our method:
 all the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ashr, due to the UA, has a mode near 0.
@@
 
-I thank J.~Lafferty for pointing out the convexity of the likelihood function, which lead to our improved implementation with interior point methods. 
+I thank J.~Lafferty for pointing out the convexity of the likelihood function, which lead to an improved implementation with interior point methods. 
 Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text

### 2016-05-04 analysis/referee_uaza.Rmd

---
title: "Assessment of referee comment regarding q value and zero assumption"
author: "Matthew Stephens"
date: 2016-03-21
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

# Simulation

The referee reports a simulation of 10,000 hypotheses, a fraction 0.8 are null ($\mu_i=0$), and a fraction 0.2 are alternative with effect $\mu_i \sim N(0,1)$. For each effect $i$ there
are n=30 observations from $x_{i,j} \sim N(\mu_i,1)$ and a one-sample t test is used to compute a $p$ value. The referee correctly notes that these simulations i) obey the
unimodal assumption, and ii) have the property that "most" $z$ scores near 0 are null.

Here we repeat these simulations and apply qvalue to check how it behaves.

```{r}
set.seed(100)
mu = c(rep(0,8000),rnorm(2000))
x = matrix(rnorm(30*10000),nrow=10000,ncol=30)+mu
ttest.pval = function(x){return(t.test(x)$p.value)}
x.p = apply(x,1,ttest.pval)
```

Here is how qvalue (arguably, implicitly) decomposes the distribution of p values into null and alternative
```{r}
source("../R/nullalthist.R")
lfdr.qv = qvalue::qvalue(x.p)$lfdr
altnullhist(x.p,lfdr.qv,main="p values: qvalue",ncz=40,xlab="p value",cex.axis=0.8)
```

And here is the corresponding plot in z score space. Note the "hole" in the distribution of $z$ scores at 0 under the alternative.
```{r}
ttest.est = function(x){return(t.test(x)$estimate)}
x.est = apply(x,1,ttest.est)
zscore = x.est/(1/sqrt(30))
nullalthist(zscore,lfdr.qv,main="qvalue's implicit partition of z \n into null and alternative",ncz=60,cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6),xlab="z score")
```

For comparison here is the decomposition, using ash, of $z$ scores into null and alternative components.
```{r}
lfdr.ash=ashr::ash(x.est,1/sqrt(30))$lfdr
nullalthist(zscore,lfdr.ash,main="ash partition of z \n into null and alternative",ncz=60,cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6),xlab="z score")
```

For completeness here is (roughly) the ``true" decomposition of $z$ scores into null and alternative components.
```{r}
trueg = ashr::normalmix(c(0.8,0.2),c(0,0),c(0,1))
lfdr.true=ashr::ash(x.est,1/sqrt(30),g=trueg,fixg=TRUE)$lfdr
nullalthist(zscore,lfdr.true,main="true partition of z \n into null and alternative",ncz=60,cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6),xlab="z score")
```

# Explanation

In estimating $\pi_0$ qvalue assumes that *all* p values near 1 are null.
This is equivalent to assuming that *all* z scores near 0 are null. Thus,
despite the fact that qvalue does not explicitly model the distribution of
p values or z scores under the alternative, its way of estimating $\pi_0$
necessarily creates a hole in the z score distribution at 0 under the alternative.

## Session information

```{r info}
sessionInfo()
```

### 2016-05-04 paper/main.tex

@@
 controlling the FDR, and not on estimation. To help provide a bridge between FDR and estimation we introduce the term
-``local false sign rate" ($\lfsr$), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
+``local false sign rate" (\lfsr), which is analogous to the ``local false discovery rate" (\lfdr) \cite{efron2008microarrays}, but which measures confidence  
 in the {\it sign} of each effect rather than confidence in each effect being non-zero. We show that in some settings, particularly those with many discoveries, 
@@
 existing methods effectively push the dark blue component until the light-blue component {\it disappears} at 0.
-See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive
-demonstration.
+See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive demonstration.
 

### 2016-05-04 analysis/estimate_mode.Rmd
add estimate_mode

---
title: "Estimate Non-zero mode"
author: "Matthew Stephens"
date: 2016-02-01
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

The purpose of this analysis is to check the performance of the non-zero mode option.

## Simple simulation

I conjecture that the mean of the posterior means should be close to the optimum for the mode...
maybe even equal to it. (That is, this would constitute a fixed point for the update. We aren't
explicitly using that directly in the current implementation; the uniform mixture uses optim to
do it numerically; the normal mixture uses a true EM I think...)

```{r}

check_mode = function(betahat, sebetahat,mixcompdist){
    z.ash=ashr::ash(betahat,sebetahat,mixcompdist=mixcompdist,nonzeromode=TRUE)
    average.posteriormean = mean(z.ash$PosteriorMean)
    fitted.mode = ashr::comp_mean(z.ash$fitted.g)[1]

    z.ash1 = ashr::ash(betahat-fitted.mode,sebetahat, mixcompdist=mixcompdist) #refit to get g
    g = z.ash1$fitted.g
    loglik = z.ash1$loglik

    loglik.down = ashr::ash(z-fitted.mode-0.01,1,g=g)$loglik
    loglik.up  = ashr::ash(z-fitted.mode+0.01,1,g=g)$loglik
    loglik.posteriormean = ashr::ash(z-average.posteriormean,1,g=g)$loglik

    return(list(fitted.mode=fitted.mode, average.posteriormean =average.posteriormean,
         loglik = c(loglik,loglik.down,loglik.up,loglik.posteriormean)))
  }
  set.seed(100)
  z = rnorm(1000)+3
  check_mode(z,1,mixcompdist="uniform")
  check_mode(z,1,mixcompdist="normal")
  check_mode(z,1,mixcompdist="halfuniform")
```

```{r}
  set.seed(100)
  beta = rexp(1000)
  betahat = beta + rnorm(1000,0,0.1)
  z.ash.hu = ashr::ash(betahat,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
  z.ash.pu = ashr::ash(betahat,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")

  z.ash.hu2 = ashr::ash(betahat-0.2,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
  z.ash.pu2 = ashr::ash(betahat-0.1,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")

  z.ash.hu.nzm = ashr::ash(betahat,0.1,mixcompdist="halfuniform",nonzeromode=TRUE,method="shrink")
  z.ash.hu$loglik
  z.ash.hu.nzm$loglik

  test.op = function(c){return(-ashr::ash(betahat-c,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")$loglik)}
  optimize(test.op,interval=c(min(betahat),max(betahat)))
#  check_mode(betahat,.1,mixcompdist="halfuniform")
```

Note to self: check the normal version works too; check nonzeromodeEMobj  as it doesn't seem needed.

## Session information

```{r info}
sessionInfo()
```

### 2016-05-04 analysis/index.Rmd
add estimate_mode

@@
 *  [Compare loglik achieved by EM vs Interior Point](checkIP.html)
+*  [Compare speed of EM vs Interior Point](IPvsEM.html)
 
@@
 * [Summarize results of null simulations](summarize_dsc_znull.html)
+
+## Check how estimating the mode performs
+
+* [Basic check of non-zero model option](estimate_mode.html)

### 2016-05-09 analysis/efron.fcr.Rmd
add fcr example

---
title: "Post-selection CI example, assymetric, not unimodal at 0"
author: "Matthew Stephens"
date: 2016-05-09
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

## Introduction

This example comes from Efron (2008) p16 when examining the false coverage rate (FCR). 
I selected this example because the distribution of the non-zero effect is
highly assymetric and not at all unimodal at zero, both issues a referee asked me
to elaborate on. Specifically, the distribution of the non-zero effects is N(-3,1).
Here I simulate data, and apply ash (with the "halfuniform" option to allow for asymmetric g).

```{r}
  set.seed(10)
  nsamp=10000
  altmean=-3
  mu0 = rep(0,nsamp)
  mu1 = rnorm(nsamp,altmean,1)
  comp = rbinom(nsamp,1,0.1)
  mu = ifelse(comp==0,mu0,mu1)
  z = rnorm(nsamp,mu,1)

  res.ash = ashr::ash(z,1,mixcompdist="halfuniform")
  CImatrix= ashr::ashci(res.ash,level=0.95)
```

The following plot compares the (symmetric-tail) 95% CIs from ash (red) for the "significant" observations with Bayes rule (green), similar to Figure 8 from Efron. Note that the lower 97.5% point is pretty accurate, but the upper 97.5% point is curtailed - presumably due, at least in part, to the short tails of the uniform mixture.
```{r}
  BayesComparePlot=function(CImatrix,altmean=-3,...){
    plot(z,mu,xlim=c(-8,0),...)
    points(z[CImatrix[,1]],CImatrix[,5],col="red")
    points(z[CImatrix[,1]],CImatrix[,4],col="red")

    fdr = 0.9*dnorm(z)/(0.9*dnorm(z)+0.1*dnorm(z,altmean,sqrt(2)))
    o=order(z)
    upper = ifelse(fdr[o]<0.025,(z[o]+altmean)/2+qnorm(0.975+fdr[o])/sqrt(2),0)
    lines(z[o],upper,col="green",lwd=2)
    lines(z[o],(z[o]+altmean)/2-qnorm(0.975)/sqrt(2),col="green",lwd=2)
    abline(v=max(z[fdr<0.05]))
  }
  BayesComparePlot(CImatrix,main="CIs for highly asymmetric and non-unimodal-at-zero data")
```

## Variational version

Although not a focus of the paper, ash does have an option to do variational inference for the mixture components (with a Dirichlet prior). In practice this approach usually
ends up spreading the posterior mass up more among the mixture components. It
seemed plausible that this might lead to slightly less extreme tail behaviour than
above (because the model will put a little more weight on the uniforms with larger variance, which are essentially set to zero in the above).

```{r}
  res.ash.VB = ashr::ash(z,1,mixcompdist="halfuniform",optmethod="mixVBEM")
  CImatrix.VB= ashr::ashci(res.ash.VB,level=0.95)
```

Again, we can compare results with Bayes rule
```{r}
  BayesComparePlot(CImatrix.VB, main="CIs for highly asymmetric and non-unimodal-at-zero data \n Variational Version")
```

## Session information

```{r info}
sessionInfo()
```

### 2016-05-09 analysis/index.Rmd
update index

@@
 *  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)
+*  [Example of Credible Interval behaviour when UA does not hold](efron.fcr.html)
 

### 2016-05-25 analysis/efron.fcr.Rmd
add examples

@@
   res.ash = ashr::ash(z,1,mixcompdist="halfuniform")
-  CImatrix= ashr::ashci(res.ash,level=0.95)
+
 ```
+
+We can also run ash with the "true" g, to allow us to compare the lfsr, lfdr etc. 
+```{r}
+  true.g = ashr::normalmix(c(0.9,0.1),c(0,-3),c(0,1))
+  res.ash.true = ashr::ash(z,1,g=true.g,fixg=TRUE)
+```
+
+Here we can see how the partition of $z$ scores compares with the truth. Note the effect of the unimodal assumption is to extend the inferred alternative distribution toward 0.
+```{r}
+  source("../R/nullalthist.R")
+  par(mfcol=c(2,1))
+  nullalthist(z,lfdr=res.ash.true$lfdr,main="true partition")
+  nullalthist(z,lfdr=res.ash$lfdr,main="inferred partition")
+```
+
+Comparing the inferred Posterior Means, lfdr, and lfsr with the true values of these quantities, we find reassuringly good correspondence.
+```{r}
+  par(mfcol=c(1,1))
+  plot(res.ash.true$PosteriorMean,res.ash$PosteriorMean,xlab="Truth", ylab="ash.hu",main="Posterior Mean (inferred vs truth)",xlim=c(-6,1),ylim=c(-6,1))
+  abline(a=0,b=1,col=2,lwd=3)
+
+    plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth", ylab="ash.hu",main="lfdr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
+  abline(a=0,b=1,col=2,lwd=3)
+
+  plot(res.ash.true$lfsr,res.ash$lfsr,xlab="Truth", ylab="ash.hu",main="lfsr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
+  abline(a=0,b=1,col=2,lwd=3)
+
+```
+
+We can also run qvalue and locfdr.
+We see that locfdr perhaps performs a bit better than ash for the decomposition here, but the estimated local fdrs are pretty similar.
+Here qvalue does less well because of the asymmetry which we didn't take account of.
+```{r}
+library(qvalue)
+library(locfdr)
+res.locfdr = locfdr(z,nulltype = 0)
+res.qvalue = qvalue(p=pchisq(z^2,df=1,lower.tail = FALSE))  
+#res.pos.qvalue = qvalue(p=pchisq(z[z>0]^2,df=1,lower.tail = FALSE))  
+#res.neg.qvalue = qvalue(p=pchisq(z[z<0]^2,df=1,lower.tail = FALSE))  
+#df1= data.frame(Truth = res.ash.true$lfdr,Estimate=res.locfdr$fdr,Method="locfdr")
+#df2= data.frame(Truth = res.ash.true$lfdr,Estimate=res.ash$lfdr,Method="ash.hu")
+#df3 =data.frame(Truth = res.ash.true$lfdr,Estimate=res.qvalue$lfdr,Method="qvalue")
+#df4 =data.frame(Truth = res.ash.true$lfdr[z<0],Estimate=res.neg.qvalue$lfdr,Method="qvalue (-)")
+#df5 =data.frame(Truth = res.ash.true$lfdr[z>0],Estimate=res.pos.qvalue$lfdr,Method="qvalue (+)")
+
+#df = rbind(df1,df2,df3,df4,df5)
+#ggplot2::qplot(data=df,x=Truth,y=Estimate,facets= .~Method)
+
+par(mfrow=c(1,3))
+plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth (lfdr)", ylab="ash.hu",main="ash.hu",xlim=c(0,1),ylim=c(0,1))
+abline(a=0,b=1,col=2,lwd=3)
+plot(res.ash.true$lfdr,res.locfdr$fdr,xlab="Truth (lfdr)", ylab="Estimate",main="locfdr",xlim=c(0,1),ylim=c(0,1))
+abline(a=0,b=1,col=2,lwd=3)
+plot(res.ash.true$lfdr,res.qvalue$lfdr,xlab="Truth (lfdr)", ylab="Estimate",main="qvalue",xlim=c(0,1),ylim=c(0,1))
+abline(a=0,b=1,col=2,lwd=3)
+
+```
+
+# Uniform tail curtails Credible Intervals
 
@@
 ```{r}
+  CImatrix= ashr::ashci(res.ash,level=0.95)
   BayesComparePlot=function(CImatrix,altmean=-3,...){

### 2016-05-25 analysis/index.Rmd
add examples

--- 2016-05-04 analysis/index.Rmd
@@
 *  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)
+*  [Comparison of methods, and example of Credible Interval behaviour when UA does not hold](efron.fcr.html)
 

### 2016-05-25 paper/main.tex
add examples

@@
 assumes that the distribution of the actual (unobserved) effects being tested is unimodal, with a mode at 0.
-This ``unimodal assumption" (UA), although natural in many contexts, is very different from assumptions usually made
-in FDR analyses, and yields more accurate inferences than existing methods provided that it holds.
-The UA also facilitates efficient and robust computation because estimating the unimodal distribution involves solving a simple convex optimization problem.
+This ``unimodal assumption" (UA), although natural in many contexts, is not usually incorporated into standard FDR analysis,
+and we demonstrate how incorporating it brings many benefits.  Specifically, the UA
+facilitates efficient and robust computation -- estimating the unimodal distribution involves solving a simple convex optimization problem --
+and enables more accurate inferences provided that it holds.
 Second, the method takes as its input two numbers for each test (an effect size estimate, and corresponding standard error),
@@
 helps account for variation in measurement precision across tests. It also facilitates the estimation of actual 
-effect sizes, and our approach provides interval estimates (credible regions) for each effect in addition to measures of significance.
+effect sizes, and unlike standard FDR analysis methods
+our approach provides interval estimates (credible regions) for each effect in addition to measures of significance.
 To provide a bridge between interval estimates and significance measures we introduce the term ``local false sign rate"
@@
 can be more robustly estimated. Our methods are implemented in an R package {\tt ashr}
-available from \href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr}.
+available from \url{http://github.com/stephens999/ashr}.
 
@@
 
-Although we have focussed here on applications to FDR, the idea of performing EB inference by flexibly esimating an underlying unimodal prior distribution
-could be useful more generally - for example in shrinkage estimation contexts such as wavelet denoising \cite{donoho:1995}. 
-Importantly, our work demonstrates that EB inference under a general unimodal assumption is, if anything, computationally simpler  
+Although we focus here on FDR applications, the idea of performing EB inference using a flexible unimodal prior distribution
+is useful more generally. For example, the methods described here can be applied directly to perform shrinkage estimation for
+wavelet denoising \cite{donoho:1995}, an idea explored in a companion paper \cite{xing:2016}. And analogous ideas can be used to perform
+EB inference for variances \cite{lu2016variance}.
+Importantly, our work demonstrates how EB inference under a general unimodal assumption is, if anything, computationally simpler  
 than commonly-used more restrictive assumptions -- such as a spike and slab or Laplace prior distribution \cite{johnstone2004needles} --  as well as being more flexible.
+
 Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at 
-\href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr}.
+\url{http://github.com/stephens999/ashr}.
 (The name comes from the fact that our EB approach using a unimodal prior naturally results in shrinkage estimation,
@@
 Code and instructions for reproducing analyses and figures in this paper are at
-\href{https://github.com/stephenslab/ash}{https://github.com/stephenslab/ash}.
+\url{https://github.com/stephenslab/ash}.
 
@@
 \end{enumerate}
-These embellishments are detailed in Detailed Methods.
+These embellishments, detailed in Detailed Methods, are all implemented in our software.
 Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions
@@
 Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}.
-(Alternatively a simple EM algorithm can also work well, particularly for modest $J$, say $J<2000$).
+(Alternatively a simple EM algorithm can also work well, particularly for modest $J$, say $J<2000$;
+see \url{http:://stephenslab.github.io/ash/analysis/IPvsEM.html}.) 
 And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
@@
 We compare results of \ashr with existing FDR-based methods implemented
-in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \href{https://cran.r-project.org/src/contrib/Archive/mixfdr/}{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).  
+in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
 In all our simulations we assume that the test statistics follow the expected theoretical distribution under the null, and we indicate this
@@
 This qualitative difference among methods in the inferred alternative distribution of the $z$ scores is quite general, and 
-not limited to this simple simulation. To understand why the alternative distribution of $z$ scores from $\locfdr$ and $\qvalue$ has a hole at zero, note that neither
+also occurs in simulations where most effects are null (e.g.~\url{http://stephenslab.github.io/ash/analysis/referee_uaza.html}).
+To understand why the alternative distribution of $z$ scores from $\locfdr$ and $\qvalue$ has a hole at zero, note that neither
 of these methods explicitly models the alternative distribution: instead they simply subtract a null distribution (of $z$ scores or $p$ values) from the observed empirical distribution, letting the alternative
@@
 existing methods effectively push the dark blue component until the light-blue component {\it disappears} at 0.
-See \href{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd}{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive demonstration.
+See \url{https://stephens999.shinyapps.io/unimodal/unimodal.Rmd} for an interactive demonstration.
 
@@
 The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than existing methods.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs. 
+Consequently \ashr will estimate smaller $\lfdr$s and FDRs, and so identify more significant discoveries at a given threshold. 
 This is desirable, provided that these estimates remain conservative: that is,
@@
  become substantially less troublesome if we use the local false sign rate $\lfsr$ (\ref{eqn:lfsr}), rather than $\lfdr$, to measure significance. 
- This is essentially because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
+ This is because $\lfsr$ is less sensitive to the estimate of $\pi_0$.
 To illustrate, Figure \ref{fig:lfsr} compares the estimated $\lfsr$ from ash.n with the true value: although the estimated $\lfsr$ continue to
@@
 then the analysis with no point mass will tend to be anti-conservative, underestimating $\lfsr$ (Figure \ref{fig:lfsr-s}). 
-We conclude that {\it if} ensuring a ``conservative" analysis is important then one must allow for a point mass at 0.
+We conclude that {\it if} ensuring a ``conservative" analysis is important then one should allow for a point mass at 0.
 
@@
 to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\href{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}{http://stephenslab.github.io/ash/analysis/check\_mixfdr\_lownoise.html}). 
+(\url{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}). 
 Of course, we would not advocate the UA in settings where multi-modality is 
@@
 as adequate for practical applications. The strongest deviations from nominal rates are noted and discussed in the table captions.
+One general issue is that the methods based on mixtures of uniform distributions often slightly curtail the tail of $g$, causing the probability of very large outlying effects to be understated;  see also \url{http://stephenslab.github.io/ash/analysis/efron.fcr.html}.
 
@@
 Both EM and IP methods are implemented in the {\tt ashr} package; results shown here are from the IP method, but graphs from EM are essentially the same.
-See \href{http://stephenslab.github.io/ash/analysis/checkIP.html}{http://stephenslab.github.io/ash/analysis/checkIP.html}  and \href{http://stephenslab.github.io/ash/analysis/IPvsEM.html}{http://stephenslab.github.io/ash/analysis/IPvsEM.html}  for comparisons.
+See \url{http://stephenslab.github.io/ash/analysis/checkIP.html}  and \url{http://stephenslab.github.io/ash/analysis/IPvsEM.html}  for comparisons.
 
@@
 prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \href{http://github.com/stephens999/ashr}{http://github.com/stephens999/ashr} and includes contributions 
+git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
 from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.

### 2016-05-31 analysis/ash_npmle.Rmd
add example of npmle

---
title: "Using ash for NPMLE"
author: "Matthew Stephens"
date: 2016-05-31
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

## The NPMLE

Here we show how we can use ash to (approximately) compute the unconstrained
NPMLE - that is, estimate the underlying distribution $g$ by maximizing the likelihood without the unimodal constraint.
See Koenker and Mizera, JASA 2014, for background.

Within ash one can approximate the npmle as a mixture of uniforms
on a dense grid of non-overlapping values - this results in a piecewise constant density with changes in the density only at the grid points.

The following example comes from the REBayes vignette by Koenker and Gu.
The underlying $g$ is a mixture of a point mass (weight 0.8) at 0 and a point mass (weight 0.2) at 2.
```{r}
  set.seed(102)
  y <- c(rep(0,800), rnorm(200, 2)) + rnorm(1000)
  z <- REBayes::GLmix(y)
```

Now we fit the NPMLE using 'ashr' and compare with the REBayes solution.
```{r}
  grid = seq(from=min(z$x),to=max(z$x),length=1000)
  k=length(grid)
  y.ash.npmle = ashr::ash(y,1,g=ashr::unimix(pi=rep(1/(k-1),(k-1)),a=grid[-k],b=grid[-1]),method="shrink")

  plot(z$x,cumsum(z$y)/sum(z$y),col=2,main="Estimated cdf",type="l")
  lines(ashr::cdf.ash(y.ash.npmle,x=z$x),type="l")
  z$logLik
  y.ash.npmle$loglik
```

## Session information

```{r info}
sessionInfo()
```

### 2016-05-31 analysis/index.Rmd
add example of npmle

--- 2016-05-04 analysis/index.Rmd
@@
 *  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)
+*  [Comparison of methods, and example of Credible Interval behaviour when UA does not hold](efron.fcr.html)
+*  [Using ash to fit the NPMLE; comparison with REBayes](ash_npmle.html)
 

### 2016-06-06 analysis/make_GOODPOOR_figs.Rmd
update

@@
 First perform a simple simulated example to illustrate high and low signal.
-True values of beta simulated from $(0.5 N(0,1) + 0.5 \delta_0)$.
+True values of beta simulated from $(0.5 N(0,2^2) + 0.5 \delta_0)$.
 ```{r}
@@
 null_alt = rbinom(ntest,1,0.5)
-beta = rnorm(ntest)
+beta = rnorm(ntest,0,sd=1)
 beta = ifelse(null_alt==1,beta,0)
@@
 res.ash.good = ash(betahat[GOOD],sebetahat[GOOD],method="fdr")
-```
-
-```{r}
-par(mai=c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
-layout(matrix(1:3,ncol=3,byrow=TRUE))
-plot(res.qvalue.good$q,res.qvalue$q[GOOD],main="qvalue",xlim=c(0,1),ylim=c(0,1),axes=F)
-axis(side=2)
-axis(side=1)
-abline(a=0,b=1,col=2)
-plot(res.locfdr.good$fdr,res.locfdr$fdr[GOOD],main="locfdr",xlim=c(0,1),ylim=c(0,1),axes=F)
-axis(side=1)
-abline(a=0,b=1,col=2)
-plot(res.ash.good$lfsr,res.ash$lfsr[GOOD],main="ashr",xlim=c(0,1),ylim=c(0,1),axes=F)
-axis(side=1)
-abline(a=0,b=1,col=2)
 ```
@@
 
+```
+
+```{r  lfsr_vs_pval_GOODPOOR_single, dev='pdf', fig.width=4.5, fig.height=3}
+#now plot one at a time
+p=plot_pval_vs_lfsr(make_df_for_pval(res.ash,method="ash"))
+print(p+ theme(axis.text.x = element_text(size = 8,angle=45)))
+
+p=plot_pval_vs_lfsr(make_df_for_pval(res.ash.ET,method="ash (p-value prior)"))
+print(p+ theme(axis.text.x = element_text(size = 8,angle=45)))
+```
+
+Compare the log-likelihoods:
+```{r}
 res.ash.ET$logLR
 res.ash$logLR
+```
 
+Plot number of true positives against false positives
+```{r tp-fp, dev='pdf', fig.width=4.5, fig.height=3}
+  df0 = cbind(make_df_for_pval(res.ash,method="Default prior (alpha=0)"),beta=beta)
+  df1 = cbind(make_df_for_pval(res.ash.ET,method="p-value prior (alpha=1)"),beta=beta)
+  #false positives and true positives
+  fp = function(df){cumsum((df$beta==0)[order(df$lfsr,decreasing = FALSE)])}
+  tp = function(df){cumsum((df$beta!=0)[order(df$lfsr,decreasing = FALSE)])}
+  df.fptp = rbind(data.frame(method="ash lfsr", tp = tp(df0), fp = fp(df0)),
+             data.frame(method="p values", tp = tp(df1), fp = fp(df1)))
+
+  ggplot(df.fptp,aes(x=fp,y=tp,col=method)) + geom_line()  + xlab("False positives") + ylab("True positives") + xlim(c(0,100)) + ylim(c(0,400)) + scale_colour_manual(values=c("black","green")) 
 ```

### 2016-06-06 analysis/makefig_FDReg.Rmd
update

@@
 plotlabel("mixfdr")
-plotlabel("ashr")
+plotlabel("ash")
 
@@
 axis(side=1, labels=FALSE,tck=-0.01)
-altnullhist(pval,res.ash$lfdr,main="ashr",ncz=ncz,xaxt='n',cex.axis=0.8)
+altnullhist(pval,res.ash$lfdr,main="ash",ncz=ncz,xaxt='n',cex.axis=0.8)
 axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
@@
 axis(side=1, labels=FALSE,tck=-0.01)
-nullalthist(zscore,res.ash$lfdr,main="ashr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
+nullalthist(zscore,res.ash$lfdr,main="ash",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
 axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)

### 2016-06-06 analysis/plot_cdf_eg.Rmd
update

@@
 
-  ggplot(df.cdf,aes(x=x,y=y,color=method),...) + colScale + geom_line() + facet_grid(.~scenario) + theme(legend.position = "bottom")
+  ggplot(df.cdf,aes(x=x,y=y,color=method),...) + colScale + geom_line(lwd=1.5,alpha=0.7) + facet_grid(.~scenario) + theme(legend.position = "bottom")
 
@@
 ```{r egcdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
-#pdf("../paper/figures/egcdf.pdf",height=3,width=9)
 plot_mean_cdf(1)
@@
 ```{r mean_cdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
-#pdf("../paper/figures/mean_cdf.pdf",height=3,width=9)
 plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth"),pi0filter=TRUE)
@@
 ```{r mean_cdf_nopen, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
-#pdf("../paper/figures/mean_cdf_nopen.pdf",height=3,width=9)
-names(myColors) <- c("mixfdr.tnull","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr","truth")
+names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr","mixfdr.tnull")
 colScale <- scale_colour_manual(name = "method",values = myColors)
 plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n.s","ash.u.s","ash.hu.s","truth"),pi0filter=TRUE)
+source("../R/set_plot_colors.R") #reset color scale for other plots
+```
+
+The following plots have fewer methods and scenarios for more clarity
+
+```{r egcdf-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
+plot_mean_cdf(1,PLOTMETHODS=c("ash.n","truth"),PLOTSCENARIOS = c("spiky","near-normal","bimodal"))
+```
+
+```{r mean_cdf-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
+plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n","truth"),PLOTSCENARIOS = c("spiky","near-normal","bimodal"),pi0filter=TRUE)
+```
+
+```{r mean_cdf_nopen-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
+names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr","mixfdr.tnull")
+colScale <- scale_colour_manual(name = "method",values = myColors)
+plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n.s","truth"),PLOTSCENARIOS = c("spiky","near-normal","bimodal"), pi0filter=TRUE)
+source("../R/set_plot_colors.R")
+```
+
+I wanted to add the npmle to the plot...
+
+```{r egcdf-reduce-with-npmle, dev='pdf',fig.height=4.5, fig.width=6, crop=TRUE}
+run_nplme=function(SEED=1,PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal")){
+  df=data.frame(seed=NULL,scenario=NULL,method=NULL,x=NULL,y=NULL,pi0=NULL)
+  for(SCENARIO in PLOTSCENARIOS){
+    temp=dscr::load_example(dsc_shrink,SEED,SCENARIO,"ash.n","../output/dsc-shrink-files/")
+    z <- REBayes::GLmix(temp$input$betahat)
+    df = rbind(df,data.frame(seed=SEED,scenario=SCENARIO,method="NPMLE",x=z$x,y=cumsum(z$y)/sum(z$y),pi0=NA))
+  }
+  return(df)
+}
+
+plot_mean_cdf_with_npmle = function(SEED=1,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
+      PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),pi0filter=FALSE,...){
+  if(length(SEED)>1){stop("plot with npmle only implemented for a single seed")}
+  PLOTNAMES=PLOTSCENARIOS
+  #set up dataframe with cdf for all methods and all datasets
+  df= expand.grid(seed=SEED,scenario=PLOTSCENARIOS, method=PLOTMETHODS,stringsAsFactors = FALSE)
+  df.cdf=plyr::ddply(df,.(seed,scenario,method), get_cdf)
+
+  df.cdf$scenario=factor(df.cdf$scenario,levels=PLOTSCENARIOS)
+  levels(df.cdf$scenario)=PLOTNAMES
+
+  df.npmle = run_nplme(SEED, PLOTSCENARIOS)
+  df.cdf = rbind(df.cdf,df.npmle)
+
+  p=ggplot(df.cdf,aes(x=x,y=y,color=method),...) +  xlim(c(-4,4)) + colScale + geom_line(lwd=1.2,alpha=0.7)  + theme(legend.position = "bottom")
+  return(p)
+}
+
+plot_mean_cdf_with_npmle(1,PLOTMETHODS=c("truth","ash.hu"),PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")) + facet_wrap(~ scenario, nrow=2)
+
 ```

### 2016-06-06 analysis/referee_uaza.Rmd
update

@@
 
-The referee reports a simulation of 10,000 hypotheses, a fraction 0.8 are null ($\mu_i=0$), and a fraction 0.2 are alternative with effect $\mu_i \sim N(0,1)$. For each effect $i$ there
+A referee reported a simulation of 10,000 hypotheses, a fraction 0.8 are null ($\mu_i=0$), and a fraction 0.2 are alternative with effect $\mu_i \sim N(0,1)$. For each effect $i$ there
 are n=30 observations from $x_{i,j} \sim N(\mu_i,1)$ and a one-sample t test is used to compute a $p$ value. The referee correctly notes that these simulations i) obey the
-unimodal assumption, and ii) have the property that "most" $z$ scores near 0 are null.
+unimodal assumption (UA), and ii) have the property that "most" $z$ scores near 0 are null (Efron's "Zero Assumption"; ZA).
+The motivation is to demonstrate that the UA and the ZA are not contradictory.
 
-Here we repeat these simulations and apply qvalue to check how it behaves.
+I agree with this, if we define the ZA as "most" $z$ scores near 0 are null.
+However, in practice the existing methods effectively make a stronger assumption
+(or at least, behave as if they do): that *all* $z$ scores near 0 are null.
+Consequently they end up creating a hole in the distribution of alternative $z$ scores at 0. Here we illustrate this by applying qvalue to this scenario.
 
@@
 despite the fact that qvalue does not explicitly model the distribution of
-p values or z scores under the alternative, its way of estimating $\pi_0$
-necessarily creates a hole in the z score distribution at 0 under the alternative.
+$p$ values or $z$ scores under the alternative, its way of estimating $\pi_0$
+necessarily creates a hole in the $z$ score distribution at 0 under the alternative.
 

### 2016-06-06 paper/main.tex
update

@@
 \def\mixfdr{{\tt mixfdr}\xspace}
+\def\ash{{\tt ash}\xspace}
 \def\ashr{{\tt ashr}\xspace}
@@
 
-We introduce a novel Empirical Bayes approach for large-scale hypothesis testing, including  
-estimating False Discovery Rates (FDRs), and estimating effect sizes. Compared with
-existing approaches to FDR analysis, the method has two key differences. First, it
-assumes that the distribution of the actual (unobserved) effects being tested is unimodal, with a mode at 0.
-This ``unimodal assumption" (UA), although natural in many contexts, is not usually incorporated into standard FDR analysis,
+We introduce a new Empirical Bayes approach for large-scale hypothesis testing, including  
+estimating False Discovery Rates (FDRs), and effect sizes. This approach has two key differences
+from existing approaches to FDR analysis. First, it
+assumes that the distribution of the actual (unobserved) effects is unimodal, with a mode at 0.
+This ``unimodal assumption" (UA), although natural in many contexts, is not usually incorporated into 
+standard FDR analysis,
 and we demonstrate how incorporating it brings many benefits.  Specifically, the UA
@@
 rather than the one number usually used ($p$ value, or $z$ score). When available, using two numbers instead of one 
-helps account for variation in measurement precision across tests. It also facilitates the estimation of actual 
-effect sizes, and unlike standard FDR analysis methods
+helps account for variation in measurement precision across tests. It also facilitates estimation of  
+effects, and unlike standard FDR methods
 our approach provides interval estimates (credible regions) for each effect in addition to measures of significance.
@@
 datta2005empirical,muralidharan2010empirical} for example.
+
 So what is the ``New Deal" here? We introduce two simple ideas that are new (at least compared with
 existing widely-used FDR pipelines) and can substantially affect
-inference. The first idea is to {\it assume that the distribution of effects is unimodal}.  We provide a very simple, fast, and stable computer 
-implementation for perfoming EB inference under this assumption, and illustrate how it can improve inference of FDR when the unimodal assumption is correct.  
+inference. The first idea is to {\it assume that the distribution of effects is unimodal}.  We provide a simple, fast, and stable computer 
+implementation for perfoming EB inference under this assumption, and illustrate how it can improve inferences 
+when the unimodal assumption is correct.  
 The second idea is to use two numbers -- effect sizes, and their standard errors -- rather than just one -- $p$ values, or $z$ scores -- 
@@
 is useful more generally. For example, the methods described here can be applied directly to perform shrinkage estimation for
-wavelet denoising \cite{donoho:1995}, an idea explored in a companion paper \cite{xing:2016}. And analogous ideas can be used to perform
+wavelet denoising \cite{donoho:1995}, an idea explored in a companion paper \cite{xing2016smoothing}. And analogous ideas can be used to perform
 EB inference for variances \cite{lu2016variance}.
-Importantly, our work demonstrates how EB inference under a general unimodal assumption is, if anything, computationally simpler  
+Importantly, and perhaps surprisingly, our work demonstrates how EB inference under a general unimodal assumption is, if anything, computationally simpler  
 than commonly-used more restrictive assumptions -- such as a spike and slab or Laplace prior distribution \cite{johnstone2004needles} --  as well as being more flexible.
 
-Our methods are implemented in an R package, \ashr (for {\bf a}daptive {\bf sh}rinkage in {\bf R}), available at 
+We refer to our EB method as {\bf a}daptive {\bf sh}rinkage, or \ash, to emphasise its key points:
+using a unimodal prior naturally results in shrinkage estimation,
+and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation. We provide implementations in an R package, \ashr, available at 
 \url{http://github.com/stephens999/ashr}.
-(The name comes from the fact that our EB approach using a unimodal prior naturally results in shrinkage estimation,
-and the shrinkage is adaptive to both the amount of signal in the data and the measurement precision of each observation.) 
 Code and instructions for reproducing analyses and figures in this paper are at
@@
 %is, we argue, often realistic, and comes with considerable benefits, both statistical and computational. 
-%Our tools are implemented in an R package \ashr (short for ``Adaptive Shrinkage") and we 
+%Our tools are implemented in an R package \ash (short for ``Adaptive Shrinkage") and we 
 %demonstrate how the incorporation of these two ideas can lead to a substantial reduction in estimated FDRs (or, equivalently,
@@
 
-%Although we motivate our tools by the need to address the FDR problem, our \ashr package, as its name suggests,  provides a general tool for shrinkage estimation. 
+%Although we motivate our tools by the need to address the FDR problem, our \ash package, as its name suggests,  provides a general tool for shrinkage estimation. 
 %As an important special case, these methods address the "multiple comparisons" setting, where interest usually focuses on which $\beta_j$ can be confidently inferred to be non-zero.
@@
 
-We compare results of \ashr with existing FDR-based methods implemented
+We compare results of \ash with existing FDR-based methods implemented
 in the R packages \qvalue (v2.1.1), \locfdr (v1.1-8), and \mixfdr (v1.0, from \url{https://cran.r-project.org/src/contrib/Archive/mixfdr/}).
@@
 \center \includegraphics[height=6in]{../analysis/figure/makefig_FDReg.Rmd/decomp_ZA-1.pdf}
-\caption{Illustration that the unimodal assumption (UA) in \ashr can produce very different results from existing methods.
+\caption{Illustration that the unimodal assumption (UA) in \ash can produce very different results from existing methods.
 The figure shows, for a single simulated dataset, the way different methods decompose $p$ values (left) and $z$ scores (right) into a null component (dark blue) and an alternative component (cyan).
@@
 The three existing methods (\qvalue, \locfdr, \mixfdr) all produce a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ashr) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
-(In this case the \ashr decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
+In contrast the method introduced here (\ash) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
+(In this case the \ash decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
 \end{figure}
@@
 but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
-\qvalue, \locfdr, \mixfdr and \ashr to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
+\qvalue, \locfdr, \mixfdr and \ash to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
 Here we are using the fact that these methods all provide an estimate of the \lfdr{} for each observation, which
@@
 The results (Figure \ref{fig:ZA}) illustrate a clear difference between the existing methods and our method:
-all the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ashr, due to the UA, has a mode near 0.
+all the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ash, due to the UA, has a mode near 0.
 %Although \mixfdr does not formally make the \sZA, its decomposition
 %exhibits a similar hole. 
-%In contrast, due to the UA, the alternative $z$ score distribution for \ashr is required to have a mode at 0, 
+%In contrast, due to the UA, the alternative $z$ score distribution for \ash is required to have a mode at 0, 
 %effectively ``filling in" the hole.  
@@
 Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ashr's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
+Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ash's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
 This would increase the null component (dark blue) at the expense of
@@
 The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than existing methods.
-Consequently \ashr will estimate smaller $\lfdr$s and FDRs, and so identify more significant discoveries at a given threshold. 
+Consequently \ash will estimate smaller $\lfdr$s and FDRs, and so identify more significant discoveries at a given threshold. 
 This is desirable, provided that these estimates remain conservative: that is,
-provided that $\pi_0$ does not underestimate the true $\pi_0$ and $\lfdr$ does not underestimate the true $\lfdr$.
+provided that $\hat\pi_0$ does not underestimate the true $\pi_0$ and $\hat\lfdr$ does not underestimate the true $\lfdr$.
 The penalty term (\ref{eqn:penalty}) aims to ensure this conservative behavior. To check its effectiveness
@@
 making reliable estimation of $\pi_0$ easier. We also include one asymmetric distribution (``skew"), and one clearly bimodal distribution (``bimodal"),
-which, although we view as generally unrealistic, we include to assess robustness of \ashr~to deviations from the UA.
+which, although we view as generally unrealistic, we include to assess robustness of \ash~to deviations from the UA.
 
@@
 \caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ashr being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ashr estimates are slightly anti-conservative.} \label{fig:pi0sims}
+When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ash being least conservative, and hence most accurate.
+When the UA does not hold (``bimodal" scenario) the \ash estimates are slightly anti-conservative.} \label{fig:pi0sims}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfdr-1.png} 
-\caption{Comparison of true and estimated $\lfdr$ from \ashr (ash.n).
+\caption{Comparison of true and estimated $\lfdr$ from \ash (ash.n).
 Black line is $y=x$ and red line is $y=2x$.
@@
 
-Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ashr ($y$ axis) with the true values ($x$ axis). 
-For \ashr we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
+Figure \ref{fig:pi0sims} compares estimates of $\pi_0$ from \qvalue, \locfdr, \mixfdr and \ash ($y$ axis) with the true values ($x$ axis). 
+For \ash we show results for $g_1$ modelled as a mixture of normal components (``ash.n") 
 and as a mixture of symmetric uniform components (``ash.u"). (Results using the asymmetric uniforms, which we refer to
 as ``half-uniforms", and denote ``ash.hu" in subsequent sections, are here generally similar to ash.u and omitted to avoid over-cluttering figures.)
-The results show that \ashr provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
-in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ashr estimates can be slightly anti-conservative.
+The results show that \ash provides the smallest more accurate, estimates for $\pi_0$, while remaining conservative
+in all scenarios where the UA holds. When the UA does not hold (``bimodal" scenario) the \ash estimates can be slightly anti-conservative.
 We view this as a minor concern in practice, since we view such a strong bimodal scenario as unlikely in most applications where FDR methods
@@
 
-The results above show that \ashr can improve on existing methods in producing smaller, more accurate, 
+The results above show that \ash can improve on existing methods in producing smaller, more accurate, 
 estimates of $\pi_0$, which will lead to more accurate estimates of FDR.
-Nonetheless, in many scenarios \ashr continues
+Nonetheless, in many scenarios \ash continues
 to substantially over-estimate $\pi_0$ (see the ``spiky" scenario for example). 
@@
 An important advantage of our EB approach based on modelling the effects $\beta_j$, rather than $p$ values or $z$ scores, is that it
-can estimate the {\it size} of each effect $\beta_j$.
+can estimate the effects $\beta_j$.
 Specifically, it provides a posterior distribution for each $\beta_j$, which can be used
-to construct interval estimates for $\beta_j$ and address question such as ``which effects exceed $T$'', for any 
-threshold $T$.
+to construct interval estimates, etc.
 %Importantly, because the posterior distribution for $\beta_j$ depends on the ``prior" effect-size distribution $g$, which is estimated from {\it all} the observations,
@@
 generally satisfactory parametric formulation..." (\cite{efron2008microarrays} rejoinder, p46).
-Our key point here is that the UA greatly simplifies the deconvolution problem.
-While not meeting Efron's desire for an entirely general nonparametric approach, we believe 
-that the UA can handle many cases of practical interest.
-
-To illustrate this, Figure \ref{fig:egcdf} compares the
-estimated $g$ from \ashr  with that from \mixfdr which does not make the UA 
-(and which models $g$ as a mixture of $J$ normal distributions, with $J=3$ by default).
-The greater reliability of estimates afforded
-by the UA is immediately apparent. In particular the estimated cdf from \mixfdr often has an almost-vertical
-segment at some non-zero location, indicative of a concentration of density in the estimated $g$ at that location. 
-The UA prevents this kind of ``irregular" behavior, effectively requiring $g$ to be somewhat smooth.
-While the UA is not the only way to achieve this, we find it an attractive, simple and effective approach.
-
-Interestingly, even in the ``bimodal" scenario $\ashr$ 
-is visually more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
-Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
-Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
-to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
-(\url{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}). 
-Of course, we would not advocate the UA in settings where multi-modality is 
-clearly visible in the observed $\bhat$.
-
-We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ashr tends to systematically
-overestimate the mass of $g$ near zero. On careful inspection, this is apparent in Figure \ref{fig:egcdf}: the estimated cdf is generally below the true cdf just to the left of zero,
-and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf}b),
-and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
-under-estimating $\pi_0$ (Figure \ref{fig:egcdf}c).
+We argue here that the UA can greatly simplify the deconvolution problem, producing both computationally and statistically stable estimates of $g$.
+
+To illustrate, we compare the
+estimated $g$ from \ash (under the UA) with the non-parametric maximum likelihood estimate (NPMLE) for 
+$g$ (i.e.~estimated entirely non-parametrically without the unimodal constraint). 
+The NPMLE is straightforward to compute, thanks to 
+\cite{koenker2014convex}. Figure \ref{fig:egcdf} shows results for one dataset under each of six different scenarios.
+The estimated cdf from \ash is generally closer to the truth, even in the bi-modal scenario.
+(\ash tends to systematically overestimate the mass of $g$ near zero; this can be avoided 
+by removing the penalty term (\ref{eqn:penalty}); Figure \ref{fig:pen-nopen}.)
+Furthermore, the estimate from \ash is also substantially more ``regular" than the  NPMLE, 
+which has several almost-vertical segments indicative of a concentration of density in the estimated $g$ at those locations. 
+Indeed the NPMLE is a discrete distribution \cite{koenker2014convex}, so this kind of concentration will always occur. 
+The UA prevents this concentration, effectively regularizing the estimated $g$. 
+While the UA is not the only way to achieve this, we view it as attractive and widely applicable (although not universally applicable).
+
+%Interestingly, even in the ``bimodal" scenario $\ash$ 
+%is more accurate than \mixfdr: although \mixfdr is capable, in principle, of fitting the multiple modes of $g$, it does not do this well here. 
+%Possibly the noise level here is sufficiently large to make reliable estimation of the multiple modes difficult. 
+%Indeed, in multi-modal simulations where the multiple modes are sufficiently well-spaced
+%to be clearly visible in the observed $\bhat$, \mixfdr fits these modes 
+%(\url{http://stephenslab.github.io/ash/analysis/check_mixfdr_lownoise.html}). 
+%Of course, we would not advocate the UA in settings where multi-modality is 
+%clearly visible in the observed $\bhat$.
+%
+%We note one caveat on the accuracy of estimated $g$: due to the penalty term (\ref{eqn:penalty}) \ash tends to systematically
+%overestimate the mass of $g$ near zero. This can be seen in Figure \ref{fig:egcdf-reduce}: the estimated cdf is generally below the true cdf just to the left of zero,
+%and above the true cdf just to the right of zero. Averaging the cdf over many replicates confirms this systematic effect (Figure \ref{fig:egcdf-reduce}b),
+%and applying our methods without the penalty term removes this systematic effect, although at the cost of sometimes 
+%under-estimating $\pi_0$ (Figure \ref{fig:egcdf-reduce}c).
 
 \begin{figure}[h!] 
-\begin{subfigure}{\textwidth}
-	\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/egcdf-1.pdf}
-	\caption{Example estimated cdfs for single data sets compared with truth. The unimodal assumption made by the ash methods effectively regularizes estimates compared with \mixfdr.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf-1.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) so $\pi_0$ is systematically overestimated.}
-\end{subfigure}
-\begin{subfigure}{\textwidth}
-\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf_nopen-1.pdf}
-\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
-Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
-\end{subfigure}
-\caption{Comparisons of estimated cdfs of $g$ and true cdf of $g$. See Figure \ref{fig:pi0sims} for simulation scenarios.} \label{fig:egcdf}
+\begin{center}
+	\includegraphics[height=4.5in]{../analysis/figure/plot_cdf_eg.Rmd/egcdf-reduce-with-npmle-1.pdf}
+	\caption{Comparison of estimated cdfs from ashr and NPMLE.
+Different ash methods perform similarly, so only {\tt ash.hu} is shown for clarity.
+Each panel shows results for a single example data set, one for each scenario in Figure \ref{fig:altdens}.
+The results illustrate how the unimodal assumption made by ash regularizes the estimated cdfs compared with the NPMLE.} \label{fig:egcdf}
+\end{center}
 \end{figure}
@@
  produce $p$ values that are approximately uniform  (Figure \ref{fig:goodpoor}a), 
- which, when combined with the good-precision measurements, dilute the overall signal (e.g. they reduce the density of $p$ values near 0).
+ which, when combined with the good-precision measurements, dilute the overall signal 
+ (e.g.~they reduce the density of $p$ values near 0).
  This is reflected in the results of FDR methods like $\qvalue$ and $\locfdr$:
 the estimated error rates ($q$-values, or $\lfdr$ values) for the good-precision observations increase when the low-precision observations are included in the analysis
-(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ashr$ 
+(Figure \ref{fig:goodpoor}b). In contrast, the results from $\ash$ 
 for the good-precision observations are unaffected by including the low-precision observations in the analysis (Figure \ref{fig:goodpoor}b).
 
-\begin{figure}
+\begin{figure}[hp!]
 \begin{subfigure}{\textwidth}
@@
     \label{fig:goodpoor-hist}
-    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations}
+    \caption{Density histograms of $p$ values for good-precision, poor-precision, and combined observations.
+The combined data show less signal than the good-precision data, due to the contamination effect of the poor-precision measurements.}
  \end{subfigure}
@@
 \centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/GOODPOOReg_scatter-1.pdf}
-    \caption{Comparison of results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
-Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ashr) of a good-precision observation under the two different analyses.} \label{fig:goodpoor-scatter}
+    \caption{Results of different methods applied to good-precision observations only ($x$ axis) and combined data ($y$ axis).
+Each point shows the ``significance" ($q$ values from \qvalue; $\lfdr$ for \locfdr; $\lfsr$ for \ash) of a good-precision observation under the two different analyses.
+For existing methods including the poor-precision observations reduces significance of good-precision observations, whereas for \ash the poor-precision observations have little effect (because they have a very flat likelihood).} \label{fig:goodpoor-scatter}
 \end{subfigure}
-\caption{Simulation illustrating how, for existing FDR methods, 
-poor-precision observations can contaminate signal from good-precision observations. The top panel (a) illustrates that when 
-$p$ values from good-precision observations (left) and from poor-precision observations (center) are combined (right), they produce
-a distribution of $p$ values with less overall signal - and so, by conventional methods, will give a higher estimated FDR at any given threshold.
-The bottom panel (b) illustrates this behavior directly for the methods \qvalue and \locfdr: the $q$-values from \qvalue and the $\lfdr$ estimates from \locfdr are higher when applied to all data than when applied to good-precision observations only.
-In contrast the methods described here (\ashr) produce effectively the same results (here, the lfsr) in the good-precision and combined data analyses.} \label{fig:goodpoor}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{../analysis/figure/make_GOODPOOR_figs.Rmd/lfsr_vs_pval_GOODPOOR_single-2.pdf}
+    \caption{The relationship between $\lfsr$ and $p$-value is different for good-precision ($s=1$) and low-precision ($s=10$) measurements: \ash assigns the low-precision measurements a higher $\lfsr$, effectively downweighting them.
+%The default prior produces better performance here, as expected because the simulations match its assumptions.
+In practice one \ash can assess which prior best matches the data by comparing likelihoods (see text).
+    } \label{fig:lfsr_pval}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+\centering\includegraphics[width=3in]{../analysis/figure/make_GOODPOOR_figs.Rmd/tp-fp-1.pdf}
+    \caption{Trade-off between true positives ($y$) vs false positives ($x$) as the significance threshold ($\lfsr$ or $p$ value) is varied.
+By downweighting the low-precision observations \ash re-orders the significance of observations, producing more true positives at a given number of false positives.
+    %The default prior produces better performance here, as expected because the simulations match its assumptions.
+In practice one \ash can assess which prior best matches the data by comparing likelihoods (see text).
+    } \label{fig:tpfp}
+\end{subfigure}
+\caption{Simulations showing how, with existing methods, but not \ash, 
+poor-precision observations can contaminate signal from good-precision observations.} \label{fig:goodpoor}
 \end{figure}
 
- \subsubsection*{Reordering of significance, and the ``$p$ value prior"}
-
-Another consequence of accounting for differences in measurement precision is that \ashr may re-order the significance of the observations compared
- with the original $p$ values or $z$ scores. This is illustrated, using the same simulation as above, in Figure \ref{fig:lfsr_pval} (left panel).
- We see that poor precision measurements are assigned a higher $\lfsr$ than good precision measurements that have the same $p$ value.
+Another consequence of incorporating measurement precision into the likelihood
+is that \ash can re-order the significance of observations compared
+ with the original $p$ values or $z$ scores. Effectively \ash downweights the poor-precision 
+ observations  by assigning them a higher $\lfsr$ than 
+ good precision measurements that have the same $p$ value
+ (Figure \ref{fig:lfsr_pval}). 
  The intuition is that, due to their poor precision, these measurements contain very little information about the sign of the effects (or indeed any other aspect of the effects),
- and so the $\lfsr$ for these poor-precision measurements is always high.
-
-The potential for Bayesian analyses to re-order the significance of observations, and specifically to down-weight imprecise observations,
-was previously discussed in \cite{guan.stephens.08}.
-However, Wakefield \cite{wakefield:2009} showed that, under a certain prior assumption, the Bayesian analysis produces the same ranking of significance as $p$ values (or their $z$ scores).
-He named this prior the ``$p$-value prior" because it can be thought of as the
-implicit prior assumption that is being made if we rank the significance of observations by their $p$ value.
-Wakefield's $p$-value prior assumes that the less precise effect estimates correspond to larger true effects, and specifically that they scale proportional
-to the standard errors $s_j$. More specifically still, it assumes a normal prior for the non-zero $\beta_j$ with mean 0 and variance $K s^2_j$ for some constant $K$.
-Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. Specifically, if, instead of assuming that $\beta_j$ is independent
-of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
-then the $\lfsr$ computed by ash.n provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
-as is illustrated in Figure \ref{fig:lfsr_pval}, right panel. (This result does not hold when using the mixtures of uniforms prior, ash.u.)
-
-This $p$-value prior assumes that the $z$ scores $z_j=\beta_j/s_j$
-are identically distributed, independent of $s_j$. This is essentially the assumption made, implicitly or explicitly, by existing methods -- like \locfdr, \mixfdr and \qvalue --
-that model the $z_j$ or $p_j$ directly.  In contrast, we have assumed up to now that the $\beta_j$ are independent of $s_j$. 
-We can set both these assumptions within a more general framework, which allows that $\beta_j/s^\alpha$ is independent of $s_j$ for some $\alpha$ [Equation (\ref{eqn:beta-alpha}))].
-Setting $\alpha=0$ implies that $\beta_j$ is
-independent of $s_j$, as we have assumed up to now, and $\alpha>0$ implies that observations with larger standard error tend to have larger
-effects (in absolute value). This latter assumption may often be qualitatively plausible: 
-for example, in gene expression studies the standard error for gene $j$ depends partly on
-the variance of its expression among samples, and genes with a larger variance may tend to be less tightly regulated and
-so be amenable to a larger shift in expression between conditions (i.e.~larger effect $\beta_j$).
-On the other hand, there is no particular reason to expect that either $\alpha=1$ or $\alpha=0$ will be the optimal choice.
-Indeed, optimal choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$,
-which will be dataset-specific. 
-Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as 
-comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that 
-likelihood-based methods can be used to select $\alpha$. For example, following the logic of the EB approach it
-would be natural to select $\alpha$ by maximum likelihood. Since $\alpha$ is a one-dimensional parameter, this can be achieved by a 1-d grid search,
-which has been implemented in our software by C. Dai.
+ and so the $\lfsr$ for these poor-precision measurements is always high; see \cite{guan.stephens.08} for related discussion.
+Here this re-ordering results in improved performance: \ash identifies more true positive effects at a given level of false positives (Figure \ref{fig:tpfp}).
+
+ \subsubsection*{Dependence of $\beta_j$ on $s_j$}
+
+ Although downweighting low-precision observations may seem intuitive, we must now
+ confess that the issues are more subtle than our treatment above suggests. 
+ Specifically, it turns out that the downweighting behaviour depends on an assumption 
+that we have made up to now, that $\beta_j$ is independent
+of $s_j$ (Equation (\ref{eqn:beta})). In practice this assumption may not hold. 
+For example, in gene expression studies, 
+genes with higher biological variance may tend to exhibit larger effects $\beta_j$ (because they are
+less constrained). These genes will also tend to have larger $s_j$, 
+inducing a dependence between $\beta_j$ and $s_j$.
+
+Motivated by this, we generalize the prior (\ref{eqn:beta}) to
+ \begin{equation} \label{eqn:beta-alpha}
+ \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
+ \end{equation}
+where $\alpha$ is to be estimated or specified. 
+Setting $\alpha=0$ yields (\ref{eqn:beta}). Setting $\alpha>0$ implies that 
+the effects with larger standard error tend to be larger (in absolute value).
+Fitting this model for any $\alpha$ is straightforward using the same methods as for $\alpha=0$ (see Methods).
+%(Note that results above that were generated with all $\shat_j=1$ would not be affected by choice of $\alpha$.)
+
+The case $\alpha=1$ in (\ref{eqn:beta-alpha}) is of special interest because it corresponds most closely
+to existing methods. Specifically, it can be shown that with $\alpha=1$ the $\lfsr_j$ values from {\tt ash.n}
+are monotonic in the $p$ values: effects with smaller $p$ values have smaller $\lfsr_j$. 
+This result generalizes a result in \cite{wakefield:2009}, 
+who referred to a prior that produces the same
+ranking as the $p$ values as a ``$p$ value prior".
+Of course, if the $\lfsr_j$ are monotonic in the $p$ values then the downweighting and
+reordering of significance illustrated in Figure \ref{fig:goodpoor} will not occur.
+The intuition is that under $\alpha=1$ the poor precision observations have larger effects sizes,
+and consequently the same power 
+as the high-precision observations -- under these conditions the poor precision observations are not ``contaminating" the high precision observations, and so downweighting them is unnecessary.
+Thus running \ash with $\alpha=1$ will produce the same significance ranking as existing methods.
+Nonetheless, it is not equivalent to them, and indeed still has the benefits outlined previously: due to the UA \ash can produce less conservative estimates of $\pi_0$ and $\lfdr_j$; and because \ash models the $\beta_j$ it can produce interval estimates.
+
+As an aside, we note that with $\alpha=1$ the \ash estimates of both $g$ and $\lfsr_j$ depend on the pairs $(\bhat_j,\shat_j)$ only through  the $z$ scores $z_j = \bhat_j/\shat_j$ (though interval estimates for $\beta_j$ still depend on
+$(\bhat_j,\shat_j)$). This means that \ash can be run (with $\alpha=1$) in settings where the $z_j$ are available but 
+the $(\bhat_j,\shat_j)$ are not. It also opens the intriguing possibility 
+of running \ash on $p$ values obtained from any test (e.g.~a permutation test), by first converting each $p$ value to a corresponding $z$ score.
+However, the meaning of and motivation for the UA may be unclear in such settings, and caution seems warranted before proceeding along these lines.
+
+In practice, appropriate choice of $\alpha$ will depend on the actual relationship between $\beta_j$ and $s_j$, which will be dataset-specific.
+Further, although we have focussed on the special cases $\alpha=0$ and $\alpha=1$, there seems no strong reason to expect that either will necessarily be optimal in practice.
+Following the logic of the EB approach we suggest selecting $\alpha$ by maximum likelihood, which is 
+implemented in our software using a simple 1-d grid search (implementation due to C. Dai).
+
+%Framing the problem in this way -- that is, as comparing different modelling assumptions for $\beta_j$, rather than as comparing ``modelling $\beta_j$" vs  ``modelling $z_j$" (or ``modelling $p_j$") -- has the important advantage that
+%likelihood-based methods can be used to select $\alpha$. For example,
+
+% He named this prior the ``$p$-value prior" because it can be thought of as the 
+%implicit prior assumption that is made when using $p$ values to rank observations.
+%Wakefield's $p$-value prior assumes that the prior for $\beta_j$ depends on its standard error $s_j$,
+%and specifically that 
+%\begin{equation}
+%\beta_j | s_j \sim N(0,K s^2_j) \text{for some constant $K$.}
+%\end{equation}
+%Here we observe that Wakefield's result extends to our mixture of (zero-mean) normal priors. 
+%Specifically, if, instead of assuming that $\beta_j$ is independent
+%of $s_j$ as we have up to now, we assume that $z_j=\beta_j/s_j$ is independent of $s_j$, and drawn from a mixture of zero-mean normal distributions,
+%then the $\lfsr$ provides the {\it same ranking} of observations as the $z$ scores and $p$ values,
+%as is illustrated in Figure \ref{fig:lfsr_pval}, right panel.
 
@@
 
-%To summarize: the behaviour of \ashr depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
+%To summarize: the behaviour of \ash depends on the value of $\alpha$ assumed in (\ref{eqn:beta-alpha}). 
 %Up to now we have assumed $\alpha=0$, which implies that effects are identically distributed, independent of their standard error.
@@
 
-\begin{figure}
-\centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/lfsr_vs_pval_GOODPOOR-1.pdf}
-\caption{Figure illustrating affects of prior assumptions on re-ordering of significance.
-Left panel shows results under our ``default prior" which assumes that effects $\beta_j$ are identically distributed, independent of $s_j$.
-Right panel shows results under the ``$p$-value prior", which assumes that $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$.} \label{fig:lfsr_pval}
-\end{figure}
+%\begin{figure}
+%\begin{subfigure}{\textwidth}
+%\centering\centering\includegraphics[width=4in]{../analysis/figure/make_GOODPOOR_figs.Rmd/lfsr_vs_pval_GOODPOOR-1.pdf} \label{fig:lfsr_pval}
+%    \caption{Illustration of how prior assumptions re-order significance.
+Left panel shows results under the ``default prior", which assumes $\beta_j$ are identically distributed, independent of $s_j$ ($\alpha=0$ in (\ref{eqn:beta-alpha})).
+Right panel shows results under the ``$p$-value prior", which assumes $z$ scores $\beta_j/s_j$ are identically distributed, independent of $s_j$ (\ref{eqn:beta-alpha}).
+This right panel is exactly the same as would be obtained by any monotonic function of the $p$ values to measure significance (e.g.
+$p$ values, $q$ values, $|z|$ scores)}
+% \end{subfigure}
+%
+%\caption{} 
+%\end{figure}
 
@@
 
-Equation (\ref{eqn:beta}) assumes that the $\beta_j$ all come from the same distribution $g$, independent of $\shat_j$.
-This can be relaxed to allow the distribution of $\beta_j$ to depend on $\shat_j$ using
- \begin{equation} \label{eqn:beta-alpha}
- \frac{\beta_j}{\shat_j^\alpha} \, \big |  \shat_j \sim g(\cdot; \pi)
- \end{equation}
-for any $\alpha$. Setting $\alpha=0$ yields (\ref{eqn:beta}), and setting $\alpha=1$ corresponds to
-assuming that the $t_j =\beta_j/ \shat_j$ have a common distribution. This case is of special interest:
-it effectively corresponds to the ``$p$ value prior" in \cite{wakefield:2009} and is, implicitly, 
-the assumption made by existing FDR methods that rank tests by their $p$ values (or $z$ or $t$ scores). See Results for further discussion.
-
 The model (\ref{eqn:beta-alpha}) for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
@@
 
+% Do NOT remove this, even if you are not including acknowledgments
+\section*{Acknowledgements}
+
+I thank J.~Lafferty for pointing out the convexity of the likelihood function, which lead to an improved implementation with interior point methods. 
+Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text
+prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
+git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
+from Chaoxing (Rick) Dai, Mengyin Lu, Nan Xiao, and Tian Sen.
+
+This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
+
+%\section*{References}
+% The bibtex filename
+\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
+
+\clearpage
 \appendix
+
+\section{Appendix: Additional Figures and Tables}
+
+\begin{figure}[h!] 
+\begin{subfigure}{\textwidth}
+\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf-1.pdf}
+\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here use penalty (\ref{eqn:penalty}) which leads to systematic overestimation of $\pi_0$ in some scenarios.}
+\end{subfigure}
+\begin{subfigure}{\textwidth}
+\includegraphics[height=2in]{../analysis/figure/plot_cdf_eg.Rmd/mean_cdf_nopen-1.pdf}
+\caption{Average estimated cdfs across $\sim10$ data sets compared with truth; methods here do not use penalty (\ref{eqn:penalty}) so $\pi_0$ is not systematically overestimated.
+Systematic differences from the truth in ``skew" and ``bimodal" scenarios highlight the effects of model mis-specification.}
+\end{subfigure}
+\caption{Comparisons of average estimated cdfs of $g$ with and without penalty term.
+See Figure \ref{fig:pi0sims} for simulation scenarios.
+In most cases the three different ash methods are very similar and so the lines lie on top of one another.} \label{fig:pen-nopen}
+\end{figure}
 
@@
 \includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s_nn-1.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
+\caption{Comparison of true and estimated $\lfsr$ when data are simulated with no point mass at zero ($\pi_0=0$), and also analyzed by \ash with no point mass on 0 (and mixture of normal components for $g$).
 Black line is $y=x$ and red line is $y=2x$.
@@
 \includegraphics[width=\textwidth]{../analysis/figure/plot_lfsr.Rmd/plot_lfsr_s-1.png} 
-\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ashr with no point mass on 0 (and mixture of normal components for $g$).
+\caption{Comparison of true and estimated $\lfsr$ when data are simulated with point mass at zero (drawn uniformly from [0,1] in each simulation), but analyzed by \ash with no point mass on 0 (and mixture of normal components for $g$).
 Black line is $y=x$ and red line is $y=2x$.
@@
 %which is close to the definition in \cite{storey.03}.
-
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgements}
-
-I thank J.~Lafferty for pointing out the convexity of the likelihood function, which lead to an improved implementation with interior point methods. 
-Statistical analyses were conducted in the {\sf R} programming language \cite{Rcore:2012}, Figures produced using the {\tt ggplot2} package \cite{ggplot2}, and text
-prepared using \LaTeX. Development of the methods in this paper was greatly enhanced by the use of the {\tt knitr} package \cite{xie2013dynamic}  within the RStudio GUI, and 
-git and github. The {\tt ashr} R package is available from \url{http://github.com/stephens999/ashr} and includes contributions 
-from Chaoxing (Rick) Dai, Mengyin Lu, and Tian Sen.
-
-This work was supported by NIH grant HG02585 and a grant from the Gordon and Betty Moore Foundation.
-
-%\section*{References}
-% The bibtex filename
-\bibliography{/Users/stephens/Dropbox/Documents/mainbib}
-
 \section*{Figure Legends}

### 2016-06-06 R/set_plot_colors.R

@@ -1,6 +1,6 @@
 library("RColorBrewer")
 library("ggplot2")
-myColors <- brewer.pal(7,"Set1")
-names(myColors) <- c("mixfdr.tnull","ash.hu","ash.n","ash.u","qvalue","locfdr","truth")
+myColors <- brewer.pal(8,"Set1")
+names(myColors) <- c("truth","ash.hu","ash.n","ash.u","qvalue","locfdr","mixfdr.tnull","NPMLE")
 colScale <- scale_colour_manual(name = "method",values = myColors)
 

### 2016-06-08 paper/main.tex
update paper

@@
 \def\FDR{\text{FDR}}
-\def\fdr{\text{\lfdr}}
 \def\FSR{\text{FSR}}
@@
 $\beta_j$ might be the difference in the mean (log) expression of gene $j$ in two conditions.
-In contexts where FDR methods are applied, interest often focuses on identifying ``significant" non-zero effects;
-that is, testing the null hypotheses $H_j:\beta_j=0$.  Here we tackle both this problem, and the  
+We tackle both the problem of testing the null hypotheses $H_j:\beta_j=0$, and the  
 more general problem of estimating, and assessing uncertainty in, $\beta_j$.
@@
 Assume that the available data are estimates $\bhat=(\bhat_1,\dots,\bhat_J)$ of the effects,
-and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given the observed data $\bhat,\shat$,
-which by Bayes theorem can be written as 
+and corresponding (estimated) standard errors $\shat=(\shat_1,\dots,\shat_J)$.  Our goal is to compute a posterior distribution for $\beta$ given $\bhat,\shat$,
+which by Bayes theorem is
 \begin{equation}
@@
 where $\delta_0(\cdot)$ denotes a point mass on 0, and $N(\cdot; \mu, \sigma^2)$ denotes the density of the normal distribution with mean $\mu$ and variance $\sigma^2$.
-Here the mixture proportions $\pi=(\pi_0,\dots,\pi_K)$ are hyper-parameters, which are non-negative and sum to one, and are to be estimated,
-while the mixture component standard deviations $\sigma_1,\dots,\sigma_K$ 
-represent a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
+Here we take $\sigma_1,\dots,\sigma_K$ to be a large and dense grid of {\it fixed} positive numbers spanning a range from very small to very big (so $K$ is fixed and large). 
 We encourage the reader to think of this grid as becoming infinitely large and dense, as a non-parametric limit,
 although of course in practice we use a finite grid -- see Implementation Details.
+The mixture proportions $\pi=(\pi_0,\dots,\pi_K)$, which are non-negative and sum to one, are 
+hyper-parameters to be estimated.
 
@@
 \begin{enumerate}
-\item  The use of a mixture of zero-mean normals (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
-More flexibility can be obtained using mixtures of uniforms [see equation (\ref{eqn:g-general})]; indeed this 
-can allow $g$ to approximate {\it any} unimodal distribution. 
+\item  The form (\ref{eqn:mixnorm}) implies that $g$ is symmetric about 0.
+More flexibility can be obtained using mixtures of uniforms [Equation (\ref{eqn:g-general})]; indeed this 
+allows $g$ to approximate {\it any} unimodal distribution. 
 \item The model (\ref{eqn:beta}) assumes that the effects are identically distributed, independent of their standard errors $\shat$.
@@
 \end{enumerate}
-These embellishments, detailed in Detailed Methods, are all implemented in our software.
-Of course there remain limitations that are harder to relax, most notably the independence and conditional independence assumptions
+These embellishments, detailed in Detailed Methods, are implemented in our software.
+Other limitations are harder to relax, most notably the independence and conditional independence assumptions
 (which are also made by most existing approaches). Correlations among tests certainly arise in practice, either due to genuine correlations 
@@
 
-Together, (\ref{eqn:beta})-(\ref{eqn:normlik}) imply that $\bhat_j$ are independent observations from a mixture of normal distributions:
+Together, (\ref{eqn:beta})-(\ref{eqn:normlik}) imply that $\bhat_j$ are independent with
 \begin{equation} \label{eqn:marginallik}
@@
 
-The usual EB approach to fitting this model would involve two simple steps:
+The usual EB approach to fitting this model would involve two steps:
 \begin{enumerate}
@@
 \end{enumerate}
-Both steps 1 and 2 are very straightforward.
+Both steps are straightforward.
 Step 1 is a convex optimization problem, and can be solved quickly and reliably using interior point methods \cite{boyd2004convex,koenker2013convex}.
-(Alternatively a simple EM algorithm can also work well, particularly for modest $J$, say $J<2000$;
-see \url{http:://stephenslab.github.io/ash/analysis/IPvsEM.html}.) 
+(Alternatively a simple EM algorithm can also work well, particularly for modest $J$; see \url{http:://stephenslab.github.io/ash/analysis/IPvsEM.html}.)
 And the conditional distributions $p(\beta_j | \bhat_j, \shat_j,\hat{\pi})$ in Step 2 are analytically available, each a mixture of a point mass on zero and $K$ normal distributions.
-(The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$,
-which may seem more natural but is not straightforward when $\shat_j$ varies among $j$. This simple device may be useful in other applications.)
+The simplicity of step 1 is due to our use of a fixed grid for $\sigma_k$ in (\ref{eqn:mixnorm}), instead of estimating $\sigma_k$. This simple device may be useful in other applications.
 
@@
 In practice it is common, and desirable, to summarize these distributions to convey the ``significance" of each observation $j$.
-One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}, which is
-the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
-In other words it is the posterior probability that $\beta_j$ is actually zero:
+One natural measure of the significance of observation $j$ is its ``local FDR" \cite{efron2008microarrays}
 \begin{equation} \label{eqn:fdr}
-\fdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
-\end{equation}
-
-The $\lfdr$, like most other measures of significance, is rooted in the hypothesis testing paradigm which focuses on 
+\lfdr_j := \Pr(\beta_j =0  |  \bhat, \shat, \hat{\pi}).
+\end{equation}
+In words, $\lfdr_j$ is the probability, given the observed data, that effect $j$ would be a false discovery, if we were to declare it a discovery.
+
+The $\lfdr$, like most measures of significance, is rooted in the hypothesis testing paradigm which focuses on 
 whether or not an effect is exactly zero. This paradigm is popular, despite the fact that many statistical practitioners have argued that it is often inappropriate because
@@
 and $\beta_j \leq 0$ in (\ref{eqn:lfsr}) include the event $\beta_j=0$).
-In this sense, as a measure of ``significance", $\lfsr$ is more conservative than $\lfdr$. More importantly, as we illustrate in Results,
-$\lfsr$ is more robust to modeling assumptions.
+In this sense $\lfsr$ is a more conservative measure of significance than $\lfdr$. 
+More importantly, $\lfsr$ is more robust to modeling assumptions (see Results).
 
@@
 And, we can define the ``$s$-value" 
-\begin{equation} \label{eqn:qvalue}
+\begin{equation} \label{eqn:svalue}
 s_j := \widehat \FSR(\{k: \lfsr_k \leq \lfsr_j\}) 
@@
 as a measure of significance analogous to Storey's $q$-value \cite{storey.03}.
-(Replacing $\lfsr_j$ with $\lfdr_j$ in these definitions gives an estimate for the usual $\FDR(\Gamma)$, and an estimated $q$ value for each observation respectively.)
+(Replacing $\lfsr_j$ with $\lfdr_j$ in (\ref{eqn:FSRhat}) and (\ref{eqn:svalue}) gives estimates for the usual $\FDR(\Gamma)$, and the $q$-values respectively.)
 
@@
 
-Among previous methods that explicitly consider the FDR and related quantities,
-our work here seems most naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
+Among previous methods that explicitly consider FDR,
+our work seems naturally compared with the EB methods of \cite{efron2008microarrays} and \cite{muralidharan2010empirical}
  (implemented in the R packages \locfdr and \mixfdr respectively) and with the widely-used methods from  \cite{storey.03} (implemented in the R package \qvalue), which although not formally an EB approach, shares some elements in common.
@@
 
-There are two key differences between our approach and all of these three existing methods. First, 
-whereas these existing methods summarize the information on $\beta_j$ by a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- 
-we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayesian tests 
-for individual null hypotheses in a similar way; see also \cite{efron1993bayes}.
-Using two numbers instead of one clearly has the potential to be more informative, and this is illustrated in Results later (Figure \ref{fig:goodpoor}).
+There are two key differences between our approach and these existing methods. First, 
+whereas these existing methods take as input a single number -- either a $z$ score (\locfdr and \mixfdr), or a $p$ value (\qvalue) -- for each effect, 
+we instead work with two numbers ($\bhat_j,\shat_j$). Here we are building on \cite{wakefield:2009}, who develops Bayes Factors testing
+individual null hypotheses in a similar way; see also \cite{efron1993bayes}.
+Using two numbers instead of one clearly has the potential to be more informative, as illustrated in Results 
+(Figure \ref{fig:goodpoor}).
 
@@
 which are analogous to the ``effects" we estimate here. 
-Common choices of uni-modal distribution for regression coefficients 
+Common choices of unimodal distribution for regression coefficients 
 include the spike and slab, Laplace, $t$, normal-gamma, normal-inverse-gamma, or horseshoe priors \cite{carvalho2010horseshoe}.
-These are all less flexible than our approach, which provides for general uni-modal distributions, and 
+These are all less flexible than our approach, which provides for general unimodal distributions, and 
 it may be fruitful to apply our methods to the regression context; indeed see \cite{moser:2015} for work in this vein. 
-Additionally, the UA can be motivated by its effect on point estimates, which is to ``shrink" the estimates towards the mode - 
-such shrinkage is desirable from several standpoints for improving estimation accuracy. Indeed most model-based approaches to shrinkage
-make parametric assumptions that obey the UA (e.g. \cite{johnstone2004needles}).
+Additionally, the UA can be motivated by its effect on point estimates, which is to ``shrink" the estimates towards the mode - such shrinkage is desirable from several standpoints for improving estimation accuracy.
+Indeed most model-based approaches to shrinkage
+make parametric assumptions that obey the UA (e.g.~\cite{johnstone2004needles}).
 
@@
 indeed, this paper includes all the elements of our approach outlined above, except for the point mass on 0 and corresponding penalty term.
-This said, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
+However, the focus is very different: \cite{cordy1997deconvolution} focuses entirely on estimating $g$, whereas our primary focus is on estimating $\beta_j$.
 Also, they provide no software implementation.
 
-More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
+More generally, the related literature is too large to review comprehensively, but relevant key-words include ``empirical Bayes", ``shrinkage", ``deconvolution", ``semi-parametric", ``shape-constrained", and ``heteroskedastic".
 Some pointers to recent papers in which other relevant
-citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex}.
-Much of the literature focusses on the homoskedastic case (i.e.
-$\shat_j$ all equal) whereas we allow for heteroskedasticity.
+citations can be found include \cite{xie2012sure, sarkar:2014, koenker2014convex, efron2016empirical}. 
+Much of the literature focusses on the homoskedastic case (i.e.~$\shat_j$ all equal) whereas we allow for heteroskedasticity.
 And much of the recent shrinkage-oriented literature focuses only on point estimation of $\beta_j$, whereas for FDR-related applications measures of uncertainty are essential.
@@
 estimate (NPMLE) for $g$.
-These methods may provide alternatives to our approach in settings where the UA assumption is considered too restrictive.
+These methods may be useful in settings where the UA assumption is considered too restrictive.
 However, the  NPMLE for $g$ is a discrete distribution, which will induce a discrete posterior distribution on $\beta_j$, and so although the NPMLE
-may perform well for point estimation, it may not adequately reflect uncertainty in $\beta_j$, and some regularization on $g$ may be necessary.
+may perform well for point estimation, it may not adequately reflect uncertainty in $\beta_j$. To address this 
+some regularization on $g$ (e.g.~as in \cite{efron2016empirical}) may be necessary.
 Indeed, one way of thinking about the UA is as a way to regularize $g$.
@@
 To isolate these effects we consider the simplest case, where every observation has the same
-standard error, $s_j=1$ and all methods are provided that information. That is,
+standard error, $s_j=1$. That is,
 $\bhat_j | \beta_j \sim N(\beta_j,1)$ and $\shat_j=s_j=1$.
-In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$, and so the primary difference among methods lies in how they estimate $g$.
+In this case the $z$ scores $z_j:=\bhat_j/\shat_j=\bhat_j$, so modelling the $z$ scores is the same as modelling the $\bhat_j$.
+Thus the primary difference among methods in this setting is that \ash makes the UA and other methods do not.
 
@@
 \item The UA can yield conservative estimates of the proportion of true nulls, $\pi_0$, and hence conservative estimates of $\lfdr$ and $\FDR$.
-\item The UA results in a stable procedure, both numerically and statistically, and is somewhat robust to deviations from unimodality.
+\item The UA yields a procedure that is numerically and statistically stable, and is somewhat robust to deviations from unimodality.
 \end{enumerate}
@@
 The three existing methods (\qvalue, \locfdr, \mixfdr) all produce a ``hole" in the alternative $z$ score distribution around 0.
-In contrast the method introduced here (\ash) makes the Unimodal Assumption -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
+In contrast \ash makes the UA -- that the effect sizes, and thus the $z$ scores, have a unimodal distribution about 0 -- which yields a very different decomposition.
 (In this case the \ash decomposition is closer to the truth: the data were simulated under a model where all of the effects are non-zero, so the ``true" decomposition would make everything cyan.)} \label{fig:ZA}
@@
 
-We illustrate the effects of the UA, and the difference with existing methods, using a simple simulated dataset
- with the true effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Note that none of the effects are truly null,
-but nonetheless there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used each of the methods
+We illustrate the effects of the UA with a simple simulation, 
+ with effects $\beta_j \sim N(0,1)$ (so with $s_j=1$, $\bhat_j  \sim N(0,2)$). Though no effects are null, there are many $p$ values near 1 and $z$ scores near 0 (Figure \ref{fig:ZA}).  We used
 \qvalue, \locfdr, \mixfdr and \ash to decompose the $z$ scores ($z_j = \bhat_j$), or their corresponding $p$ values, into null and alternative components.
-Here we are using the fact that these methods all provide an estimate of the \lfdr{} for each observation, which
-implies such a decomposition;  specifically the average \lfdr{} within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component.
-The results (Figure \ref{fig:ZA}) illustrate a clear difference between the existing methods and our method:
-all the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ash, due to the UA, has a mode near 0.
+Here we are using the fact that these methods all provide an estimated $\lfdr_j$ for each observation $j$, which
+implies such a decomposition;  specifically the average $\lfdr_j$ within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component.
+The results (Figure \ref{fig:ZA}) illustrate a clear difference between \ash and existing methods:
+the existing methods have a ``hole"  in the alternative $z$ score distribution near 0, whereas \ash, due to the UA, has a mode near 0.
 %Although \mixfdr does not formally make the \sZA, its decomposition
@@
 %effectively ``filling in" the hole.  
-(Of course the null distribution also has a peak at 0, and the local fdr 
+(Of course the null distribution also has a peak at 0, and the \lfdr{} 
  under the UA is still smallest for $z$ scores that are far from zero -- i.e. large $z$ scores remain the ``most significant".)
 
-This qualitative difference among methods in the inferred alternative distribution of the $z$ scores is quite general, and 
+This qualitative difference among methods is quite general, and 
 also occurs in simulations where most effects are null (e.g.~\url{http://stephenslab.github.io/ash/analysis/referee_uaza.html}).
@@
 
-Figure \ref{fig:ZA} may also be helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
-Specifically, consider the panel of Figure \ref{fig:ZA} that shows \ash's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
+Figure \ref{fig:ZA} is also helpful in understanding the interacting role of the UA and the penalty term (\ref{eqn:penalty}) that attempts to make $\pi_0$ as ``large as possible" while remaining consistent with the UA.
+Specifically, consider the panel that shows \ash's decomposition of $z$ scores, and imagine increasing $\pi_0$ further.
 This would increase the null component (dark blue) at the expense of
@@
 
-The illustration in Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than existing methods.
+Figure \ref{fig:ZA} suggests that the UA will produce smaller estimates of $\pi_0$ than existing methods.
 Consequently \ash will estimate smaller $\lfdr$s and FDRs, and so identify more significant discoveries at a given threshold. 
@@
 \caption{Comparison of true and estimated values of $\pi_0$.
-When the UA holds all methods yield conservative (over-)estimates for $\pi_0$, with \ash being least conservative, and hence most accurate.
-When the UA does not hold (``bimodal" scenario) the \ash estimates are slightly anti-conservative.} \label{fig:pi0sims}
+When the UA holds all methods typically yield conservative (over-)estimates for $\pi_0$, with \ash being least conservative, and hence most accurate.
+\qvalue is sometimes anti-conservative when $\pi_0 \approx 1$.
+When the UA does not hold (``bimodal" scenario) the \ash estimates are slightly anti-conservative.
+} \label{fig:pi0sims}
 \end{subfigure}
@@
 
-%\subsubsection{The UA produces less conservative estimates of $\pi_0$}
-%
-%The data in Figure \ref{fig:ZA} were simulated assuming that the true effects are standard normal, $\beta_j \sim N(0,1)$,
-%and so all effects are non-zero and no nulls hold. This scenario is rather different
-%from the ``mostly null" scenarios that originally motivated FDR methods, where most tests are null,
-%and relatively few are non-null. It could be argued that the existing FDR methods are not well adapted to the 
-%``mostly non-null"
-%scenario, and we would agree! On the other hand, the distribution of $p$ values
-%in Figure \ref{fig:ZA} looks like data that, before this work, we would have been happy to apply
-%existing FDR methods to. Further, we believe that both ``mostly non-null" and ``mostly null" scenarios likely occur in practical applications.
-%Thus it seems desirable to have methods that perform well in both cases, and this is
-%an important feature of the methods we introduce here.
-%
-%TODO: put FDRs on graph? Simulate with 50\% null?
-
 \subsubsection*{The $\lfsr$ is more robust than \lfdr}
@@
 $\pi_0$ can be grossly overestimated.  
-(Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.)
+Of course other methods will be similarly affected by this: those that more grossly overestimate $\pi_0$, will more grossly overestimate $\lfdr$ and FDR/$q$-values.
 
@@
 
-Note that, compared with previous debates regarding testing, 
+Compared with previous debates, 
 this section advances an additional reason for focussing on the sign of the effect, rather than just testing whether it is 0. 
@@
 nothing about the sign of the effect.
-Thus, there are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g. it applies
+Thus, there are two reasons to use the $\lfsr$ instead of the $\lfdr$: it answers a question that is more generally meaningful (e.g.~it applies
 whether or not zero effects truly exist),  and estimation of $\lfsr$ is more robust.
@@
 $g$ (i.e.~estimated entirely non-parametrically without the unimodal constraint). 
-The NPMLE is straightforward to compute, thanks to 
-\cite{koenker2014convex}. Figure \ref{fig:egcdf} shows results for one dataset under each of six different scenarios.
+The NPMLE is straightforward to compute in R using the {\tt REBayes::GLmix} function 
+\cite{koenker2014convex}. Figure \ref{fig:egcdf} shows results under six different scenarios.
 The estimated cdf from \ash is generally closer to the truth, even in the bi-modal scenario.
@@
 The UA prevents this concentration, effectively regularizing the estimated $g$. 
-While the UA is not the only way to achieve this, we view it as attractive and widely applicable (although not universally applicable).
+While the UA is not the only way to achieve this (e.g.~\cite{efron2016empirical}), we view it as attractive and widely applicable.
 
@@
 	\includegraphics[height=4.5in]{../analysis/figure/plot_cdf_eg.Rmd/egcdf-reduce-with-npmle-1.pdf}
-	\caption{Comparison of estimated cdfs from ashr and NPMLE.
+	\caption{Comparison of estimated cdfs from \ash and the NPMLE.
 Different ash methods perform similarly, so only {\tt ash.hu} is shown for clarity.
@@
 The model (\ref{eqn:beta-alpha}) for general $\alpha$ can be fitted using the algorithm for $\alpha=0$. 
-To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b} := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
+To see this, define $b_j:= \beta_j/ \shat_j^\alpha$, and $\hat{b}_j := \bhat_j/\shat_j^\alpha$. Then $\hat{b}_j$ is an estimate of $b_j$ with
  standard error $\shat'_j:=\shat_j^{1-\alpha}$. 
@@
 \end{table}
-%In simple settings (e.g. symmetric effects, with the same standard error) $\fdr_j$ is a monotonic function of the $p$ value $p_j$, and so
+%In simple settings (e.g. symmetric effects, with the same standard error) $\lfdr_j$ is a monotonic function of the $p$ value $p_j$, and so
 %\begin{equation}
@@
 %which is close to the definition in \cite{storey.03}.
-\section*{Figure Legends}
-
-\clearpage
-
-\section*{Tables}
-
-\section*{Supporting Information Legends}
-
-Supplementary material can be found in {\bf Supplementary Information S1.}
+%\section*{Figure Legends}
+%
+%
+%\clearpage
+%
+%\section*{Tables}
+%
+%\section*{Supporting Information Legends}
+%
+%Supplementary material can be found in {\bf Supplementary Information S1.}
 %\begin{table}[!ht]

### 2016-06-08 analysis/referee.response.Rmd
add referee response

---
title: "Response to referees"
author: "Matthew Stephens"
date: 2016-03-21
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

# Referee 1

I agree with a lot of what the referee says, although there are some points of disagreement. The most fundamental disagreement regards Figure 1. This Figure shows
not the detectable effects (as the referee believed) but the decomposition
of *all* effects into null and alternative components. Since this seems
to be an important misunderstanding I have revised the text to try to make it clearer how this figure was derived. Specifically the revised text says:

“We used each of the methods qvalue, locfdr, mixfdr and ash to decompose the $z$ scores ($z_j = \hat{\beta}_j$), or their corresponding p values, into null and alternative components. Here we are using the fact that these methods all provide an estimate of the lfdr for each observation, which implies such a decomposition; specifically the average lfdr within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component.”

Since the referee objected to the presentation of the ZA and UA as opposing assumptions I have removed all explicit reference to the ZA. (This also 
solves the issue that I had defined the ZA differently than Efron’s original definition — see below — which may have created confusion.) 
Instead, as the referee 
suggested, I have presented the UA as an additional assumption “not made by other methods”, and - as the editor suggested —
I have focussed on contrasting the behavior of ash with the behavior 
of the other methods (which, as Figure 1 demonstrates, is to create a hole at 0 in the distribution of the alternative Z scores) rather than contrasting their assumptions.

However, it seems unsatisfactory to merely point out this behavioral
difference between methods without giving any explanation for the reason behind it.
So I have provided my best attempt to give the reason. Essentially the reason is that the existing methods (locfdr and qvalue), when estimating $\pi_0$, make the assumption that  *all* p values near 1 are null, or *all* z scores near 0 are null. 
(This was how I defined the ZA in the original paper; however this is stronger than Efron’s original statement of the ZA, which the referee - understandably - uses in his/her report, that ``most” z scores near 0 are null. So I don’t refer to this
as the ZA any more.) This assumption necessarily creates a hole in the
implied distribution of z scores under the alternative, and explains the behavior.

To underscore that this behavior is general, and not just a feature of the particular
simulation I did in Figure 1, I provide results using qvalue under the setting
suggested by the referee at (http://stephenslab.github.io/ash/analysis/referee_uaza.html).

## Minor Concerns

- I added the citation, thank you.

- I took up the suggestion to define FSR-hat and the s-value, thank you.

- I agree that returning the length J results as a dataframe would have been better, but
unfortunately changing this would break other packages that depend on ashr. So instead I have provided a method `as.data.frame` to extract the most important components of the ash object into a dataframe.

- I fixed the typo, thank you.

# Referee 2

## Major Concern

The referee says: "it would be interesting to
include distributions that favor other models, such as a bimodal distribution with no mass at zero, an assymmetrical distribution,. . . The unimodality is a strong assumption so it would be good to see how anti-conservative the results would get under different alternative distributions."

I found this puzzling since both an assymetrical distribution and a strongly bimodal distribution are already included in the simulation results (Figure 2, Figure 3, Table 1). I believe these existing results address the referee comment. I also want to emphasise that the unimodal assumption is actually less strong than the assumptions usually made in variable selection for regression, which is directly analogous to the fdr context. For example, it is common to assume that effects are a mixture of a point mass at zero and a single normal, which is considerably more restrictive than the general unimodal assumption.

However, to help reassure the referee, I have added another example to illustrate an assymetric situation where the unimodal-at-zero assumption is badly contradicted [here](http://stephenslab.github.io/ash/analysis/efron.fcr.html). As
the referee will see, the resulting estimates of the lfdr from ashr are
reasonably close to the true values (and also to those from locfdr).

## Minor Concern

- Is there a way to model the distribution as a mixture of both uniforms and normals? If not: why?

This is straightforward in principle. However, it would require 
non-trivial code modifications because of the way I implemented the mixture class (assuming that all components  come from the same family.) 
Another way to achieve an assymetric distribution without
the "hard" tails of the uniform might be to use a mixture of truncated normals. I have a student investigating this option. However, the uniform mixture has the advantage of allowing not only the assymetry, but also the $t$ likelihood. Although the paper
does not focus on this, I believe this is likely to be an important issue in practice.
So I suspect that the mixture of uniforms is often going to be the method of choice in practice.

- How would the model perform with moderate correlations?

There seems to be no reason to think that
moderate correlations among tests will have a substantial detrimental effect. 
But I do not want
to focus on this because I think it will give the reader the wrong message:
I think that users of both ashr and other fdr methods
should very much worry about correlations in practice (which I believe will
often be large, and not moderate).

- In equation 3, the null component (I think) is represented as delta0, but delta0 hasn’t been defined. What is assumed for the distribution of delta0?

Thank you for noticing this. I now define $\delta_0$ (which represents a point mass on 0).

- Figure 3: the truth is hard to distinguish from the different methods. And in general this figure is hard to read.

I agree. I have completely rethought this figure, which now has fewer, larger,
panels, and compares just two  methods (ash.hu and the NPMLE) against the truth on a few examples. The less important point - the effect of the penalty term on pi0 -is now illustrated in an additional figure included in the appendix.

- I particularly like the idea that the ordering of the tests differs from the classical p-values. Which influence does this have on the balance between sensitivity and specificity (for example with ROC curves)? It should change, but does it get better?

Thanks for the question. This is now addressed in the new Figure 4d.

## Typos and grammar

- page 14, line 20. The first mentioned pi0 and lfdr should have a hat.

Thanks - fixed.

## Session information

```{r info}
sessionInfo()
```

### 2016-06-08 analysis/index.Rmd
re-run analysis, finalize edits

--- 2016-05-09 analysis/index.Rmd
@@
 
-*  [Illustrate Zero Assumption](makefig_FDReg.html)
+*  [Figures illustrating FDR](makefig_FDReg.html)
 *  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
@@
 *  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)
-*  [Example of Credible Interval behaviour when UA does not hold](efron.fcr.html)
+*  [Comparison of methods, and example of Credible Interval behaviour when UA does not hold](efron.fcr.html)
+*  [Using ash to fit the NPMLE; comparison with REBayes](ash_npmle.html)
 

### 2016-06-08 analysis/summarize_dsc_znull.Rmd
re-run analysis, finalize edits

@@
 ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
-  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
+  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
 ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
-  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
+  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
 ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.n",method)), aes(sample = logLR)) + facet_grid(scenario~method) +
-  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
+  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
 

### 2016-07-03 analysis/CompareEEvsETmodel.Rmd
update EE vs ET

@@
 simply specify model="ET". (Internally, this causes ash to 
-replaces betahat with the standardized betahat, $\hat\beta_j/s_j$, and the standard errors
-for these standarized betahat values with 1.) It is also straightforward to compare
+replace betahat with the standardized betahat, $\hat\beta_j/s_j$, and the standard errors for these standarized betahat values with 1; there there is some bookkeeping to be done to make sure we return the right likelihoods and posteriors for the original beta, and not for these standardized values...
+ash takes care of this.) It is also straightforward to compare
 the two competing modelling assumptions (EE vs ET) by computing the log likelihood ratio, $log[p_{EE}(\hat\beta | s, \hat g_{EE})/p_{ET}(\hat\beta | s, \hat g_{ET})]$.
@@
 And here is how we can perform the ET analysis,
-by replacing betahat with the standardized betahat, and replacing the standard error with 1:
 ```{r}
@@
 ```
-Note that the posterior distributions computed by ash will now be for the standardized
-effects, and not for the effects $\beta_j$. However, the local false sign rate (lfsr) and
-related quantities are unaffected by this because the standardized effect is positive if and only if the effect is positive.
 

### 2016-07-24 analysis/CompareEEvsETmodel.Rmd
update from model parameter to alpha parameter

@@
 
-A key modelling assumption is that the effects $\beta_j$ are identically distributed
+The default model is that the effects $\beta_j$ are identically distributed
 from a unimodal distribution $g$. 
@@
 
-Here we consider an alternative modelling assumption:
-$$\beta_j/s_j | s_j \sim g(\cdot)   \quad (**).$$
-Under this alternative assumption
+Here we consider the more general model:
+$$\beta_j/(s_j)^\alpha | s_j \sim g(\cdot)   \quad (**).$$
+The case $\alpha=0$ is the model (*) above. The case $\alpha=1$ assumes that 
 the *t statistics* $\betahat_j/s_j$ are identically distributed from a unimodal
@@
 It is straightforward to use the ashr package to perform analysis under the ET model:
-simply specify model="ET". (Internally, this causes ash to 
+simply specify alpha=1. (Internally, this causes ash to 
 replace betahat with the standardized betahat, $\hat\beta_j/s_j$, and the standard errors for these standarized betahat values with 1; there there is some bookkeeping to be done to make sure we return the right likelihoods and posteriors for the original beta, and not for these standardized values...
@@
   library(ashr)
-  ashEE.res = ash(betahat, betahat.se,method="fdr",model="EE")
+  ashEE.res = ash(betahat, betahat.se,method="fdr",alpha=0)
 ```
@@
 ```{r}
-  ashET.res = ash(betahat, betahat.se,method="fdr", model="ET")
+  ashET.res = ash(betahat, betahat.se,method="fdr", alpha=1)
 ```
@@
   pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
-  ashEE.res = ash(betahat, betahat.se,method="fdr",model="EE")
-  ashET.res = ash(betahat, betahat.se, method="fdr",model="ET")
+  ashEE.res = ash(betahat, betahat.se,method="fdr",alpha=0)
+  ashET.res = ash(betahat, betahat.se, method="fdr",alpha=1)
   ashEE.res$loglik-ashET.res$loglik

### 2016-09-28 analysis/adaptive_shrinkage.Rmd
update

@@
   df = data.frame(betahat = c(sim1$input$betahat,sim2$input$betahat),
-                beta_est = c(ash1$PosteriorMean, ash2$PosteriorMean),
-                lfsr = c(ash1$lfsr, ash2$lfsr),
+                beta_est = c(get_pm(ash1), get_pm(ash2)),
+                lfsr = c(get_lfsr(ash1), get_lfsr(ash2)),
                 s=c(sim1$input$sebetahat,sim2$input$sebetahat),
@@
 ```{r run-ash-ET}
-beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,model="ET",mixcompdist="normal")
-beta.spiky.ash.ET = ash(sim.spiky$input$betahat, s,model="ET",mixcompdist="normal")
+beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,alpha=1,mixcompdist="normal")
+beta.spiky.ash.ET = ash(sim.spiky$input$betahat, s,alpha=1,mixcompdist="normal")
 df.ET = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ET,beta.bignormal.ash.ET)

### 2016-09-28 analysis/efron.fcr.Rmd
update example
update

@@
   par(mfcol=c(2,1))
-  nullalthist(z,lfdr=res.ash.true$lfdr,main="true partition")
-  nullalthist(z,lfdr=res.ash$lfdr,main="inferred partition")
+  nullalthist(z,lfdr=ashr::get_lfdr(res.ash.true),main="true partition")
+  nullalthist(z,lfdr=ashr::get_lfdr(res.ash),main="inferred partition")
 ```
@@
   par(mfcol=c(1,1))
-  plot(res.ash.true$PosteriorMean,res.ash$PosteriorMean,xlab="Truth", ylab="ash.hu",main="Posterior Mean (inferred vs truth)",xlim=c(-6,1),ylim=c(-6,1))
+  plot(ashr::get_pm(res.ash.true),ashr::get_pm(res.ash),xlab="Truth", ylab="ash.hu",main="Posterior Mean (inferred vs truth)",xlim=c(-6,1),ylim=c(-6,1))
   abline(a=0,b=1,col=2,lwd=3)
 
-    plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth", ylab="ash.hu",main="lfdr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
+    plot(ashr::get_lfdr(res.ash.true),ashr::get_lfdr(res.ash),xlab="Truth", ylab="ash.hu",main="lfdr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
   abline(a=0,b=1,col=2,lwd=3)
 
-  plot(res.ash.true$lfsr,res.ash$lfsr,xlab="Truth", ylab="ash.hu",main="lfsr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
+  plot(ashr::get_lfsr(res.ash.true),ashr::get_lfsr(res.ash),xlab="Truth", ylab="ash.hu",main="lfsr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
   abline(a=0,b=1,col=2,lwd=3)
@@
 ```
+
+# Comparison with qvalue and locfdr
 
@@
 par(mfrow=c(1,3))
-plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth (lfdr)", ylab="ash.hu",main="ash.hu",xlim=c(0,1),ylim=c(0,1))
+plot(ashr::get_lfdr(res.ash.true),ashr::get_lfdr(res.ash),xlab="Truth (lfdr)", ylab="ash.hu",main="ash.hu",xlim=c(0,1),ylim=c(0,1))
 abline(a=0,b=1,col=2,lwd=3)
-plot(res.ash.true$lfdr,res.locfdr$fdr,xlab="Truth (lfdr)", ylab="Estimate",main="locfdr",xlim=c(0,1),ylim=c(0,1))
+plot(ashr::get_lfdr(res.ash.true),res.locfdr$fdr,xlab="Truth (lfdr)", ylab="Estimate",main="locfdr",xlim=c(0,1),ylim=c(0,1))
 abline(a=0,b=1,col=2,lwd=3)
-plot(res.ash.true$lfdr,res.qvalue$lfdr,xlab="Truth (lfdr)", ylab="Estimate",main="qvalue",xlim=c(0,1),ylim=c(0,1))
+plot(ashr::get_lfdr(res.ash.true),res.qvalue$lfdr,xlab="Truth (lfdr)", ylab="Estimate",main="qvalue",xlim=c(0,1),ylim=c(0,1))
 abline(a=0,b=1,col=2,lwd=3)
@@
     plot(z,mu,xlim=c(-8,0),...)
-    points(z[CImatrix[,1]],CImatrix[,5],col="red")
-    points(z[CImatrix[,1]],CImatrix[,4],col="red")
+    points(z,CImatrix[,1],col="red")
+    points(z,CImatrix[,2],col="red")
 

### 2016-09-28 analysis/estimate_mode.Rmd
update

@@
 ```{r}
+library(ashr)
+check_mode = function(betahat, sebetahat,mixcompdist){
+    z.ash=ash(betahat,sebetahat,mixcompdist=mixcompdist,mode="estimate")
+    average.posteriormean = mean(get_pm(z.ash))
+    fitted.mode = comp_mean(get_fitted_g(z.ash))[1]
 
-check_mode = function(betahat, sebetahat,mixcompdist){
-    z.ash=ashr::ash(betahat,sebetahat,mixcompdist=mixcompdist,nonzeromode=TRUE)
-    average.posteriormean = mean(z.ash$PosteriorMean)
-    fitted.mode = ashr::comp_mean(z.ash$fitted.g)[1]
+    z.ash1 = ash(betahat-fitted.mode,sebetahat, mixcompdist=mixcompdist) #refit to get g
+    g = get_fitted_g(z.ash1)
+    loglik = get_loglik(z.ash1)
 
-    z.ash1 = ashr::ash(betahat-fitted.mode,sebetahat, mixcompdist=mixcompdist) #refit to get g
-    g = z.ash1$fitted.g
-    loglik = z.ash1$loglik
-
-    loglik.down = ashr::ash(z-fitted.mode-0.01,1,g=g)$loglik
-    loglik.up  = ashr::ash(z-fitted.mode+0.01,1,g=g)$loglik
-    loglik.posteriormean = ashr::ash(z-average.posteriormean,1,g=g)$loglik
+    loglik.down = ash(z-fitted.mode-0.01,1,g=g)$loglik
+    loglik.up  = ash(z-fitted.mode+0.01,1,g=g)$loglik
+    loglik.posteriormean = ash(z-average.posteriormean,1,g=g)$loglik
 
@@
   betahat = beta + rnorm(1000,0,0.1)
-  z.ash.hu = ashr::ash(betahat,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
-  z.ash.pu = ashr::ash(betahat,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")
+  z.ash.hu = ash(betahat,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
+  z.ash.pu = ash(betahat,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")
 
-  z.ash.hu2 = ashr::ash(betahat-0.2,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
-  z.ash.pu2 = ashr::ash(betahat-0.1,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")
+  z.ash.hu2 = ash(betahat-0.2,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
+  z.ash.pu2 = ash(betahat-0.1,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")
 
-  z.ash.hu.nzm = ashr::ash(betahat,0.1,mixcompdist="halfuniform",nonzeromode=TRUE,method="shrink")
+  z.ash.hu.nzm = ash(betahat,0.1,mixcompdist="halfuniform",mode="estimate",method="shrink")
   z.ash.hu$loglik
@@
 
-  test.op = function(c){return(-ashr::ash(betahat-c,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")$loglik)}
-  optimize(test.op,interval=c(min(betahat),max(betahat)))
 #  check_mode(betahat,.1,mixcompdist="halfuniform")

### 2016-09-28 analysis/make_GOODPOOR_figs.Rmd
update

@@
             data.frame(x=res.locfdr.good$fdr,y=res.locfdr$fdr[GOOD],type='locfdr'), 
-            data.frame(x=res.ash.good$lfsr,y=res.ash$lfsr[GOOD],type="ashr"))
+            data.frame(x=get_lfsr(res.ash.good),y=get_lfsr(res.ash)[GOOD],type="ashr"))
 
@@
 make_df_for_pval=function(ash,method="default"){
-  df = data.frame(p=pnorm(-abs(ash$data$betahat/ash$data$sebetahat)),lfsr=ash$lfsr,s=ash$data$sebetahat,method=method)
+  df = data.frame(p=pnorm(-abs(ash$data$x/ash$data$s)),lfsr=get_lfsr(ash),s=ash$data$s,method=method)
 }
@@
 
-res.ash.ET = ash(betahat,sebetahat,method="fdr",model="ET",mixcompdist="normal",outputlevel=4)
+res.ash.ET = ash(betahat,sebetahat,method="fdr",alpha=1,mixcompdist="normal",outputlevel=4)
 p=plot_pval_vs_lfsr(rbind(

### 2016-09-28 analysis/makefig_FDReg.Rmd
update

@@
 
-Here's an example of how qvalue decomposes $p$ values into null and alternative components
+Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:
 ```{r}
 #pdf("../figures/FDReg_hist.pdf",width=6,height=3)
-par(mai=c(0.5,0.5,0.2,0.2),mgp = c(3, 0.5, 0))
-layout(matrix(1:2,ncol=2,byrow=TRUE))
+#par(mai=c(0.5,0.5,0.2,0.2),mgp = c(3, 0.5, 0))
+#layout(matrix(1:2,ncol=2,byrow=TRUE))
 plot_FDReg_hist(pval,res.qvalue$pi0,type=1,title="p values",cex.axis=0.8,cex.main=0.8)
+
+```
+
+Now the decomposition:
+```{r}
 plot_FDReg_hist(pval,res.qvalue$pi0,type=4,cex.axis=0.8,yaxt='n',textsize=0.9,cex.main=0.8,title="Decomposition into null/alternative")
@@
 axis(side=1, labels=FALSE,tck=-0.01)
-altnullhist(pval,res.ash$lfdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
+altnullhist(pval,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8)
 axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
@@
 axis(side=1, labels=FALSE,tck=-0.01)
-nullalthist(zscore,res.ash$lfdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
+nullalthist(zscore,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
 axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
@@
 axis(side=1, labels=FALSE,tck=-0.01)
-altnullhist(pval,res.ash$lfdr,main="ash",ncz=ncz,xaxt='n',cex.axis=0.8)
+altnullhist(pval,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8)
 axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
@@
 axis(side=1, labels=FALSE,tck=-0.01)
-nullalthist(zscore,res.ash$lfdr,main="ash",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
+nullalthist(zscore,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
 axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)

### 2016-09-28 analysis/referee_uaza.Rmd
update

@@
 ```{r}
-lfdr.ash=ashr::ash(x.est,1/sqrt(30))$lfdr
+library(ashr)
+lfdr.ash=get_lfdr(ash(x.est,1/sqrt(30)))
 nullalthist(zscore,lfdr.ash,main="ash partition of z \n into null and alternative",ncz=60,cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6),xlab="z score")
@@
 trueg = ashr::normalmix(c(0.8,0.2),c(0,0),c(0,1))
-lfdr.true=ashr::ash(x.est,1/sqrt(30),g=trueg,fixg=TRUE)$lfdr
+lfdr.true=get_lfdr(ash(x.est,1/sqrt(30),g=trueg,fixg=TRUE))
 nullalthist(zscore,lfdr.true,main="true partition of z \n into null and alternative",ncz=60,cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6),xlab="z score")

### 2016-09-28 code/dsc-shrink/methods/ash.wrapper.R

@@ -34,7 +34,7 @@ bayes.wrapper = function(input, meta, args=NULL){
 }
 
 ash2beta_est =function(output){
-  return (list(beta_est=output$PosteriorMean))
+  return (list(beta_est=get_pm(output)))
 } 
 
 ash2pi0_est =function(output){
@@ -42,5 +42,5 @@ ash2pi0_est =function(output){
 } 
 
 ash2fitted.g = function(output){
-  return (list(fitted.g=output$fitted.g))
+  return (list(fitted.g=get_fitted_g(output)))
 }

### 2016-09-29 analysis/adaptive_shrinkage.Rmd
Working through adaptive_shrinkage.Rmd.

@@
 
-```{r chunk-options, include=FALSE}
+The goal here is to illustrate the "adaptive" nature of the adaptive
+shrinkage. The shrinkage is adaptive in two senses. First, the amount
+of shrinkage depends on the distribution $g$ of the true effects,
+which is learned from the data: when $g$ is very peaked about zero
+then ash learns this and deduces that signals should be more strongly
+shrunk towards zero than when g is less peaked about zero.  Second,
+the amount of shrinkage of each observation depends on its standard
+error: the smaller the standard error, the more informative the data,
+and so the less shrinkage that occurs. From an Empirical Bayesian
+perspective both of these points are entirely natural: the posterior
+depends on both the prior and the likelihood; the prior, $g$, is
+learned from the data, and the likelihood incorporates the standard
+error of each observation.
+
+TO DO: explain what this chunk does.
+
+```{r chunk-options,include = FALSE}
 source("chunk-options.R")
@@
 
-The goal here is to illustrate the "adaptive" nature of the adaptive shrinkage. The shrinkage is adaptive in two senses. First,
-the amount of shrinkage depends on the distribution $g$ of the true effects, which is learned from the data: when $g$ is very peaked about zero
-then ash learns this and deduces that signals should be more strongly shrunk towards zero than when g is less peaked about zero.
-Second, the amount of shrinkage of each observation depends on its standard error: the smaller the standard error, the more informative
-the data, and so the less shrinkage that occurs. From an Empirical Bayesian perspective both of these 
-points are entirely natural: the posterior depends on both the prior
-and the likelihood; the prior, $g$, is learned from the data, and the likelihood incorporates the standard error of each observation.
+We simulate from two scenarios, the first with effects more peaked
+about zero, the second with effects less peaked.
 
-We simulate from two scenarios, the first with effects more peaked about zero, the second with effects less peaked.
 ```{r initialize}
@@
 library(scales)
-#load the simulation function, (rnormmix_datamaker)
 source("../code/dsc-shrink/datamakers/datamaker.R")

### 2016-09-30 analysis/adaptive_shrinkage.Rmd
Made some adjustments to adaptive_shrinkage.Rmd so that it renders correctly in atom.

@@
 
-TO DO: explain what this chunk does.
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
 
-```{r chunk-options,include = FALSE}
+```r
+library(ashr)
+library(ggplot2)
+library(scales)
 source("chunk-options.R")
@@
 
-We simulate from two scenarios, the first with effects more peaked
-about zero, the second with effects less peaked.
+We simulate from two scenarios: in the first scenario, the effects are more
+peaked about zero (**sim.spiky**); in the second scenario, the effects are
+less peaked (**sim.bignormal**).
 
-```{r initialize}
+```r
 set.seed(100)
-library(ashr)
-library(ggplot2)
-library(scales)
 source("../code/dsc-shrink/datamakers/datamaker.R")
 
-NSAMP=1000
+NSAMP = 1000
 s     = 1/rgamma(NSAMP,5,5)
-sim.bignormal = rnormmix_datamaker(args=list(g=normalmix(c(1),c(0),c(4)),
-                        min_pi0=0,
-                        max_pi0=0,
-                        nsamp=NSAMP,
-                        betahatsd=s
-                      ))
 
-sim.spiky = rnormmix_datamaker(args=list(g=normalmix(c(.4,.2,.2,.2),c(0,0,0,0),c(.25,.5,1,2)),
-                        min_pi0=0,
-                        max_pi0=0,
-                        nsamp=NSAMP,
-                        betahatsd=s
-                      ))
+sim.spiky =
+  rnormmix_datamaker(args = list(g = normalmix(c(0.4,0.2,0.2,0.2),
+                                               c(0,0,0,0),
+                                               c(0.25,0.5,1,2)),
+                                  min_pi0   = 0,
+                                  max_pi0   = 0,
+                                  nsamp     = NSAMP,
+                                  betahatsd = s))
 
+sim.bignormal =
+  rnormmix_datamaker(args = list(g         = normalmix(1,0,4),
+                                 min_pi0   = 0,
+                                 max_pi0   = 0,
+                                 nsamp     = NSAMP,
+                                 betahatsd = s))
 ```
 
-Now run ash on both:
-```{r run-ash}
-beta.bignormal.ash = ash(sim.bignormal$input$betahat, s)
-beta.spiky.ash = ash(sim.spiky$input$betahat, s)
+Now we run ash on both data sets.
+
+```r
+beta.spiky.ash     = ash(sim.spiky$input$betahat,s)
+beta.bignormal.ash = ash(sim.bignormal$input$betahat,s)
 ```
@@
 
-```{r plotting}
+```r
 make_df_for_ashplot = function(sim1,sim2,ash1,ash2,name1="spiky",name2="big-normal"){
@@
 
-```{r}
+```r
 pval_plot=function(df){

### 2016-09-30 analysis/chunk-options.R

@@ -7,4 +7,4 @@ opts_chunk$set(tidy = FALSE,
                fig.align = "center",
                fig.path = paste0("figure/", current_input(), "/"))
 
-knit_hooks$set(crop=hook_pdfcrop)
+knit_hooks$set(crop = hook_pdfcrop)

### 2016-09-30 analysis/adaptive_shrinkage.Rmd
Removed curly brace syntax in adaptive_shrinkage.Rmd that causes problems with rendering.

@@
 author: "Matthew Stephens"
-date: 2015-10-26
+date: 2016-09-29
 ---
@@
 Here we plot the shrunken estimates against the observed values, colored
-according to the (square root of) precision: precise esitmates being colored red, and less precise estimates being blue. Two key features of the plots
-illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large standard error (blue) are shrunk more than estimates with small standard error (red) illustrating that shrinkage adapts to measurement precision.
+according to the (square root of) precision: precise esitmates being colored
+red, and less precise estimates being blue. Two key features of the plots
+illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky
+scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large
+standard error (blue) are shrunk more than estimates with small standard
+error (red) illustrating that shrinkage adapts to measurement precision.
 
@@
 
-```{r plotting lfsr}
+```r
 
@@
 the standardized beta are exchangeable (rather than the beta being exchangeable).
-```{r run-ash-ET}
+```r
 beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,alpha=1,mixcompdist="normal")
@@
 In this case the significance by lfsr is not quite the same as cutting off at a given $p$ value (you can see that the decision boundary is not quite the same as drawing a horizontal line), but also not that different, presumably because the standard errors, although varying across observations, do not vary greatly.
-```{r volcano}
+```r
   ggplot(df,aes(x=betahat,y=-log10(2*pnorm(-abs(betahat/s))),col=signif)) + geom_point(alpha=1,size=1.75) +  facet_grid(.~scenario) + theme(legend.position = "none") +
@@
 
-```{r sehist}
+```r
 hist(s,main="histogram of standard errors")
@@
 
-```{r info}
+```r
 sessionInfo()

### 2016-09-30 analysis/adaptive_shrinkage.Rmd
Added back chunk labels to adaptive_shrinkage.Rmd.

--- 2016-09-30 analysis/adaptive_shrinkage.Rmd
@@
 then ash learns this and deduces that signals should be more strongly
-shrunk towards zero than when g is less peaked about zero.  Second,
+shrunk towards zero than when $g$ is less peaked about zero.  Second,
 the amount of shrinkage of each observation depends on its standard
@@
 
-```r
+```{r chunk_options, include=FALSE}
 library(ashr)
@@
 
-```r
+```{r initialize}
 set.seed(100)
@@
 
-```r
+```{r run_ash}
 beta.spiky.ash     = ash(sim.spiky$input$betahat,s)
@@
 Here we plot the shrunken estimates against the observed values, colored
-according to the (square root of) precision: precise esitmates being colored red, and less precise estimates being blue. Two key features of the plots
-illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large standard error (blue) are shrunk more than estimates with small standard error (red) illustrating that shrinkage adapts to measurement precision.
+according to the (square root of) precision: precise esitmates being colored
+red, and less precise estimates being blue. Two key features of the plots
+illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky
+scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large
+standard error (blue) are shrunk more than estimates with small standard
+error (red) illustrating that shrinkage adapts to measurement precision.
 
-```r
+```{r plotting}
 make_df_for_ashplot = function(sim1,sim2,ash1,ash2,name1="spiky",name2="big-normal"){
@@
 
-```{r plotting lfsr}
+```{r plotting_lfsr}
 
@@
 
-```r
+```{r plot_pvalues}
 pval_plot=function(df){
@@
 the standardized beta are exchangeable (rather than the beta being exchangeable).
-```{r run-ash-ET}
+```{r run_ash_ET}
 beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,alpha=1,mixcompdist="normal")

### 2016-09-30 analysis/adaptive_shrinkage.Rmd
Updated code in adaptive_shrinkage.Rmd up to plot_shrunk_vs_obs.

@@
 peaked about zero (**sim.spiky**); in the second scenario, the effects are
-less peaked (**sim.bignormal**).
+less peaked at zero (**sim.bignormal**). A summary of the two data sets is
+printed at the end of this chunk.
 
@@
                                  betahatsd = s))
+
+cat("Summary of observed beta-hats:\n")
+print(rbind(spiky     = quantile(sim.spiky$input$betahat,seq(0,1,0.1)),
+            bignormal = quantile(sim.bignormal$input$betahat,seq(0,1,0.1))),
+      digits = 3)
+
 ```
@@
 
-Here we plot the shrunken estimates against the observed values, colored
-according to the (square root of) precision: precise esitmates being colored
+Next we plot the shrunken estimates against the observed values, colored
+according to the (square root of) precision: precise estimates being colored
 red, and less precise estimates being blue. Two key features of the plots
@@
 
-```{r plotting}
-make_df_for_ashplot = function(sim1,sim2,ash1,ash2,name1="spiky",name2="big-normal"){
-  df = data.frame(betahat = c(sim1$input$betahat,sim2$input$betahat),
+```{r plot_shrunk_vs_obs}
+make_df_for_ashplot =
+  function (sim1, sim2, ash1, ash2, name1 = "spiky", name2 = "big-normal") {
+    n    = length(sim1$input$betahat)
+    lfsr = c(get_lfsr(ash1),get_lfsr(ash2))
+    return(data.frame(betahat  = c(sim1$input$betahat,sim2$input$betahat),
                       beta_est = c(get_pm(ash1), get_pm(ash2)),
-                lfsr = c(get_lfsr(ash1), get_lfsr(ash2)),
-                s=c(sim1$input$sebetahat,sim2$input$sebetahat),
-                scenario=c(rep(name1,NSAMP),rep(name2,NSAMP)))
-  df$signif = df$lfsr<0.05
-  return(df)
+                      lfsr     = lfsr,
+                      s        = c(sim1$input$sebetahat,sim2$input$sebetahat),
+                      scenario = c(rep(name1,n),rep(name2,n)),
+                      signif   = lfsr < 0.05))
   }
 
-ashplot=function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate"){
-  ggplot(df,aes(x=betahat,y=beta_est,color=1/s)) +
-    xlab(xlab) +
-    ylab(ylab) + geom_point() +  facet_grid(.~scenario) +   geom_abline(intercept=0,slope=1) +
-  scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab")  +coord_fixed(ratio=1)
+ashplot = function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate") {
+  ggplot(df,aes(x = betahat,y = beta_est,color = 1/s)) +
+    xlab(xlab) + ylab(ylab) + geom_point() + facet_grid(.~scenario) +   geom_abline(intercept = 0,slope = 1,linetype = "dotted") +
+    scale_colour_gradient2(midpoint = median(1/s),low = "blue",mid = "white",
+                           high = "red",space = "Lab") +
+    coord_fixed(ratio = 1)
  }
-df=make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash,beta.bignormal.ash)
-ashplot(df)
+
+df = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash,
+                         beta.bignormal.ash)
+print(ashplot(df))
 ```
 
-Now plot lfsr against z scores, colored
-according to the (square root of) precision
+Now plot lfsr against z scores, colored according to the (square root of)
+precision.
 
-```{r plotting_lfsr}
+```{r plot_lfsr}
 
@@
 
-z_lfsr_plot(df)
+print(z_lfsr_plot(df))
 ```

### 2016-09-30 code/dsc-shrink/datamakers/datamaker.R

@@ -2,13 +2,13 @@ library(ashr)
 
 #' @title datamaker for FDR DSC
 #'
-#' @description Simulates data for a DSC for methods to compute FDR/q values  
+#' @description Simulates data for a DSC for methods to compute FDR/q values
 #' @details None
-#' 
+#'
 #' @param args A list of the remaining arguments, which in this case is
 #' \item{nsamp}{The number of samples to create}
-#' \item{g}{An object of class normalmix specifying the mixture distribution from which non-null beta values 
-#' are to be simulated}
+#' \item{g}{An object of class normalmix specifying the mixture distribution from
+#'          which non-null beta values are to be simulated}
 #' \item{min_pi0}{The minimum value of pi0, the proportion of true nulls}
 #' \item{max_pi0}{The maximum value of pi0, the proportion of true null}
 #' \item{betahatsd}{The standard deviation of betahat to use}
@@ -16,11 +16,11 @@ library(ashr)
 #' @return a list with the following elements
 #' \item{meta}{A list containing the meta data. In this case beta}
 #' \item{input}{A list containing the input data; in this case the set of betahat values and their standard errors}
-#' 
+#'
 rnormmix_datamaker = function(args){
   #here is the meat of the function that needs to be defined for each dsc to be done
   pi0 = runif(1,args$min_pi0,args$max_pi0) #generate the proportion of true nulls randomly
-  
+
   k = ncomp(args$g)
   comp = sample(1:k,args$nsamp,mixprop(args$g),replace=TRUE) #randomly draw a component
   isnull = (runif(args$nsamp,0,1) < pi0)
@@ -29,13 +29,11 @@ rnormmix_datamaker = function(args){
   betahat = beta + rnorm(args$nsamp,0,sebetahat)
   meta=list(g1=args$g,beta=beta,pi0=pi0)
   input=list(betahat=betahat,sebetahat=sebetahat,df=NULL)
-  
+
   #end of meat of function
-  
+
   data = list(meta=meta,input=input)
-  
+
   return(data)
-  
+
 }
-
-

### 2016-10-03 analysis/adaptive_shrinkage.Rmd
Made a few slight adjustments to adaptive_shrinkage.Rmd, mainly formatting of the code. It works with 'general' branch of ashr.

@@
 illustrate the ideas of adaptive shrinkage: i) the estimates under the spiky
-scenario are shrunk more strongly, illustrating that shrinkage adapts to the underlying distribution of beta; ii) in both cases, estimates with large
+scenario are shrunk more strongly, illustrating that shrinkage adapts to the
+underlying distribution of beta; ii) in both cases, estimates with large
 standard error (blue) are shrunk more than estimates with small standard
@@
     n = length(sim1$input$betahat)
-    lfsr = c(get_lfsr(ash1),get_lfsr(ash2))
+    x = c(get_lfsr(ash1),get_lfsr(ash2))
     return(data.frame(betahat  = c(sim1$input$betahat,sim2$input$betahat),
-                      beta_est = c(get_pm(ash1), get_pm(ash2)),
-                      lfsr     = lfsr,
+                      beta_est = c(get_pm(ash1),get_pm(ash2)),
+                      lfsr     = x,
                       s        = c(sim1$input$sebetahat,sim2$input$sebetahat),
                       scenario = c(rep(name1,n),rep(name2,n)),
-                      signif   = lfsr < 0.05))
+                      signif   = x < 0.05))
   }
 
-ashplot = function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate") {
+ashplot = function(df,xlab="Observed beta-hat",ylab="Shrunken beta estimate")
   ggplot(df,aes(x = betahat,y = beta_est,color = 1/s)) +
-    xlab(xlab) + ylab(ylab) + geom_point() + facet_grid(.~scenario) +   geom_abline(intercept = 0,slope = 1,linetype = "dotted") +
-    scale_colour_gradient2(midpoint = median(1/s),low = "blue",mid = "white",
-                           high = "red",space = "Lab") +
+    xlab(xlab) + ylab(ylab) + geom_point() +
+    facet_grid(.~scenario) +
+    geom_abline(intercept = 0,slope = 1,linetype = "dotted") +
+    scale_colour_gradient2(midpoint = median(1/s),low = "blue",
+                           mid = "white",high = "red",space = "Lab") +
     coord_fixed(ratio = 1)
- }
 
@@
 
-z_lfsr_plot=function(df,ylab="Observed Z score",xlab="lfsr"){
-  ggplot(df,aes(x=lfsr,y=betahat/s,color=1/s)) +
-    xlab(xlab) +
-    ylab(ylab) + geom_point() +  facet_grid(.~scenario) + scale_colour_gradient2(midpoint = median(1/s),low="blue", mid="white",high="red",space="Lab")
-}
+z_lfsr_plot = function(df,ylab = "Observed Z score",xlab = "lfsr")
+  ggplot(df,aes(x = lfsr,y = betahat/s,color = 1/s)) +
+    xlab(xlab) + ylab(ylab) + geom_point() + facet_grid(.~scenario) +
+    scale_colour_gradient2(midpoint = median(1/s),low = "blue",
+                           mid = "white",high = "red",space = "Lab")
 
@@
 
-A related consequence is that significance of each observation is no longer monotonic with $p$ value.
+A related consequence is that significance of each observation is no longer
+monotonic with $p$ value.
+
+**Note:** I find it confusing that you are using the same colour scheme here as
+above, but you are showing `log(s)`, whereas above you are showing `1/s`. It
+seems that it would be better to reverse the colour scheme for the `pval_plot`
+to better correspond to the above plots. –Peter
 
 ```{r plot_pvalues}
-pval_plot=function(df){
-  ggplot(df,aes(x=pnorm(-abs(betahat/s)),y=lfsr,color=log(s))) + geom_point() +  facet_grid(.~scenario) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr") +
-  scale_colour_gradient2(midpoint = 0,low="blue", mid="white",high="red")
-}
-pval_plot(df)
+pval_plot = function (df)
+  ggplot(df,aes(x = pnorm(-abs(betahat/s)),y = lfsr,color = log(s))) +
+  geom_point() + facet_grid(.~scenario) + xlim(c(0,0.025)) +
+  xlab("p value") + ylab("lfsr") +
+  scale_colour_gradient2(midpoint = 0,low = "blue",
+                         mid = "white",high = "red")
+
+print(pval_plot(df))
 ```
@@
 Let's see how these are affected by changing the modelling assumptions so that
-the standardized beta are exchangeable (rather than the beta being exchangeable).
+the *standardized* beta are exchangeable (rather than the beta being exchangeable).
+
 ```{r run_ash_ET}
-beta.bignormal.ash.ET = ash(sim.bignormal$input$betahat, s,alpha=1,mixcompdist="normal")
-beta.spiky.ash.ET = ash(sim.spiky$input$betahat, s,alpha=1,mixcompdist="normal")
-df.ET = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ET,beta.bignormal.ash.ET)
-ashplot(df.ET,ylab="Shrunken beta estimate (ET model)")
+beta.bignormal.ash.ET =
+  ash(sim.bignormal$input$betahat,s,alpha = 1,mixcompdist = "normal")
+beta.spiky.ash.ET =
+  ash(sim.spiky$input$betahat,s,alpha = 1,mixcompdist = "normal")
+df.ET = make_df_for_ashplot(sim.spiky,sim.bignormal,beta.spiky.ash.ET,
+                            beta.bignormal.ash.ET)
+ashplot(df.ET,ylab = "Shrunken beta estimate (ET model)")
 pval_plot(df.ET)
@@
 
-This is a "volcano plot" showing effect size against p value.
-The blue points are "significant" in that they have lfsr<0.05.
-In this case the significance by lfsr is not quite the same as cutting off at a given $p$ value (you can see that the decision boundary is not quite the same as drawing a horizontal line), but also not that different, presumably because the standard errors, although varying across observations, do not vary greatly.
+This is a "volcano plot" showing effect size against p value. The blue points
+are "significant" in that they have lfsr < 0.05.
+
 ```{r volcano}
-  ggplot(df,aes(x=betahat,y=-log10(2*pnorm(-abs(betahat/s))),col=signif)) + geom_point(alpha=1,size=1.75) +  facet_grid(.~scenario) + theme(legend.position = "none") +
-  xlim(c(-10, 10)) + ylim(c(0, 15)) +
-  xlab("Effect (beta)") + ylab("-log10 p-value")
+print(ggplot(df,aes(x = betahat,y = -log10(2*pnorm(-abs(betahat/s))),
+                    col = signif)) +
+  geom_point(alpha = 1,size = 1.75) + facet_grid(.~scenario) +
+  theme(legend.position = "none") + xlim(c(-10,10)) + ylim(c(0,15)) +
+  xlab("Effect (beta)") + ylab("-log10 p-value"))
 ```
 
+In this case the significance by lfsr is not quite the same as cutting off
+at a given p value (you can see that the decision boundary is not quite the
+same as drawing a horizontal line), but also not that different, presumably
+because the standard errors, although varying across observations, do not
+vary greatly.
+
 ```{r sehist}
-hist(s,main="histogram of standard errors")
-summary(s)
+hist(s,main = "histogram of standard errors")
+print(summary(s))
 ```
 
-## Session information
+Session information.
 
 ```{r info}
-sessionInfo()
+print(sessionInfo())
 ```

### 2016-10-03 analysis/adaptive_shrinkage.Rmd
In adaptive_shrinkage.Rmd, I found it confusing that the same colour scheme was used in plot_pvalues as above, but this plot shows `log(s)`, whereas above the plots show `1/s`. It seems that it would be better to reverse the colour scheme for the `pval_plot` to better correspond to the above plots.

@@
 
-**Note:** I find it confusing that you are using the same colour scheme here as
-above, but you are showing `log(s)`, whereas above you are showing `1/s`. It
-seems that it would be better to reverse the colour scheme for the `pval_plot`
-to better correspond to the above plots. –Peter
-
 ```{r plot_pvalues}
@@
   xlab("p value") + ylab("lfsr") +
-  scale_colour_gradient2(midpoint = 0,low = "blue",
-                         mid = "white",high = "red")
+  scale_colour_gradient2(midpoint = 0,low = "red",
+                         mid = "white",high = "blue")
 

### 2016-10-14 analysis/IPvsEM.Rmd
Encountered error in running 'ash' with REBayes package using stephens999/ashr with 'general' branch.

@@
 
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
 ```{r chunk-options, include=FALSE}
@@
 
-The following simulation assumes $\beta \sim N(0,betasd^2)$ and with $\hat{\beta}$ having standard deviation 1.
-```{r}
-timed_sims = function(ash.args,nsim=20,nmin = 100,nmax = 1000,betasd=1){
+The following simulation assumes $\beta \sim N(0,\sigma_{\sf betasd}^2)$ and
+with $\hat{\beta}$ having standard deviation 1.
+
+```{r generate_data}
+timed_sims = function(ash.args,nsim = 20,nmin = 100,nmax = 1000,betasd = 1){
   n=10^seq(log10(nmin),log10(nmax),length=nsim)

### 2016-10-14 analysis/adaptive_shrinkage.Rmd
Encountered error in running 'ash' with REBayes package using stephens999/ashr with 'general' branch.

@@
 ```{r plot_lfsr}
-
 z_lfsr_plot = function(df,ylab = "Observed Z score",xlab = "lfsr")
@@
 
-Session information.
+## Session information.
 

### 2016-10-14 analysis/efron.fcr.Rmd
Encountered error in running 'ash' with REBayes package using stephens999/ashr with 'general' branch.

@@
   true.g = ashr::normalmix(c(0.9,0.1),c(0,-3),c(0,1))
-  res.ash.true = ashr::ash(z,1,g=true.g,fixg=TRUE)
+  res.ash.true = ashr::ash(z,1,g=true.g,fixg = TRUE)
 ```
@@
 res.locfdr = locfdr(z,nulltype = 0)
-res.qvalue = qvalue(p=pchisq(z^2,df=1,lower.tail = FALSE))  
+res.qvalue = qvalue(p = pchisq(z^2,df=1,lower.tail = FALSE))
 #res.pos.qvalue = qvalue(p=pchisq(z[z>0]^2,df=1,lower.tail = FALSE))

### 2016-10-14 analysis/estimate_mode.Rmd
Encountered error in running 'ash' with REBayes package using stephens999/ashr with 'general' branch.

@@
 
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
 ```{r chunk-options, include=FALSE}
+library(ashr)
 source("chunk-options.R")
@@
 
-The purpose of this analysis is to check the performance of the non-zero mode option.
+The purpose of this analysis is to check the performance of the non-zero mode
+option.
 
@@
 
-I conjecture that the mean of the posterior means should be close to the optimum for the mode...
-maybe even equal to it. (That is, this would constitute a fixed point for the update. We aren't
-explicitly using that directly in the current implementation; the uniform mixture uses optim to
-do it numerically; the normal mixture uses a true EM I think...)
+I conjecture that the mean of the posterior means should be close to the
+optimum for the mode... maybe even equal to it. (That is, this would constitute
+a fixed point for the update. We aren't explicitly using that directly in the
+current implementation; the uniform mixture uses `optim` to do it numerically;
+the normal mixture uses a true EM I think...)
 
 ```{r}
-library(ashr)
-check_mode = function(betahat, sebetahat,mixcompdist){
-    z.ash=ash(betahat,sebetahat,mixcompdist=mixcompdist,mode="estimate")
+check_mode = function(betahat, sebetahat, mixcompdist) {
+  z.ash = ash(betahat,sebetahat,mixcompdist = mixcompdist,mode = "estimate")
   average.posteriormean = mean(get_pm(z.ash))
@@
 
-    z.ash1 = ash(betahat-fitted.mode,sebetahat, mixcompdist=mixcompdist) #refit to get g
+  # Refit to get g.
+  z.ash1 = ash(betahat - fitted.mode,sebetahat,mixcompdist = mixcompdist)
   g      = get_fitted_g(z.ash1)
@@
 
-    loglik.down = ash(z-fitted.mode-0.01,1,g=g)$loglik
-    loglik.up  = ash(z-fitted.mode+0.01,1,g=g)$loglik
-    loglik.posteriormean = ash(z-average.posteriormean,1,g=g)$loglik
+  loglik.down          = ash(z - fitted.mode - 0.01,1,g = g)$loglik
+  loglik.up            = ash(z - fitted.mode + 0.01,1,g = g)$loglik
+  loglik.posteriormean = ash(z - average.posteriormean,1,g = g)$loglik
 
-    return(list(fitted.mode=fitted.mode, average.posteriormean =average.posteriormean,
+  return(list(fitted.mode = fitted.mode,
+              average.posteriormean = average.posteriormean,
               loglik = c(loglik,loglik.down,loglik.up,loglik.posteriormean)))
 }
+
 set.seed(100)
-  z = rnorm(1000)+3
-  check_mode(z,1,mixcompdist="uniform")
-  check_mode(z,1,mixcompdist="normal")
-  check_mode(z,1,mixcompdist="halfuniform")
+z = rnorm(1000) + 3
+print(check_mode(z,1,mixcompdist="uniform"))
+print(check_mode(z,1,mixcompdist="normal"))
+print(check_mode(z,1,mixcompdist="halfuniform"))
 ```
@@
 
-## Session information
+## Session information.
 

### 2016-10-14 analysis/plot_lfsr.Rmd
remove pdf files
Encountered error in running 'ash' with REBayes package using stephens999/ashr with 'general' branch.

@@
 
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
 ```{r chunk-options, include=FALSE}
+library(dplyr)
+library(ggplot2)
 source("chunk-options.R")
 ```
+
+TO DO: Add note here about having to first run script to generate data sets
+required to run the analysis.
 
@@
 source("../R/set_plot_colors.R")
-library(dplyr)
 
@@
 
-## Session information
+## Session information.
 
 ```{r info}
-sessionInfo()
+print(sessionInfo())
 ```

### 2016-10-19 analysis/IPvsEM.Rmd
Working on updates to IPvsEM.Rmd.

@@
 ```{r chunk-options, include=FALSE}
+library(REBayes)
+library(ashr)
+library(ggplot2)
 source("chunk-options.R")
@@
 
-```{r generate_data}
-timed_sims = function(ash.args,nsim = 20,nmin = 100,nmax = 1000,betasd = 1){
-  n=10^seq(log10(nmin),log10(nmax),length=nsim)
-  elapsed.time= rep(0,nsim)
-  for(i in 1:nsim){
+```{r simulation_setup}
+timed_sims = function(ash.args,nsim = 20,nmin = 100,nmax = 1000,betasd = 1) {
+  n = 10^seq(log10(nmin),log10(nmax),length = nsim)
+  elapsed.time = rep(0,nsim)
+  for (i in 1:nsim) {
     set.seed(i)
-    betahat = rnorm(n[i],sd=sqrt(1+betasd^2))
-    elapsed.time[i] = system.time(do.call(ashr::ash,args = modifyList(ash.args,list(betahat = betahat, sebetahat=1))))[3]
+    with(ash.args,cat(sprintf("%6s %13s %3d\n",method,optmethod,i)))
+    betahat = rnorm(n[i],sd = sqrt(1 + betasd^2))
+    elapsed.time[i] =
+      system.time(do.call(ash,args = modifyList(ash.args,
+        list(betahat = betahat, sebetahat = 1))))[3]
   }
-  return(data.frame(elapsed.time=elapsed.time,seed=1:nsim,n = n))
+  return(data.frame(elapsed.time = elapsed.time,seed = 1:nsim,n = n))
 }
@@
 
-Now run a simulation study, for $n$ (number of tests; $p$ in paper) in range 10 to 100,000.
-Note that the warnings are being generated by the EM algorithm in big problems due to lack of convergence.
+Now run a simulation study, for $n$ (number of tests; $p$ in paper) in
+range 10 to 100,000. Note that the warnings are being generated by the EM
+algorithm in big problems due to lack of convergence.
 
-```{r}
-  df=data.frame()
-  for(method in c("fdr","shrink")){
-    for(optmethod in c("mixIP","cxxMixSquarem")){
-      df = rbind(df, data.frame(method=method, optmethod=optmethod,  timed_sims(list(method=method,optmethod=optmethod),nsim=50,nmin=10,nmax=100000)))
+```{r model_fitting}
+df = data.frame()
+cat("method --optmethod-- sim\n")
+for(method in c("fdr","shrink")) {
+  for(optmethod in c("mixIP","cxxMixSquarem")) {
+    df = rbind(df,data.frame(method = method,optmethod = optmethod,
+           timed_sims(list(method = method,optmethod = optmethod),
+                           nsim = 50,nmin = 10,nmax = 100000)))
   }
 }
-
+cat("\n")
 ```
@@
 
-Now plot the time as a function of $n$
-```{r}
-  ggplot2::qplot(x=n,y=elapsed.time,data=df,col=optmethod,facets=.~method,ylab="elapsed time (s)")
+Now plot time as a function of $n$:
+
+```{r plot_results1}
+qplot(x = n,y = elapsed.time,data = df,col=optmethod,facets=.~method,ylab="elapsed time (s)")
 ```
@@
 And a zoom-in on "small" problems with $n<5000$.
-```{r}
+
+```{r plot_results2}
 ggplot2::qplot(x=n,y=elapsed.time,data=dplyr::filter(df,n<5000),col=optmethod,facets=.~method,ylab="elapsed time (s)")

### 2016-10-19 analysis/adaptive_shrinkage.Rmd
Working on updates to IPvsEM.Rmd.

@@
 ```{r chunk_options, include=FALSE}
+library(REBayes)
 library(ashr)
@@
       digits = 3)
-
 ```

### 2016-10-19 analysis/makefig_FDReg.Rmd
Working on updates to IPvsEM.Rmd.

@@
 
-```{r chunk-options, include=FALSE}
-source("chunk-options.R")
-```
-
 This file makes two figures:
 
-1. illustrates basic idea of way q value does FDR analysis in p value space
-2. compares methods in the way they decompose p values and z scores into two groups
+1. Illustrates basic idea of way q value does FDR analysis in p value space.
 
-```{r}
+2. Compares methods in the way they decompose p values and z scores into
+two groups.
+
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk-options, include=FALSE}
 require(ashr)
@@
 require(mixfdr)
-source("../R/plot_FDReg_hist.R") #some plotting functions
+source("chunk-options.R")
+source("../R/plot_FDReg_hist.R")
 source("../R/nullalthist.R")
+```
 
-############################ FDReg_hist.pdf ##################################################################
+## Simple simulated example
 
-#simple simulated example
-ncz = 100 # number of bins in z score histograms
+Generates DReg_hist.pdf?
+
+```{r add_label_here}
+
+# ncz is the number of bins in the z score histograms.
+ncz   = 100
 ntest = 10000
@@
 
-#simulate (with all tests alternative, true effects $\beta \sim N(0,1)$)
+# Simulate (with all tests alternative, true effects $\beta \sim N(0,1)$).
 beta      = rnorm(ntest)
 sebetahat = 1
-betahat = beta + rnorm(ntest, 0, sebetahat)
+betahat   = beta + rnorm(ntest,0,sebetahat)
 zscore    = betahat/sebetahat
-pval = pchisq(zscore^2,df=1,lower.tail=F)
+pval      = pchisq(zscore^2,df = 1,lower.tail = FALSE)
 
-#apply the different methods
+# Apply the different methods.
 res.qvalue = qvalue(pval)
-res.locfdr = locfdr(zscore,nulltype=0,plot=0)
-res.mixfdr = mixFdr(zscore,noiseSD=1,theonull=TRUE,plot=FALSE)
-res.ash = ash(betahat,1,method="fdr")
+res.locfdr = locfdr(zscore,nulltype = 0,plot = 0)
+res.mixfdr = mixFdr(zscore,noiseSD = 1,theonull = TRUE,plot = FALSE)
+res.ash    = ash(betahat,1,method = "fdr")
 

### 2016-10-19 analysis/IPvsEM.Rmd
IPvsEM.Rmd runs correctly; made a few minor adjustments to code and text.

@@
 cat("\n")
+
+# Summarize the results in a table.
+cat("Average computation time (seconds):\n")
+print(as.table(by(df,df[c("method","optmethod")],
+                  function (x) mean(x$elapsed.time))))
 ```
@@
 
-```{r plot_results1}
-qplot(x = n,y = elapsed.time,data = df,col=optmethod,facets=.~method,ylab="elapsed time (s)")
+```{r plot_results}
+qplot(x = n,y = elapsed.time,data = df,col = optmethod,facets = .~method,
+      ylab = "elapsed time (s)")
 ```
 
-And a zoom-in on "small" problems with $n<5000$.
+Zoom-in on "small" problems with $n < 5000$.
 
-```{r plot_results2}
-ggplot2::qplot(x=n,y=elapsed.time,data=dplyr::filter(df,n<5000),col=optmethod,facets=.~method,ylab="elapsed time (s)")
+```{r plot_results_magnified}
+qplot(x = n,y = elapsed.time, data = subset(df,n < 5000),
+      col = optmethod,facets = .~method,ylab = "elapsed time (s)")
 ```
@@
 
-The IP method clearly scales to large problem much better than EM. It is faster, and also more reliable (sometimes reaches higher log-likelihood than EM, never smaller); see
-[checkIP.html](checkIP.html).
+The IP method clearly scales to large problem much better than EM. It is faster,
+and also more reliable (sometimes reaches higher log-likelihood than EM,
+never smaller); see [checkIP.html](checkIP.html).
 
-However, for small problems (n<5000) the EM is adequate for many practical purposes, solving within a few seconds. This is particularly true for the penalty term (method=fdr), which
- helps the EM converge, presumably because it is helping identifiability, removing the large flat parts of the likelihood objective function.
+However, for small problems (n < 5000) the EM is adequate for many practical
+purposes, solving within a few seconds. This is particularly true for the
+penalty term (method = fdr), which helps the EM converge, presumably because
+it is helping identifiability, removing the large flat parts of the likelihood
+objective function.
 
-Indeed, for small problems with the penalty term (n<2000) EM with squarem acceleration is a little faster in these comparisons than IP.
+Indeed, for small problems with the penalty term (n < 2000) EM with squarem
+acceleration is a little faster in these comparisons than IP.
 

### 2016-10-19 analysis/CompareEEvsETmodel.Rmd
Working on revisions to CompareEEvsETmodel.Rmd.

@@
 
-The ashr package implements a shrinkage-based Empirical Bayes
-method for estimating the values of effects $\beta_j$ based
-on estimates of the effects ($\hat\beta_j$) and their standard errors ($s_j$).
+The ashr package implements a shrinkage-based Empirical Bayes method for
+estimating the values of effects $\beta_j$ based on estimates of the effects
+($\hat{\beta}_j$) and their standard errors ($s_j$).
 
 The default model is that the effects $\beta_j$ are identically distributed
-from a unimodal distribution $g$. 
-In particular, ashr assumes that $\beta_j$ is independent of the standard error $s_j$.
-That is,
+from a unimodal distribution $g$. In particular, ashr assumes that $\beta_j$
+is independent of the standard error $s_j$. That is,
+
 $$\beta_j | s_j \sim g(\cdot) \quad (*).$$
@@
 Here we consider the more general model:
-$$\beta_j/(s_j)^\alpha | s_j \sim g(\cdot)   \quad (**).$$
-The case $\alpha=0$ is the model (*) above. The case $\alpha=1$ assumes that 
-the *t statistics* $\betahat_j/s_j$ are identically distributed from a unimodal
-distribution.
-Under this assumption the expected size of the *unstandardized* effect $\beta_j$ depends on the standard error $s_j$: the larger $s_j$ is, the larger (in absolute value) $\beta_j$ is expected to be.
 
-By analogy with Wen and Stephens (AoAS), we refer to the model (*) as the EE model ("exchangeable effects"), and the model (**)
-as the ET model ("exchangeable T statistics model")
+$$\beta_j/s_j^\alpha | s_j \sim g(\cdot)   \quad (**).$$
 
-What might motivate the ET model? 
-We can provide two distinct motivations. First, suppose for concreteness that the effects
-$\beta_j$ reflect differences in gene expression between two conditions. One factor that affects
-$s_j$ is the variance of gene $j$ within each condition.
-One could imagine that, perhaps, genes with a larger variance within each condition are less tightly regulated, and therefore more likely to show a large difference between conditions (i.e.
-large $\beta_j$) than genes with a small variance.
-This provides
+The case $\alpha = 0$ is the model (*) above. The case $\alpha = 1$ assumes
+that the *t* statistics $\beta_j/s_j$ are identically distributed from
+a unimodal distribution. Under this assumption the expected size of the
+*unstandardized* effect $\beta_j$ depends on the standard error $s_j$; the
+larger $s_j$ is, the larger (in absolute value) $\beta_j$ is expected to be.
+
+By analogy with Wen and Stephens (AoAS), we refer to the model (*) as the
+EE model ("exchangeable effects"), and the model (**) as the ET model
+("exchangeable T statistics model").
+
+What might motivate the ET model? We can provide two distinct motivations.
+First, suppose for concreteness that the effects $\beta_j$ reflect differences
+in gene expression between two conditions. One factor that affects $s_j$ is
+the variance of gene $j$ within each condition. One could imagine that, perhaps,
+genes with a larger variance within each condition are less tightly regulated,
+and therefore more likely to show a large difference between conditions
+(*i.e.* large $\beta_j$) than genes with a small variance. This provides
 a biological motivation for the possibility that larger $s_j$ might
-correlate with larger $\beta_j$ (although of course not for the exact functional form above).
-A second motivation is more statistical: it turns out that this assumption is in some sense
-the implicit assumption made by existing methods to fdr analysis based on p values. 
-More specifically, under this alternative assumption, when 
-attempting to identify ``significant" effects, the empirical Bayes approach will rank the genes 
-in the same way as the usual $p$ values computed from $\hat\beta_j/s_j$.
+correlate with larger $\beta_j$ (although of course not for the exact
+functional form above).
 
-It is straightforward to use the ashr package to perform analysis under the ET model:
-simply specify alpha=1. (Internally, this causes ash to 
-replace betahat with the standardized betahat, $\hat\beta_j/s_j$, and the standard errors for these standarized betahat values with 1; there there is some bookkeeping to be done to make sure we return the right likelihoods and posteriors for the original beta, and not for these standardized values...
+A second motivation is more statistical: it turns out that this assumption is
+in some sense the implicit assumption made by existing methods to fdr analysis
+based on p values. More specifically, under this alternative assumption, when
+attempting to identify "significant" effects, the empirical Bayes approach
+will rank the genes in the same way as the usual $p$ values computed from
+$\hat{\beta}_j/s_j$.
+
+It is straightforward to use the ashr package to perform analysis under the
+ET model: simply specify alpha = 1. (Internally, this causes ash to replace
+betahat with the standardized betahat, $\hat{\beta}_j/s_j$, and the standard
+errors for these standarized betahat values with 1; there there is some
+bookkeeping to be done to make sure we return the right likelihoods and
+posteriors for the original beta, and not for these standardized values...
 ash takes care of this.) It is also straightforward to compare
-the two competing modelling assumptions (EE vs ET) by computing the log likelihood ratio, $log[p_{EE}(\hat\beta | s, \hat g_{EE})/p_{ET}(\hat\beta | s, \hat g_{ET})]$.
+the two competing modelling assumptions (EE vs ET) by computing the
+log-likelihood ratio,
 
-We now illustrate by a simulated example. We assume that the standard errors come
-from a gamma distribution, and then generate the effects $\beta_j$ under the ET model so that genes with bigger standard errors tend to have bigger effects.
-```{r}
+$$\log\{p_{EE}(\hat{\beta} | s, \hat{g}_{EE}) / p_{ET}(\hat{\beta} | s, \hat{g}_{ET})\}.$$
+
+We now illustrate by a simulated example. We assume that the standard errors
+come from a gamma distribution, and then generate the effects $\beta_j$ under
+the ET model so that genes with bigger standard errors tend to have bigger
+effects.
+
+```{r generate_data}
 set.seed(1234)
-  nsamp=1000
+nsamp      = 1000
 betahat.se = rgamma(nsamp,1,1)
@@
 pval       = pchisq(zscore^2,df=1,lower.tail=FALSE)
-  plot(betahat, -log(pval))
+plot(betahat,-log(pval))
 ```

### 2016-10-19 analysis/CompareEEvsETmodel.Rmd
Completed revisions of CompareEEvsETmodel.Rmd; it is working with the new version of ashr.

@@
 betahat.se = rgamma(nsamp,1,1)
-beta       = betahat.se * rnorm(nsamp) #simulate effects under the alternative assumption
+beta       = betahat.se * rnorm(nsamp)
 betahat    = rnorm(nsamp,beta,betahat.se)
@@
 pval       = pchisq(zscore^2,df=1,lower.tail=FALSE)
-plot(betahat,-log(pval))
+plot(betahat,-log(pval),pch = 20)
 ```
@@
 Here is the EE analysis.
-```{r}
+
+```{r analysis_ee}
 library(ashr)
-  ashEE.res = ash(betahat, betahat.se,method="fdr",alpha=0)
+ashEE.res = ash(betahat,betahat.se,method = "fdr",alpha = 0)
 ```
 
-And here is how we can perform the ET analysis,
-```{r}
-  ashET.res = ash(betahat, betahat.se,method="fdr", alpha=1)
+And here is how we can perform the ET analysis.
+
+```{r analysis_et}
+ashET.res = ash(betahat,betahat.se,method = "fdr",alpha = 1)
 ```
@@
 Now we compare the EE vs ET models:
-```{r}
-  ashEE.res$loglik-ashET.res$loglik
+```{r compare_analyses}
+print(ashEE.res$loglik - ashET.res$loglik)
 ```
 
-Then the log likelihood ratio is loglikEE-loglikET=`r ashEE.res$loglik-ashET.res$loglik`. This highly negative
-loglikelihood indicate that the data strongly favor
-the ET model, which is expected because the data were generated under this model.
-(One might be tempted to ask whether the log likelihood ratio is ``significant". We don't
-know how to address this question, but suggest in practice it doesn't matter: if the loglikelihood ratio is positive then use EE, if it is negative then use ET.)
+Then the log likelihood ratio is loglikEE - loglikET =
+`r ashEE.res$loglik - ashET.res$loglik`. This highly negative loglikelihood
+indicates that the data strongly favor the ET model, which is expected
+because the data were generated under this model. (One might be tempted to ask
+whether the log likelihood ratio is "significant". We don't know how to address
+this question, but suggest in practice it doesn't matter: if the loglikelihood
+ratio is positive then use EE, if it is negative then use ET.)
 
-The above illustrates these ideas on simulations from the ET model. For comparison,
-we now provide results for simulations under the EE model.
-```{r}
+The above illustrates these ideas on simulations from the ET model. For
+comparison, we now provide results for simulations under the EE model.
+
+```{r generate_data_from_ee_model}
 set.seed(1234)
-  nsamp=1000
+samp       = 1000
 betahat.se = rgamma(nsamp,1,1)
-  beta = rnorm(nsamp) #simulate effects under the original assumption
+beta       = rnorm(nsamp)
 betahat    = rnorm(nsamp,beta,betahat.se)
 zscore     = betahat/betahat.se
-  pval = pchisq(zscore^2,df=1,lower.tail=FALSE)
-  ashEE.res = ash(betahat, betahat.se,method="fdr",alpha=0)
-  ashET.res = ash(betahat, betahat.se, method="fdr",alpha=1)
-  ashEE.res$loglik-ashET.res$loglik
+pval       = pchisq(zscore^2,df = 1,lower.tail = FALSE)
+ashEE.res  = ash(betahat,betahat.se,method = "fdr",alpha = 0)
+ashET.res  = ash(betahat,betahat.se,method = "fdr",alpha = 1)
+print(ashEE.res$loglik - ashET.res$loglik)
 ```
-So here the log likelihood ratio is positive, indicating that the EE model is preferred
-(which is expected since the data were generated under that model).
+
+So here the log likelihood ratio is positive, indicating that the EE model
+is preferred (which is expected since the data were generated under that
+model).

### 2016-10-20 analysis/IPvsEM.Rmd
Working on revisions to CompareEEvsETmodel.Rmd.
Revised ash_nplme.Rmd.

@@
 
-```{r chunk-options, include=FALSE}
+```{r chunk_options, include=FALSE}
 library(REBayes)
@@
 
-# Set up simulation
+## Set up simulation
 
@@
 # Summarize the results in a table.
-cat("Average computation time (seconds):\n")
+cat("Average computation time (in seconds):\n")
 print(as.table(by(df,df[c("method","optmethod")],
@@
 
-# Results
+## Results
 
@@
 
-# Summary
+## Summary
 

### 2016-10-20 analysis/ash_npmle.Rmd
Revised ash_nplme.Rmd.

@@
 
-```{r chunk-options, include=FALSE}
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
+library(REBayes)
+library(ashr)
 source("chunk-options.R")
@@
 Here we show how we can use ash to (approximately) compute the unconstrained
-NPMLE - that is, estimate the underlying distribution $g$ by maximizing the likelihood without the unimodal constraint.
-See Koenker and Mizera, JASA 2014, for background.
+NPMLE - that is, estimate the underlying distribution $g$ by maximizing the
+likelihood without the unimodal constraint. See Koenker and Mizera, JASA 2014,
+for background.
 
-Within ash one can approximate the npmle as a mixture of uniforms
-on a dense grid of non-overlapping values - this results in a piecewise constant density with changes in the density only at the grid points.
+Within ash one can approximate the npmle as a mixture of uniforms on a dense
+grid of non-overlapping values - this results in a piecewise constant density
+with changes in the density only at the grid points.
 
-The following example comes from the REBayes vignette by Koenker and Gu.
-The underlying $g$ is a mixture of a point mass (weight 0.8) at 0 and a point mass (weight 0.2) at 2.
-```{r}
+The following example comes from the REBayes vignette by Koenker and Gu. The
+underlying $g$ is a mixture of a point mass (weight 0.8) at 0 and a point
+mass (weight 0.2) at 2.
+```{r fit_glmix_model}
 set.seed(102)
-  y <- c(rep(0,800), rnorm(200, 2)) + rnorm(1000)
-  z <- REBayes::GLmix(y)
+y <- c(rep(0,800),rnorm(200,2)) + rnorm(1000)
+z <- GLmix(y)
 ```
@@
 Now we fit the NPMLE using 'ashr' and compare with the REBayes solution.
-```{r}
-  grid = seq(from=min(z$x),to=max(z$x),length=1000)
-  k=length(grid)
-  y.ash.npmle = ashr::ash(y,1,g=ashr::unimix(pi=rep(1/(k-1),(k-1)),a=grid[-k],b=grid[-1]),method="shrink")
+```{r fit_ash_model}
+grid = seq(from = min(z$x),to = max(z$x),length = 1000)
+k    = length(grid)
+y.ash.npmle =
+  ash(y,1,g = unimix(pi=rep(1/(k-1),(k-1)),a = grid[-k],b = grid[-1]),
+      method = "shrink")
 
-  plot(z$x,cumsum(z$y)/sum(z$y),col=2,main="Estimated cdf",type="l")
-  lines(ashr::cdf.ash(y.ash.npmle,x=z$x),type="l")
-  z$logLik
-  y.ash.npmle$loglik
+# Plot the model fitted using GLmix against the fitted ash model.
+plot(z$x,cumsum(z$y)/sum(z$y),col = 2,main = "Estimated cdf",type = "l")
+lines(cdf.ash(y.ash.npmle,x = z$x),type = "l",lty = "dashed",lwd = 2)
+
+# Compare the likelihood given the GLmix model and given ash model.
+print(z$logLik,digits = 12)
+print(y.ash.npmle$loglik,digits = 12)
 ```

### 2016-10-20 analysis/makefig_FDReg.Rmd
Revised ash_nplme.Rmd.

@@
 source("chunk-options.R")
-source("../R/plot_FDReg_hist.R")
+source("../R/plot_FDReg_hist.R")t/
 source("../R/nullalthist.R")

### 2016-10-20 code/dsc-shrink/methods/ash.wrapper.R

@@ -34,7 +34,7 @@ bayes.wrapper = function(input, meta, args=NULL){
 }
 
 ash2beta_est =function(output){
-  return (list(beta_est=get_pm(output)))
+    return (list(beta_est=get_pm(output)))
 } 
 
 ash2pi0_est =function(output){

### 2016-10-21 analysis/IPvsEM.Rmd
Revised	efron.fcr.Rmd.

@@
 ```{r simulation_setup}
-timed_sims = function(ash.args,nsim = 20,nmin = 100,nmax = 1000,betasd = 1) {
+timed_sims = function(ash.args,nsim = 20,nmin = 100,nmax = 1e3,betasd = 1) {
   n = 10^seq(log10(nmin),log10(nmax),length = nsim)
@@
            timed_sims(list(method = method,optmethod = optmethod),
-                           nsim = 50,nmin = 10,nmax = 100000)))
+                           nsim = 50,nmin = 10,nmax = 1e5)))
   }

### 2016-10-21 analysis/ash_npmle.Rmd
Revised	efron.fcr.Rmd.

@@
 y.ash.npmle =
-  ash(y,1,g = unimix(pi=rep(1/(k-1),(k-1)),a = grid[-k],b = grid[-1]),
+  ash(y,1,g = unimix(pi = rep(1/(k-1),(k-1)),a = grid[-k],b = grid[-1]),
       method = "shrink")

### 2016-10-21 analysis/efron.fcr.Rmd
Revised	efron.fcr.Rmd.

--- 2016-05-25 analysis/efron.fcr.Rmd
@@
 
-```{r chunk-options, include=FALSE}
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
+library(ashr)
+library(qvalue)
+library(locfdr)
 source("chunk-options.R")
@@
 
-This example comes from Efron (2008) p16 when examining the false coverage rate (FCR). 
-I selected this example because the distribution of the non-zero effect is
-highly assymetric and not at all unimodal at zero, both issues a referee asked me
-to elaborate on. Specifically, the distribution of the non-zero effects is N(-3,1).
-Here I simulate data, and apply ash (with the "halfuniform" option to allow for asymmetric g).
+This example comes from Efron (2008) p16 when examining the false coverage rate
+(FCR). I selected this example because the distribution of the non-zero effects
+is highly assymetric and not at all unimodal at zero, both issues a referee
+asked me to elaborate on. Specifically, the distribution of the non-zero
+effects is N(-3,1). Here I simulate data, and apply ash (with the "halfuniform"
+option to allow for asymmetric g).
 
-```{r}
+```{r fit_halfuniform_model}
 set.seed(10)
-  nsamp=10000
-  altmean=-3
+nsamp   = 1e4
+altmean = -3
 mu0     = rep(0,nsamp)
@@
 comp    = rbinom(nsamp,1,0.1)
-  mu = ifelse(comp==0,mu0,mu1)
+mu      = ifelse(comp == 0,mu0,mu1)
 z       = rnorm(nsamp,mu,1)
 
-  res.ash = ashr::ash(z,1,mixcompdist="halfuniform")
-
+# Fit the model.
+res.ash = ash(z,1,mixcompdist = "halfuniform")
 ```
 
-We can also run ash with the "true" g, to allow us to compare the lfsr, lfdr etc. 
-```{r}
-  true.g = ashr::normalmix(c(0.9,0.1),c(0,-3),c(0,1))
-  res.ash.true = ashr::ash(z,1,g=true.g,fixg=TRUE)
+We can also run ash with the "true" g, to allow us to compare the lfsr, lfdr,
+etc.
+
+```{r fit_normalmix_model}
+true.g       = normalmix(c(0.9,0.1),c(0,-3),c(0,1))
+res.ash.true = ash(z,1,g = true.g,fixg = TRUE)
 ```
 
-Here we can see how the partition of $z$ scores compares with the truth. Note the effect of the unimodal assumption is to extend the inferred alternative distribution toward 0.
-```{r}
+Here we can see how the partition of $z$ scores compares with the truth. Note
+the effect of the unimodal assumption is to extend the inferred alternative
+distribution toward 0. Here, `nullalthist` plots a histogram of z scores,
+highlighting the alternative distribution of z scores that is implied by
+the localfdr values lfdr.
+
+```{r plot_zscore_hist}
 source("../R/nullalthist.R")
-  par(mfcol=c(2,1))
-  nullalthist(z,lfdr=res.ash.true$lfdr,main="true partition")
-  nullalthist(z,lfdr=res.ash$lfdr,main="inferred partition")
+par(mfcol = c(2,1),cex.main = 1,font.main = 1)
+nullalthist(z,lfdr = get_lfdr(res.ash.true),main = "true partition (res.ash.true)")
+nullalthist(z,lfdr = get_lfdr(res.ash),main = "inferred partition (res.ash)")
 ```
 
-Comparing the inferred Posterior Means, lfdr, and lfsr with the true values of these quantities, we find reassuringly good correspondence.
-```{r}
-  par(mfcol=c(1,1))
-  plot(res.ash.true$PosteriorMean,res.ash$PosteriorMean,xlab="Truth", ylab="ash.hu",main="Posterior Mean (inferred vs truth)",xlim=c(-6,1),ylim=c(-6,1))
-  abline(a=0,b=1,col=2,lwd=3)
+Comparing the inferred posterior means, lfdr, and lfsr with the true values of
+these quantities, we find reassuringly good correspondence.
 
-    plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth", ylab="ash.hu",main="lfdr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
-  abline(a=0,b=1,col=2,lwd=3)
+```{r scatterplots}
+par(mfrow = c(1,3))
+plot(get_pm(res.ash.true),get_pm(res.ash),xlab = "Truth",ylab = "ash.hu",
+     main = "Posterior Mean (inferred vs truth)",pch = 20,cex = 0.6,
+     xlim = c(-6,1),ylim = c(-6,1))
+abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")
 
-  plot(res.ash.true$lfsr,res.ash$lfsr,xlab="Truth", ylab="ash.hu",main="lfsr (inferred vs truth)",xlim=c(0,1),ylim=c(0,1))
-  abline(a=0,b=1,col=2,lwd=3)
+plot(get_lfdr(res.ash.true),get_lfdr(res.ash),xlab = "Truth", ylab = "ash.hu",
+     main = "lfdr (inferred vs truth)",pch = 20,cex = 0.6,
+     xlim = c(0,1),ylim = c(0,1))
+abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")
 
+plot(get_lfsr(res.ash.true),get_lfsr(res.ash),xlab = "Truth", ylab = "ash.hu",
+     main = "lfsr (inferred vs truth)",pch = 20,cex = 0.6,
+     xlim = c(0,1),ylim = c(0,1))
+abline(a = 0,b = 1,col = "orangered",lwd = 2,lty = "dotted")
 ```
 
-We can also run qvalue and locfdr.
-We see that locfdr perhaps performs a bit better than ash for the decomposition here, but the estimated local fdrs are pretty similar.
-Here qvalue does less well because of the asymmetry which we didn't take account of.
-```{r}
-library(qvalue)
-library(locfdr)
+## Comparison with qvalue and locfdr
+
+We can also run qvalue and locfdr. We see that locfdr perhaps performs a bit
+better than ash for the decomposition here, but the estimated local fdrs are
+pretty similar. Here qvalue does less well because of the asymmetry which it
+didn't take account of.
+
+```{r comparison_to_other_methods}
 res.locfdr = locfdr(z,nulltype = 0)
-res.qvalue = qvalue(p=pchisq(z^2,df=1,lower.tail = FALSE))  
-#res.pos.qvalue = qvalue(p=pchisq(z[z>0]^2,df=1,lower.tail = FALSE))  
-#res.neg.qvalue = qvalue(p=pchisq(z[z<0]^2,df=1,lower.tail = FALSE))  
-#df1= data.frame(Truth = res.ash.true$lfdr,Estimate=res.locfdr$fdr,Method="locfdr")
-#df2= data.frame(Truth = res.ash.true$lfdr,Estimate=res.ash$lfdr,Method="ash.hu")
-#df3 =data.frame(Truth = res.ash.true$lfdr,Estimate=res.qvalue$lfdr,Method="qvalue")
-#df4 =data.frame(Truth = res.ash.true$lfdr[z<0],Estimate=res.neg.qvalue$lfdr,Method="qvalue (-)")
-#df5 =data.frame(Truth = res.ash.true$lfdr[z>0],Estimate=res.pos.qvalue$lfdr,Method="qvalue (+)")
+res.qvalue = qvalue(p = pchisq(z^2,df = 1,lower.tail = FALSE))
 
-#df = rbind(df1,df2,df3,df4,df5)
-#ggplot2::qplot(data=df,x=Truth,y=Estimate,facets= .~Method)
+par(mfrow = c(1,3))
+plot(get_lfdr(res.ash.true),get_lfdr(res.ash),pch = 20,cex = 0.6,
+     xlab = "Truth (lfdr)",ylab = "ash.hu",main = "ash.hu",
+     xlim = c(0,1),ylim = c(0,1))
+abline(a = 0,b = 1,col = "orangered",lty = "dotted",lwd = 2)
 
-par(mfrow=c(1,3))
-plot(res.ash.true$lfdr,res.ash$lfdr,xlab="Truth (lfdr)", ylab="ash.hu",main="ash.hu",xlim=c(0,1),ylim=c(0,1))
-abline(a=0,b=1,col=2,lwd=3)
-plot(res.ash.true$lfdr,res.locfdr$fdr,xlab="Truth (lfdr)", ylab="Estimate",main="locfdr",xlim=c(0,1),ylim=c(0,1))
-abline(a=0,b=1,col=2,lwd=3)
-plot(res.ash.true$lfdr,res.qvalue$lfdr,xlab="Truth (lfdr)", ylab="Estimate",main="qvalue",xlim=c(0,1),ylim=c(0,1))
-abline(a=0,b=1,col=2,lwd=3)
+plot(get_lfdr(res.ash.true),res.locfdr$fdr,pch = 20,cex = 0.6,
+     xlab = "Truth (lfdr)",ylab = "Estimate",main = "locfdr",
+     xlim = c(0,1),ylim = c(0,1))
+abline(a = 0,b = 1,lwd = 2,col = "orangered",lty = "dotted")
 
+plot(get_lfdr(res.ash.true),res.qvalue$lfdr,pch = 20,cex = 0.6,
+     xlab = "Truth (lfdr)",ylab = "Estimate",main = "qvalue",
+     xlim = c(0,1),ylim = c(0,1))
+abline(a = 0,b = 1,lwd = 2,col = "orangered",lty = "dotted")
 ```
 
-# Uniform tail curtails Credible Intervals
+## Uniform tail curtails Credible Intervals
 
-The following plot compares the (symmetric-tail) 95% CIs from ash (red) for the "significant" observations with Bayes rule (green), similar to Figure 8 from Efron.
-Note that the lower 97.5% point is pretty accurate, but the upper 97.5% point is curtailed - presumably due, at least in part, to the short tails of the uniform mixture.
-```{r}
-  CImatrix= ashr::ashci(res.ash,level=0.95)
-  BayesComparePlot=function(CImatrix,altmean=-3,...){
-    plot(z,mu,xlim=c(-8,0),...)
-    points(z[CImatrix[,1]],CImatrix[,5],col="red")
-    points(z[CImatrix[,1]],CImatrix[,4],col="red")
+The following plot compares the (symmetric-tail) 95% CIs from ash (red) for
+the "significant" observations with Bayes rule (green), similar to Figure 8
+from Efron. Note that the lower 97.5% point is pretty accurate, but the
+upper 97.5% point is curtailed - presumably due, at least in part, to the
+short tails of the uniform mixture.
 
-    fdr = 0.9*dnorm(z)/(0.9*dnorm(z)+0.1*dnorm(z,altmean,sqrt(2)))
-    o=order(z)
-    upper = ifelse(fdr[o]<0.025,(z[o]+altmean)/2+qnorm(0.975+fdr[o])/sqrt(2),0)
-    lines(z[o],upper,col="green",lwd=2)
-    lines(z[o],(z[o]+altmean)/2-qnorm(0.975)/sqrt(2),col="green",lwd=2)
-    abline(v=max(z[fdr<0.05]))
+```{r plot_credints}
+CImatrix = ashci(res.ash,level = 0.95)
+BayesComparePlot = function(CImatrix, altmean = -3,...) {
+  plot(z,mu,xlim = c(-8,0),pch = 20,cex = 0.6,...)
+  points(z,CImatrix[,1],col = "orangered",cex = 0.6,pch = 20)
+  points(z,CImatrix[,2],col = "orangered",cex = 0.6,pch = 20)
+
+  fdr   = 0.9*dnorm(z)/(0.9*dnorm(z) + 0.1*dnorm(z,altmean,sqrt(2)))
+  o     = order(z)
+  upper = ifelse(fdr[o] < 0.025,
+                 (z[o] + altmean)/2 + qnorm(0.975 + fdr[o])/sqrt(2),0)
+  lines(z[o],upper,col = "limegreen",lwd = 2)
+  lines(z[o],(z[o] + altmean)/2 - qnorm(0.975)/sqrt(2),col = "limegreen",lwd = 2)
+  abline(v = max(z[fdr < 0.05]))
 }
-  BayesComparePlot(CImatrix,main="CIs for highly asymmetric and non-unimodal-at-zero data")
+par(mfrow = c(1,1))
+BayesComparePlot(CImatrix,
+  main = "CIs for highly asymmetric and non-unimodal-at-zero data")
 ```
@@
 
-Although not a focus of the paper, ash does have an option to do variational inference for the mixture components (with a Dirichlet prior). In practice this approach usually
-ends up spreading the posterior mass up more among the mixture components. It
-seemed plausible that this might lead to slightly less extreme tail behaviour than
-above (because the model will put a little more weight on the uniforms with larger variance, which are essentially set to zero in the above).
+Although not a focus of the paper, ash does have an option to do variational
+inference for the mixture components (with a Dirichlet prior). In practice
+this approach usually ends up spreading the posterior mass up more among the
+mixture components. It seemed plausible that this might lead to slightly
+less extreme tail behaviour than above (because the model will put a little
+more weight on the uniforms with larger variance, which are essentially set
+to zero in the above).
 
-```{r}
-  res.ash.VB = ashr::ash(z,1,mixcompdist="halfuniform",optmethod="mixVBEM")
-  CImatrix.VB= ashr::ashci(res.ash.VB,level=0.95)
+```{r vb_model_fitting}
+res.ash.VB  = ash(z,1,mixcompdist = "halfuniform",optmethod = "mixVBEM")
+CImatrix.VB = ashci(res.ash.VB,level = 0.95)
 ```
 
-Again, we can compare results with Bayes rule
-```{r}
-  BayesComparePlot(CImatrix.VB, main="CIs for highly asymmetric and non-unimodal-at-zero data \n Variational Version")
+Again, we can compare results with Bayes rule:
+
+```{r plot_vb_credints}
+BayesComparePlot(CImatrix.VB,
+  main = paste("CIs for highly asymmetric and non-unimodal-at-zero data\n",
+              "Variational Version"))
 ```

### 2016-10-21 analysis/estimate_mode.Rmd
Revised	efron.fcr.Rmd.
Revised estimate_mode.Rmd.

@@
 
+The purpose of this analysis is to check the performance of the non-zero mode
+option.
+
 First, we load the necessary libraries, and specify settings for
@@
 ```
-
-The purpose of this analysis is to check the performance of the non-zero mode
-option.
 
@@
 
-```{r}
+```{r evaluate_mode_estimates}
 check_mode = function(betahat, sebetahat, mixcompdist) {
@@
 z = rnorm(1000) + 3
-print(check_mode(z,1,mixcompdist="uniform"))
-print(check_mode(z,1,mixcompdist="normal"))
-print(check_mode(z,1,mixcompdist="halfuniform"))
+print(check_mode(z,1,mixcompdist = "uniform"))
+print(check_mode(z,1,mixcompdist = "normal"))
+print(check_mode(z,1,mixcompdist = "halfuniform"))
 ```
+
+An additional experiment:
 
@@
 
-  z.ash.hu2 = ash(betahat-0.2,0.1,mixcompdist="halfuniform",outputlevel=4,method="shrink")
-  z.ash.pu2 = ash(betahat-0.1,0.1,mixcompdist="+uniform",outputlevel=4,method="shrink")
-
-  z.ash.hu.nzm = ash(betahat,0.1,mixcompdist="halfuniform",mode="estimate",method="shrink")
-  z.ash.hu$loglik
-  z.ash.hu.nzm$loglik
-
-#  check_mode(betahat,.1,mixcompdist="halfuniform")
+z.ash.hu2    = ash(betahat - 0.2,0.1,mixcompdist = "halfuniform",outputlevel = 4,
+                   method = "shrink")
+z.ash.pu2    = ash(betahat - 0.1,0.1,mixcompdist = "+uniform",outputlevel = 4,
+                   method = "shrink")
+z.ash.hu.nzm = ash(betahat,0.1,mixcompdist = "halfuniform",mode = "estimate",
+                   method = "shrink")
+print(z.ash.hu$loglik,digits = 8)
+print(z.ash.hu.nzm$loglik,digits = 8)
 ```
 
-Note to self: check the normal version works too; check nonzeromodeEMobj  as it doesn't seem needed.
+*Note to self:* check that the normal version works too; check
+nonzeromodeEMobj as it doesn't seem needed.
 
-## Session information.
+## Session information
 

### 2016-10-22 analysis/check_mixfdr_lownoise.Rmd
Revised check_mixfdr_lownoise.Rmd---not much to do there.

@@
 
-```{r chunk-options, include=FALSE}
+Here we check that mixfdr actually does OK at fitting the bimodal distribution
+when noise is small compared with bimodality.
+
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
 source("chunk-options.R")
@@
 
-Here we check that mixfdr actually does OK at fitting the bimodal distribution when 
-noise is small compared with bimodality.
+Simulate bimodal data with well-separated peaks. The histogram shows the
+multi-modal nature of the underlying g is clear.
 
-First simulate bimodal data with well-separated peaks. The histogram
-shows the multi-modal nature of the underlying g is clear.
-```{r}
+```{r sim_data}
 source("../code/dsc-shrink/datamakers/datamaker.R")
-  sim.bimodal=rnormmix_datamaker(
-              args=list(g=normalmix(c(0.5,0.5),c(-5,5),c(1,1)),
-              min_pi0=0.4,
-              max_pi0=0.4,
-              nsamp=1000,
-              betahatsd=1)
-            )
-  hist(sim.bimodal$input$betahat,nclass=20,xlim=c(-10,10),prob=TRUE,main="histogram of simulated betahat values")
+sim.bimodal =
+  rnormmix_datamaker(args = list(g = normalmix(c(0.5,0.5),c(-5,5),c(1,1)),
+                                 min_pi0   = 0.4,
+                                 max_pi0   = 0.4,
+                                 nsamp     = 1000,
+                                 betahatsd = 1))
+hist(sim.bimodal$input$betahat,nclass = 20,xlim = c(-10,10),prob = TRUE,
+     main="histogram of simulated betahat values")
 ```
 
-Now run mixfdr on it. Note that the fit captures the multi-modal nature of g, 
-although mixfdr continues to overestimate pi0.
-```{r}
+Now run mixfdr on the simulated data. Note that the fit captures the
+multi-modal nature of g, although mixfdr continues to overestimate pi0.
+
+```{r fit_models}
 source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
-  sim.bimodal.mixfdroutput = mixfdr.wrapper(sim.bimodal$input,args=list(theonull=TRUE))
+sim.bimodal.mixfdroutput =
+  mixfdr.wrapper(sim.bimodal$input,args = list(theonull = TRUE))
 g = mixfdr2fitted.g(sim.bimodal.mixfdroutput)$fitted.g
-  x=seq(-10,10,length=100)
-  plot(ecdf(sim.bimodal$meta$beta),main="Cdf of true beta: empirical (black) and fitted (red)")
-  lines(x,mixcdf(g,x),col=2)
+x = seq(-10,10,length = 100)
+par(cex.main = 1,font.main = 1)
+plot(ecdf(sim.bimodal$meta$beta),cex = 0.6,
+     main = "CDF of true beta: empirical (black) and fitted (red)")
+lines(x,mixcdf(g,x),lwd = 2,lty = "dotted",col = "orangered")
 ```

### 2016-10-25 analysis/metaplot_examples.Rmd
add some plot functions for talks

---
title: "DDD plots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simulate Z scores

```{r}
library(rmeta)
library(ashr)
n=50
set.seed(1)
beta = c(rep(0,n/2),rnorm(n/2,0,2))
betahat = beta+rnorm(n)
beta.ash=ashr::ash(betahat,1)
o = order(betahat)
```

# Meta Plots

```{r unshrunk}
rmeta::metaplot(betahat[o],rep(1,n),ylab="",xlab="",xlim=c(-6,6),colors = meta.colors(box=c(rep("red",n/2),rep("black",n/2))[o]))
```

```{r shrunk}
rmeta::metaplot(get_pm(beta.ash)[o],get_psd(beta.ash)[o],ylab="",xlab="",xlim=c(-6,6),colors = meta.colors(box=c(rep("red",n/2),rep("black",n/2))[o]))
```

### 2016-10-25 analysis/index.Rmd
update meta-analysis plot example

--- 2016-05-25 analysis/index.Rmd
@@
 
-*  [Illustrate Zero Assumption](makefig_FDReg.html)
+*  [Figures illustrating FDR](makefig_FDReg.html)
 *  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
 *  [Adaptive Shrinkage](adaptive_shrinkage.html)
+*  [Adaptive Shrinkage: shrunken interval estimates](metaplot_examples.html)
+
 *  [Comparing models: EE vs ET models](CompareEEvsETmodel.html)
@@
 *  [Comparison of methods, and example of Credible Interval behaviour when UA does not hold](efron.fcr.html)
+*  [Using ash to fit the NPMLE; comparison with REBayes](ash_npmle.html)
 

### 2016-12-06 analysis/makefig_FDReg.Rmd
Working on makefig_FDRreg.Rmd.

--- 2016-10-19 analysis/makefig_FDReg.Rmd
@@
 
-```{r add_label_here}
+```{r qvalue_analysis}
 
@@
 
-#roughly compute a local fdr for qvalue (to aid the plotting of the decomposition in zscore space)
-temp=hist(pval,nclass=50)
-bin_fdr = res.qvalue$pi0/temp$density #compute local fdr in each bin of histogram
+# Roughly compute a local fdr for qvalue (to aid the plotting of the
+# decomposition in zscore space).
+temp     = hist(pval,nclass = 50)
+bin_fdr  = res.qvalue$pi0/temp$density # compute local fdr in each bin of histogram
 qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
-qval_fdr = pmin(1,qval_fdr) #threshold at 1, so fdr never bigger than 1
+qval_fdr = pmin(1,qval_fdr) # threshold at 1, so fdr never bigger than 1
 qval_fdr = qvalue::lfdr(pval)
 ```
+Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:
 
-Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:
-```{r}
-#pdf("../figures/FDReg_hist.pdf",width=6,height=3)
-#par(mai=c(0.5,0.5,0.2,0.2),mgp = c(3, 0.5, 0))
-#layout(matrix(1:2,ncol=2,byrow=TRUE))
-plot_FDReg_hist(pval,res.qvalue$pi0,type=1,title="p values",cex.axis=0.8,cex.main=0.8)
+```{r plot_pvalues}
+plot_FDReg_hist(pval,res.qvalue$pi0,type = 1,title = "p values",
+                cex.axis = 0.8,cex.main = 0.8)
 
@@
 Now the decomposition:
-```{r}
-plot_FDReg_hist(pval,res.qvalue$pi0,type=4,cex.axis=0.8,yaxt='n',textsize=0.9,cex.main=0.8,title="Decomposition into null/alternative")
-axis(side=2, labels=FALSE,tck=-0.01)
+
+```{r decompose_pvalues}
+plot_FDReg_hist(pval,res.qvalue$pi0,type = 4,cex.axis = 0.8,yaxt = 'n',
+                textsize = 0.9,cex.main = 0.8,
+                title = "Decomposition into null/alternative")
+axis(side = 2,labels = FALSE,tck = -0.01)
 ```
 
-############################## decomp_ZA.pdf #############################################
 ```{r decomp_ZA, dev='pdf', fig.width=6, fig.height=6, crop=TRUE}
-#pdf("../figures/decomp_ZA.pdf",width=6,height=6)
-layout(matrix(1:12,ncol=3,byrow=FALSE))
-plotlabel=function(label,cex=1.5){
-  plot(0,0,type="n",axes=FALSE,xlab="",ylab="")
-  text(0,0,label,cex=cex)
+layout(matrix(1:12,ncol = 3,byrow = FALSE))
+plotlabel = function(label,cex = 1.5){
+  plot(0,0,type = "n",axes = FALSE,xlab = "",ylab = "")
+  text(0,0,label,cex = cex)
 }
@@
 
-#p value histograms
+# p value histograms
 altnullhist(pval,qval_fdr,main="p values",ncz=50,xaxt='n',cex.axis=0.8)
-#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
 axis(side=1, labels=FALSE,tck=-0.01)
-#mtext(side=3,"p values",line=1)
 altnullhist(pval,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
@@
 axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
-#mtext(side=1,"p values",line=2)
 
-#z score histograms
+# z score histograms
 nullalthist(zscore,qval_fdr,main="z scores",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
 axis(side=1, labels=FALSE,tck=-0.01)
-#mtext(side=3,"z scores",line=1)
 nullalthist(zscore,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
@@
 axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
-#mtext(side=1,"z scores",line=2)
-
-#dev.off()
-
 ```

### 2016-12-07 analysis/IPvsEM.Rmd
I hid (long) output when knitting IPvsEM.Rmd.

@@
 
-```{r model_fitting}
+```{r model_fitting, results='hide'}
 df = data.frame()
@@
 cat("\n")
+```
 
-# Summarize the results in a table.
+Summarize the model fitting results in a table.
+
+```{r model_fitting_summary}
 cat("Average computation time (in seconds):\n")

### 2016-12-07 code/dsc-shrink/run_dsc.R

@@ -4,12 +4,13 @@ sessionInfo()
 
 dsc_shrink=new_dsc("shrink","../../output/dsc-shrink-files")
 source("add_named_scenarios.R")
-add_named_scenarios(dsc_shrink,c("spiky","near-normal","flat-top","skew","big-normal","bimodal","near-normal-nn"))
+add_named_scenarios(dsc_shrink,c("spiky","near-normal","flat-top","skew","big-nqormal","bimodal","near-normal-nn"))
 add_named_scenarios(dsc_shrink,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),min_pi0=0,max_pi0=0,suffix="-nn")
 
 source("add_methods.R")
 
-add_output_parser(dsc_shrink,"ash2beta",ash2beta_est,"ash_output","beta_est_output")
+add_output_parser(dsc_shrink,"ash2beta",ash2beta_est,"ash_output",
+                  "beta_est_output")
 add_output_parser(dsc_shrink,"mixfdr2beta",mixfdr2beta_est,"mixfdr_output","beta_est_output")
 
 add_output_parser(dsc_shrink,"ash2pi0",ash2pi0_est,"ash_output","pi0_est_output")
@@ -19,8 +20,6 @@ add_output_parser(dsc_shrink,"qvalue2pi0",qvalue2pi0_est,"qvalue_output","pi0_es
 
 add_output_parser(dsc_shrink,"ash2fitted.g",ash2fitted.g,"ash_output","g_output")
 add_output_parser(dsc_shrink,"mixfdr2fitted.g",mixfdr2fitted.g,"mixfdr_output","g_output")
-
-
 
 source("score.R")
 add_score(dsc_shrink,score,"beta_err","beta_est_output")
@@ -33,8 +32,5 @@ add_score(dsc_shrink,score_betahat,"betahat","mixfdr_output") #just extracts the
 add_score(dsc_shrink,score_lfsr,"lfsr","ash_output") #just extracts the lfsr
 add_score(dsc_shrink,score_lfdr,"lfdr","ash_output") #just extracts the lfdr
 
-
 res=run_dsc(dsc_shrink)
 save(res,dsc_shrink,file="../../output/dsc-shrink-files/res.RData")
-
-

### 2016-12-08 analysis/make_GOODPOOR_figs.Rmd
Made a few small adjustments to make_GOODPOOR_figs.Rmd.

@@
 
+First, we load the necessary libraries, and specify settings for displaying 
+the plots in the rendered document.
+
 ```{r chunk-options, include=FALSE}
+library(ashr)
+library(qvalue)
+library(locfdr)
+library(mixfdr)
 source("chunk-options.R")
+source("../R/plot_FDReg_hist.R")
 ```
 
-First perform a simple simulated example to illustrate high and low signal.
-True values of beta simulated from $(0.5 N(0,2^2) + 0.5 \delta_0)$.
-```{r}
+Create a simple simulated data set to illustrate high and low signal. True 
+values of beta simulated from $(0.5 N(0,2^2) + 0.5 \delta_0)$.
+
+```{r sim_data}
 ntest = 10000
 set.seed(112)
-require(ashr)
-require(qvalue)
-require(locfdr)
-require(mixfdr)
-source("../R/plot_FDReg_hist.R")
 
@@
 beta      = ifelse(null_alt==1,beta,0)
-GOOD= 1:(ntest/2)
+GOOD      = 1:(ntest/2)
 sebetahat = rep(1,ntest)
@@
 
-Show how poor precision observations dilute good ones:
-```{r, GOODPOOReg_hist, dev='pdf',fig.width=6.5, fig.height=3}
-par(mai=c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
-layout(matrix(1:3,ncol=3,byrow=TRUE))
-plot_FDReg_hist(pval[GOOD],1,type=1,title="Good-precision observations",ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
-plot_FDReg_hist(pval[-GOOD],1,type=1,title="Poor-precision observations",ylab="",nc=20,yaxt='n',cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
-axis(side=2, labels=FALSE,tck=-0.01)
-plot_FDReg_hist(pval,1,type=1,title="Combined",yaxt='n',ylab="",nc=20,cex.axis=1,cex.main=1.2,ylim=c(0,2.5))
-axis(side=2, labels=FALSE,tck=-0.01)
+Show how poor precision observations dilute good ones.
+
+```{r GOODPOOReg_hist, dev='pdf', fig.width=6.5, fig.height=3}
+par(mai = c(0.3,0.3,0.2,0.2),mgp = c(3, 0.5, 0))
+layout(matrix(1:3,ncol = 3,byrow = TRUE))
+plot_FDReg_hist(pval[GOOD],1,type = 1,title = "Good-precision observations", 
+                ylab = "",nc = 20,cex.axis = 1,cex.main = 1.2,ylim = c(0,2.5))
+plot_FDReg_hist(pval[-GOOD],1,type = 1,
+                title = "Poor-precision observations",ylab = "",nc = 20,
+                yaxt = 'n',cex.axis = 1,cex.main = 1.2,ylim = c(0,2.5))
+axis(side = 2, labels = FALSE,tck = -0.01)
+plot_FDReg_hist(pval,1,type = 1,title = "Combined",yaxt = 'n',ylab = "",
+                nc = 20,cex.axis = 1,cex.main = 1.2,ylim = c(0,2.5))
+axis(side = 2, labels = FALSE,tck = -0.01)
 ```

### 2016-12-09 analysis/make_GOODPOOR_figs.Rmd
Rendered make_GOODPOOR_figs.html after some minor alterations throughout.

@@
 
-First, we load the necessary libraries, and specify settings for displaying 
-the plots in the rendered document.
+First, we load the necessary libraries and function definitions, and specify 
+settings for displaying the plots in the rendered document.
 
@@
 library(mixfdr)
+library(ggplot2)
 source("chunk-options.R")
@@
 null_alt  = rbinom(ntest,1,0.5)
-beta      = rnorm(ntest,0,sd=1)
-beta      = ifelse(null_alt==1,beta,0)
+beta      = rnorm(ntest,0,sd = 1)
+beta      = ifelse(null_alt == 1,beta,0)
 GOOD      = 1:(ntest/2)
@@
 zscore  = betahat/sebetahat
-pval    = pchisq(zscore^2,df=1,lower.tail=F)
+pval    = pchisq(zscore^2,df = 1,lower.tail = F)
 ```
@@
 
-Apply different methods
-```{r}
+Apply alternative methods to the same data set:
+
+```{r other_methods}
 res.qvalue = qvalue(pval)
-res.locfdr = locfdr(zscore,nulltype=0,plot=0)
-res.ash = ash(betahat,sebetahat,method="fdr",outputlevel=4)
+res.locfdr = locfdr(zscore,nulltype = 0,plot = 0)
+res.ash    = ash(betahat,sebetahat,method = "fdr",outputlevel = 4)
 
 res.qvalue.good = qvalue(pval[GOOD])
-res.locfdr.good = locfdr(zscore[GOOD],nulltype=0,plot=0)
-res.ash.good = ash(betahat[GOOD],sebetahat[GOOD],method="fdr")
+res.locfdr.good = locfdr(zscore[GOOD],nulltype = 0,plot = 0)
+res.ash.good    = ash(betahat[GOOD],sebetahat[GOOD],method = "fdr")
 ```
 
+Compare the ash's accuracy against the alternative approaches.
+
 ```{r GOODPOOReg_scatter, dev='pdf', fig.width=6.5, fig.height=3}
-res = rbind(data.frame(x= res.qvalue.good$qvalues, y = res.qvalue$qvalues[GOOD],type="qvalue"),
-            data.frame(x=res.locfdr.good$fdr,y=res.locfdr$fdr[GOOD],type='locfdr'), 
-            data.frame(x=get_lfsr(res.ash.good),y=get_lfsr(res.ash)[GOOD],type="ashr"))
+res = rbind(data.frame(x    = res.qvalue.good$qvalues,
+                       y    = res.qvalue$qvalues[GOOD],
+                       type = "qvalue"),
+            data.frame(x    = res.locfdr.good$fdr,
+                       y    = res.locfdr$fdr[GOOD],
+                       type = 'locfdr'), 
+            data.frame(x    = get_lfsr(res.ash.good),
+                       y    = get_lfsr(res.ash)[GOOD],
+                       type = "ashr"))
 
-library("ggplot2")
-pp= ggplot(data=res,aes(x,y)) +geom_point(shape=1) +
+pp = ggplot(data = res,aes(x,y)) + geom_point(shape = 1) +
   facet_grid(. ~ type) +
@@
   ylab("Analysing combined data")
-
-print(pp +scale_y_continuous(limits=c(0,1)) +
-        scale_x_continuous(limits=c(0,1))  +
-        coord_equal(ratio=1))
-
+print(pp + scale_y_continuous(limits = c(0,1)) +
+      scale_x_continuous(limits = c(0,1))  +
+      coord_equal(ratio = 1))
 ```
 
+Compare the LFSR against the p-values with different prior choices, and at
+different levels of precision in the observations.
+
 ```{r lfsr_vs_pval_GOODPOOR, dev='pdf', fig.width=6.5, fig.height=3}
-make_df_for_pval=function(ash,method="default"){
-  df = data.frame(p=pnorm(-abs(ash$data$x/ash$data$s)),lfsr=get_lfsr(ash),s=ash$data$s,method=method)
+make_df_for_pval = function(ash,method = "default")
+  data.frame(p      = pnorm(-abs(ash$data$x/ash$data$s)),
+             lfsr   = get_lfsr(ash),
+             s      = ash$data$s,
+             method = method)
+
+plot_pval_vs_lfsr=function(df,plot.it=TRUE){
+  if (length(unique(df$s)) > 2) {
+    df$s = log(df$s)
+  } else { 
+    df$s = as.factor(df$s)
   }
 
-plot_pval_vs_lfsr=function(df,plot.it=TRUE){
-  if(length(unique(df$s))>2){df$s = log(df$s)} else{ df$s=as.factor(df$s)}
+  p = ggplot(df,aes(x = p,y = lfsr,color = s)) + geom_point() +
+      facet_grid(. ~ method) + xlim(c(0, 0.025)) + xlab("p value") + 
+      ylab("lfsr")
 
-  p=ggplot(df,aes(x=p,y=lfsr,color=s)) + geom_point() +facet_grid(. ~ method) +xlim(c(0, 0.025)) + xlab("p value") + ylab("lfsr")
-
-  if(length(unique(df$s))>2){
-    p = p+scale_colour_gradient2(midpoint = 1, low="blue", mid="white", high="red", space="Lab")
-  } 
-  if(plot.it){print(p)}
+  if (length(unique(df$s)) > 2)
+    p = p + scale_colour_gradient2(midpoint = 1,low = "blue",mid = "white", 
+                                   high = "red",space = "Lab")
+  if (plot.it)
+    print(p)
   return(p)
@@
 
-res.ash.ET = ash(betahat,sebetahat,method="fdr",alpha=1,mixcompdist="normal",outputlevel=4)
-p=plot_pval_vs_lfsr(rbind(
-  make_df_for_pval(res.ash,method="Default prior (alpha=0)"),
-  make_df_for_pval(res.ash.ET,method="p-value prior (alpha=1)")))
-print(p+ theme(axis.text.x = element_text(size = 8,angle=45)) + 
-        scale_colour_manual(values=c("blue","red")))
-
+res.ash.ET = ash(betahat,sebetahat,method = "fdr",alpha = 1,
+                 mixcompdist = "normal",outputlevel = 4)
+p = plot_pval_vs_lfsr(rbind(
+  make_df_for_pval(res.ash,method = "Default prior (alpha=0)"),
+  make_df_for_pval(res.ash.ET,method = "p-value prior (alpha=1)")))
+print(p + theme(axis.text.x = element_text(size = 8,angle = 45)) + 
+      scale_colour_manual(values = c("blue","red")))
 ```
 
+Now plot one at a time.
+
 ```{r  lfsr_vs_pval_GOODPOOR_single, dev='pdf', fig.width=4.5, fig.height=3}
-#now plot one at a time
-p=plot_pval_vs_lfsr(make_df_for_pval(res.ash,method="ash"))
-print(p+ theme(axis.text.x = element_text(size = 8,angle=45)))
+p = plot_pval_vs_lfsr(make_df_for_pval(res.ash,method =" ash"))
+print(p + theme(axis.text.x = element_text(size = 8,angle = 45)))
 
-p=plot_pval_vs_lfsr(make_df_for_pval(res.ash.ET,method="ash (p-value prior)"))
-print(p+ theme(axis.text.x = element_text(size = 8,angle=45)))
+p = plot_pval_vs_lfsr(make_df_for_pval(res.ash.ET,method = "ash (p-value prior)"))
+print(p + theme(axis.text.x = element_text(size = 8,angle = 45)))
 ```
@@
 Compare the log-likelihoods:
-```{r}
+
+```{r compare_ash_likelihoods}
 res.ash.ET$logLR
@@
 
-Plot number of true positives against false positives
-```{r tp-fp, dev='pdf', fig.width=4.5, fig.height=3}
-  df0 = cbind(make_df_for_pval(res.ash,method="Default prior (alpha=0)"),beta=beta)
-  df1 = cbind(make_df_for_pval(res.ash.ET,method="p-value prior (alpha=1)"),beta=beta)
-  #false positives and true positives
-  fp = function(df){cumsum((df$beta==0)[order(df$lfsr,decreasing = FALSE)])}
-  tp = function(df){cumsum((df$beta!=0)[order(df$lfsr,decreasing = FALSE)])}
-  df.fptp = rbind(data.frame(method="ash lfsr", tp = tp(df0), fp = fp(df0)),
-             data.frame(method="p values", tp = tp(df1), fp = fp(df1)))
+Plot number of true positives against false positives (an "ROC curve"):
 
-  ggplot(df.fptp,aes(x=fp,y=tp,col=method)) + geom_line()  + xlab("False positives") + ylab("True positives") + xlim(c(0,100)) + ylim(c(0,400)) + scale_colour_manual(values=c("black","green")) 
+```{r tp_vs_fp, dev='pdf', fig.width=4.5, fig.height=3}
+df0 = cbind(make_df_for_pval(res.ash,method = "Default prior (alpha=0)"), 
+            beta = beta)
+df1 = cbind(make_df_for_pval(res.ash.ET,method = "p-value prior (alpha=1)"),
+            beta = beta)
+
+# False positives (fp) and true positives (tp).
+fp = function(df) cumsum((df$beta == 0)[order(df$lfsr,decreasing = FALSE)])
+tp = function(df) cumsum((df$beta != 0)[order(df$lfsr,decreasing = FALSE)])
+df.fptp = rbind(data.frame(method = "ash lfsr",tp = tp(df0),fp = fp(df0)),
+                data.frame(method = "p values",tp = tp(df1),fp = fp(df1)))
+ggplot(df.fptp,aes(x = fp,y = tp,col = method)) + geom_line() + 
+  xlab("False positives") + ylab("True positives") + xlim(c(0,100)) + 
+  ylim(c(0,400)) + scale_colour_manual(values = c("black","green")) 
 ```

### 2016-12-09 analysis/makefig_FDReg.Rmd
knit makefig_FDReg.Rmd (note that this is in RStudio 1.0.44).

@@
 ```{r chunk-options, include=FALSE}
-require(ashr)
-require(qvalue)
-require(locfdr)
-require(mixfdr)
+library(ashr)
+library(qvalue)
+library(locfdr)
+library(mixfdr)
 source("chunk-options.R")
@@
 
-Generates DReg_hist.pdf?
+Generate a small data set, and run the different methods on this data set.
 
@@
 # Roughly compute a local fdr for qvalue (to aid the plotting of the
-# decomposition in zscore space).
+# decomposition in zscore space) in each bin of histogram. I set the 
+# threshold at 1 so that fdr is never larger than 1.
 temp     = hist(pval,nclass = 50)
-bin_fdr  = res.qvalue$pi0/temp$density # compute local fdr in each bin of histogram
+bin_fdr  = res.qvalue$pi0/temp$density 
 qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
-qval_fdr = pmin(1,qval_fdr) # threshold at 1, so fdr never bigger than 1
+qval_fdr = pmin(1,qval_fdr)
 qval_fdr = qvalue::lfdr(pval)
 ```
+
 Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:
@@
                 cex.axis = 0.8,cex.main = 0.8)
-
 ```
 
-Now the decomposition:
+Now plot the decomposition:
 
@@
 ```
+
+Show the distribution of the p-values and the z-scores, stratified by their
+classification into the "null" and "alternative".
 
@@
 # p value histograms
-altnullhist(pval,qval_fdr,main="p values",ncz=50,xaxt='n',cex.axis=0.8)
-axis(side=1, labels=FALSE,tck=-0.01)
-altnullhist(pval,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
-axis(side=1, labels=FALSE,tck=-0.01)
-altnullhist(pval,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
-axis(side=1, labels=FALSE,tck=-0.01)
-altnullhist(pval,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8)
-axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
+altnullhist(pval,qval_fdr,main = "p values",ncz = 50,xaxt = 'n',cex.axis = 0.8)
+axis(side = 1, labels = FALSE,tck = -0.01)
+altnullhist(pval,res.locfdr$fdr,main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
+axis(side = 1, labels = FALSE,tck = -0.01)
+altnullhist(pval,res.mixfdr$fdr,main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
+axis(side = 1, labels = FALSE,tck = -0.01)
+altnullhist(pval,get_lfdr(res.ash),main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
+axis(side = 1, labels = TRUE,tck = -0.01,cex.axis = 0.8)
 
 # z score histograms
-nullalthist(zscore,qval_fdr,main="z scores",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
-axis(side=1, labels=FALSE,tck=-0.01)
-nullalthist(zscore,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
-axis(side=1, labels=FALSE,tck=-0.01)
-nullalthist(zscore,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
-axis(side=1, labels=FALSE,tck=-0.01)
-nullalthist(zscore,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
-axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
+nullalthist(zscore,qval_fdr,main = "z scores",ncz = 50,xaxt = 'n',
+            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
+axis(side = 1,labels = FALSE,tck = -0.01)
+nullalthist(zscore,res.locfdr$fdr,main = "",ncz = 50,xaxt = 'n',
+            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
+axis(side = 1,labels = FALSE,tck = -0.01)
+nullalthist(zscore,res.mixfdr$fdr,main = "",ncz = 50,xaxt = 'n',
+            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
+axis(side = 1,labels = FALSE,tck = -0.01)
+nullalthist(zscore,get_lfdr(res.ash),main = "",ncz = 50,xaxt = 'n',
+            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
+axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
 ```
@@
 ```{r decomp_ZA_poster, dev='pdf', fig.width=6, fig.height=2, crop=TRUE}
-#pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
-layout(matrix(1:8,ncol=4,byrow=TRUE))
+# pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
+layout(matrix(1:8,ncol = 4,byrow = TRUE))
 par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))
-ncz=25
+ncz = 25
 
-#p value histograms
-altnullhist(pval,qval_fdr,main="p values: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8)
-#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
-axis(side=1, labels=FALSE,tck=-0.01)
-#mtext(side=3,"p values",line=1)
-altnullhist(pval,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
-axis(side=1, labels=FALSE,tck=-0.01)
-altnullhist(pval,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
-axis(side=1, labels=FALSE,tck=-0.01)
-altnullhist(pval,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8)
-axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
-#mtext(side=1,"p values",line=2)
+# p value histograms
+altnullhist(pval,qval_fdr,main = "p values: qvalue",ncz = ncz,
+            xaxt = 'n',cex.axis = 0.8)
+# plot_FDReg_hist(pval,res.qvalue$pi0,type = 2,title = "p values", 
+#                 xaxt = 'n',cex.axis = 0.8)
+axis(side = 1,labels = FALSE,tck = -0.01)
+# mtext(side = 3,"p values",line = 1)
+altnullhist(pval,res.locfdr$fdr,main = "locfdr",ncz = ncz,xaxt = 'n',
+            cex.axis = 0.8)
+axis(side = 1,labels = FALSE,tck = -0.01)
+altnullhist(pval,res.mixfdr$fdr,main = "mixfdr",ncz = ncz,xaxt = 'n',
+            cex.axis = 0.8)
+axis(side = 1,labels = FALSE,tck = -0.01)
+altnullhist(pval,get_lfdr(res.ash),main = "ash",ncz = ncz,xaxt = 'n',
+            cex.axis = 0.8)
+axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
+# mtext(side = 1,"p values",line = 2)
 
 #z score histograms
-nullalthist(zscore,qval_fdr,main="z scores: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
-axis(side=1, labels=FALSE,tck=-0.01)
-#mtext(side=3,"z scores",line=1)
-nullalthist(zscore,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
-axis(side=1, labels=FALSE,tck=-0.01)
-nullalthist(zscore,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
-axis(side=1, labels=FALSE,tck=-0.01)
-nullalthist(zscore,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
-axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
-#mtext(side=1,"z scores",line=2)
-
-dev.off()
+nullalthist(zscore,qval_fdr,main = "z scores: qvalue",ncz = ncz,xaxt = 'n',
+            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
+axis(side = 1,labels = FALSE,tck = -0.01)
+# mtext(side = 3,"z scores",line = 1)
+nullalthist(zscore,res.locfdr$fdr,main = "locfdr",ncz = ncz,xaxt = 'n',
+            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
+axis(side = 1,labels = FALSE,tck = -0.01)
+nullalthist(zscore,res.mixfdr$fdr,main = "mixfdr",ncz = ncz,xaxt = 'n',
+            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
+axis(side = 1,labels = FALSE,tck = -0.01)
+nullalthist(zscore,get_lfdr(res.ash),main = "ash",ncz = ncz,xaxt = 'n',
+            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
+axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
+# mtext(side = 1,"z scores",line = 2)
+# dev.off()
 ```

### 2016-12-09 analysis/investigate_hu_badcoverage.Rmd
update
Rendered referee_uaza.html and fixed chunk-options label to chunk_options.

@@
 
-```{r chunk-options, include=FALSE}
+```{r chunk_options, include=FALSE}
 source("chunk-options.R")
@@
   eg=load_example(dsc_shrink,42,"flat-top",method="ash.hu.s","../code/dsc-shrink")
-  eg$output$fitted.g
+  eg$output$fitted_g
 ```

### 2016-12-09 analysis/make_GOODPOOR_figs.Rmd
Rendered referee_uaza.html and fixed chunk-options label to chunk_options.

@@
 
-```{r chunk-options, include=FALSE}
+```{r chunk_options, include=FALSE}
 library(ashr)

### 2016-12-09 analysis/referee_uaza.Rmd
Revised	efron.fcr.Rmd.
Rendered referee_uaza.html and fixed chunk-options label to chunk_options.

@@
 
-```{r chunk-options, include=FALSE}
+A referee reported a simulation of 10,000 hypotheses, a fraction 0.8 are null ($\mu_i = 0$), and a fraction 0.2 are alternative with effect $\mu_i \sim 
+N(0,1)$. For each effect $i$ there are n=30 observations from $x_{i,j} \sim N(\mu_i,1)$ and a one-sample t test is used to compute a $p$ value. The 
+referee correctly notes that these simulations i) obey the unimodal 
+assumption (UA), and ii) have the property that "most" $z$ scores near 0 are 
+null (Efron's "Zero Assumption"; ZA). The motivation is to demonstrate that 
+the UA and the ZA are not contradictory.
+
+I agree with this, if we define the ZA as "most" $z$ scores near 0 are null.
+However, in practice the existing methods effectively make a stronger 
+assumption (or at least, behave as if they do): that *all* $z$ scores near 
+0 are null. Consequently they end up creating a hole in the distribution of alternative $z$ scores at 0. Here we illustrate this by applying qvalue to 
+this scenario.
+
+First, we load the necessary libraries, and specify settings for displaying 
+the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
 source("chunk-options.R")
@@
 
-# Simulation
+## Simulation
 
-A referee reported a simulation of 10,000 hypotheses, a fraction 0.8 are null ($\mu_i=0$), and a fraction 0.2 are alternative with effect $\mu_i \sim N(0,1)$. For each effect $i$ there
-are n=30 observations from $x_{i,j} \sim N(\mu_i,1)$ and a one-sample t test is used to compute a $p$ value. The referee correctly notes that these simulations i) obey the
-unimodal assumption (UA), and ii) have the property that "most" $z$ scores near 0 are null (Efron's "Zero Assumption"; ZA).
-The motivation is to demonstrate that the UA and the ZA are not contradictory.
+First, we generate a small data set.
 
-I agree with this, if we define the ZA as "most" $z$ scores near 0 are null.
-However, in practice the existing methods effectively make a stronger assumption
-(or at least, behave as if they do): that *all* $z$ scores near 0 are null.
-Consequently they end up creating a hole in the distribution of alternative $z$ scores at 0. Here we illustrate this by applying qvalue to this scenario.
-
-```{r}
+```{r sim_data}
 set.seed(100)
@@
 
-# Explanation
+## Explanation
 

### 2016-12-12 analysis/referee.response.Rmd
Revised	efron.fcr.Rmd.
Rendered referee_uaza.html and fixed chunk-options label to chunk_options.
A few formatting adjustments to referee.response.Rmd.

@@
 
-# Referee 1
+## Referee 1
 
-I agree with a lot of what the referee says, although there are some points of disagreement. The most fundamental disagreement regards Figure 1. This Figure shows
-not the detectable effects (as the referee believed) but the decomposition
-of *all* effects into null and alternative components. Since this seems
-to be an important misunderstanding I have revised the text to try to make it clearer how this figure was derived. Specifically the revised text says:
+I agree with a lot of what the referee says, although there are some points of 
+disagreement. The most fundamental disagreement regards Figure 1. This Figure 
+shows not the detectable effects (as the referee believed) but the decomposition
+of *all* effects into null and alternative components. Since this seems to be an 
+important misunderstanding I have revised the text to try to make it clearer how
+this figure was derived. Specifically the revised text says:
 
-“We used each of the methods qvalue, locfdr, mixfdr and ash to decompose the $z$ scores ($z_j = \hat{\beta}_j$), or their corresponding p values, into null and alternative components.
-Here we are using the fact that these methods all provide an estimate of the lfdr for each observation, which implies such a decomposition; specifically the average lfdr within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component.”
+"We used each of the methods qvalue, locfdr, mixfdr and ash to decompose the $z$ 
+scores ($z_j = \hat{\beta}_j$), or their corresponding p values, into null and 
+alternative components. Here we are using the fact that these methods all 
+provide an estimate of the lfdr for each observation, which implies such a 
+decomposition; specifically the average lfdr within each histogram bin estimates
+the fraction of observations in that bin that come from the null vs the 
+alternative component.”
 
-Since the referee objected to the presentation of the ZA and UA as opposing assumptions I have removed all explicit reference to the ZA. (This also 
-solves the issue that I had defined the ZA differently than Efron’s original definition — see below — which may have created confusion.) 
-Instead, as the referee 
-suggested, I have presented the UA as an additional assumption “not made by other methods”, and - as the editor suggested —
-I have focussed on contrasting the behavior of ash with the behavior 
-of the other methods (which, as Figure 1 demonstrates, is to create a hole at 0 in the distribution of the alternative Z scores) rather than contrasting their assumptions.
+Since the referee objected to the presentation of the ZA and UA as opposing 
+assumptions I have removed all explicit reference to the ZA. (This also
+solves the issue that I had defined the ZA differently than Efron’s original 
+definition — see below — which may have created confusion.) Instead, as the 
+referee suggested, I have presented the UA as an additional assumption “not 
+made by other methods”, and - as the editor suggested — I have focussed on 
+contrasting the behavior of ash with the behavior of the other methods (which, 
+as Figure 1 demonstrates, is to create a hole at 0 in the distribution of the 
+alternative Z scores) rather than contrasting their assumptions.
 
-However, it seems unsatisfactory to merely point out this behavioral
-difference between methods without giving any explanation for the reason behind it.
-So I have provided my best attempt to give the reason.
-Essentially the reason is that the existing methods (locfdr and qvalue), when estimating $\pi_0$, make the assumption that  *all* p values near 1 are null, or *all* z scores near 0 are null.
-(This was how I defined the ZA in the original paper; however this is stronger than Efron’s original statement of the ZA, which the referee - understandably - uses in his/her report, that ``most” z scores near 0 are null.
-So I don’t refer to this
-as the ZA any more.) This assumption necessarily creates a hole in the
-implied distribution of z scores under the alternative, and explains the behavior.
+However, it seems unsatisfactory to merely point out this behavioral difference
+between methods without giving any explanation for the reason behind it. So I 
+have provided my best attempt to give the reason. Essentially the reason is 
+that the existing methods (locfdr and qvalue), when estimating $\pi_0$, make 
+the assumption that *all* p values near 1 are null, or *all* z scores near 0 
+are null. (This was how I defined the ZA in the original paper; however this 
+is stronger than Efron’s original statement of the ZA, which the referee - 
+understandably - uses in his/her report, that ``most” z scores near 0 are null. 
+So I don’t refer to this as the ZA any more.) This assumption necessarily 
+creates a hole in the implied distribution of z scores under the alternative,
+and explains the behavior.
 
-To underscore that this behavior is general, and not just a feature of the particular
-simulation I did in Figure 1, I provide results using qvalue under the setting
-suggested by the referee at (http://stephenslab.github.io/ash/analysis/referee_uaza.html).
+To underscore that this behavior is general, and not just a feature of the 
+particular simulation I did in Figure 1, I provide results using qvalue under
+the setting suggested by the referee at; see  (http://stephenslab.github.io/ash/analysis/referee_uaza.html).
 
-## Minor Concerns
+### Minor Concerns
 
@@
 
-- I agree that returning the length J results as a dataframe would have been better, but
-unfortunately changing this would break other packages that depend on ashr.
-So instead I have provided a method `as.data.frame` to extract the most important components of the ash object into a dataframe.
+- I agree that returning the length J results as a dataframe would have been 
+better, but unfortunately changing this would break other packages that 
+depend on ashr. So instead I have provided a method `as.data.frame` to 
+extract the most important components of the ash object into a dataframe.
 
@@
 
-# Referee 2
+## Referee 2
 
-## Major Concern
+### Major Concern
 
-The referee says: "it would be interesting to
-include distributions that favor other models, such as a bimodal distribution with no mass at zero, an assymmetrical distribution,.
-.
-.
-The unimodality is a strong assumption so it would be good to see how anti-conservative the results would get under different alternative distributions."
+The referee says: "it would be interesting to include distributions that 
+favor other models, such as a bimodal distribution with no mass at zero, an 
+assymmetrical distribution,. . . The unimodality is a strong assumption so it 
+would be good to see how anti-conservative the results would get under 
+different alternative distributions."
 
-I found this puzzling since both an assymetrical distribution and a strongly bimodal distribution are already included in the simulation results (Figure 2, Figure 3, Table 1).
-I believe these existing results address the referee comment.
-I also want to emphasise that the unimodal assumption is actually less strong than the assumptions usually made in variable selection for regression, which is directly analogous to the fdr context.
-For example, it is common to assume that effects are a mixture of a point mass at zero and a single normal, which is considerably more restrictive than the general unimodal assumption.
+I found this puzzling since both an assymetrical distribution and a strongly 
+bimodal distribution are already included in the simulation results (Figure 2, 
+Figure 3, Table 1). I believe these existing results address the referee comment. 
+I also want to emphasise that the unimodal assumption is actually less strong 
+than the assumptions usually made in variable selection for regression, which 
+is directly analogous to the fdr context. For example, it is common to assume 
+that effects are a mixture of a point mass at zero and a single normal, which 
+is considerably more restrictive than the general unimodal assumption.
 
-However, to help reassure the referee, I have added another example to illustrate an assymetric situation where the unimodal-at-zero assumption is badly contradicted [here](http://stephenslab.github.io/ash/analysis/efron.fcr.html).
-As
+However, to help reassure the referee, I have added another example to 
+illustrate an assymetric situation where the unimodal-at-zero assumption is 
+badly contradicted 
+[here](http://stephenslab.github.io/ash/analysis/efron.fcr.html). As
 the referee will see, the resulting estimates of the lfdr from ashr are
@@
 
-## Minor Concern
+### Minor Concern
 
-- Is there a way to model the distribution as a mixture of both uniforms and normals? If not: why?
+*Is there a way to model the distribution as a mixture of both uniforms and 
+normals? If not: why?*
 
-This is straightforward in principle. However, it would require 
-non-trivial code modifications because of the way I implemented the mixture class (assuming that all components  come from the same family.) 
-Another way to achieve an assymetric distribution without
-the "hard" tails of the uniform might be to use a mixture of truncated normals.
-I have a student investigating this option.
-However, the uniform mixture has the advantage of allowing not only the assymetry, but also the $t$ likelihood.
-Although the paper
-does not focus on this, I believe this is likely to be an important issue in practice.
-So I suspect that the mixture of uniforms is often going to be the method of choice in practice.
+This is straightforward in principle. However, it would require non-trivial 
+code modifications because of the way I implemented the mixture class (assuming 
+that all components  come from the same family.) Another way to achieve an 
+assymetric distribution without the "hard" tails of the uniform might be to use 
+a mixture of truncated normals. I have a student investigating this option. 
+However, the uniform mixture has the advantage of allowing not only the 
+assymetry, but also the $t$ likelihood. Although the paper does not focus on 
+this, I believe this is likely to be an important issue in practice. So I 
+suspect that the mixture of uniforms is often going to be the method of choice 
+in practice.
 
-- How would the model perform with moderate correlations?
+*How would the model perform with moderate correlations?*
 
-There seems to be no reason to think that
-moderate correlations among tests will have a substantial detrimental effect. 
-But I do not want
-to focus on this because I think it will give the reader the wrong message:
-I think that users of both ashr and other fdr methods
-should very much worry about correlations in practice (which I believe will
-often be large, and not moderate).
+There seems to be no reason to think that moderate correlations among tests 
+will have a substantial detrimental effect. But I do not want to focus on this 
+because I think it will give the reader the wrong message: I think that users 
+of both ashr and other fdr methods should very much worry about correlations 
+in practice (which I believe will often be large, and not moderate).
 
-- In equation 3, the null component (I think) is represented as delta0, but delta0 hasn’t been defined. What is assumed for the distribution of delta0?
+*In equation 3, the null component (I think) is represented as delta0, but 
+delta0 hasn’t been defined. What is assumed for the distribution of delta0?*
 
-Thank you for noticing this. I now define $\delta_0$ (which represents a point mass on 0).
+Thank you for noticing this. I now define $\delta_0$ (which represents a point 
+mass on 0).
 
-- Figure 3: the truth is hard to distinguish from the different methods. And in general this figure is hard to read.
+*Figure 3: the truth is hard to distinguish from the different methods. And in
+general this figure is hard to read.*
 
 I agree. I have completely rethought this figure, which now has fewer, larger,
-panels, and compares just two  methods (ash.hu and the NPMLE) against the truth on a few examples.
-The less important point - the effect of the penalty term on pi0 -is now illustrated in an additional figure included in the appendix.
+panels, and compares just two  methods (ash.hu and the NPMLE) against the truth 
+on a few examples. The less important point - the effect of the penalty term on 
+pi0 -is now illustrated in an additional figure included in the appendix.
 
-- I particularly like the idea that the ordering of the tests differs from the classical p-values.
-Which influence does this have on the balance between sensitivity and specificity (for example with ROC curves)?
-It should change, but does it get better?
+*I particularly like the idea that the ordering of the tests differs from the
+classical p-values. Which influence does this have on the balance between
+sensitivity and specificity (for example with ROC curves)? It should change, 
+but does it get better?*
 
@@
 
-## Typos and grammar
+### Typos and grammar
 
-- page 14, line 20. The first mentioned pi0 and lfdr should have a hat.
+*page 14, line 20. The first mentioned pi0 and lfdr should have a hat.*
 

### 2016-12-16 code/dsc-opt/run_dsc_opt.R
Fixed bugs in functions used by run_dsc.R; I was able to successfully generate res.RData.

@@ -8,33 +8,28 @@ dsc_opt=new_dsc("opt","../../output/dsc-opt-files")
 ###### Add Scenarios #####
 dscr::source_dir("../dsc-shrink/datamakers")
 source("../dsc-shrink/add_named_scenarios.R")
-add_named_scenarios(dsc_opt,c("spiky","near-normal","flat-top","skew","big-normal","bimodal"))
-
+add_named_scenarios(dsc_opt,c("spiky","near-normal","flat-top","skew",
+                              "big-normal","bimodal"))
 
 ###### Add Methods #####
 
-add_method(dsc_opt,"ash.u",ash.multiopt.wrapper,args=list(mixcompdist="uniform",method="shrink"))
-add_method(dsc_opt,"ash.hu",ash.multiopt.wrapper,args=list(mixcompdist="halfuniform",method="shrink"))
-add_method(dsc_opt,"ash.n",ash.multiopt.wrapper,args=list(mixcompdist="normal",method="shrink"))
-
-
+add_method(dsc_opt,"ash.u",ash.multiopt.wrapper,
+           args = list(mixcompdist = "uniform",method = "shrink"))
+add_method(dsc_opt,"ash.hu",ash.multiopt.wrapper,
+           args = list(mixcompdist = "halfuniform",method = "shrink"))
+add_method(dsc_opt,"ash.n",ash.multiopt.wrapper,
+           args = list(mixcompdist = "normal",method = "shrink"))
 
 ####### Define Score and Add it #######
 
-score = function(data, output){
-  x=output$loglik
-  return(list(diff1 = x[1]-x[2])) 
+score = function(data, output) {
+  x = output$loglik
+  return(list(diff1 = x[1] - x[2])) 
 }
 
 add_score(dsc_opt,score)
 
 ######## Run the DSC #################
 
-res_opt=run_dsc(dsc_opt)
-save(dsc_opt,file="../../output/dsc-opt-files/dsc_opt.RData")
-
-
-
-
-
-
+res_opt = run_dsc(dsc_opt)
+save(dsc_opt,file = "../../output/dsc-opt-files/dsc_opt.RData")

### 2016-12-16 code/dsc-shrink/add_methods.R

@@ -1,6 +1,6 @@
 dscr::source_dir("methods")
 
-add_method(dsc_shrink,name="ash.hu.nocxx",fn =ash.wrapper,args=list(mixcompdist="halfunif",cxx=FALSE),outputtype = "ash_output")
+add_method(dsc_shrink,name="ash.hu.nocxx",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
 add_method(dsc_shrink,name="ash.hu",fn =ash.wrapper,args=list(mixcompdist="halfunif"),outputtype = "ash_output")
 add_method(dsc_shrink,name="ash.u",fn =ash.wrapper,args=list(mixcompdist="unif"),outputtype = "ash_output")
 add_method(dsc_shrink,name="ash.n",fn =ash.wrapper,args=list(mixcompdist="normal"),outputtype = "ash_output")

### 2016-12-16 code/dsc-shrink/methods/ash.wrapper.R

@@ -34,7 +34,7 @@ bayes.wrapper = function(input, meta, args=NULL){
 }
 
 ash2beta_est =function(output){
-    return (list(beta_est=get_pm(output)))
+  return (list(beta_est=get_pm(output)))
 } 
 
 ash2pi0_est =function(output){

### 2016-12-16 code/dsc-shrink/score.R

@@ -23,7 +23,8 @@ score2 = function(data, output){
 
 score3 = function(data, output){
   return(c(S=pcdf_post(output$fitted.g,data$meta$beta,
-              data$input$betahat,data$input$sebetahat,v=NULL)))
+                       set_data(data$input$betahat,
+                                data$input$sebetahat))))
 }
 
 score_neg = function(data, output){

### 2016-12-16 code/dsc-shrink/run_dsc_znull.R
Fixed bug when running calc_logLR; run_dsc_znull.R now runs to completion succesfully (no errors, only warnings).

@@ -10,7 +10,6 @@ source("add_methods.null.R")
 source("score.R")
 add_score(dsc_znull,score_logLR,"logLR","ash_output")
 
-
 res=run_dsc(dsc_znull)
 save(res,dsc_znull,file="../../output/dsc-znull-files/res.znull.RData")
 

### 2016-12-16 code/dsc-shrink/score.R

@@ -52,5 +52,6 @@ score_betahat = function(data, output){
 }
 
 score_logLR = function(data,output){
-  return(c(logLR=ashr::calc_logLR(output,data$input$betahat,data$input$sebetahat,NULL)))
+  return(c(logLR = ashr::calc_logLR(output,set_data(data$input$betahat,
+                                                    data$input$sebetahat))))
 }

### 2016-12-20 analysis/checkIP.Rmd
Working on revisions to CompareEEvsETmodel.Rmd.
Successfully knit summarize_dsc_znull.Rmd using simulation results from res.znull.RData.

@@
 
-```{r chunk-options, include=FALSE}
+Here we compare the likelihoods achieved by the newly implemented
+interior point method against the EM algorithm.
+
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
+library(ggplot2)
+library(dplyr)
+library(reshape2)
 source("chunk-options.R")
@@
 
-We want to compare the likelihoods achieved by the newly implemented interior point method with the EM algorithm.
-
 ## Log-likelihood comparison
 
-The 'dsc-opt' simply runs both methods (IP and EM) on the same data sets, 
-and compares the difference in the log-likelihoods. We can see that
-in >90% of cases the difference is negligible. Then there is a tail of
-cases where the IP loglik is higher - presumably where the EM doesn't really converge so well.
+The 'dsc-opt' script simply runs both methods (IP and EM) on the same data sets,
+and compares the difference in the log-likelihoods. We can see that in >90% of 
+cases the difference is negligible. Then there is a tail of cases where the IP 
+log-likelihood is higher - presumably where the EM doesn't really converge so
+well.
 
-```{r}
+```{r compare_likelihoods}
 load("../output/dsc-opt-files/dsc_opt.RData")
-  plot(ecdf(dsc_opt$res$diff1),main="ECDF of log-likelihood difference (IP-EM)")
+plot(ecdf(dsc_opt$res$diff1),
+     main = "ECDF of log-likelihood difference (IP - EM)")
 ```
@@
 
-However, what we really care about is inference quantities, such as the lfsr, rather than the log-likelihood. Here we compare the lfsr for the simulations where
-we ran both the IP method and EM method (nocxx) in dsc-shrink.
+However, what we really care about is inference quantities, such as the lfsr, 
+rather than the log-likelihood. Here we compare the lfsr for the simulations 
+where we ran both the IP method and EM method (nocxx) in dsc-shrink.
 
-```{r}
+```{r compare_lfsr}
 load("../output/dsc-shrink-files/res.RData")
-  library("ggplot2")
-  library("dplyr")
-  library("reshape2")
-  # select out the lfsr estimates for each method
 
-  df1 = res$lfsr %>% dplyr::filter(method %in% c("ash.hu.nocxx")) %>% select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    melt(id.vars=c("method","scenario","seed",".id"),value.name="lfsr.nocxx")
-  df2 = res$lfsr %>% dplyr::filter(method %in% c("ash.hu")) %>% select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    melt(id.vars=c("method","scenario","seed",".id"),value.name="lfsr.IP")
-
-  df = inner_join(df1,df2,by=c("scenario","seed","variable"))
-  ggplot(df,aes(lfsr.nocxx,lfsr.IP)) +geom_point(shape=1) +
-  facet_wrap(~ scenario, nrow=2) + 
+# Select out the lfsr estimates for each method.
+df1 = res$lfsr %>% filter(method %in% c("ash.hu.nocxx")) %>% 
+        select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
+        melt(id.vars = c("method","scenario","seed",".id"),
+             value.name = "lfsr.nocxx")
+df2 = res$lfsr %>% filter(method %in% c("ash.hu")) %>%
+        select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
+        melt(id.vars = c("method","scenario","seed",".id"),
+             value.name = "lfsr.IP")
+df = inner_join(df1,df2,by = c("scenario","seed","variable"))
+ggplot(df,aes(lfsr.nocxx,lfsr.IP)) + geom_point(shape = 1) +
+facet_wrap(~ scenario,nrow = 2) +
 geom_abline(colour = "black") +

### 2016-12-20 analysis/investigate_hu_badcoverage.Rmd
Successfully knit summarize_dsc_znull.Rmd using simulation results from res.znull.RData.

@@
 
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
 ```{r chunk_options, include=FALSE}
+library(ashr)
+library(reshape2)
+library(ggplot2)
+library(dplyr)
 source("chunk-options.R")
@@
 
-library("dscr")
-library("ashr")
-library(reshape2)
-library(ggplot2)
-library(dplyr)
-
+out <- res$negprob
 neglong = 
-  res$negprob %>% 
+  out %>% 
     select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    melt(id.vars=c("method","scenario","seed",".id"),value.name="negprob") %>%
+    melt(id.vars=c("method","scenario","seed",".id"),value.name = "negprob") %>%
     filter(negprob > 0.95)
@@
     select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    melt(id.vars=c("method","scenario","seed",".id"),value.name="posprob") %>%
+    melt(id.vars = c("method","scenario","seed",".id"),value.name="posprob") %>%
     filter(posprob > 0.95)
@@
     select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%    
-    melt(id.vars=c("method","scenario","seed",".id"))
+    melt(id.vars = c("method","scenario","seed",".id"))
 

### 2016-12-20 analysis/summarize_dsc_znull.Rmd
Rendered referee_uaza.html and fixed chunk-options label to chunk_options.
Successfully knit summarize_dsc_znull.Rmd using simulation results from res.znull.RData.

--- 2016-01-27 analysis/summarize_dsc_znull.Rmd
@@
 author: "Matthew Stephens"
-date: 2015-10-26
+date: '2015-10-26'
 ---
@@
 
-```{r chunk-options, include=FALSE}
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
+library(ggplot2)
+library(dplyr)
 source("chunk-options.R")
@@
 
-```{r}
-library(ggplot2)
-library(dplyr)
+Load and summarize the results of the data simulations.
+
+```{r load_sim_results}
 load("../output/dsc-znull-files/res.znull.RData")
-temp=ungroup(res %>% group_by(scenario,method) %>% summarise(gt0=mean(logLR>0)))
-xtabs(gt0~method+scenario,temp)
-
+out <- ungroup(res %>% group_by(scenario,method) %>% 
+                  summarise(gt0 = mean(logLR > 0)))
+xtabs(gt0 ~ method + scenario,out)
 ```
 
-Note that the ash.n.s conforms closely to the null expectation
-of 0.5 chi2_0 +0.5 chi2_1 from Stram and Lee, Biometrics
+Note that the ash.n.s conforms closely to the null expectation of 0.5 
+chi2_0 + 0.5 chi2_1 from Stram and Lee, Biometrics.
 
-conjecture: the uniform has the same asymptotic behaviour
-and the half uniform is like the sum of two of these?
+*Conjecture:* the uniform has the same asymptotic behaviour and the half 
+uniform is like the sum of two of these?
 
@@
 
-```{r}
-ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
-  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
-ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
-  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
-ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.n",method)), aes(sample = logLR)) + facet_grid(scenario~method) +
-  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
+```{r scatterplots}
+ggplot(res %>% filter(logLR > 0) %>% filter(grepl("ash.hu",method)),
+       aes(sample = logLR)) + 
+  facet_grid(scenario ~ method) +
+  stat_qq(distribution = qchisq,dparams = list(df=1)) + 
+  geom_abline() + 
+  ggtitle("qqplot under null vs chisq-1; line slope=1")
 
+ggplot(res %>% filter(logLR > 0) %>% filter(grepl("ash.u",method)),
+       aes(sample = logLR)) + 
+  facet_grid(scenario~method) +
+  stat_qq(distribution = qchisq,dparams = list(df = 1)) + 
+  geom_abline(intercept = 0,slope = 0.5) + 
+  ggtitle("qqplot under null vs chisq-1; line slope=0.5")
+
+ggplot(res %>% filter(logLR > 0) %>% filter(grepl("ash.n",method)),
+       aes(sample = logLR)) + 
+  facet_grid(scenario ~ method) +
+  stat_qq(distribution = qchisq,dparams = list(df = 1)) + 
+  geom_abline(intercept = 0,slope = 0.5) + 
+  ggtitle("qqplot under null vs chisq-1; line slope=0.5")
 ```

### 2016-12-20 analysis/plot_cdf_eg.Rmd
plot_pi0est.Rmd script succesfully runs; next need to knit.

@@
 
-```{r chunk-options, include=FALSE}
-source("chunk-options.R")
-```
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
 
-```{r}
+```{r chunk_options, include=FALSE}
 library(plyr)
@@
 library(ggplot2)
+source("chunk-options.R")
 source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
 source("../R/set_plot_colors.R")
+```
+
+```{r}
 load("../output/dsc-shrink-files/res.RData")

### 2016-12-20 analysis/plot_lfsr.Rmd
plot_pi0est.Rmd script succesfully runs; next need to knit.

@@
 
-```{r chunk-options, include=FALSE}
+```{r chunk_options, include=FALSE}
 library(dplyr)
@@
 source("chunk-options.R")
+source("../R/set_plot_colors.R")
 ```
@@
 load("../output/dsc-shrink-files/res.RData")
-source("../R/set_plot_colors.R")
 

### 2016-12-20 analysis/plot_pi0est.Rmd
Rendered referee_uaza.html and fixed chunk-options label to chunk_options.
plot_pi0est.Rmd script succesfully runs; next need to knit.

@@
 
-```{r chunk-options, include=FALSE}
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
+library(magrittr)
+library(dplyr)
 source("chunk-options.R")
+source("../R/set_plot_colors.R")
 ```
+
+*Explain here what this chunk does.*
 
@@
 load("../output/dsc-shrink-files/res.RData")
-source("../R/set_plot_colors.R")
-library(dplyr)
 
-PLOTMETHODS= c("mixfdr.tnull","ash.n","ash.u","qvalue","locfdr")
-PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
+PLOTMETHODS   = c("mixfdr.tnull","ash.n","ash.u","qvalue","locfdr")
+PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","bimodal")
 PLOTNAMES     = PLOTSCENARIOS
@@
 
-df = res$pi0_score %>% filter(scenario %in% PLOTSCENARIOS) %>% filter(method %in% PLOTMETHODS)
+df = res$pi0_score %>% 
+       filter(scenario %in% PLOTSCENARIOS) %>% 
+       filter(method %in% PLOTMETHODS)
 df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
-levels(df$scenario)= PLOTNAMES
+levels(df$scenario) = PLOTNAMES
 
-pi0_plot=ggplot(df,
-         aes(pi0,pi0_est,colour=method,alpha=ALPHALEVEL)) +geom_point(shape=1) +
+pi0_plot = ggplot(df,aes(pi0,pi0_est,colour = method,alpha = ALPHALEVEL)) +
+  geom_point(shape = 1) +
   facet_grid(. ~ scenario) + 
-  guides(alpha=FALSE) +
+  guides(alpha = FALSE) +
   geom_abline(colour = "black") +
@@
   ylab("Estimated pi0") 
-print(pi0_plot +scale_y_continuous(limits=c(0,1)) +
-        scale_x_continuous(limits=c(0,1)) +
-        #scale_colour_manual(values=cbbPalette,breaks=breaks,labels=labels) +
-        coord_equal(ratio=1) + colScale + theme(legend.position = "top",axis.text.x = element_text(size = 8,angle=45)))
-
+print(pi0_plot + scale_y_continuous(limits = c(0,1.01)) +
+        scale_x_continuous(limits = c(0,1.01)) +
+        coord_equal(ratio = 1) + 
+        colScale + 
+        theme(legend.position = "top",
+              axis.text.x     = element_text(size = 8,angle = 45)))
 ```

### 2016-12-20 analysis/plot_egdens.Rmd
Successfully re-knit plot_egdens.Rmd using ashr 2.0.4.

--- 2015-10-26 analysis/plot_egdens.Rmd
@@
 
-```{r chunk-options, include=FALSE}
+```{r chunk_options, include=FALSE}
+library(ashr)
+library(ggplot2)
+library(dplyr)
 source("chunk-options.R")
@@
 
-```{r}
-library("ashr")
-library("ggplot2")
-library("dplyr")
-library("dscr")
+Load the results of the simulation experiments, and generate the density
+data for all the simulation scenarios in a single data frame ("df").
+
+```{r chunk_label}
 load("../output/dsc-shrink-files/res.RData")
-PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
+PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","bimodal")
 PLOTNAMES     = PLOTSCENARIOS
-#PLOTFILE = "../paper/figures/scenarios_density.pdf"
 
-df=data.frame()
-
-for(i in PLOTSCENARIOS){
-  s=dsc_shrink$scenarios[[i]]
-  g=s$args$g
-  x = seq(-6,6,length=100)
+df = data.frame()
+for(i in PLOTSCENARIOS) {
+  s  = dsc_shrink$scenarios[[i]]
+  g  = s$args$g
+  x  = seq(-6,6,length = 100)
   y  = as.numeric(dens(g,x))
-  df = rbind(df,data.frame(x=x,y=y,scenario=i))
+  df = rbind(df,data.frame(x = x,y = y,scenario = i))
 }
 
-df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
-levels(df$scenario)= PLOTNAMES
+df$scenario         = factor(df$scenario,levels = PLOTSCENARIOS)
+levels(df$scenario) = PLOTNAMES
 ```
 
-```{r scenario_density}
-#pdf(PLOTFILE,height=3,width=9)
-ggplot(df, aes(x=x,y=y)) + geom_line(size=1.2,linetype=1) + facet_grid(.~scenario) + ylab("density")
-#dev.off()
+Generate density plots using ggplot.
+
+```{r scenario_density, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
+ggplot(df,aes(x = x,y = y)) + 
+  geom_line(size = 1.2,linetype = 1) + 
+  facet_grid(.~scenario) + 
+  ylab("density")
 ```

### 2016-12-20 analysis/plot_cdf_eg.Rmd
Stopping here with revisions to analysis/*.Rmd files due to issue with missing columns in data frames in res.RData; will post issue.

@@
 library(dplyr)
-library(dscr)
 library(magrittr)
@@
 library(ggplot2)
+library(dscr)
 source("chunk-options.R")
@@
 
-```{r}
+*Explain here what this chunk does.*
+
+```{r chunk_label}
 load("../output/dsc-shrink-files/res.RData")
@@
 plot_mean_cdf = function(SEEDS,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
-      PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),pi0filter=FALSE,...){
-  PLOTNAMES=PLOTSCENARIOS
+      PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","bimodal"),pi0filter=FALSE,...){
+  PLOTNAMES = PLOTSCENARIOS
   #set up dataframe with cdf for all methods and all datasets
@@
 plot_mean_cdf_with_npmle(1,PLOTMETHODS=c("truth","ash.hu"),PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")) + facet_wrap(~ scenario, nrow=2)
-
 ```

### 2016-12-20 analysis/plot_egdens.Rmd
Stopping here with revisions to analysis/*.Rmd files due to issue with missing columns in data frames in res.RData; will post issue.

@@
 
-```{r chunk_label}
+```{r retrieve_sim_results}
 load("../output/dsc-shrink-files/res.RData")

### 2016-12-20 analysis/plot_lfsr.Rmd
Stopping here with revisions to analysis/*.Rmd files due to issue with missing columns in data frames in res.RData; will post issue.

@@
                                                  GOLDMETHOD="bayes",PLOTSEEDS=1:100,
-                                                 PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),
-                                                 PLOTNAMES=PLOTSCENARIOS){
-  df %<>% filter(seed %in% PLOTSEEDS) %>% filter(scenario %in% PLOTSCENARIOS) %>% filter(method %in% c(PLOTMETHODS,GOLDMETHOD))
+                                                   PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
+  PLOTNAMES=PLOTSCENARIOS
+  df %<>% filter(seed %in% PLOTSEEDS) %>% 
+    filter(scenario %in% PLOTSCENARIOS) %>% 
+    filter(method %in% c(PLOTMETHODS,GOLDMETHOD))
   df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)

### 2016-12-20 analysis/summarize_coverage.Rmd
Rendered referee_uaza.html and fixed chunk-options label to chunk_options.
Stopping here with revisions to analysis/*.Rmd files due to issue with missing columns in data frames in res.RData; will post issue.

@@
 
-```{r chunk-options, include=FALSE}
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
+library(dscr)
+library(ashr)
+library(reshape2)
+library(ggplot2)
+library(dplyr)
 source("chunk-options.R")
@@
 
----
-title: "Summarize Coverage"
-output: html_document
----
-
-```{r}
+```{chunk_label r}
 load("../output/dsc-shrink-files/res.RData")
 coverthresh = 0.05 # threshold at which we look at coverage
-findthresh=0.95 #threshold at we define a discovery significant
-library("dscr")
-library("ashr")
-library(reshape2)
-library(ggplot2)
-library(dplyr)
+findthresh  = 0.95 # threshold at we define a discovery significant
 

### 2016-12-21 code/dsc-opt/run_dsc_opt.R
update dsc to deal with updates to ashr, v2

@@ -22,8 +22,8 @@ add_method(dsc_opt,"ash.n",ash.multiopt.wrapper,args=list(mixcompdist="normal",m
 ####### Define Score and Add it #######
 
 score = function(data, output){
-  x=output$loglik
-  return(list(diff1 = x[1]-x[2])) 
+  x=output$loglik #this loglik is a vector of length 2 with logliks from two different optimizations
+  return(list(diff1 = x[1]-x[2]))
 }
 
 add_score(dsc_opt,score)

### 2016-12-21 code/dsc-shrink/run_dsc.R

@@ -29,7 +29,7 @@ add_score(dsc_shrink,score3,"cdf_score","g_output")
 add_score(dsc_shrink,score_neg,"negprob","ash_output") #just extract the negativeprobs
 add_score(dsc_shrink,score_pos,"posprob","ash_output") #just extracts the positiveprobs
 add_score(dsc_shrink,score_fdr,"fdr","mixfdr_output") #just extracts the fdr
-add_score(dsc_shrink,score_betahat,"betahat","mixfdr_output") #just extracts the fdr
+add_score(dsc_shrink,score_betahat,"betahat","mixfdr_output") #just extracts betahat
 add_score(dsc_shrink,score_lfsr,"lfsr","ash_output") #just extracts the lfsr
 add_score(dsc_shrink,score_lfdr,"lfdr","ash_output") #just extracts the lfdr
 

### 2016-12-21 code/dsc-shrink/score.R

@@ -1,11 +1,11 @@
 #' @title compute score for shrinkage DSC
 #'
-#' @description Outputs the RMSE and MAE of estimated beta values 
+#' @description Outputs the RMSE and MAE of estimated beta values
 #' @details None
-#' 
+#'
 #' @param data
 #' @param output
-#' 
+#'
 #' @return score a list with
 #' \item{RMSE}{root mean squared error of estimated beta values}
 #' \item{MAE}{Median absolute error of estimated beta values}
@@ -27,23 +27,23 @@ score3 = function(data, output){
 }
 
 score_neg = function(data, output){
-  return(c(S=output$NegativeProb))
+  return(c(S=get_np(output)))
 }
 
 score_pos = function(data, output){
-  return(c(S=output$PositiveProb))
+  return(c(S=get_pp(output)))
 }
 
-score_fdr = function(data, output){
-  return(c(S=output$res$fdr))
+score_fdr = function(data, output){ #for mixfdr output
+  return(c(S=output$fdr))
 }
 
 score_lfsr = function(data, output){
-  return(c(S=output$lfsr))
+  return(c(S=get_lfsr(a)))
 }
 
 score_lfdr = function(data, output){
-  return(c(S=output$lfdr))
+  return(c(S=get_lfdr(a)))
 }
 
 score_betahat = function(data, output){

### 2016-12-21 code/dsc-shrink/score.R
fix bug with a instead of output

@@ -39,11 +39,11 @@ score_fdr = function(data, output){ #for mixfdr output
 }
 
 score_lfsr = function(data, output){
-  return(c(S=get_lfsr(a)))
+  return(c(S=get_lfsr(output)))
 }
 
 score_lfdr = function(data, output){
-  return(c(S=get_lfdr(a)))
+  return(c(S=get_lfdr(output)))
 }
 
 score_betahat = function(data, output){

### 2016-12-21 analysis/metaplot_examples.Rmd

@@
 title: "DDD plots"
-output: html_document
+author: "Matthew Stephens"
+date: 2016-10-25
 ---
+
+**Last updated:** `r Sys.Date()`
+
+**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`
+
+```{r chunk-options, include=FALSE}
+source("chunk-options.R")
+```
 
@@
 
-# Simulate Z scores
+# Introduction
+
+This file produces ``meta-analysis" style plots that illustrate the shrinkage behaviour of ash on some simple simulated data.
+
+# Simulate Data
+
+We simulate 50 effects ($\beta$), half of which are truly "null" (ie 0) and the other half are $N(0,2^2)$. The observed effects ($\hat\beta$) all have standard error 1 here.
 
@@
 beta.ash=ashr::ash(betahat,1)
-o = order(betahat)
 ```
@@
 
+Here we plot the standard Confidence Intervals, ordered by the observed effect sizes. The red colors indicate observations correspondign to effects that were truly 0.
 ```{r unshrunk}
+o = order(betahat)
 rmeta::metaplot(betahat[o],rep(1,n),ylab="",xlab="",xlim=c(-6,6),colors = meta.colors(box=c(rep("red",n/2),rep("black",n/2))[o]))
@@
 
+Here we plot the shrunken Confidence Intervals (more formally, posterior Credible Intervals), in the same order, with same color code.
+(We use the approximation of the posterior mean $\pm$ 2 posterior standard deviation; really we should compute the posterior intervals using `ashci`, but this simple approximation suffices for illustrating the idea.)
+
+Notice how the CIs are strongly shrunk towards 0, except for those that correspond to large effects, where the shrinkage is more modest.
+What is happening here is that ash recognizes that although many effects are null, others are quite big, so it avoids shrinking too much the intervals whose data are sufficiently strong to convince it that they are non-null.
 ```{r shrunk}

### 2016-12-21 analysis/plot_egdens2.Rmd

---
title: "Plot two densities for strong vs weak shrinkage"
author: "Matthew Stephens"
date: 2016-10-25
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

```{r}
library("ashr")
library("ggplot2")
library("dplyr")
library("dscr")
load("../output/dsc-shrink-files/res.RData")
PLOTSCENARIOS = c("spiky","big-normal")
PLOTNAMES = c("Strong Shrinkage","Weak Shrinkage")

df=data.frame()

for(i in PLOTSCENARIOS){
  s=dsc_shrink$scenarios[[i]]
  g=s$args$g
  x = seq(-6,6,length=100)
  y = as.numeric(dens(g,x))
  df = rbind(df,data.frame(x=x,y=y,scenario=i))
}

df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
levels(df$scenario)= PLOTNAMES
```

```{r scenario_density_2, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
#pdf(PLOTFILE,height=3,width=9)
ggplot(df, aes(x=x,y=y)) + geom_line(size=1.2,linetype=1) + facet_grid(.~scenario) + ylab("density")
#dev.off()
```

## Session information

```{r info}
sessionInfo()
```

### 2016-12-21 code/dsc-opt/run_dsc_opt.R
Mistakenly deleted a couple lines from run_dsc_opt.R. Fixed this.

@@ -25,7 +25,9 @@ add_method(dsc_opt,"ash.n",ash.multiopt.wrapper,
 score = function(data, output) {
   # This loglik is a vector of length 2 with logliks from two
   # different optimizations.
-  x = output$loglik 
+  x = output$loglik
+  return(list(diff1 = x[1] - x[2]))  
+}
 
 add_score(dsc_opt,score)
 

### 2017-01-01 analysis/metaplot_examples.Rmd
Successfully re-knit metaplot_examples using new (CRAN) version of ashr.

@@
 
-```{r chunk-options, include=FALSE}
+This file produces ``meta-analysis" style plots that illustrate the shrinkage behaviour of ash on some simple simulated data.
+
+First, we load the necessary libraries, and specify settings for
+displaying the plots in the rendered document.
+
+```{r chunk_options, include=FALSE}
+library(knitr)
+library(rmeta)
+library(ashr)
 source("chunk-options.R")
+opts_chunk$set(echo = TRUE)
 ```
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+## Simulate Data
+
+We simulate 50 effects ($\beta$), half of which are truly "null" (ie 0) and 
+the other half are $N(0,2^2)$. The observed effects ($\hat\beta$) all have 
+standard error 1 here.
+
+```{r chunk_label}
+set.seed(1)
+n        = 50
+beta     = c(rep(0,n/2),rnorm(n/2,0,2))
+betahat  = beta + rnorm(n)
+beta.ash = ash(betahat,1)
 ```
 
-# Introduction
+## Meta Plots
 
-This file produces ``meta-analysis" style plots that illustrate the shrinkage behaviour of ash on some simple simulated data.
+Here we plot the standard Confidence Intervals, ordered by the observed effect 
+sizes. The red colors indicate observations corresponding to effects that were 
+truly 0.
 
-# Simulate Data
-
-We simulate 50 effects ($\beta$), half of which are truly "null" (ie 0) and the other half are $N(0,2^2)$. The observed effects ($\hat\beta$) all have standard error 1 here.
-
-```{r}
-library(rmeta)
-library(ashr)
-n=50
-set.seed(1)
-beta = c(rep(0,n/2),rnorm(n/2,0,2))
-betahat = beta+rnorm(n)
-beta.ash=ashr::ash(betahat,1)
+```{r plot_unshrunk}
+i = order(betahat)
+metaplot(betahat[i],rep(1,n),ylab = "",xlab = "",xlim = c(-6,6),
+         colors = meta.colors(box = c(rep("red",n/2),rep("black",n/2))[i]))
 ```
 
-# Meta Plots
+Here we plot the *shrunken* Confidence Intervals (more formally, posterior 
+Credible Intervals), in the same order, with same color code. (We use the
+approximation of the posterior mean $\pm$ 2 posterior standard deviation; 
+really we should compute the posterior intervals using `ashci`, but this 
+simple approximation suffices for illustrating the idea.)
 
-Here we plot the standard Confidence Intervals, ordered by the observed effect sizes. The red colors indicate observations correspondign to effects that were truly 0.
-```{r unshrunk}
-o = order(betahat)
-rmeta::metaplot(betahat[o],rep(1,n),ylab="",xlab="",xlim=c(-6,6),colors = meta.colors(box=c(rep("red",n/2),rep("black",n/2))[o]))
+Notice how the CIs are strongly shrunk towards 0, except for those that 
+correspond to large effects, where the shrinkage is more modest. What is 
+happening here is that ash recognizes that although many effects are null, 
+others are quite big, so it avoids shrinking too much the intervals whose 
+data are sufficiently strong to convince it that they are non-null.
+
+```{r shrunk}
+metaplot(get_pm(beta.ash)[i],get_psd(beta.ash)[i],
+         ylab = "",xlab = "",xlim = c(-6,6),
+         colors = meta.colors(box = c(rep("red",n/2),rep("black",n/2))[i]))
 ```
-
-Here we plot the shrunken Confidence Intervals (more formally, posterior Credible Intervals), in the same order, with same color code.
-(We use the approximation of the posterior mean $\pm$ 2 posterior standard deviation; really we should compute the posterior intervals using `ashci`, but this simple approximation suffices for illustrating the idea.)
-
-Notice how the CIs are strongly shrunk towards 0, except for those that correspond to large effects, where the shrinkage is more modest.
-What is happening here is that ash recognizes that although many effects are null, others are quite big, so it avoids shrinking too much the intervals whose data are sufficiently strong to convince it that they are non-null.
-```{r shrunk}
-rmeta::metaplot(get_pm(beta.ash)[o],get_psd(beta.ash)[o],ylab="",xlab="",xlim=c(-6,6),colors = meta.colors(box=c(rep("red",n/2),rep("black",n/2))[o]))
-```

### 2017-01-01 analysis/metaplot_examples.Rmd
Revised chunk labels for metaplot_examples.Rmd.

@@
 
-```{r shrunk}
+```{r plot_shrunk}
 metaplot(get_pm(beta.ash)[i],get_psd(beta.ash)[i],

### 2017-01-02 analysis/summarize_coverage.Rmd
Successfully re-knit summarize_coverage.Rmd after addressing bugs in dsc-shrink-files output.

@@
 
+Note that rendering this RMarkdown document may take a few minutes because 
+it involves loading and processing tables with millions of rows.
+
 First, we load the necessary libraries, and specify settings for
@@
 library(ggplot2)
+library(magrittr)
 library(dplyr)
+library(xtable)
 source("chunk-options.R")
@@
 
-```{chunk_label r}
+Compile the results on the simulated data sets for the summaries below.
+
+```{r compile_tables}
 load("../output/dsc-shrink-files/res.RData")
@@
     select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    melt(id.vars=c("method","scenario","seed",".id"),value.name="negprob") %>%
+    melt(id.vars = c("method","scenario","seed",".id"),
+         value.name = "negprob") %>%
     filter(negprob > findthresh)
@@
     select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    melt(id.vars=c("method","scenario","seed",".id"),value.name="posprob") %>%
+    melt(id.vars = c("method","scenario","seed",".id"),
+         value.name = "posprob") %>%
     filter(posprob > findthresh)
@@
 reslong.neg = inner_join(reslong,neglong)
-dim(reslong.pos)
-dim(reslong.neg)
-
 ```
@@
 
-Table of lower tail for all observations
-```{r}
-xtabs(lt~method+scenario,reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
+Table of lower tail for all observations:
+
+```{r table_all_obs}
+print(xtabs(lt ~ method + scenario,
+            reslong %>% group_by(scenario,method) %>% 
+            summarize(lt = mean(value < coverthresh))) %>% round(2))
 ```
@@
 Table of lower tail of positive findings. Because of the unimodal assumption
-and the favoritism toward the null this should assess problems
-with``over shrinkage" toward 0.
-```{r}
-xtabs(lt~method+scenario,reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
+and the favoritism toward the null, this should assess problems with "over 
+shrinkage" toward 0.
+
+```{r table_pos}
+print(xtabs(lt ~ method + scenario,
+            reslong.pos %>% group_by(scenario,method) %>% 
+              summarize(lt = mean(value < coverthresh))) %>% round(2))
 ```
 
-Table of lower tail of negative findings. This should indicate problems with tail behaviour of $g$. The uniform methods tend to over-shrink.
-```{r}
-xtabs(lt~method+scenario,reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh))) %>% round(2)
+Table of lower tail of negative findings. This should indicate problems with 
+tail behaviour of $g$. The uniform methods tend to over-shrink.
+
+```{r table_neg}
+print(xtabs(lt ~ method + scenario,reslong.neg %>% 
+            group_by(scenario,method) %>% 
+            summarize(lt = mean(value < coverthresh))) %>% round(2))
 ```
 
-```{r}
-save_latex_coverage_table=function(df,methodnames,filename,SCENARIONAMES=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),switch=FALSE){
-  library(xtable)
-  df$method=factor(df$method,levels=methodnames)
-  df$scenario=factor(df$scenario,levels=SCENARIONAMES)
+Compile some more summary tables in Latex format.
+
+```{r write_latex_tables}
+save_latex_coverage_table = 
+  function(df, methodnames, filename,
+           SCENARIONAMES = c("spiky","near-normal","flat-top","skew",
+                             "big-normal","bimodal"),
+           switch = FALSE) {
+  df$method   <- factor(df$method,levels = methodnames)
+  df$scenario <- factor(df$scenario,levels = SCENARIONAMES)
   mat         <- as.matrix(xtabs(lt~method+scenario,df))
-  if(switch){
-    mat=1-mat
-  }
-  mat <- xtable(mat,digits=rep(2,ncol(mat)+1))
-  write(print(mat, 
-        sanitize.text.function = function(x){x},
-        floating=FALSE, 
-        hline.after=NULL, 
-        add.to.row=list(pos=list(-1,0, nrow(mat)), 
-                        command=c('\\toprule ',
+  if (switch)
+    mat <- 1 - mat
+  mat <- xtable(mat,digits = rep(2,ncol(mat) + 1))
+  write(print(mat,sanitize.text.function = function (x) x,
+        floating = FALSE,hline.after = NULL, 
+        add.to.row = list(pos = list(-1,0,nrow(mat)), 
+                          command = c('\\toprule ',
                                       '\\midrule ',
-                                  '\\bottomrule '))
-      ),file=filename)
-
+                                      '\\bottomrule '))),file = filename)
 }
 
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/coverage_neg.tex")
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/coverage_pos.tex")
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/coverage_all.tex")
-
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/coverage_neg_nopen.tex")
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/coverage_pos_nopen.tex")
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/coverage_all_nopen.tex")
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n","ash.u","ash.hu"),
+                          "table/coverage_neg.tex")
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n","ash.u","ash.hu"),
+                          "table/coverage_pos.tex")
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n","ash.u","ash.hu"),
+                          "table/coverage_all.tex")
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n.s","ash.u.s","ash.hu.s"),
+                          "table/coverage_neg_nopen.tex")
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n.s","ash.u.s","ash.hu.s"),
+                          "table/coverage_pos_nopen.tex")
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n.s","ash.u.s","ash.hu.s"),
+                          "table/coverage_all_nopen.tex")
 ```
 
-These tables show 1-the left tail instead of the left tail.
-```{r}
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/scoverage_neg.tex",switch=TRUE)
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/scoverage_pos.tex",switch=TRUE)
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n","ash.u","ash.hu"),"table/scoverage_all.tex",switch=TRUE)
+These tables show right tail instead of the left tail.
 
-save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/scoverage_neg_nopen.tex",switch=TRUE)
-save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/scoverage_pos_nopen.tex",switch=TRUE)
-save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% summarize(lt = mean(value<coverthresh)),c("ash.n.s","ash.u.s","ash.hu.s"),"table/scoverage_all_nopen.tex",switch=TRUE)
+```{r write_latex_tables_right}
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n","ash.u","ash.hu"),
+                          "table/scoverage_neg.tex",switch = TRUE)
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n","ash.u","ash.hu"),
+                          "table/scoverage_pos.tex",switch = TRUE)
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n","ash.u","ash.hu"),
+                          "table/scoverage_all.tex",switch=TRUE)
+
+save_latex_coverage_table(reslong.neg %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n.s","ash.u.s","ash.hu.s"),
+                          "table/scoverage_neg_nopen.tex",switch = TRUE)
+save_latex_coverage_table(reslong.pos %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n.s","ash.u.s","ash.hu.s"),
+                          "table/scoverage_pos_nopen.tex",switch = TRUE)
+save_latex_coverage_table(reslong %>% group_by(scenario,method) %>% 
+                            summarize(lt = mean(value < coverthresh)),
+                          c("ash.n.s","ash.u.s","ash.hu.s"),
+                          "table/scoverage_all_nopen.tex",switch = TRUE)
 ```

### 2017-01-02 analysis/checkIP.Rmd
Successfully rendered checkIP.Rmd using new dsc-opt and dsc-shrink output.

@@
 
+**Note:** the last code chunk ("compare_lfsr") may take several minutes to 
+render in R or RStudio because it uses ggplot to plot more than a million data 
+points. Also, it it may not be possible to run the last code chunk 
+interactively in RStudio.
+
 First, we load the necessary libraries, and specify settings for
@@
 
-The 'dsc-opt' script simply runs both methods (IP and EM) on the same data sets,
-and compares the difference in the log-likelihoods. We can see that in >90% of 
-cases the difference is negligible. Then there is a tail of cases where the IP 
-log-likelihood is higher - presumably where the EM doesn't really converge so
-well.
+The 'dsc-opt' script simply runs both methods (IP and EM) on the same
+data sets, and compares the difference in the log-likelihoods. We can
+see that in >90% of cases the difference is negligible. Then there is
+a tail of cases where the IP log-likelihood is higher - presumably
+where the EM doesn't really converge so well.
 
@@
 
-However, what we really care about is inference quantities, such as the lfsr, 
-rather than the log-likelihood. Here we compare the lfsr for the simulations 
-where we ran both the IP method and EM method (nocxx) in dsc-shrink.
+However, what we really care about is inference quantities, such as
+the lfsr, rather than the log-likelihood. Here we compare the lfsr for
+the simulations where we ran both the IP method and EM method (nocxx)
+in dsc-shrink.
 
@@
 df = inner_join(df1,df2,by = c("scenario","seed","variable"))
-ggplot(df,aes(lfsr.nocxx,lfsr.IP)) + geom_point(shape = 1) +
+df = transform(df,scenario = factor(scenario))
+print(ggplot(df,aes(lfsr.nocxx,lfsr.IP)) + 
+      geom_point(shape = 1) +
       facet_wrap(~ scenario,nrow = 2) +
@@
       xlab("EM algorithm (no cxx)") +
-ylab("IP method")
+      ylab("IP method"))
 ```

### 2017-01-03 analysis/investigate_hu_badcoverage.Rmd
Added chunk descriptions and succesfully rendered investigate_hu_badcoverage.Rmd after fixing problems with code in /output.

@@
 library(dplyr)
+library(dscr)
 source("chunk-options.R")
@@
 
-```{r}
+Compile the tables which we will use to examine the results of the simulation
+experiments.
+
+```{r compile_tables}
 load("../output/dsc-shrink-files/res.RData")
 
-out <- res$negprob
 neglong = 
-  out %>% 
+  res$negprob %>% 
     select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    melt(id.vars=c("method","scenario","seed",".id"),value.name = "negprob") %>%
+    melt(id.vars = c("method","scenario","seed",".id"),value.name = "negprob") %>%
     filter(negprob > 0.95)
@@
     select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    melt(id.vars = c("method","scenario","seed",".id"),value.name="posprob") %>%
+    melt(id.vars = c("method","scenario","seed",".id"),value.name = "posprob") %>%
     filter(posprob > 0.95)
@@
 
+ash.hu.ft = reslong.pos %>% filter(method == "ash.hu.s" & 
+                                   scenario == "flat-top")
 ```
 
-```{r}
-ash.hu.ft = reslong.pos %>% dplyr::filter(method=="ash.hu.s" & scenario=="flat-top")
-head(ash.hu.ft)
+Extract an example to illustrate coverage of ash.hu.
+
+```{r show_example}
+eg = load_example(dsc_shrink,42,"flat-top",methodname = "ash.hu.s",
+                  homedir = "../code/dsc-shrink")
+out        <- eg$output$fitted_g
+class(out) <- "list"
+out        <- as.data.frame(out)
+print(subset(out,pi > 0.01))
 ```
 
-```{r}
-  eg=load_example(dsc_shrink,42,"flat-top",method="ash.hu.s","../code/dsc-shrink")
-  eg$output$fitted_g
+Notice how almost all the inferred weight is on a small positive component. 
+As a result, false sign rate will be small, and there will be a strong tendency
+to overestimate zero effects. This leads to coverage problems observed.
+
+Now let's look at an example with u and spiky which seems to be somewhat 
+badly calibrated for the negative discoveries.
+
+```{r show_more_examples}
+ash.u.s.spiky = reslong.neg %>% filter(method == "ash.u.s" & 
+                                       scenario == "spiky")
+ash.n.s.spiky = reslong.neg %>% filter(method == "ash.n.s" & 
+                                         scenario == "spiky")
+hist(ash.u.s.spiky$value,nclass = 100,
+     main = paste("histogram of quantile where observation falls in its CI"))
+hist(ash.n.s.spiky$value,nclass = 100,
+     main = paste("histogram of quantile where observation falls in its CI"))
 ```
 
-Notice how almost all the inferred weight is on a small positive component. As a result false sign rate will be small, and there will be a strong tendency
-to overestimate 0 effects. This leads to coverage problems observed.
+So what seems to be happening here is that the uniform tail is too short; 
+when observation falls outside of this tail it gets a zero quantile of 
+posterior interval.
 
-Now let's look at an example with u and spiky which seems to be somewhat badly calibrated for the negative discoveries. 
-```{r}
-ash.u.s.spiky = reslong.neg %>% dplyr::filter(method=="ash.u.s" & scenario=="spiky")
-ash.n.s.spiky = reslong.neg %>% dplyr::filter(method=="ash.n.s" & scenario=="spiky")
-hist(ash.u.s.spiky$value,nclass=100,main="histogram of quantile where observation falls in its CI; significant negative discoveries")
-hist(ash.n.s.spiky$value,nclass=100,main="histogram of quantile where observation falls in its CI; significant negative discoveries")
+Can we find an example illustrating this trend? (Note: the next chunk below
+did not work for me, so I se* `eval=FALSE, include=FALSE`. Perhaps this can be 
+fixed at some point. -Peter)
+
+```{r example_short_tail, eval=FALSE, include=FALSE}
+table((ash.u.s.spiky %>% filter(value == 0))$seed)
+eg = load_example(dsc_shrink,seed = 39,scenario = "spiky",
+                  methodname = "ash.u.s","../code/dsc-shrink")
 ```
 
-So what seems to be happening here is that the uniform tail is too short; when observation falls outside of this tail it gets a zero quantile of posterior interval.
+For comparison, here are the positive discoveries; here they are not too bad.
 
-Can we find an example?
-```{r}
-  table((ash.u.s.spiky %>% filter(value==0))$seed)
-  eg =load_example(dsc_shrink,seed = 39,method="ash.u.s",scenario="spiky","../code/dsc-shrink")
-```
-
-For comparison, here are the positive discoveries; here there is not too bad.
-```{r}
-ash.u.s.spiky = reslong.pos %>% dplyr::filter(method=="ash.u.s" & scenario=="spiky")
-ash.n.s.spiky = reslong.pos %>% dplyr::filter(method=="ash.n.s" & scenario=="spiky")
+```{r histograms_pos}
+ash.u.s.spiky = reslong.pos %>% filter(method == "ash.u.s" & scenario == "spiky")
+ash.n.s.spiky = reslong.pos %>% filter(method == "ash.n.s" & scenario == "spiky")
 hist(ash.n.s.spiky$value)

### 2017-01-03 analysis/plot_lfsr.Rmd
Succesfully re-knit plot_lfsr.Rmd, and switched to png format for the plots to reduce the size of the HTML output.

@@
 library(dplyr)
+library(reshape2)
 library(ggplot2)
@@
 
-TO DO: Add note here about having to first run script to generate data sets
-required to run the analysis.
+Load the results of the simulation experiments generated by 
+code/dsc-shrink/run_dsc.R, and prepare the posterior statistics for 
+plotting.
 
-```{r}
+```{r chunk_label}
 load("../output/dsc-shrink-files/res.RData")
@@
 #' @return tall dataframe with columns of scores for each method and the "goldmethod" against which plot is to be made
-process_score_for_plotting_against_gold=function(df,PLOTMETHODS=c("ash.n","ash.u","ash.hu"),
-                                                 GOLDMETHOD="bayes",PLOTSEEDS=1:100,
-                                                   PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal")
-  PLOTNAMES=PLOTSCENARIOS
+process_score_for_plotting_against_gold = 
+  function(df, 
+           PLOTMETHODS   = c("ash.n","ash.u","ash.hu"),
+           GOLDMETHOD    = "bayes",PLOTSEEDS = 1:100,
+           PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew",
+                             "big-normal","bimodal"),
+           PLOTNAMES     = PLOTSCENARIOS) {
   df %<>% filter(seed %in% PLOTSEEDS) %>% 
@@
     filter(method %in% c(PLOTMETHODS,GOLDMETHOD))
-  df$scenario = factor(df$scenario,levels=PLOTSCENARIOS)
-  levels(df$scenario)= PLOTNAMES
-  #create tall version of dataframe
-  df %<>% dplyr::select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
-    reshape2::melt(id.vars=c("method","scenario","seed",".id"),value.name="val")
+  df$scenario = factor(df$scenario,levels = PLOTSCENARIOS)
+  levels(df$scenario) = PLOTNAMES
+
+  # Create "tall"" version of dataframe.
+  df %<>% select(-user.self,-sys.self,-elapsed,-user.child,-sys.child) %>%
+    melt(id.vars = c("method","scenario","seed",".id"),value.name = "val")
+
   #separate bayes and remainder
-  df.bayes = df %>% filter(method==GOLDMETHOD)
-  df.rest = df %>% filter(method!=GOLDMETHOD)
+  df.bayes = df %>% filter(method == GOLDMETHOD)
+  df.rest  = df %>% filter(method != GOLDMETHOD)
 
-#join bayes with others, so each line has both the bayes and the non-bayes version
-  return(inner_join(df.bayes, df.rest, by=c("scenario","seed","variable")))
+  # Join bayes with others, so each line has both the bayes and the 
+  # non-bayes version.
+  return(inner_join(df.bayes,df.rest,by = c("scenario","seed","variable")))
 }
 
-plot_lfsr=function(lfsr,xlab="True lfsr",ylab="Estimated lfsr",xlim=c(0,0.2),ylim=c(0,0.2),legend.position="bottom"){
-  p=ggplot(lfsr,
-         aes(val.x,val.y,colour=method.y)) +
+plot_lfsr = function(lfsr,xlab = "True lfsr",ylab = "Estimated lfsr",
+                     xlim = c(0,0.2),ylim = c(0,0.2),
+                     legend.position = "bottom")
+  ggplot(lfsr,aes(val.x,val.y,colour = method.y)) +
     facet_grid(. ~ scenario) +
-    guides(alpha=FALSE) +
+    guides(alpha = FALSE) +
     geom_abline(colour = "black") +
-    geom_abline(colour= "red", slope=2) +
-    xlab(xlab) +
-    ylab(ylab) +
-    geom_point(shape=1,size=0.1,alpha=0.2)
+    geom_abline(colour = "red",slope = 2) +
+    xlab(xlab) + ylab(ylab) +
+    geom_point(shape = 1,size = 0.1,alpha = 0.2) +
+    scale_y_continuous(limits = ylim) +
+    scale_x_continuous(limits = xlim)
 
-  p +scale_y_continuous(limits=ylim) +
-        scale_x_continuous(limits=xlim)
+lfsr = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS = 1:100,
+                                               PLOTMETHODS = "ash.n")
+lfdr = process_score_for_plotting_against_gold(res$lfdr,PLOTSEEDS = 1:100,
+                                               PLOTMETHODS = "ash.n")
 
-}
-
-lfsr = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n")
-lfdr = process_score_for_plotting_against_gold(res$lfdr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n")
-
-p1=plot_lfsr(lfsr,ylim=c(0,1),xlim=c(0,0.2))
-p2=plot_lfsr(lfdr,ylim=c(0,1),xlim=c(0,0.2),xlab="True lfdr",ylab="Estimated lfdr")
-
-# cowplot command: plot_grid(p2+theme_gray()+theme(legend.position="none"),p1+theme_gray()+theme(legend.position="none"),nrow=2,align="v")
-
+p1 = plot_lfsr(lfsr,ylim = c(0,1),xlim = c(0,0.2))
+p2 = plot_lfsr(lfdr,ylim = c(0,1),xlim = c(0,0.2),
+               xlab = "True lfdr",ylab = "Estimated lfdr")
 ```
 
-```{r plot_lfsr, dev='pdf', fig.width=9, fig.height=3, crop=TRUE}
-print(p1+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
-      +coord_equal(ratio=1/5) + colScale)
-#ggsave("../paper/figures/lfsr_est.pdf",height=3,width=9)
+Separately for each model, create a scatterplot comparing the estimated LFSR 
+against the "gold-standard" LFSR.
+
+```{r plot_lfsr, dev='png', fig.width=9, fig.height=3, crop=TRUE}
+print(p1 + theme(legend.position = "none",
+                 axis.text.x = element_text(size = 8,angle = 45)) +
+      coord_equal(ratio = 1/5) + colScale)
 ```
 
-```{r plot_lfdr, dev='pdf', fig.width=9, fig.height=3, crop=TRUE}
-print(p2+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
-      +coord_equal(ratio=1/5) + colScale)
-#ggsave("../paper/figures/lfdr_est.png",height=3,width=9)
-#ggsave("../paper/figures/lfdr_est.pdf",height=3,width=9)
+Separately for each model, create a scatterplot comparing the estimated LFDR 
+against the "gold-standard" LFDR.
+
+```{r plot_lfdr, dev='png', fig.width=9, fig.height=3, crop=TRUE}
+print(p2 + theme(legend.position = "none",
+                 axis.text.x = element_text(size = 8,angle = 45)) +
+      coord_equal(ratio = 1/5) + colScale)
 ```
 
-```{r plot_lfsr_s, dev='pdf', fig.width=9, fig.height=3, crop=TRUE}
-lfsr.s = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n.s")
-p1.s=plot_lfsr(lfsr.s,ylim=c(0,1),xlim=c(0,0.2))
+Separately for each of the ash.n.s methods, create a scatterplot comparing the 
+estimated LFSR against the "gold-standard" LFSR.
 
-print(p1.s+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
-      +coord_equal(ratio=1/5))
-#ggsave("../paper/figures/lfsr_est_s.png",height=3,width=9)
-#ggsave("../paper/figures/lfsr_est_s.pdf",height=3,width=9)
+```{r plot_lfsr_s, dev='png', fig.width=9, fig.height=3, crop=TRUE}
+lfsr.s = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS = 1:100,
+                                                 PLOTMETHODS = "ash.n.s")
+p1.s   = plot_lfsr(lfsr.s,ylim = c(0,1),xlim = c(0,0.2))
+print(p1.s + theme(legend.position = "none",
+                   axis.text.x = element_text(size = 8,angle = 45)) +
+      coord_equal(ratio = 1/5))
 ```
 
-```{r plot_lfsr_s_nn, dev='pdf', fig.width=9, fig.height=3, crop=TRUE}
-lfsr.s.nn = process_score_for_plotting_against_gold(
-    res$lfsr,PLOTSEEDS=1:100,PLOTMETHODS="ash.n.s",
-    PLOTSCENARIOS=paste0(c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),"-nn"))
-p1.s.nn=plot_lfsr(lfsr.s.nn,ylim=c(0,1),xlim=c(0,0.2))
+Separately for each of the ash.n.s methods with the -nn data sets, 
+create a scatterplot comparing the estimated LFSR against the 
+"gold-standard" LFSR.
 
-print(p1.s.nn+theme(legend.position="none",axis.text.x = element_text(size = 8,angle=45))
-      +coord_equal(ratio=1/5))
-#ggsave("../paper/figures/lfsr_est_s_nn.png",height=3,width=9)
-
+```{r plot_lfsr_s_nn, dev='png', fig.width=9, fig.height=3, crop=TRUE}
+lfsr.s.nn = 
+  process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS = 1:100,
+    PLOTMETHODS = "ash.n.s",
+    PLOTSCENARIOS = paste0(c("spiky","near-normal","flat-top","skew",
+                             "big-normal","bimodal"),"-nn"))
+p1.s.nn = plot_lfsr(lfsr.s.nn,ylim = c(0,1),xlim = c(0,0.2))
+print(p1.s.nn + theme(legend.position = "none",
+                      axis.text.x = element_text(size = 8,angle = 45)) +
+      coord_equal(ratio = 1/5))
 ```

### 2017-01-03 analysis/plot_cdf_eg.Rmd
Now generate both PDFs and PNGs in plot_lfsr.Rmd; this keeps HTML file small while also creating PDFs. Thanks to JB for suggestion.

@@
 library(dscr)
+library(REBayes)
 source("chunk-options.R")
@@
 
-*Explain here what this chunk does.*
+TO DO: Explain here what this chunk does.
 
@@
 
-#df is a list with components method,seed, scenario
-#cdf is evaluated at x
-#returns list of x,y,pi0, where cdf values are in y
-get_cdf = function(df,dsc=dsc_shrink,x=seq(-6,6,length=100),homedir="../code/dsc-shrink"){
-  m=df$method
-  if(m=="truth"){m="ash.n"}
-  temp=load_example(dsc,df$seed,df$scenario,m,homedir)
-  pi0=temp$meta$pi0
+# df is a list with components method,seed, scenario
+# cdf is evaluated at x
+# returns list of x,y,pi0, where cdf values are in y
+get_cdf = function (df,dsc = dsc_shrink,x = seq(-6,6,length = 100),
+                    homedir = "../code/dsc-shrink") {
+  m = df$method
+  if (m == "truth")
+    m = "ash.n"
+  temp = load_example(dsc,df$seed,df$scenario,m,homedir)
+  pi0  = temp$meta$pi0
 
-  if(df$method=="truth"){
-    s=dsc$scenarios[[df$scenario]]
-    galt=s$args$g
-    g=normalmix(c(pi0,(1-pi0)*mixprop(galt)),
+  if (df$method == "truth"){
+    s    = dsc$scenarios[[df$scenario]]
+    galt = s$args$g
+    g    = normalmix(c(pi0,(1 - pi0) * mixprop(galt)),
                      c(0,comp_mean(galt)),
@@
   } else {
-    if(grepl("mixfdr",df$method)){
-      temp$output= mixfdr2fitted.g(temp$output)
+    if (grepl("mixfdr",df$method))
+      temp$output = mixfdr2fitted.g(temp$output)
+    g = temp$output$fitted.g
   }
-    g=temp$output$fitted.g
-  }
-  data.frame(x=x,y=as.numeric(mixcdf(g,x)),pi0=pi0)
+  return(data.frame(x = x,y = as.numeric(mixcdf(g,x)),pi0 = pi0))
 }
 
-plot_mean_cdf = function(SEEDS,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
-      PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","bimodal"),pi0filter=FALSE,...){
+plot_mean_cdf = 
+  function(SEEDS,
+           PLOTMETHODS   = c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
+           PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","bimodal"),
+           pi0filter     = FALSE,...) {
   PLOTNAMES = PLOTSCENARIOS
+
   #set up dataframe with cdf for all methods and all datasets
-  df= expand.grid(seed=SEEDS,scenario=PLOTSCENARIOS, method=PLOTMETHODS,stringsAsFactors = FALSE)
-  df.cdf=ddply(df,.(seed,scenario,method), get_cdf)
+  df     = expand.grid(seed = SEEDS,scenario = PLOTSCENARIOS,
+                       method = PLOTMETHODS,stringsAsFactors = FALSE)
+  df.cdf = ddply(df,.(seed,scenario,method),get_cdf)
 
-  if(pi0filter==TRUE){
-    df.cdf %<>% filter(pi0<0.55 & pi0>0.45)
-  }
-  if(length(SEEDS)>1){
-    df.cdf %<>% group_by(x,method,scenario) %>% dplyr::summarise(y=mean(y))
-  }
+  if (pi0filter)
+    df.cdf %<>% filter(pi0 < 0.55 & pi0 > 0.45)
+  if (length(SEEDS) > 1)
+    df.cdf %<>% group_by(x,method,scenario) %>% summarise(y = mean(y))
 
-  df.cdf$scenario=factor(df.cdf$scenario,levels=PLOTSCENARIOS)
-  levels(df.cdf$scenario)=PLOTNAMES
+  df.cdf$scenario         = factor(df.cdf$scenario,levels = PLOTSCENARIOS)
+  levels(df.cdf$scenario) = PLOTNAMES
 
-  ggplot(df.cdf,aes(x=x,y=y,color=method),...) + colScale + geom_line(lwd=1.5,alpha=0.7) + facet_grid(.~scenario) + theme(legend.position = "bottom")
-
+  return(ggplot(df.cdf,aes(x = x,y = y,color = method),...) + colScale + 
+    geom_line(lwd = 1.5,alpha = 0.7) + facet_grid(.~scenario) + 
+      theme(legend.position = "bottom"))
 }
 ```
+
+TO DO: Explain here what this chunk does.
 
@@
 
+TO DO: Explain here what this chunk does.
+
 ```{r mean_cdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
@@
 
+TO DO: Explain here what this chunk does.
+
 ```{r mean_cdf_nopen, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
-names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr","mixfdr.tnull")
+names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue",
+                     "locfdr","mixfdr.tnull")
 colScale <- scale_colour_manual(name = "method",values = myColors)
-plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n.s","ash.u.s","ash.hu.s","truth"),pi0filter=TRUE)
-source("../R/set_plot_colors.R") #reset color scale for other plots
+plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n.s","ash.u.s","ash.hu.s","truth"),
+              pi0filter = TRUE)
+
+# Reset color scale for other plots.
+source("../R/set_plot_colors.R") 
 ```
@@
 ```{r egcdf-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
-plot_mean_cdf(1,PLOTMETHODS=c("ash.n","truth"),PLOTSCENARIOS = c("spiky","near-normal","bimodal"))
+plot_mean_cdf(1,PLOTMETHODS = c("ash.n","truth"),
+              PLOTSCENARIOS = c("spiky","near-normal","bimodal"))
 ```
 
+TO DO: Explain here what this chunk does.
+
 ```{r mean_cdf-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
-plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n","truth"),PLOTSCENARIOS = c("spiky","near-normal","bimodal"),pi0filter=TRUE)
+plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n","truth"),
+              PLOTSCENARIOS = c("spiky","near-normal","bimodal"),
+              pi0filter = TRUE)
 ```
 
+TO DO: Explain here what this chunk does.
+
 ```{r mean_cdf_nopen-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
-names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr","mixfdr.tnull")
+names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr",
+                     "mixfdr.tnull")
 colScale        <- scale_colour_manual(name = "method",values = myColors)
-plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n.s","truth"),PLOTSCENARIOS = c("spiky","near-normal","bimodal"), pi0filter=TRUE)
+plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n.s","truth"),
+              PLOTSCENARIOS = c("spiky","near-normal","bimodal"),
+              pi0filter = TRUE)
 source("../R/set_plot_colors.R")
@@
 ```{r egcdf-reduce-with-npmle, dev='pdf',fig.height=4.5, fig.width=6, crop=TRUE}
-run_nplme=function(SEED=1,PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal")){
-  df=data.frame(seed=NULL,scenario=NULL,method=NULL,x=NULL,y=NULL,pi0=NULL)
-  for(SCENARIO in PLOTSCENARIOS){
-    temp=dscr::load_example(dsc_shrink,SEED,SCENARIO,"ash.n","../output/dsc-shrink-files/")
-    z <- REBayes::GLmix(temp$input$betahat)
+run_nplme = function(SEED = 1,
+                     PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew",
+                                       "big-normal","bimodal")) {
+  df = data.frame(seed = NULL,scenario = NULL,method = NULL,x = NULL,
+                  y = NULL,pi0 = NULL)
+  for(SCENARIO in PLOTSCENARIOS) {
+    temp = dscr::load_example(dsc_shrink,SEED,SCENARIO,"ash.n","../output/dsc-shrink-files/")
+    z    = REBayes::GLmix(temp$input$betahat)
     df = rbind(df,data.frame(seed=SEED,scenario=SCENARIO,method="NPMLE",x=z$x,y=cumsum(z$y)/sum(z$y),pi0=NA))

### 2017-01-03 analysis/plot_lfsr.Rmd
Now generate both PDFs and PNGs in plot_lfsr.Rmd; this keeps HTML file small while also creating PDFs. Thanks to JB for suggestion.

@@
 
-```{r plot_lfsr, dev='png', fig.width=9, fig.height=3, crop=TRUE}
+```{r plot_lfsr, dev=c('png','pdf'), fig.width=9, fig.height=3, crop=TRUE}
 print(p1 + theme(legend.position = "none",
@@
 
-```{r plot_lfdr, dev='png', fig.width=9, fig.height=3, crop=TRUE}
+```{r plot_lfdr, dev=c('png','pdf'), fig.width=9, fig.height=3, crop=TRUE}
 print(p2 + theme(legend.position = "none",
@@
 
-```{r plot_lfsr_s, dev='png', fig.width=9, fig.height=3, crop=TRUE}
+```{r plot_lfsr_s, dev=c('png','pdf'), fig.width=9, fig.height=3, crop=TRUE}
 lfsr.s = process_score_for_plotting_against_gold(res$lfsr,PLOTSEEDS = 1:100,
@@
 
-```{r plot_lfsr_s_nn, dev='png', fig.width=9, fig.height=3, crop=TRUE}
+```{r plot_lfsr_s_nn, dev=c('png','pdf'), fig.width=9, fig.height=3, crop=TRUE}
 lfsr.s.nn = 

### 2017-01-03 analysis/plot_cdf_eg.Rmd
Fixed bugs in function get_cdf in plot_cdf_eg.Rmd.

@@
 library(REBayes)
+library(ashr)
 source("chunk-options.R")
@@
   } else {
-    if (grepl("mixfdr",df$method))
-      temp$output = mixfdr2fitted.g(temp$output)
-    g = temp$output$fitted.g
+    if (grepl("mixfdr",df$method)) {
+      temp$output$fitted_g = mixfdr2fitted.g(temp$output)$fitted.g
+    }
+    g = temp$output$fitted_g
   }
@@
 ```{r mean_cdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
-plot_mean_cdf(1:100,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth"),pi0filter=TRUE)
+plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n","ash.u","ash.hu","truth"),
+              pi0filter = TRUE)
 ```
@@
 
-```{r egcdf-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
+```{r egcdf-reduce, fig.height=3, fig.width=6, crop=TRUE, dev='pdf'}
 plot_mean_cdf(1,PLOTMETHODS = c("ash.n","truth"),
@@
 
-```{r mean_cdf_nopen-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
+```{r mean_cdf_nopen-reduce, fig.height=3, fig.width=6, crop=TRUE, dev='pdf'}
 names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr",
@@
 
+TO DO (maybe): Add more details here about this chunk?
+
 ```{r egcdf-reduce-with-npmle, dev='pdf',fig.height=4.5, fig.width=6, crop=TRUE}
@@
   for(SCENARIO in PLOTSCENARIOS) {
-    temp = dscr::load_example(dsc_shrink,SEED,SCENARIO,"ash.n","../output/dsc-shrink-files/")
-    z    = REBayes::GLmix(temp$input$betahat)
-    df = rbind(df,data.frame(seed=SEED,scenario=SCENARIO,method="NPMLE",x=z$x,y=cumsum(z$y)/sum(z$y),pi0=NA))
+    temp = load_example(dsc_shrink,SEED,SCENARIO,"ash.n","../output/dsc-shrink-files/")
+    z    = GLmix(temp$input$betahat)
+    df = rbind(df,data.frame(seed = SEED,scenario = SCENARIO,method = "NPMLE",
+                             x = z$x,y = cumsum(z$y)/sum(z$y),pi0 = NA))
   }
@@
 
-plot_mean_cdf_with_npmle = function(SEED=1,PLOTMETHODS=c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
-      PLOTSCENARIOS=c("spiky","near-normal","flat-top","skew","big-normal","bimodal"),pi0filter=FALSE,...){
-  if(length(SEED)>1){stop("plot with npmle only implemented for a single seed")}
-  PLOTNAMES=PLOTSCENARIOS
-  #set up dataframe with cdf for all methods and all datasets
-  df= expand.grid(seed=SEED,scenario=PLOTSCENARIOS, method=PLOTMETHODS,stringsAsFactors = FALSE)
-  df.cdf=plyr::ddply(df,.(seed,scenario,method), get_cdf)
+plot_mean_cdf_with_npmle = 
+  function(SEED          = 1,
+           PLOTMETHODS   = c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
+           PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew",
+                             "big-normal","bimodal"),
+           pi0filter     = FALSE,...) {
+  if (length(SEED) > 1)
+    stop("plot with npmle only implemented for a single seed")
+  PLOTNAMES = PLOTSCENARIOS
 
-  df.cdf$scenario=factor(df.cdf$scenario,levels=PLOTSCENARIOS)
-  levels(df.cdf$scenario)=PLOTNAMES
+  # Set up dataframe with cdf for all methods and all datasets.
+  df = expand.grid(seed = SEED,scenario = PLOTSCENARIOS,method = PLOTMETHODS,
+                   stringsAsFactors = FALSE)
+  df.cdf = ddply(df,.(seed,scenario,method),get_cdf)
+
+  df.cdf$scenario         = factor(df.cdf$scenario,levels = PLOTSCENARIOS)
+  levels(df.cdf$scenario) = PLOTNAMES
 
@@
 
-  p=ggplot(df.cdf,aes(x=x,y=y,color=method),...) +  xlim(c(-4,4)) + colScale + geom_line(lwd=1.2,alpha=0.7)  + theme(legend.position = "bottom")
+  p = ggplot(df.cdf,aes(x = x,y = y,color = method),...) + 
+    xlim(c(-4,4)) + colScale + geom_line(lwd = 1.2,alpha = 0.7) + 
+    theme(legend.position = "bottom")
   return(p)
@@
 
-plot_mean_cdf_with_npmle(1,PLOTMETHODS=c("truth","ash.hu"),PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","big-normal","bimodal")) + facet_wrap(~ scenario, nrow=2)
+plot_mean_cdf_with_npmle(1,PLOTMETHODS = c("truth","ash.hu"),
+                         PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew",
+                                           "big-normal","bimodal")) + 
+  facet_wrap(~ scenario, nrow = 2)
 ```

### 2017-01-03 analysis/plot_cdf_eg.Rmd
Successfully rendered plot_cdf_eg.Rmd, but only after dropping the 'big-normal' scenarios from the plots; it seems that the big-normal results are no longer included.

@@
 
-TO DO: Explain here what this chunk does.
+Define a couple functions that compile CDF data in such a way that it is 
+easy to plot the CDFs using ggplot2.
 
@@
 
-  if (df$method == "truth"){
+  if (df$method == "truth") {
     s    = dsc$scenarios[[df$scenario]]
@@
 
-TO DO: Explain here what this chunk does.
+Show the CDFs for all methods and all scenarios, based on a single data set.
 
@@
 
-TO DO: Explain here what this chunk does.
+Show the CDFs for all scenarios, this time averaged over 100 simulated data 
+sets.
 
@@
 
-TO DO: Explain here what this chunk does.
+Same as above, but with custom colours.
 
@@
 
-TO DO: Explain here what this chunk does.
+Same as above, but this time the CDFs are averaged over 100 simulated data sets.
 
@@
 
-TO DO: Explain here what this chunk does.
+Now with custom colours.
 
@@
 
-I wanted to add the npmle to the plot...
-
-TO DO (maybe): Add more details here about this chunk?
+I wanted to add the npmle method (fit using the GLmix function) to the plot...
 
 ```{r egcdf-reduce-with-npmle, dev='pdf',fig.height=4.5, fig.width=6, crop=TRUE}
-run_nplme = function(SEED = 1,
-                     PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew",
-                                       "big-normal","bimodal")) {
+run_nplme = function (SEED = 1,PLOTSCENARIOS = c("spiky","near-normal",
+                                                 "flat-top","skew","bimodal")) {
   df = data.frame(seed = NULL,scenario = NULL,method = NULL,x = NULL,
                   y = NULL,pi0 = NULL)
-  for(SCENARIO in PLOTSCENARIOS) {
+  for (SCENARIO in PLOTSCENARIOS) {
     temp = load_example(dsc_shrink,SEED,SCENARIO,"ash.n","../output/dsc-shrink-files/")
@@
            PLOTMETHODS   = c("ash.n","ash.u","ash.hu","truth","mixfdr.tnull"),
-           PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew",
-                             "big-normal","bimodal"),
+           PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew","bimodal"),
            pi0filter     = FALSE,...) {
@@
 
-  df.npmle = run_nplme(SEED, PLOTSCENARIOS)
+  df.npmle = run_nplme(SEED,PLOTSCENARIOS)
   df.cdf   = rbind(df.cdf,df.npmle)
@@
 
-plot_mean_cdf_with_npmle(1,PLOTMETHODS = c("truth","ash.hu"),
-                         PLOTSCENARIOS = c("spiky","near-normal","flat-top","skew",
-                                           "big-normal","bimodal")) + 
+plot_mean_cdf_with_npmle(1,PLOTMETHODS = c("ash.hu","truth"),
+                         PLOTSCENARIOS = c("spiky","near-normal","flat-top",
+                                           "skew","bimodal")) +
   facet_wrap(~ scenario, nrow = 2)

### 2017-01-15 analysis/adaptive_shrinkage.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered adaptive_shrinkage.Rmd, ash_npmle.Rmd, checkIP.Rmd, check_mixfdr_lownoise.Rmd, efron.fcr.Rmd, estimate_mode.Rmd and template.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(REBayes)
@@
 library(scales)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/ash_npmle.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered adaptive_shrinkage.Rmd, ash_npmle.Rmd, checkIP.Rmd, check_mixfdr_lownoise.Rmd, efron.fcr.Rmd, estimate_mode.Rmd and template.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries.
+
+```{r packages}
+library(REBayes)
+library(ashr)
+```
 
 ```{r chunk_options, include=FALSE}
-library(REBayes)
-library(ashr)
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/checkIP.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered adaptive_shrinkage.Rmd, ash_npmle.Rmd, checkIP.Rmd, check_mixfdr_lownoise.Rmd, efron.fcr.Rmd, estimate_mode.Rmd and template.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(ggplot2)
@@
 library(reshape2)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/efron.fcr.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered adaptive_shrinkage.Rmd, ash_npmle.Rmd, checkIP.Rmd, check_mixfdr_lownoise.Rmd, efron.fcr.Rmd, estimate_mode.Rmd and template.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(ashr)
@@
 library(locfdr)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/estimate_mode.Rmd
Rendered referee_uaza.html and fixed chunk-options label to chunk_options.
Split setting chunk options and loading packages into separate chunks; re-rendered adaptive_shrinkage.Rmd, ash_npmle.Rmd, checkIP.Rmd, check_mixfdr_lownoise.Rmd, efron.fcr.Rmd, estimate_mode.Rmd and template.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries.
 
-```{r chunk-options, include=FALSE}
+```{r packages}
 library(ashr)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/investigate_hu_badcoverage.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 library(dscr)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/make_GOODPOOR_figs.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-First, we load the necessary libraries and function definitions, and specify 
-settings for displaying the plots in the rendered document.
+First, we load the necessary libraries and function definitions.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(ashr)
@@
 library(ggplot2)
+source("../R/plot_FDReg_hist.R")
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")
-source("../R/plot_FDReg_hist.R")
 ```

### 2017-01-15 analysis/makefig_FDReg.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries and function definitions.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(ashr)
@@
 library(mixfdr)
-source("chunk-options.R")
 source("../R/plot_FDReg_hist.R")
 source("../R/nullalthist.R")
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
+source("chunk-options.R")
 ```

### 2017-01-15 analysis/metaplot_examples.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(knitr)
@@
 library(ashr)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/plot_cdf_eg.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries and useful function definitions.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(plyr)
@@
 library(ashr)
-source("chunk-options.R")
 source("../code/dsc-shrink/methods/mixfdr.wrapper.R")
 source("../R/set_plot_colors.R")
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
+source("chunk-options.R")
 ```

### 2017-01-15 analysis/plot_egdens.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-```{r chunk_options, include=FALSE}
+First, we load the necessary libraries.
+
+```{r packages}
 library(ashr)
@@
 library(dplyr)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/plot_lfsr.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries and other useful function definitions.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(dplyr)
@@
 library(ggplot2)
+source("../R/set_plot_colors.R")
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")
-source("../R/set_plot_colors.R")
 ```

### 2017-01-15 analysis/plot_pi0est.Rmd
Successfully re-knit plot_pi0est.Rmd.
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(magrittr)
 library(dplyr)
-source("chunk-options.R")
 source("../R/set_plot_colors.R")
@@
 
-*Explain here what this chunk does.*
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
+source("chunk-options.R")
+```
+
+Generate scatterplots compare the estimated estimate of pi0 against the 
+ground-truth value, colored by the method used, and split by the scenario
+type.
 

### 2017-01-15 analysis/summarize_coverage.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries.
 
-```{r chunk_options, include=FALSE}
+```{r packages}
 library(dscr)
@@
 library(xtable)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/summarize_dsc_znull.Rmd
Split setting chunk options and loading packages into separate chunks; re-rendered investigate_hu_badcoverage.Rmd, make_GOODPOOR_figs.Rmd, makefig_FDReg.Rmd, metaplot_examples.Rmd, plot_cdf_eg.Rmd, plot_egdens.Rmd, plot_lfsr.Rmd and plot_pi0est.Rmd.

@@
 
-First, we load the necessary libraries, and specify settings for
-displaying the plots in the rendered document.
+First, we load the necessary libraries.
+
+```{r packages}
+library(ggplot2)
+library(dplyr)
+```
 
 ```{r chunk_options, include=FALSE}
-library(ggplot2)
-library(dplyr)
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-01-15 analysis/IPvsEM.Rmd
Re-rendered summarize_coverage.Rmd and summarize_dsc_znull.Rmd after updates given in above commits.

@@
 library(ggplot2)
+```
+
+```{r chunk_options, include=FALSE}
+# Specify settings for displaying the plots in the rendered document.
 source("chunk-options.R")

### 2017-04-18 analysis/plot_cdf_eg.Rmd
Fixed plots in plot_cdf_eg.Rmd so that they render as PNGs.

@@
 
-```{r egcdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
+```{r egcdf, fig.height=3, fig.width=9}
+# These chunk options were used to create the figure for the paper:
+# dev='pdf', fig.height=3, fig.width=9, crop=TRUE
 plot_mean_cdf(1)
@@
 
-```{r mean_cdf, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
+```{r mean_cdf, fig.height=3, fig.width=9}
+# These chunk options were used to create the figure for the paper:
+# dev='pdf', fig.height=3, fig.width=9, crop=TRUE
 plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n","ash.u","ash.hu","truth"),
@@
 
-```{r mean_cdf_nopen, dev='pdf',fig.height=3, fig.width=9, crop=TRUE}
+```{r mean_cdf_nopen, fig.height=3, fig.width=9}
+# These chunk options were used to create the figure for the paper:
+# dev='pdf', fig.height=3, fig.width=9, crop=TRUE
 names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue",
@@
 
-```{r egcdf-reduce, fig.height=3, fig.width=6, crop=TRUE, dev='pdf'}
+```{r egcdf-reduce, fig.height=3.5, fig.width=7}
+# These chunk options were used to create the figure for the paper:
+# fig.height=3, fig.width=6, crop=TRUE, dev='pdf'
 plot_mean_cdf(1,PLOTMETHODS = c("ash.n","truth"),
@@
 
-```{r mean_cdf-reduce, dev='pdf',fig.height=3, fig.width=6, crop=TRUE}
+```{r mean_cdf-reduce, fig.height=3.5, fig.width=7}
+# These chunk options were used to create the figure for the paper:
+# dev='pdf', fig.height=3, fig.width=6, crop=TRUE
 plot_mean_cdf(1:100,PLOTMETHODS = c("ash.n","truth"),
@@
 
-```{r mean_cdf_nopen-reduce, fig.height=3, fig.width=6, crop=TRUE, dev='pdf'}
+```{r mean_cdf_nopen-reduce, fig.height=3.5, fig.width=7}
+# These chunk options were used to create the figure for the paper:
+# fig.height=3, fig.width=6, crop=TRUE, dev='pdf'
 names(myColors) <- c("truth","ash.hu.s","ash.n.s","ash.u.s","qvalue","locfdr",
@@
 
-```{r egcdf-reduce-with-npmle, dev='pdf',fig.height=4.5, fig.width=6, crop=TRUE}
+```{r egcdf-reduce-with-npmle, fig.height=5.25, fig.width=7}
+# These chunk options were used to create the figure for the paper:
+# dev='pdf', fig.height=4.5, fig.width=6, crop=TRUE
 run_nplme = function (SEED = 1,PLOTSCENARIOS = c("spiky","near-normal",
